### 2013-11-21 manuscript/first draft/ms_main.tex
Initial commit

\section{Abstract}

\section{Introduction}

Arid rangelands are at high risk of sudden desertification. This process of persistent vegetation loss under livestock grazing is hardly understood but evidently an issue of the system's bistability. Under particular environmental conditions and grazing regimes, two ecosystem states are likewise resilient to perturbations: One state with a robust vegetation cover of shrubs and grasses as well as a desert with no significant vegetation cover (Scheffer 2001, Aguiar and Sala 1999). After crossing critical thresholds, or tipping-points, the system transites between the two stable states.

For arid ecosystems, the major mechanism behind this bistability was found to be local facilitation of plant individuals. The presence of a plant provides its direct local environment with abilities of water retention, organic matter accumulation and protection against external stressors. As a consequence, the vegetation grows in patches (Aguiar and Sala 1999).  When implemented in a simple spatially explicit model, this positive neighborhood effect provides true bistability on the landscape scale (K\'efi et al 2007a,b).

This parent model was used before to investigate the consequences of increased plant mortality. This simple formulation of additional plant stress can be interpreted as the effects of increased grazing intensity by large herbivores (K\'efi et al 2007a). However, the model is lacking some important features of grazing: the additional individual plant mortality (1) did not vary with vegetation density, and (2) it applied globally, affecting all plants likewise. 
Both assumptions are justified by the need for model simplicity, but disallow a translation of the model parameters to real livestock rates. The relationship between livestock rates and the stability of vegetation cover is a crucial information for management decisions. The present study aims at increasing the realism and applicability of the model by applying two extensions that enable simulations of a grazing gradient on drylands.

%Grazing depends on vegetation cover
The first extension adds a contrasting assumption on the plant density dependence of feeding to the parent model. The prior model defined grazing as a constant risk of additional plant death. On the population level, grazing mortality increased with vegetation cover (K\'efi et al. 2007). However, grazers forage in the landscape and consume to fulfill their daily demands. In other words, the number of plants suffering from grazing is rather constant with vegetation cover. As a consequence, at low vegetation cover the grazing risk on individual plants will be high, whereas when vegetation cover is high, risk will be shared among many plants.
In reality, the total feeding of large herbivores in the landscape is not entirely independent of vegetation cover but follows a non-linear saturating curve (type-II functional response, Holling 1959, Spalinger and Hobbs 1992). Thus, particularly at low vegetation cover grazing is limited by the search success of the grazer, which is resource-density dependent. In rangeland systems, however, the functional response at low vegetation cover depends on factors on a larger scale: the animals move between landscapes motivated by internal decision patterns or by external management decisions (Senft 1989). We assume therefore, that smooth type-II functional responses, as described for wild ungulates, are unlikely to happen in intensive rangelands within the spatial scale adressed in the present study.

Because of these uncertainties and for reasons of simplicity we contrast the parent assumption of density independent individual risk with density independent global feeding. We expect constant global feeding to consolidate both stable states: the desert state will be more resilient against recolonisation of single plants while the vegetated state will provide better protection to the plant individual. As a consequence, the reversibility of catastrophic shifts is expected to deteriorate.

%Spalinger and Hobbs 1992

%Grazing is spatialy differentiated
The second extension implements a local differentiation of grazing pressure. %For some time, the relationship between vegetation structure and grazing is acknowledged to be interdependent. Grazing affects the mortality of plants as well as plant recruitment, soil suitability, and other factors in positive as well as negative ways. Vegetation cover is a 
%Senft et al 1987
%Bailey 1996
Particularly, the importance of direct neighborhood interactions on the individual plant scale (local scale) in explaining spatial structure was demonstrated by Adler et al. (2001). The authors showed in a theoretical framework similar to the present study, that positive local interactions are superior to environmental heterogeneity or heterogeneous stress in explaining the patterns of plant spatial heterogeneity (Adler et al 2001). 
To differentiate the grazing locally, we implement an associative protection of plant individuals growing next to each other. This mechanism is observed frequently among shrubs in dryland ecosystems and considered to be a major driver of spatial heterogeneity of stress on local scales.

 %Bisigato et al 2005 Ecography
%Milchunas and Noy-Meir 2002

Here, the individual vulnerability against grazing is highest for plants growing isolated. The more direct neighbors a plant associates to, the lower is the effect of grazing. A plant in the patch center is not affected by grazing. We expect this mechanism to benefit the desert state, because at low densities only few plants are associated.

% comparison and main objectives/questions

 We provide a full factorial comparison of the consequences of these alternative models on vegetation structure and the ecosystems' bistability properties.

\section{Methods}
We used a cellular automaton to investigate the consequences of different grazing assumptions on the structural properties of the vegetation. This model was used before (Kefi et al 2007a, b) and can be resolved numerically in spatially explicit simulations. It defines landscape as a grid of cells, each can potentially take one of three discrete cell states: 'vegetated' cells are occupied by a plant individual (annotated as '+' in equations; black cells in figures); 'enriched' cells are suitable for seeds to germinate and colonise the cell, but do not contain adult plants right now  ('0' , grey cells); 'degraded' cells represent bare ground with lacking organic matter and bad water retention, and therefore can not be colonised by arriving seeds  ('$-$', white cells).
Transitions of cells states are only possible between vegetated and enriched (plant death / recolonisation) as well as enriched and degraded (degradation / regeneration). In biological terms, a degraded spot needs to be enriched first, before a plant can grow. Vice versa, when a plant dies, it leaves the spot in an enriched state, which might become degraded later. The rates for these transitions are defined in the following paragraphs. They might be constant values or functions of the global or local plant vegetation density (= $\rho_+$, global vegetation cover; $n_+$) .

\subsection{Facilitation model}
The original model by K\'efi et al mimics local facilitation of plants in drylands by defining the regeneration rate of degraded cells, $w_{ \left\{-,0 \right\} }$, as dependent on the density of vegetated cells in the nearest neighborhood, $\nu_+$ (assessing 'von Neumann'-neighborhood of range 1, i.e. the 4 nearest cells, $\nu_+ \in \left\{ 0, 0.25, 0.5, 0.75, 1 \right\}; $ K\'efi et al 2007b).
\begin{equation}
	w_{ \left\{-,0 \right\} } = r + \nu_{+} f
\end{equation}
% this is defining the rates of change of the population of cells in state '-'. 
%express rates rather as "probability of cell i,j to change into state x given that it is in state y".
The recolonisation of enriched cells takes into account that the majority of seeds arriving on a cell are from the plants in the direct neighborhood (local seed dispersal), while the probability of a successful plant establishment is limited by the competition on global resources.
\begin{equation}
	w_{ \left\{0,+ \right\} } = \left( \delta\rho_+ + \left( 1 - \delta \right)q_{+|0}\right) \left(b-c\rho_+ \right)
\end{equation}

%detailled description of parameters
The degradation of enriched cells is defined as a constant rate	
\begin{equation}
w_{ \left\{0,- \right\} } = d.
\label{eq:}
\end{equation}
Finally, in the original model, the intrinsic mortality of vegetated cells also is defined as constant $w_{ \left\{+,0 \right\} } = m$. Since it is defined as the probability of plant death per year, it also can be interpreted as inverse of the average lifespan of plants. However, the authors assume that this mortality might be increased by external stressors such as grazing. In the following paragraph, we explore different assumptions on how grazing affects the mortality of plant individuals.

\subsection{Grazer Models}
The \textbf{parent model} was used to explore consequences of increased plant mortality due to grazing. It simply adds grazing as a constant to individual mortality rate, affecting all plants homogenously.
\begin{equation}
	w_{ \left\{ +,0 \right\} }  = m_0 + g
\end{equation}
Therefore, $g$ .

On the landscape scale, this means that vegetation loss due to grazing scales linearly with plant vegetation cover (Fig. 1a). However, while some stressors might change the intrinsic mortality of all plant individuals, independendly of the vegetation cover (e.g. fungal pests, drought), this is certainly not true for grazing.

Instead, in reality the impact of grazers varies strongly with plant vegetation cover. In the \textbf{livestock model}, we approximate the behaviourial responses of grazers to vegetation cover by the following assumption: The number of plants dying due to grazing should correspond to the biomass consumed by a number of grazers in the landscape, while being independent from vegetation cover  (Fig. 1b). Thus, individual risk is high, when few plants are present in the landscape and it is low for high vegetation cover. The correlation between vegetation density and individual risk is as follows. 
\begin{equation}
	w_{ \left\{ +,0 \right\} }  = m_0 + g_0 / \rho_+
\end{equation}
The two models therefore produce differing mortalities in cases departing from  $ \rho_+ = g_0 / g$. For densities $ \rho_+ < g_0 / g$, the livestock model produces higher mortality than the parent model. For $ \rho_+ > g_0 / g$, the opposite is the case. Since density is dynamic over time, with attractors depending on total individual ...

To compare the two models, we define them to have equal mortality at intermediate vegetation cover, $\rho_+ = 0.5$. Therefore, in the parent model $g$ is set to be $g_0/0.5$.

\subsection{Associative Protection}
Plants benefit from associative growth in an environment with physical stressors like grazing. Shrubs in particular shield each other from grazers when growing in direct neighborhood. Therefore plants growing isolated suffer more from grazing than plants growing within patches.
The models from the previous paragraph do not account for this locally differentiated impacts of grazing. To add \textbf{associative protection} to the model we need to define individual vulnerability, 
\begin{equation}
v = 1 - \nu_+ ,
\end{equation}
to decrease with the density of vegetated neighboring cells $\nu_+$. It is highest for isolated plants with no neighbors ($\nu_+ = 0$) and decreases linearly with the density of vegetated neighboring cells. A plant with four vegetated neighbors ($\nu_+ = 1$) is not vulnerable to grazing. 
To keep global grazing in accord to the parent model or the livestock model we normalize vulnerability against the average vulnerability of all vegetated cells,

\begin{equation}
\widehat{v} =  \frac{ \sum\limits_i{ 1 - {\nu_+i}}  } {n_+} ,
\label{eq:vul}
\end{equation}
yielding
\begin{equation}
	w_{ \left\{ +,0 \right\} }  = m_0 + g_0 / \rho_+ \frac{v}{\widehat{v}}.
\label{eq:}
\end{equation}

Therefore, we compare two spatial variants of grazing (spatially homogenous \textit{vs.} associative protection) for each grazer model (parent model with constant risk \textit{vs.} lifestock model with density dependent risk; see Table 1).

\begin{table}[t!]
\label{tab:models}
\caption{the models' substitutions for grazing,  $g$, in the mortality term $w_{ \left\{ +,0 \right\} }  = m_0 + g$.  }
\centering
\begin{tabular}{ccc}

\toprule
 & parent model & lifestock model \\ \cmidrule(rl){2-2} \cmidrule(rl){3-3}
homogenous grazing &  $g_0/0.5$ & $g_0/\rho_+$\\
associative protection & $g_0/0.5 \frac{v}{\hat{v}}$  & $g_0/\rho_+ \frac{v}{\hat{v}}$ \\
	\bottomrule
\multicolumn{3}{p{9.5cm}}{\footnotesize $m_0$ : intrinsic mortality rate of vegetation, i.e. inverse of average lifespan, $g_0$: grazing intensity, $\rho_+$ : vegetation cover , $v$ : vulnerability, $\hat{v}$: mean vulnerability of all vegetated cells }
	\end{tabular}
\end{table}
%$n_+$ : number of vegetated cells on the lattice \par
%$\rho_+$ : global density of vegetated cells \par
%$\nu_n$ : density of neighbors in state n \par
%$g$ : grazing pressure

\subsection{Numerical simulations}

We applied these rules on a grid of 100 $\times$ 100 cells of 0.25m$^2$. Each timestep is defined as one year. The applicable transition probabilities for each single cell given in equations 1, 2, 3 and 4 are compared against uniform random numbers between 0 and 1 to determine if a transition occurs or if the cell remains unchanged. To initialize the grid with a standardized patchy vegetation structure, each model simulation was run for 500 years without any grazing (g = 0) using an initial environmental value $b_ini \in \left\{ 0.25, 3, 0.8, 0.9 \right\}$ before grazing is taking action. %(Note: I want to complete this to a highly resoluted gradient, to investigate the unstable equilibria). 
Starting from the preconditioned grids, the four grazing models (Table 1) with varying grazing intensity ($g \in \left{ 0, 0.025, 0.05, 0.075 , 0.10 \right}$) were run over a gradient of environmental quality ($b \in \left{ 0.2, ... , 1.0 \right}$) for 1.500 timesteps.

An average vegetation cover, $\rho_+$, as well as the average cell mortality,$\hat{m}$ , is determined as the average density of all grids of the final 500 timesteps.

\subsection{Choice of parameters}

To investigate the bistability properties of the vegetation cover, we simulate a gradient in environmental quality b ranging from 0.2 to 1 with a step-width of 0.025. The models of this study inherit all parameters from the parent model in K\'efi et al. 2007a ($r = 0.01$, $f = 0.9$, $\delta = 0.1$, $c = 0.2$, $d = 0.1$), except the ones regarding mortality. We assume intrinsic mortality m0 for all models to be 0.5, re
ecting an average individual lifespan of 20 years, if no additional mortality due to grazing is taking effect. We investigate, how the stability properties and patch size distributions change over a grazing gradient.

\section{Results}

\newpage

\begin{figure}[h]%
\includegraphics[width=\columnwidth]{figures/stability.pdf}%
\caption{Stable states of the vegetation cover, $\rho_+$, over gradient of environmental quality, $b$, under different grazing models. a) parent model with static, global grazing mortality, $g_0/0.5$. b) livestock model with density dependend, global grazing mortality, $g_0 / \rho_+$. c) associative protection model with differentiated grazing at patch border, $g_0/0.5 \frac{v}{\hat{v}}$ . d) livestock \& associative protection model with density dependend, differentiated grazing at patch border,  $g_0/ \rho_+ \frac{v}{\hat{v}}$ . Gradient of grazing intensity from grey: $g_0 = 0$, black to red: $g_0 \in \left\{ 0.025, 0.05, 0.075, 0.1 \right\}$.}%
\label{}%
\end{figure}

\begin{figure}[h]%
\includegraphics[width=\columnwidth]{figures/patterns.pdf}%
\caption{Stable states of the vegetation cover, $\rho_+$, over gradient of environmental quality, $b$, under different grazing models. a) parent model with static, global grazing mortality, $g_0/0.5$. b) livestock model with density dependend, global grazing mortality, $g_0 / \rho_+$. c) associative protection model with differentiated grazing at patch border, $g_0/0.5 \frac{v}{\hat{v}}$ . d) livestock \& associative protection model with density dependend, differentiated grazing at patch border,  $g_0/ \rho_+ \frac{v}{\hat{v}}$ . Gradient of grazing intensity from grey: $g_0 = 0$, black to red: $g_0 \in \left\{ 0.025, 0.05, 0.075, 0.1 \right\}$.}%
\label{}%
\end{figure}

\section{Discussion}

Caveats
Functional response of large herbivores; Individual based, informed foraging

In reality, however, the functional response of large grazers on vegetation, particularly at low vegetation cover, is dynamic. In rangelands, management decisions and spatial responses of the grazers will reduce feeding pressure at very low densities.

In fact, the local effects of large herbivores on plants are manifold and qualitatively different. While direct feeding reduces biomass in the first place, it might lead to growth enhancement and increased productivity. On the local scale, vegetation benefit from the presence of herbivores through fertilization, inoculation with seeds, 
In our model, however, we simplify these factors to a linear negative impact of grazer presence on the local vegetation. The

\section{Conclusion}
Undifferentiated model underestimates consequences of grazing with regard to the stability properties of the landscape. 
A linear increase in grazing causes shifts much

\section{APPENDIX}

\subsection{simulations}

The grid was defined to have periodic borders, connecting the east border to the west, the north to the south.
Prerun for 100 years. 
during pre-run, environmental conditions $b_ini \in \left\{ 0.250, 0.274, 0.300, 0.329, 0.360, 0.395, 0.433, 0.474, 0.520, 0.570, 0.624, 0.684, 0.749, 0.821, 0.900 \right\} $ emerge a patch structure ranging from low to high vegetation cover. Trajectories towards a stable vegetated state or towards desert are documented and used to approximate the unstable equilibrium.

, the whole grid is updated synchronously multiple times, to increase stochasticity
each time step is devided by the number of updates per year ( = 5) and

\subsection{model fitting}
 Four models were compared on each snapshot of t

\begin{table}[t!]
\label{tab:fitmodels}
\caption{ Models describing patch size distribution }
\centering
\begin{tabular}{lll}

\toprule
 & function & log-transform \\ \cmidrule(rl){2-2} \cmidrule(rl){3-3}
Power Law &  $n = a s^{- \alpha}$ & $ \log{n} =\log{a} - \alpha \log s$\\
Limited Power Law &  $n = b + a s^{- \alpha} $ & $ \log{n} = \log a - \alpha \log{s} + \log(1+\frac{b}{a s^{-\alpha} })$\\
Truncated Power Law & $n = a s^{- \alpha} \mathrm{e}^{\frac{s}{S_x}}$  & $\log{n} = \log a  - \alpha  \log s  - \frac{s}{Sx}  $ \\
Exponential &  $n = a \mathrm{e}^{-\varepsilon s} $ & $ \log{n} =  \log{a} - \varepsilon s $\\
	\bottomrule
\multicolumn{3}{p{12.5cm}}{\footnotesize  }
	\end{tabular}
\end{table}

table with log-transformed models

### 2013-11-21 manuscript/manuscript/ms_main.tex
Initial commit

--- 2013-11-21 manuscript/first draft/ms_main.tex
@@
 
-Arid rangelands are at high risk of sudden desertification.
-This process of persistent vegetation loss under livestock grazing is hardly understood but evidently an issue of the system's bistability.
-Under particular environmental conditions and grazing regimes, two ecosystem states are likewise resilient to perturbations: One state with a robust vegetation cover of shrubs and grasses as well as a desert with no significant vegetation cover (Scheffer 2001, Aguiar and Sala 1999).
+Drylands are at high risk of sudden desertification.
+This process of persistent vegetation loss is hardly understood but evidently a consequence of bistability.
+Under particular environmental conditions, two ecosystem states are likewise resilient to perturbations: One state with high vegetation cover of shrubs and grasses as well as a desert with no significant vegetation cover (Scheffer 2001).
 After crossing critical thresholds, or tipping-points, the system transites between the two stable states.
 
-For arid ecosystems, the major mechanism behind this bistability was found to be local facilitation of plant individuals.
-The presence of a plant provides its direct local environment with abilities of water retention, organic matter accumulation and protection against external stressors.
-As a consequence, the vegetation grows in patches (Aguiar and Sala 1999).
-When implemented in a simple spatially explicit model, this positive neighborhood effect provides true bistability on the landscape scale (K\'efi et al 2007a,b).
+For the dryland ecosystem, the major mechanism behind this bistability was found to be local facilitation of plant individuals.
+The presence of a plant provides its direct local environment with abilities of water retention, organic matter accumulation and protection against external stresses.
+When implemented in simple spatial models, This positive neighborhood effect provides true bistability on the landscape scale (K\'efi et al 2007a,b).
 
@@
 This simple formulation of additional plant stress can be interpreted as the effects of increased grazing intensity by large herbivores (K\'efi et al 2007a).
-However, the model is lacking some important features of grazing: the additional individual plant mortality (1) did not vary with vegetation density, and (2) it applied globally, affecting all plants likewise.
-Both assumptions are justified by the need for model simplicity, but disallow a translation of the model parameters to real livestock rates.
-The relationship between livestock rates and the stability of vegetation cover is a crucial information for management decisions.
-The present study aims at increasing the realism and applicability of the model by applying two extensions that enable simulations of a grazing gradient on drylands.
-
-%Grazing depends on vegetation cover
-The first extension adds a contrasting assumption on the plant density dependence of feeding to the parent model.
-The prior model defined grazing as a constant risk of additional plant death.
-On the population level, grazing mortality increased with vegetation cover (K\'efi et al.
-2007).
-However, grazers forage in the landscape and consume to fulfill their daily demands.
-In other words, the number of plants suffering from grazing is rather constant with vegetation cover.
-As a consequence, at low vegetation cover the grazing risk on individual plants will be high, whereas when vegetation cover is high, risk will be shared among many plants.
-In reality, the total feeding of large herbivores in the landscape is not entirely independent of vegetation cover but follows a non-linear saturating curve (type-II functional response, Holling 1959, Spalinger and Hobbs 1992).
-Thus, particularly at low vegetation cover grazing is limited by the search success of the grazer, which is resource-density dependent.
-In rangeland systems, however, the functional response at low vegetation cover depends on factors on a larger scale: the animals move between landscapes motivated by internal decision patterns or by external management decisions (Senft 1989).
-We assume therefore, that smooth type-II functional responses, as described for wild ungulates, are unlikely to happen in intensive rangelands within the spatial scale adressed in the present study.
-
-Because of these uncertainties and for reasons of simplicity we contrast the parent assumption of density independent individual risk with density independent global feeding.
-We expect constant global feeding to consolidate both stable states: the desert state will be more resilient against recolonisation of single plants while the vegetated state will provide better protection to the plant individual.
-As a consequence, the reversibility of catastrophic shifts is expected to deteriorate.
-
-%Spalinger and Hobbs 1992
-
-%Grazing is spatialy differentiated
-The second extension implements a local differentiation of grazing pressure.
-%For some time, the relationship between vegetation structure and grazing is acknowledged to be interdependent.
-Grazing affects the mortality of plants as well as plant recruitment, soil suitability, and other factors in positive as well as negative ways.
-Vegetation cover is a
-%Senft et al 1987
-%Bailey 1996
-Particularly, the importance of direct neighborhood interactions on the individual plant scale (local scale) in explaining spatial structure was demonstrated by Adler et al.
-(2001).
-The authors showed in a theoretical framework similar to the present study, that positive local interactions are superior to environmental heterogeneity or heterogeneous stress in explaining the patterns of plant spatial heterogeneity (Adler et al 2001).
-To differentiate the grazing locally, we implement an associative protection of plant individuals growing next to each other.
-This mechanism is observed frequently among shrubs in dryland ecosystems and considered to be a major driver of spatial heterogeneity of stress on local scales.
-
- %Bisigato et al 2005 Ecography
-%Milchunas and Noy-Meir 2002
-
+However, it is lacking some important features of grazing: the added plant mortality (1) applied globally, affecting all plants likewise, and (2) did not vary with vegetation density.
+The present study aims at increasing the realism of the model by applying two extensions that produce a more realistic gradient of grazing pressure on drylands.
+We provide a full factorial comparison of the consequences of these alternative models on vegetation structure and the ecosystems' bistability properties.
+The first extension adds the assumption of density dependent feeding.
+Grazers forage in the landscape and consume to fulfill their demands.
+At low density the grazing risk on individual plants will be high, while it will be shared among many plants when vegetation cover is high.
+For reasons of simplicity, we contrast the prior assumption of density independent individual risk with density independent global feeding.
+We expect this to consolidate the desert state by making it more resilient against recolonisation of single plants.
+The second extension implements an associative protection of plant individuals growing next to each other, a mechanism observed frequently among shrubs in dryland ecosystems.
 Here, the individual vulnerability against grazing is highest for plants growing isolated.
@@
 
-% comparison and main objectives/questions
-
- We provide a full factorial comparison of the consequences of these alternative models on vegetation structure and the ecosystems' bistability properties.
-
 \section{Methods}
 We used a cellular automaton to investigate the consequences of different grazing assumptions on the structural properties of the vegetation.
-This model was used before (Kefi et al 2007a, b) and can be resolved numerically in spatially explicit simulations.
+This model was used before (Kefi et al 2007a, b) and can be resolved numerically in spatially explicit simulations but also analytically in differential equations.
 It defines landscape as a grid of cells, each can potentially take one of three discrete cell states: 'vegetated' cells are occupied by a plant individual (annotated as '+' in equations; black cells in figures); 'enriched' cells are suitable for seeds to germinate and colonise the cell, but do not contain adult plants right now  ('0' , grey cells); 'degraded' cells represent bare ground with lacking organic matter and bad water retention, and therefore can not be colonised by arriving seeds  ('$-$', white cells).
@@
 
-We applied these rules on a grid of 100 $\times$ 100 cells of 0.25m$^2$.
-Each timestep is defined as one year.
-The applicable transition probabilities for each single cell given in equations 1, 2, 3 and 4 are compared against uniform random numbers between 0 and 1 to determine if a transition occurs or if the cell remains unchanged.
-To initialize the grid with a standardized patchy vegetation structure, each model simulation was run for 500 years without any grazing (g = 0) using an initial environmental value $b_ini \in \left\{ 0.25, 3, 0.8, 0.9 \right\}$ before grazing is taking action.
-%(Note: I want to complete this to a highly resoluted gradient, to investigate the unstable equilibria).
-Starting from the preconditioned grids, the four grazing models (Table 1) with varying grazing intensity ($g \in \left{ 0, 0.025, 0.05, 0.075 , 0.10 \right}$) were run over a gradient of environmental quality ($b \in \left{ 0.2, ...
-, 1.0 \right}$) for 1.500 timesteps.
-
-An average vegetation cover, $\rho_+$, as well as the average cell mortality,$\hat{m}$ , is determined as the average density of all grids of the final 500 timesteps.
+We apply these rules on a grid of 100 $\times$  100 cells of 0.25m$^2$ .
+The grid is defined to have periodic borders, connecting the east border to the west, the north to the south.
+Each timestep, which is defined as one year, the whole grid is updated synchronously multiple times, to increase stochasticity.
+At each update, the applicable transition probabilities for each single cell given in equations 1, 2, 3 and 4 are devided by the number of updates per year ( = 5) and tested against a uniform random number between 0 and 1.
+To initialize the grid with a standardized patch structure, each model simulation runs for 1.000 timesteps without any grazing using an initial environmental value $b_ini \in \left\{ 0.25, 3, 0.8, 0.9 \right\}$ before grazing is taking action (Note: I want to complete this to a highly resoluted gradient, to investigate the unstable equilibria).
+On those preconditioned grids, the four grazing models (Table ) run for 1.500 timesteps.
+An average vegetation cover, $\rho_+$, as well as the average cell mortality,$\hat{m}$ , is determined from the grids of the final 100 timesteps.
 
@@
 
-\begin{figure}[h]%
+\begin{figure}%
 \includegraphics[width=\columnwidth]{figures/stability.pdf}%
-\caption{Stable states of the vegetation cover, $\rho_+$, over gradient of environmental quality, $b$, under different grazing models.
-a) parent model with static, global grazing mortality, $g_0/0.5$.
-b) livestock model with density dependend, global grazing mortality, $g_0 / \rho_+$.
-c) associative protection model with differentiated grazing at patch border, $g_0/0.5 \frac{v}{\hat{v}}$ .
-d) livestock \& associative protection model with density dependend, differentiated grazing at patch border,  $g_0/ \rho_+ \frac{v}{\hat{v}}$ .
-Gradient of grazing intensity from grey: $g_0 = 0$, black to red: $g_0 \in \left\{ 0.025, 0.05, 0.075, 0.1 \right\}$.}%
-\label{}%
-\end{figure}
-
-\begin{figure}[h]%
-\includegraphics[width=\columnwidth]{figures/patterns.pdf}%
 \caption{Stable states of the vegetation cover, $\rho_+$, over gradient of environmental quality, $b$, under different grazing models.
@@
 
-In reality, however, the functional response of large grazers on vegetation, particularly at low vegetation cover, is dynamic.
-In rangelands, management decisions and spatial responses of the grazers will reduce feeding pressure at very low densities.
-
-In fact, the local effects of large herbivores on plants are manifold and qualitatively different.
-While direct feeding reduces biomass in the first place, it might lead to growth enhancement and increased productivity.
-On the local scale, vegetation benefit from the presence of herbivores through fertilization, inoculation with seeds,
-In our model, however, we simplify these factors to a linear negative impact of grazer presence on the local vegetation. The
-
 \section{Conclusion}
@@
 A linear increase in grazing causes shifts much
-
-\section{APPENDIX}
-
-\subsection{simulations}
-
-The grid was defined to have periodic borders, connecting the east border to the west, the north to the south.
-Prerun for 100 years. 
-during pre-run, environmental conditions $b_ini \in \left\{ 0.250, 0.274, 0.300, 0.329, 0.360, 0.395, 0.433, 0.474, 0.520, 0.570, 0.624, 0.684, 0.749, 0.821, 0.900 \right\} $ emerge a patch structure ranging from low to high vegetation cover.
-Trajectories towards a stable vegetated state or towards desert are documented and used to approximate the unstable equilibrium.
-
-, the whole grid is updated synchronously multiple times, to increase stochasticity
-each time step is devided by the number of updates per year ( = 5) and
-
-\subsection{model fitting}
- Four models were compared on each snapshot of t
-
-\begin{table}[t!]
-\label{tab:fitmodels}
-\caption{ Models describing patch size distribution }
-\centering
-\begin{tabular}{lll}
-
-\toprule
- & function & log-transform \\ \cmidrule(rl){2-2} \cmidrule(rl){3-3}
-Power Law &  $n = a s^{- \alpha}$ & $ \log{n} =\log{a} - \alpha \log s$\\
-Limited Power Law &  $n = b + a s^{- \alpha} $ & $ \log{n} = \log a - \alpha \log{s} + \log(1+\frac{b}{a s^{-\alpha} })$\\
-Truncated Power Law & $n = a s^{- \alpha} \mathrm{e}^{\frac{s}{S_x}}$  & $\log{n} = \log a  - \alpha  \log s  - \frac{s}{Sx}  $ \\
-Exponential &  $n = a \mathrm{e}^{-\varepsilon s} $ & $ \log{n} =  \log{a} - \varepsilon s $\\
-	\bottomrule
-\multicolumn{3}{p{12.5cm}}{\footnotesize  }
-	\end{tabular}
-\end{table}
-
-table with log-transformed models

### 2013-11-21 code/automatafunctions.r

@@ -1,0 +1,151 @@
+#################################################
+#
+#		model: cellular automata helper functions
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		benchmarking tool
+#		implement test for stable equilibrium
+#		helper functions: randomnumber, count function
+# 		plotting functions 
+# 		conut patch size and distributions
+# 		
+#
+#################################################
+
+
+# plotting function for objects of class "landscape".
+plot.landscape <- function(x, grid = FALSE, axis = FALSE, cols = "auto", add = FALSE, ani = FALSE, ...) {
+  lvls <- levels(x$cells) 
+  nlev <- length(lvls)
+  if(cols[1] == "auto") cols = greyscale(nlev) # default color value
+  
+  if(ani & Sys.info()[['sysname']] == "Windows") adj = -0.5 else adj = 0 #this adjustment constant is added when producing a pixel accurate png or gif file, requires TRUE when the function is used to plot animated figures. 
+  
+  if(!add) plot(NA,NA, xlim = c(0.5+adj, x$dim[1]+0.5+adj), ylim = c( x$dim[2]+0.5+adj, 0+0.5+adj), bty = c("n", "o")[grid+1], xaxs = "i", yaxs = "i",xlab = "", ylab = "", xaxt = "n", yaxt = "n", ... ) 
+  
+  if(axis && !add) axis(3) 
+  if(axis && !add) axis(2)
+  
+  if(grid) border = "grey80" else border = cols[as.numeric(x$cells)]
+  
+  rect(rep(1:x$dim[1], times = x$dim[2])-.5, rep(1:x$dim[2], each = x$dim[1])-.5, rep(1:x$dim[1], times = x$dim[2])+.5, rep(1:x$dim[2], each = x$dim[1])+.5, col = cols[as.numeric(x$cells)], border = border)
+  
+  if(grid) box()
+ }
+
+greyscale <- colorRampPalette(c( "black", "white"), space = "rgb")
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor[1]
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+
[…]

### 2013-11-21 code/benchmarking.r

@@ -1,0 +1,283 @@
+library(rbenchmark)
+
+
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
+			# cumulate the result for each of the four neighboring cells (for(k in interact))
+			} else neighbors <- 0
+			
+			
+			mean(neighbors/4, na.rm = TRUE)	
+			# devide by 4 to get density
+			# average over all cells
+}
+
+
+# get patch size and patchsize distribution
+patches <- function(x, state, cumulative = TRUE) {
+	pattern <- x$cells
+	pattern <- pattern %in% state
+	map <- rep(NA, times = prod(x$dim))
+	old <- rep(99, times = prod(x$dim)) 
+	
+	while(!identical(old[pattern], map[pattern])) {
+		old <- map
+		count = as.integer(1)
+		for(i in which(pattern)) {
[…]

### 2013-11-21 code/facilitationmodel.r

@@ -1,0 +1,263 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+
+rm(list=ls())
+
+# plotting function for objects of class "landscape".
+plot.landscape <- function(x, grid = FALSE, axis = FALSE, cols = "auto", add = FALSE, ani = FALSE, ...) {
+  lvls <- levels(x$cells) 
+  nlev <- length(lvls)
+  if(cols[1] == "auto") cols = greyscale(nlev) # default color value
+  
+  if(ani & Sys.info()[['sysname']] == "Windows") adj = -0.5 else adj = 0 #this adjustment constant is added when producing a pixel accurate png or gif file, requires TRUE when the function is used to plot animated figures. 
+  
+  if(!add) plot(NA,NA, xlim = c(0.5+adj, x$dim[1]+0.5+adj), ylim = c( x$dim[2]+0.5+adj, 0+0.5+adj), bty = c("n", "o")[grid+1], xaxs = "i", yaxs = "i",xlab = "", ylab = "", xaxt = "n", yaxt = "n", ... ) 
+  
+  if(axis && !add) axis(3) 
+  if(axis && !add) axis(2)
+  
+  if(grid) border = "grey80" else border = cols[as.numeric(x$cells)]
+  
+  rect(rep(1:x$dim[1], times = x$dim[2])-.5, rep(1:x$dim[2], each = x$dim[1])-.5, rep(1:x$dim[1], times = x$dim[2])+.5, rep(1:x$dim[2], each = x$dim[1])+.5, col = cols[as.numeric(x$cells)], border = border)
+  
+  if(grid) box()
+ }
+
+count  <- function(x, neighbor) {
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+			}
+	
+# function which counts the density of a state within neighboring cells of a particular state
[…]

### 2013-11-21 code/grazing_density.r

@@ -1,0 +1,397 @@
+
+
+
+# specify lattice
+width = 100
+height = 100
+
+states = c("+","0","-")
+color <- c("black","grey80", "white") # define colors for the cell state levels
+
+parms_temp <- list(
+	m = 0.05,
+	g = 0.099,
+		b = .6 ,		# beta*eps 
+		d = 0.1,		# degradation
+		c_ = 0.01, 		# beta*g  
+		del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+		r = 0.01, 	# regeneration rate
+		f = 0.9 		# local fascilitation
+)
+
+
+# initial cell states
+states = c("+","0","-")
+prob = c(6/10,2/10,2/10)
+color <- c("black","grey80", "white") # define colors for the cell state levels
+
+
+# derive helper vectors for counting: 
+# transformation vector for evaluation at the border of the grid
+	# set evaluation matrix 
+	X <- matrix(as.integer(1:(width*height)), ncol = width, byrow =TRUE)
+	# setting the border of the evaluation matrix X
+	X <- cbind(X[,width], X, X[,1] )  
+	X <- rbind(X[height,], X, X[1,] ) 
+	# transformation vector which adds the border to the lattice:
+	x_with_border <- as.integer(t(X))
+	
+	# from the matrix X (with copied border cells), which cells are the actual cells (reverse transformation vector of the previous lines) 
+	x_to_evaluate <- sort(matrix(1:prod(dim(X)), ncol = dim(X)[2], byrow =TRUE)[-c(1, dim(X)[1]), -c(1,dim(X)[2])]	)		
+
+# defining the neighborhood which is to be evaluated	
+	# set interaction matrix
+	I <- matrix(c(0,1,0,1,NA,1,0,1,0), ncol = 3, byrow = TRUE)	
+	# coordinates of neighbours in Interaction matrix I: 
+	neighbours_in_I <- which(is.finite(abs(I)/abs(I)), arr.in = TRUE)
+	# coordinates relative to the evaluated cell (=  which(is.na(I) ) 
+	relrow <- neighbours_in_I[,1]-which(is.na(I), arr.ind = TRUE)[1]
+	relcol <- neighbours_in_I[,2]-which(is.na(I), arr.ind = TRUE)[2]
+	
+	# relative position of the four direct neighbours of a cell
[…]

### 2013-11-21 code/grazing_density_v2.r

@@ -1,0 +1,558 @@
+
+
+rm(list=ls())
+
+
+# specify lattice
+width = 200
+height = 200
+
+states = c("+","0","-")
+color <- c("black","grey80", "white") # define colors for the cell state levels
+
+parms_temp <- list(
+	m = 0.05,
+	g = 0.03,
+		b = .5 ,		# beta*eps 
+		d = 0.1,		# degradation
+		c_ = 0.01, 		# beta*g  
+		del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+		r = 0.01, 	# regeneration rate
+		f = 0.9 		# local fascilitation
+)
+
+
+# initial cell states
+states = c("+","0","-")
+prob = c(6/10,2/10,2/10)
+color <- c("black","grey80", "white") # define colors for the cell state levels
+
+color <- c("black","grey80", "white") 
+# plotting function for objects of class "landscape".
+plot.landscape <- function(x, grid = FALSE, axis = FALSE, cols = "auto", add = FALSE, ani = FALSE, ...) {
+  lvls <- levels(x$cells) 
+  nlev <- length(lvls)
+  if(cols[1] == "auto") cols = c("black","grey80", "white")  # default color value
+  
+  if(ani & Sys.info()[['sysname']] == "Windows") adj = -0.5 else adj = 0 #this adjustment constant is added when producing a pixel accurate png or gif file, requires TRUE when the function is used to plot animated figures. 
+  
+  if(!add) plot(NA,NA, xlim = c(0.5+adj, x$dim[1]+0.5+adj), ylim = c( x$dim[2]+0.5+adj, 0+0.5+adj), bty = c("n", "o")[grid+1], xaxs = "i", yaxs = "i",xlab = "", ylab = "", xaxt = "n", yaxt = "n", ... ) 
+  
+  if(axis && !add) axis(3) 
+  if(axis && !add) axis(2)
+  
+  if(grid) border = "grey80" else border = cols[as.numeric(x$cells)]
+  
+  rect(rep(1:x$dim[1], times = x$dim[2])-.5, rep(1:x$dim[2], each = x$dim[1])-.5, rep(1:x$dim[1], times = x$dim[2])+.5, rep(1:x$dim[2], each = x$dim[1])+.5, col = cols[as.numeric(x$cells)], border = border)
+  
+  if(grid) box()
+ }
+
[…]

### 2013-11-21 code/grazing_density_v3.r

@@ -1,0 +1,553 @@
+
+
+rm(list=ls())
+
+
+# specify lattice
+width = 200
+height = 200
+
+states = c("+","0","-")
+color <- c("black","grey80", "white") # define colors for the cell state levels
+
+parms_temp <- list(
+	m = 0.05,
+	g = 0.03,
+		b = .5 ,		# beta*eps 
+		d = 0.1,		# degradation
+		c_ = 0.01, 		# beta*g  
+		del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+		r = 0.01, 	# regeneration rate
+		f = 0.9 		# local fascilitation
+)
+
+
+# initial cell states
+states = c("+","0","-")
+prob = c(6/10,2/10,2/10)
+color <- c("black","grey80", "white") # define colors for the cell state levels
+
+color <- c("black","grey80", "white") 
+# plotting function for objects of class "landscape".
+plot.landscape <- function(x, grid = FALSE, axis = FALSE, cols = "auto", add = FALSE, ani = FALSE, ...) {
+  lvls <- levels(x$cells) 
+  nlev <- length(lvls)
+  if(cols[1] == "auto") cols = c("black","grey80", "white")  # default color value
+  
+  if(ani & Sys.info()[['sysname']] == "Windows") adj = -0.5 else adj = 0 #this adjustment constant is added when producing a pixel accurate png or gif file, requires TRUE when the function is used to plot animated figures. 
+  
+  if(!add) plot(NA,NA, xlim = c(0.5+adj, x$dim[1]+0.5+adj), ylim = c( x$dim[2]+0.5+adj, 0+0.5+adj), bty = c("n", "o")[grid+1], xaxs = "i", yaxs = "i",xlab = "", ylab = "", xaxt = "n", yaxt = "n", ... ) 
+  
+  if(axis && !add) axis(3) 
+  if(axis && !add) axis(2)
+  
+  if(grid) border = "grey80" else border = cols[as.numeric(x$cells)]
+  
+  rect(rep(1:x$dim[1], times = x$dim[2])-.5, rep(1:x$dim[2], each = x$dim[1])-.5, rep(1:x$dim[1], times = x$dim[2])+.5, rep(1:x$dim[2], each = x$dim[1])+.5, col = cols[as.numeric(x$cells)], border = border)
+  
+  if(grid) box()
+ }
+
[…]

### 2013-11-21 code/grazing_gradient_model_analysis.r

@@ -1,0 +1,768 @@
+
+highlight <- function(x, colrange = c("black", "red2"), steps = NULL, range = "auto"){
+
+if(is.null(steps)) steps = length(unique(x))
+if(range == "auto") {
+	min_val = min(x, na.rm = TRUE)
+	max_val = max(x, na.rm = TRUE)
+	} else {
+	min_val = range[1]
+	max_val = range[2]
+	}
+
+colorscale <- colorRampPalette(colrange, space = "rgb")
+cols <- colorscale( steps)
+
+cols[as.integer((x-min_val)/(max_val-min_val)*(steps-1)+1)]
+
+}
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
[…]

### 2013-11-21 code/grazing_gradient_model_analysis_v10.r

@@ -1,0 +1,769 @@
+
+highlight <- function(x, colrange = c("black", "red2"), steps = NULL, range = "auto"){
+
+if(is.null(steps)) steps = length(unique(x))
+if(range == "auto") {
+	min_val = min(x, na.rm = TRUE)
+	max_val = max(x, na.rm = TRUE)
+	} else {
+	min_val = range[1]
+	max_val = range[2]
+	}
+
+colorscale <- colorRampPalette(colrange, space = "rgb")
+cols <- colorscale( steps)
+
+cols[as.integer((x-min_val)/(max_val-min_val)*(steps-1)+1)]
+
+}
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
[…]

### 2013-11-21 code/grazing_gradient_model_comp.r

@@ -1,0 +1,492 @@
+#################################################
+#
+#		model: facilitation model plus grazing
+#
+#
+#author: Flo
+#date: 12.09.2013
+#
+#todo: funktionen auslagern
+#  		gradienten definieren
+#################################################
+
+rm(list=ls())
+
+########################################################################################
+	
+
+highlight <- function(x, colrange = c("black", "red2"), steps = NULL, range = "auto"){
+
+if(is.null(steps)) steps = length(unique(x))
+if(range == "auto") {
+	min_val = min(x, na.rm = TRUE)
+	max_val = max(x, na.rm = TRUE)
+	} else {
+	min_val = range[1]
+	max_val = range[2]
+	}
+
+colorscale <- colorRampPalette(colrange, space = "rgb")
+cols <- colorscale( steps)
+
+cols[as.integer((x-min_val)/(max_val-min_val)*(steps-1)+1)]
+
+}
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
[…]

### 2013-11-21 code/grazing_gradient_model_comp_v10.r

@@ -1,0 +1,721 @@
+#################################################
+#
+#		model: facilitation model plus grazing
+#
+#
+#author: Flo
+#date: 12.09.2013
+#
+#todo: funktionen auslagern
+#  		gradienten definieren
+#################################################
+
+rm(list=ls())
+
+########################################################################################
+	
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
+			# cumulate the result for each of the four neighboring cells (for(k in interact))
+			} else neighbors <- 0
+			
+			
+			mean(neighbors/4, na.rm = TRUE)	
+			# devide by 4 to get density
+			# average over all cells
+}
+
+
+# get patch size and patchsize distribution
[…]

### 2013-11-21 code/grazing_gradient_model_comp_v10b.r

@@ -1,0 +1,727 @@
+#################################################
+#
+#		model: facilitation model plus grazing
+#
+#
+#author: Flo
+#date: 12.09.2013
+#
+#todo: funktionen auslagern
+#  		gradienten definieren
+#################################################
+
+rm(list=ls())
+
+########################################################################################
+	
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
+			# cumulate the result for each of the four neighboring cells (for(k in interact))
+			} else neighbors <- 0
+			
+			
+			mean(neighbors/4, na.rm = TRUE)	
+			# devide by 4 to get density
+			# average over all cells
+}
+
+
+# get patch size and patchsize distribution
[…]

### 2013-11-21 code/grazing_gradient_model_comp_v10c.r

@@ -1,0 +1,754 @@
+#################################################
+#
+#		model: facilitation model plus grazing
+#
+#
+#author: Flo
+#date: 12.09.2013
+#
+#todo: funktionen auslagern
+#  		gradienten definieren
+#################################################
+
+rm(list=ls())
+
+########################################################################################
+	
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
+			# cumulate the result for each of the four neighboring cells (for(k in interact))
+			} else neighbors <- 0
+			
+			
+			mean(neighbors/4, na.rm = TRUE)	
+			# devide by 4 to get density
+			# average over all cells
+}
+
+
[…]

### 2013-11-21 code/grazing_gradient_model_comp_v10c_2.r

@@ -1,0 +1,754 @@
+#################################################
+#
+#		model: facilitation model plus grazing
+#
+#
+#author: Flo
+#date: 12.09.2013
+#
+#todo: funktionen auslagern
+#  		gradienten definieren
+#################################################
+
+rm(list=ls())
+
+########################################################################################
+	
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
+			# cumulate the result for each of the four neighboring cells (for(k in interact))
+			} else neighbors <- 0
+			
+			
+			mean(neighbors/4, na.rm = TRUE)	
+			# devide by 4 to get density
+			# average over all cells
+}
+
+
[…]

### 2013-11-21 code/grazing_gradient_model_comp_v10c_3.r

@@ -1,0 +1,755 @@
+#################################################
+#
+#		model: facilitation model plus grazing
+#
+#
+#author: Flo
+#date: 12.09.2013
+#
+#todo: funktionen auslagern
+#  		gradienten definieren
+#################################################
+
+rm(list=ls())
+
+########################################################################################
+	
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
+			# cumulate the result for each of the four neighboring cells (for(k in interact))
+			} else neighbors <- 0
+			
+			
+			mean(neighbors/4, na.rm = TRUE)	
+			# devide by 4 to get density
+			# average over all cells
+}
+
+
[…]

### 2013-11-21 code/grazing_gradient_model_comp_v9.r

@@ -1,0 +1,444 @@
+#################################################
+#
+#		model: facilitation model plus grazing
+#
+#
+#author: Flo
+#date: 12.09.2013
+#
+#todo: funktionen auslagern
+#  		gradienten definieren
+#################################################
+
+rm(list=ls())
+
+########################################################################################
+	
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
+			# cumulate the result for each of the four neighboring cells (for(k in interact))
+			} else neighbors <- 0
+			
+			
+			mean(neighbors/4, na.rm = TRUE)	
+			# devide by 4 to get density
+			# average over all cells
+}
+
+
+# get patch size and patchsize distribution
[…]

### 2013-11-21 code/grazing_gradient_model_comp_v9a.r

@@ -1,0 +1,444 @@
+#################################################
+#
+#		model: facilitation model plus grazing
+#
+#
+#author: Flo
+#date: 12.09.2013
+#
+#todo: funktionen auslagern
+#  		gradienten definieren
+#################################################
+
+rm(list=ls())
+
+########################################################################################
+	
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
+			# cumulate the result for each of the four neighboring cells (for(k in interact))
+			} else neighbors <- 0
+			
+			
+			mean(neighbors/4, na.rm = TRUE)	
+			# devide by 4 to get density
+			# average over all cells
+}
+
+
+# get patch size and patchsize distribution
[…]

### 2013-11-21 code/grazing_gradient_repair_v10b.r

@@ -1,0 +1,205 @@
+# repair run 10
+
+
+rm(list=ls())
+
+##########
+
+
+################ starting parallel backend
+	
+library(foreach)
+library(doSNOW)
+
+#workerlist <- rep(list(ubuWorker), times = 2)
+workerlist <- rep("localhost", times = 23)
+
+cl <- makeSOCKcluster(workerlist)
+
+registerDoSNOW(cl)
+
+########################
+
+
+################ parameter settings
+first_ID = 1  #
+env <- 	seq(0.2,1,.01) #seq(0.20,0.98,.02)+.01
+graz <- seq(.0, .1, 0.025) #.025 #
+mort <- 0.05 #c(0.05, 0.1) #seq(0.05,.15, 0.05)
+init <- round(exp(seq(log(0.25), log(0.9), length = 33))[seq(1,33,2)], digits = 3)
+global <- c(FALSE, TRUE)
+stock <- c(FALSE, TRUE)
+
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+linit <- length(init)
+lglobal <- length(global)
+lstock <- length(stock)
+replication <- 1
+
+
+# defining parameter set
+parameters = data.frame(
+	ID = first_ID:(lgraz*lmort*lglobal*linit*lenv*lstock*replication+first_ID-1), 
+	global = rep(global, each = lgraz*lenv*lmort*linit*lstock*replication),
+	stock =  rep(rep(stock, each = lgraz*lenv*lmort*linit*replication), times = lstock),
+	starting = rep(rep(init, each = lgraz*lenv*lmort*replication), times = lglobal*lstock),
+	m = rep(rep(mort, each = lgraz*lenv*replication), times = lglobal*linit*lstock), 		# intrinsic mortality
+	g = rep(rep(graz, each = lenv*replication), times = lmort*lglobal*linit*lstock),		#grazing
+	b = rep(rep(env, each = replication) , times = lgraz*lmort*lglobal*linit*lstock), 		# beta*eps 
+	d = 0.1,		# degradation
+	c_ = 0.2, 		# beta*g  
+	del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+	r = 0.01, 	# regeneration rate
+	f = 0.9		# local fascilitation
+)
+
[…]

### 2013-11-21 code/models.r

@@ -1,0 +1,21 @@
+size <- exp(seq(log(1), log(10000), length = 200))
+
+a <- 0.8
+alpha <- 1
+b <- .01
+Sx <- 100
+
+PL <-  exp(  log(a) - alpha * log(size) )
+TPLup <-  exp(  log(a) - alpha * log(size) + log(1+b/(a*size^(-alpha)))) 
+TPLdown <-  exp(    log(a) - alpha * log(size) - (size/Sx)  )
+
+eps = .3
+
+EXP <-   exp( log(a) - (eps*size) )
+plot(NA, NA, xlim = c(1, 10000) , ylim = c(0.0001, 1), log = "xy")
+lines(size, PL, lwd = 2)
+lines(size, TPLup)
+lines(size, TPLdown)
+lines(size, EXP)
+
+

### 2013-11-21 code/old code/fitting_psd.r

@@ -1,0 +1,133 @@
+
+setwd("E:\\Eigene Dokumente\\Uni\\projects\\2013 CASCADE grazing\\sim4")
+
+output <- read.csv("output.csv")
+
+output[(output$b == 0.52 & output$globalgrazing == TRUE & output$starting == "high"),]
+
+load(paste("results\\result", 777, sep = "_"))
+
+
+
+
+
+pdf("E:\\SkyDrive\\Uni\\talks\\patchdist_graz.pdf", height = 5, width = 6, paper = "special")
+par(mfrow = c(1,1))
+
+for(j in seq(0,0.025, 0.005)) {
+	plot(NA, NA, xlim = c(1,10000), ylim = c(1,1000), log= "xy", main = paste("g =", j))
+
+    select <- with(output, ID[which(g == j & b == 0.52 & globalgrazing == TRUE & starting == "high")])
+	load(paste("results/result_", select, sep = ""), envir = globalenv() )
+	if(all(!is.na(result$cumpatch))) points(mean~size, pch = 20, col = "#968376", data = result$cumpatch)
+
+	
+    select <- with(output, ID[which(g == j & b == 0.52 & globalgrazing == FALSE & starting == "high")])
+	load(paste("results/result_", select, sep = ""), envir = globalenv() )
+	if(all(!is.na(result$cumpatch))) points(mean~size, pch = 20, col = "blue3", data = result$cumpatch)
+
+
+}
+dev.off()
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+sort(unlist(result$patches))
+
+result$cumpatch$size
+plot(mean~size, data = result$cumpatch, log= "xy", pch = 20)
+
+lmean <- mean(log(sort(unlist(result$patches))))
+lstd  <- sd(log(sort(unlist(result$patches))))
+
+logncdf <- function(x, lmean, lstd) 1/(std*sqrt(2*pi)*sum(exp(-(log())) seq())) 
+P <- ecdf(sort(unlist(result$patches)))
+
+qlnorm(sort(unlist(result$patches)))
+
+
+
+	bins <- sort(unique(unlist(result$patches[-1])))
+	bins <- seq(1,501, 10)
+	bins <- 10^seq(log10(1),log10(10000), .1)
+	plot(NA,NA, xlim = c(1,10000), ylim = c(1,1000), log = "xy")
+	for(i in 3:22) {
[…]

### 2013-11-21 code/old code/herbivorymodel2_baseline.r

@@ -1,0 +1,293 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+
+rm(list=ls())
+source("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+source("E:\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+setwd("E:\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+
+# specify lattice
+width = 50
+height = 50
+
+# initial cell states
+states = c("+","0","-")
+prob = c(1/10,1/10,8/10)
+
+# time and resolution of simulation
+timesteps = 750
+addgrazing = 200
+delta = 1/10
+
+
+# derive helper vectors for counting: 
+# transformation vector for evaluation at the border of the grid
+	# set evaluation matrix 
+	X <- matrix(as.integer(1:(width*height)), ncol = width, byrow =TRUE)
+	# setting the border of the evaluation matrix X
+	X <- cbind(X[,width], X, X[,1] )  
+	X <- rbind(X[height,], X, X[1,] ) 
+	# transformation vector which adds the border to the lattice:
+	x_with_border <- as.integer(t(X))
+	
+	# from the matrix X (with copied border cells), which cells are the actual cells (reverse transformation vector of the previous lines) 
+	x_to_evaluate <- sort(matrix(1:prod(dim(X)), ncol = dim(X)[2], byrow =TRUE)[-c(1, dim(X)[1]), -c(1,dim(X)[2])]	)		
+
+# defining the neighborhood which is to be evaluated	
+	# set interaction matrix
+	I <- matrix(c(0,1,0,1,NA,1,0,1,0), ncol = 3, byrow = TRUE)	
[…]

### 2013-11-21 code/old code/herbivorymodel2_baseline_low.r

@@ -1,0 +1,293 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+
+rm(list=ls())
+source("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+source("E:\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+setwd("E:\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+
+# specify lattice
+width = 50
+height = 50
+
+# initial cell states
+states = c("+","0","-")
+prob = c(1/10,1/10,8/10)
+
+# time and resolution of simulation
+timesteps = 200
+addgrazing = 200
+delta = 1/10
+
+
+# derive helper vectors for counting: 
+# transformation vector for evaluation at the border of the grid
+	# set evaluation matrix 
+	X <- matrix(as.integer(1:(width*height)), ncol = width, byrow =TRUE)
+	# setting the border of the evaluation matrix X
+	X <- cbind(X[,width], X, X[,1] )  
+	X <- rbind(X[height,], X, X[1,] ) 
+	# transformation vector which adds the border to the lattice:
+	x_with_border <- as.integer(t(X))
+	
+	# from the matrix X (with copied border cells), which cells are the actual cells (reverse transformation vector of the previous lines) 
+	x_to_evaluate <- sort(matrix(1:prod(dim(X)), ncol = dim(X)[2], byrow =TRUE)[-c(1, dim(X)[1]), -c(1,dim(X)[2])]	)		
+
+# defining the neighborhood which is to be evaluated	
+	# set interaction matrix
+	I <- matrix(c(0,1,0,1,NA,1,0,1,0), ncol = 3, byrow = TRUE)	
[…]

### 2013-11-21 code/old code/herbivorymodel2_corrected feeding.r

@@ -1,0 +1,276 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+
+rm(list=ls())
+source("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+source("E:\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+setwd("E:\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+
+# specify lattice
+width = 50
+height = 50
+
+# initial cell states
+states = c("+","0","-")
+prob = c(1/10,1/10,8/10)
+
+# time and resolution of simulation
+timesteps = 1000
+addgrazing = 200
+delta = 1/10
+
+
+# derive helper vectors for counting: 
+# transformation vector for evaluation at the border of the grid
+	# set evaluation matrix 
+	X <- matrix(as.integer(1:(width*height)), ncol = width, byrow =TRUE)
+	# setting the border of the evaluation matrix X
+	X <- cbind(X[,width], X, X[,1] )  
+	X <- rbind(X[height,], X, X[1,] ) 
+	# transformation vector which adds the border to the lattice:
+	x_with_border <- as.integer(t(X))
+	
+	# from the matrix X (with copied border cells), which cells are the actual cells (reverse transformation vector of the previous lines) 
+	x_to_evaluate <- sort(matrix(1:prod(dim(X)), ncol = dim(X)[2], byrow =TRUE)[-c(1, dim(X)[1]), -c(1,dim(X)[2])]	)		
+
+# defining the neighborhood which is to be evaluated	
+	# set interaction matrix
+	I <- matrix(c(0,1,0,1,NA,1,0,1,0), ncol = 3, byrow = TRUE)	
[…]

### 2013-11-21 code/old code/simulation_grazing_corrected.r

@@ -1,0 +1,473 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+rm(list=ls())
+
+setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+
+env <- 	seq(0.2,1,.05)
+graz <- seq(0,.9, 0.3)
+mort <- seq(0.1,.3, 0.1)
+init <- c("low", "high")
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+
+env <-	rep(rep(env, times = lgraz), times = lmort)
+graz <-	 rep(rep(graz, each = lenv), times = lmort)
+mort <-  rep(mort, each = lgraz*lenv)
+
+replicates = 2
+init <- rep(init, each = length(mort) )
+env <- rep(env, times = replicates)
+graz <- rep(graz, times = replicates)
+mort <- rep(mort, times = replicates)
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1:length(graz), 
+	starting = init,
+	g = graz,		#grazing
+	m = mort, 		# intrinsic mortality
+	b = env, 		# beta*eps 
+	d = 0.1,		# degradation
+	c_ = 0.2, 		# beta*g  
+	del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+	r = 0.01, 	# regeneration rate
+	f = 0.9		# local fascilitation
+)
+
+
+# defining parameter set
+parameters_ini = list(
+	low = list(
+		m = 0.1, 		# intrinsic mortality
+		b = 0.25, 		# beta*eps 
+		d = 0.1,		# degradation
+		c_ = 0.2, 		# beta*g  
+		del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+		r = 0.01, 	# regeneration rate
+		f = 0.9, 		# local fascilitation
+		g = 0		#grazing
+		), 
+	high = list(
[…]

### 2013-11-21 code/old code/transitions.r

@@ -1,0 +1,241 @@
+#################################################
+#
+#		model: iterator for cellular automata
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#
+#################################################
+
+
+
+rm(list=ls())
+source("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+setwd()
+library(foreach)
+library(doSNOW)
+
+cl <- makeCluster(rep("localhost", times = 11), type = "SOCK")
+registerDoSNOW(cl)
+
+# specify lattice
+width = 50
+height = 50
+
+
+# time and resolution of simulation
+timesteps = 10
+addgrazing = 1
+delta = 1/10
+
+
+# derive helper vectors for counting: 
+# transformation vector for evaluation at the border of the grid
+	# set evaluation matrix 
+	X <- matrix(as.integer(1:(width*height)), ncol = width, byrow =TRUE)
+	# setting the border of the evaluation matrix X
+	X <- cbind(X[,width], X, X[,1] )  
+	X <- rbind(X[height,], X, X[1,] ) 
+	# transformation vector which adds the border to the lattice:
+	x_with_border <- as.integer(t(X))
+	
+	# from the matrix X (with copied border cells), which cells are the actual cells (reverse transformation vector of the previous lines) 
+	x_to_evaluate <- sort(matrix(1:prod(dim(X)), ncol = dim(X)[2], byrow =TRUE)[-c(1, dim(X)[1]), -c(1,dim(X)[2])]	)		
+
+# defining the neighborhood which is to be evaluated	
+	# set interaction matrix
+	I <- matrix(c(0,1,0,1,NA,1,0,1,0), ncol = 3, byrow = TRUE)	
+	# coordinates of neighbours in Interaction matrix I: 
+	neighbours_in_I <- which(is.finite(abs(I)/abs(I)), arr.in = TRUE)
+	# coordinates relative to the evaluated cell (=  which(is.na(I) ) 
+	relrow <- neighbours_in_I[,1]-which(is.na(I), arr.ind = TRUE)[1]
+	relcol <- neighbours_in_I[,2]-which(is.na(I), arr.ind = TRUE)[2]
+	
+	# relative position of the four direct neighbours of a cell
+	interact <- (relrow * dim(X)[2] + relcol)
+
+
+# defining parameter set
+parameters = list(
+	m = 0.1, 		# intrinsic mortality
+	b = 0.5, 		# beta*eps 
[…]

### 2013-11-21 code/patchcounting.r

@@ -1,0 +1,115 @@
+
+x <- map
+
+#topo.colors(6, alpha = 1) rainbow(6)
+rgb.palette <- colorRampPalette(c("red", "orange", "blue", "green"), space = "rgb")
+cols = rgb.palette(6)[sample(1:6, 10000, replace = TRUE)]  # default color value
+  
+# plotting function for objects of class "landscape".
+plot.map <- function(x,   dims = c(100,100), grid = FALSE, axis = FALSE, add = FALSE, ani = FALSE, ...) {
+  lvls <- unique(x) 
+ 
+  nlev <- length(lvls)
+  
+  if(ani & Sys.info()[['sysname']] == "Windows") adj = -0.5 else adj = 0 #this adjustment constant is added when producing a pixel accurate png or gif file, requires TRUE when the function is used to plot animated figures. 
+  
+  if(!add) plot(NA,NA, xlim = c(0.5+adj, dims[1]+0.5+adj), ylim = c( dims[2]+0.5+adj, 0+0.5+adj), xaxs = "i", yaxs = "i",xlab = "", ylab = "", xaxt = "n", yaxt = "n") 
+ 
+  rect(rep(1:dims[1], times = dims[2])-.5, rep(1:dims[2], each = dims[1])-.5, rep(1:dims[1], times = dims[2])+.5, rep(1:dims[2], each = dims[1])+.5, col = cols[as.numeric(x)], border = NA)
+  
+ }
+
+ 
+ plot(map, add = TRUE)
+ plot(x_new)
+
+ 
+x <- x_new
+state = "+"
+
+# get patch size and patchsize distribution
+patches <- function(x, state, cumulative = TRUE) {
+	pattern <- x$cells
+	pattern <- pattern %in% state
+	map <- rep(NA, times = prod(x$dim))
+	old <- rep(99, times = prod(x$dim)) 
+	
+	class(map) <- c("numeric","map")
+	class(old) <- c("numeric","map")
+	
+	while(!identical(old[pattern], map[pattern])) {
+		old <- map
+		count = as.integer(1)
+		for(i in which(pattern)) {
+			neighbors <- map[x_with_border][x_to_evaluate[i]+interact]
+			if(all(is.na(neighbors)) ) { 
+				map[i] <- count
+			} else {
+				map[i] <- min(neighbors, na.rm = TRUE)
+			}
+				count <- count +1
+			}
+ plot(map, add =TRUE)
+
+		}
+	
+	map <- as.factor(map)
[…]

### 2013-11-21 code/patchdist_update.r

@@ -1,0 +1,233 @@
+
+#setwd("E:\\Eigene Dokumente\\Uni\\projects\\2013 CASCADE grazing\\sim9\\")
+setwd("C:\\Users\\SCHNEIDER\\Documents\\projects\\CASCADE\\2013 CASCADE grazing\\sim9\\")
+
+
+
+output <- read.csv("output.csv")
+output <- rbind(output, read.csv("output2.csv"))
+
+
+
+library(foreach)
+library(doSNOW)
+
+workerlist <- rep("localhost", times = 11)
+
+cl <- makeSOCKcluster(workerlist)
+
+registerDoSNOW(cl)
+
+
+foreach(i = output$ID) %dopar% {
+#i =   2603  #which(! 1:2640  %in% read.csv("output_fit.csv")$ID)
+
+# output$ID[which(output$rho_plus > 0.4  & output$rho_plus < 0.5)] 
+#i = 1405
+# output$ID[which(output$rho_plus > 0.68  & output$rho_plus < 0.7)] 
+#i = 2360 
+# output$ID[which(output$rho_plus > 0.02  & output$rho_plus < 0.1)] 
+#i = 900# 328 # 2220 # 
+
+load(paste("results\\result_", i, sep = ""))
+
+	logbins <-  10^seq(log10(1),log10(10000), .1)
+	result$logbin <- data.frame(size = logbins)
+	
+	for(j in 2:length(result$patches)) {
+		result$logbin[paste(((c(NA,snapshots)-1)*delta)[j])] <- sapply(1:length(logbins), function(k) length(which(result$patches[[j]] >= logbins[k] & result$patches[[j]] < logbins[k+1])) )
+	}
+	
+	result$logbin$mean <- apply(result$logbin[,2:length(result$patches)], 1, mean)
+	result$logbin$sd <- apply(result$logbin[,2:length(result$patches)], 1, sd)
+	
+	result$cumpatch <- list()
+	
+for(j in 2:length(result$patches)) {
+
+	if( !is.na(result$patch[j])) {
+	cumbins <- sort(unique(unlist(result$patches[j]))) 
+	#bins <- seq(1,10001, 10)
+
+		result$cumpatch[[j]] <- data.frame(size = cumbins)
[…]

### 2013-11-21 code/test_foreach.r

@@ -1,0 +1,30 @@
+
+rm(list=ls())
+
+	
+library(foreach)
+library(doSNOW)
+
+
+dir.create("results")
+
+workerlist <- rep("localhost", times = 23)
+
+cl <- makeSOCKcluster(workerlist)
+
+registerDoSNOW(cl)
+
+
+foreach(iteration = 1:46) %dopar% {
+#
+temp <- rnorm(20)
+save(temp, file = paste("results/temp", iteration, sep = "_"))
+	
+
+gc() 
+}
+
+
+
+stopCluster(cl)
+

### 2013-11-21 code/test_ind_grazing.r

@@ -1,0 +1,441 @@
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1, 
+	global = FALSE,
+	stock =  TRUE,
+	m = .1, 		# intrinsic mortality
+	g = .05,		#grazing
+	b = .7, 		# beta*eps 
+	d = 0.1,		# degradation
+	c_ = 0.2, 		# beta*g  
+	del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+	r = 0.01, 	# regeneration rate
+	f = 0.9		# local fascilitation
+)
+
+# specify lattice
+width = 100
+height = 100
+
+# initial cell states
+states = c("+","0","-")
+prob = c(6/10,2/10,2/10)
+color <- c("black","grey80", "white") # define colors for the cell state levels
+
+
+# derive helper vectors for counting: 
+# transformation vector for evaluation at the border of the grid
+	# set evaluation matrix 
+	X <- matrix(as.integer(1:(width*height)), ncol = width, byrow =TRUE)
+	# setting the border of the evaluation matrix X
+	X <- cbind(X[,width], X, X[,1] )  
+	X <- rbind(X[height,], X, X[1,] ) 
+	# transformation vector which adds the border to the lattice:
+	x_with_border <- as.integer(t(X))
+	
+	# from the matrix X (with copied border cells), which cells are the actual cells (reverse transformation vector of the previous lines) 
+	x_to_evaluate <- sort(matrix(1:prod(dim(X)), ncol = dim(X)[2], byrow =TRUE)[-c(1, dim(X)[1]), -c(1,dim(X)[2])]	)		
+
+# defining the neighborhood which is to be evaluated	
+	# set interaction matrix
+	I <- matrix(c(0,1,0,1,NA,1,0,1,0), ncol = 3, byrow = TRUE)	
+	# coordinates of neighbours in Interaction matrix I: 
+	neighbours_in_I <- which(is.finite(abs(I)/abs(I)), arr.in = TRUE)
+	# coordinates relative to the evaluated cell (=  which(is.na(I) ) 
+	relrow <- neighbours_in_I[,1]-which(is.na(I), arr.ind = TRUE)[1]
+	relcol <- neighbours_in_I[,2]-which(is.na(I), arr.ind = TRUE)[2]
+	
+	# relative position of the four direct neighbours of a cell
+	interact <- (relrow * dim(X)[2] + relcol)
+
+	
[…]

### 2013-11-21 code/unstable_equilibria.r

@@ -1,0 +1,268 @@
+
+
+###########################################
+###########################################
+###########################################
+###########################################
+###########################################
+###########################################
+# unstable equilibria
+# 
+
+
+
+rm(list=ls())
+
+########################################################################################
+	
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# plotting function for objects of class "landscape".
+plot.landscape <- function(x, grid = FALSE, axis = FALSE, cols = "auto", add = FALSE, ani = FALSE, ...) {
+  lvls <- levels(x$cells) 
+  nlev <- length(lvls)
+  if(cols[1] == "auto") cols = c("black","grey80", "white")  # default color value
+  
+  if(ani & Sys.info()[['sysname']] == "Windows") adj = -0.5 else adj = 0 #this adjustment constant is added when producing a pixel accurate png or gif file, requires TRUE when the function is used to plot animated figures. 
+  
+  if(!add) plot(NA,NA, xlim = c(0.5+adj, x$dim[1]+0.5+adj), ylim = c( x$dim[2]+0.5+adj, 0+0.5+adj), bty = c("n", "o")[grid+1], xaxs = "i", yaxs = "i",xlab = "", ylab = "", xaxt = "n", yaxt = "n", ... ) 
+  
+  if(axis && !add) axis(3) 
+  if(axis && !add) axis(2)
+  
+  if(grid) border = "grey80" else border = cols[as.numeric(x$cells)]
+  
+  rect(rep(1:x$dim[1], times = x$dim[2])-.5, rep(1:x$dim[2], each = x$dim[1])-.5, rep(1:x$dim[1], times = x$dim[2])+.5, rep(1:x$dim[2], each = x$dim[1])+.5, col = cols[as.numeric(x$cells)], border = border)
+  
+  if(grid) box()
+ }
+
+# specify lattice
+width = 100
[…]

### 2013-11-21 code/unstable_equilibria_realstructure.r

@@ -1,0 +1,251 @@
+
+
+###########################################
+###########################################
+###########################################
+###########################################
+###########################################
+###########################################
+# unstable equilibria
+# 
+
+
+
+rm(list=ls())
+
+########################################################################################
+	
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# plotting function for objects of class "landscape".
+plot.landscape <- function(x, grid = FALSE, axis = FALSE, cols = "auto", add = FALSE, ani = FALSE, ...) {
+  lvls <- levels(x$cells) 
+  nlev <- length(lvls)
+  if(cols[1] == "auto") cols = c("black","grey80", "white")  # default color value
+  
+  if(ani & Sys.info()[['sysname']] == "Windows") adj = -0.5 else adj = 0 #this adjustment constant is added when producing a pixel accurate png or gif file, requires TRUE when the function is used to plot animated figures. 
+  
+  if(!add) plot(NA,NA, xlim = c(0.5+adj, x$dim[1]+0.5+adj), ylim = c( x$dim[2]+0.5+adj, 0+0.5+adj), bty = c("n", "o")[grid+1], xaxs = "i", yaxs = "i",xlab = "", ylab = "", xaxt = "n", yaxt = "n", ... ) 
+  
+  if(axis && !add) axis(3) 
+  if(axis && !add) axis(2)
+  
+  if(grid) border = "grey80" else border = cols[as.numeric(x$cells)]
+  
+  rect(rep(1:x$dim[1], times = x$dim[2])-.5, rep(1:x$dim[2], each = x$dim[1])-.5, rep(1:x$dim[1], times = x$dim[2])+.5, rep(1:x$dim[2], each = x$dim[1])+.5, col = cols[as.numeric(x$cells)], border = border)
+  
+  if(grid) box()
+ }
+
+# specify lattice
+width = 100
[…]

### 2013-11-21 data/analyses/grazing_density.r

@@ -1,0 +1,391 @@
+
+
+
+# specify lattice
+width = 100
+height = 100
+
+states = c("+","0","-")
+color <- c("black","grey80", "white") # define colors for the cell state levels
+
+parms_temp <- list(
+	m = 0.05,
+	g = 0.099,
+		b = .6 ,		# beta*eps 
+		d = 0.1,		# degradation
+		c_ = 0.01, 		# beta*g  
+		del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+		r = 0.01, 	# regeneration rate
+		f = 0.9 		# local fascilitation
+)
+
+
+
+# time and resolution of simulation
+timesteps = 5
+delta = 1/10
+
+
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+
+
+# derive helper vectors for counting: 
+# transformation vector for evaluation at the border of the grid
+	# set evaluation matrix 
+	X <- matrix(as.integer(1:(width*height)), ncol = width, byrow =TRUE)
+	# setting the border of the evaluation matrix X
+	X <- cbind(X[,width], X, X[,1] )  
+	X <- rbind(X[height,], X, X[1,] ) 
+	# transformation vector which adds the border to the lattice:
+	x_with_border <- as.integer(t(X))
+	
+	# from the matrix X (with copied border cells), which cells are the actual cells (reverse transformation vector of the previous lines) 
+	x_to_evaluate <- sort(matrix(1:prod(dim(X)), ncol = dim(X)[2], byrow =TRUE)[-c(1, dim(X)[1]), -c(1,dim(X)[2])]	)		
+
+# defining the neighborhood which is to be evaluated	
+	# set interaction matrix
+	I <- matrix(c(0,1,0,1,NA,1,0,1,0), ncol = 3, byrow = TRUE)	
+	# coordinates of neighbours in Interaction matrix I: 
+	neighbours_in_I <- which(is.finite(abs(I)/abs(I)), arr.in = TRUE)
[…]

### 2013-11-21 data/analyses/grazing_timeseries.r

@@ -1,0 +1,518 @@
+
+rm(list=ls())
+
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+#setwd("/home/schneider/herbivory/")
+#setwd("C:\\Users\\SCHNEIDER\\Documents\\CASCADE\\2013_grazingmodel\\corrected_grazing_vul\\sim5")
+#setwd("E:\\Eigene Dokumente\\Uni\\projects\\2013 CASCADE grazing")
+
+env <- 	unique(c(seq(0.2,1,.02)) )#seq(0.2,1,.025)
+graz <- c(.05, .15, .25)*0.66  #c(0.0375, 0.1125, 0.1875)
+#graz <- seq(0,.1, 0.005)#seq(0,.9, 0.075)
+mort <- 0.05 #seq(0.1,.3, 0.1)
+init <- c("low", "high")
+global <- c(TRUE, FALSE)
+stock <- c(TRUE, FALSE)
+
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+linit <- 2
+lglobal <- 2
+lstock <- 2
+
+
+#global <- rep(global, each = lgraz*lenv*lmort*linit)
+#init <- rep(rep(init, each = lgraz*lenv*lmort), times = lglobal)
+#mort <-	rep(rep(mort, each = lgraz*lenv), times = lglobal*linit)
+#graz <-	 rep(rep(graz, each = lenv), times = lmort*lglobal*linit)
+#env <-  rep(env, times = lgraz*lmort*lglobal*linit)
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1:(lgraz*lmort*lglobal*linit*lenv*lstock), 
+	global = rep(global, each = lgraz*lenv*lmort*linit*lstock),
+	stock =  rep(rep(stock, each = lgraz*lenv*lmort*linit), times = lstock),
+	starting = rep(rep(init, each = lgraz*lenv*lmort), times = lglobal*lstock),
+	m = rep(rep(mort, each = lgraz*lenv), times = lglobal*linit*lstock), 		# intrinsic mortality
+	g = rep(rep(graz, each = lenv), times = lmort*lglobal*linit*lstock),		#grazing
+	b = rep(env, times = lgraz*lmort*lglobal*linit*lstock), 		# beta*eps 
+	d = 0.1,		# degradation
[…]

### 2013-11-21 data/old_runs/corrected_grazing/sim1/simulation_grazing_corrected.r

@@ -1,0 +1,595 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+rm(list=ls())
+
+setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\corrected_grazing")
+
+env <- 	seq(0.2,1,.05)
+graz <- seq(0,.9, 0.075)
+mort <- seq(0.1,.3, 0.1)
+init <- c("low", "high")
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+
+env <-	rep(rep(env, times = lgraz), times = lmort)
+graz <-	 rep(rep(graz, each = lenv), times = lmort)
+mort <-  rep(mort, each = lgraz*lenv)
+
+replicates = 2
+init <- rep(init, each = length(mort) )
+env <- rep(env, times = replicates)
+graz <- rep(graz, times = replicates)
+mort <- rep(mort, times = replicates)
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1:length(graz), 
+	starting = init,
+	g = graz,		#grazing
+	m = mort, 		# intrinsic mortality
+	b = env, 		# beta*eps 
+	d = 0.1,		# degradation
+	c_ = 0.2, 		# beta*g  
+	del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+	r = 0.01, 	# regeneration rate
+	f = 0.9		# local fascilitation
+)
+
+
+# defining parameter set
+parameters_ini = list(
+	low = list(
+		m = 0.1, 		# intrinsic mortality
+		b = 0.25, 		# beta*eps 
+		d = 0.1,		# degradation
+		c_ = 0.2, 		# beta*g  
+		del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+		r = 0.01, 	# regeneration rate
+		f = 0.9, 		# local fascilitation
+		g = 0		#grazing
+		), 
[…]

### 2013-11-21 data/old_runs/corrected_grazing/sim2/simulation_grazing_corrected.r

@@ -1,0 +1,635 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+rm(list=ls())
+
+setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\corrected_grazing")
+
+env <- 	seq(0.2,1,.025)
+graz <- seq(0,.5, 0.025)
+mort <- 0.1 # seq(0.1,.3, 0.1)
+init <- c("low", "high")
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+
+env <-	rep(rep(env, times = lgraz), times = lmort)
+graz <-	 rep(rep(graz, each = lenv), times = lmort)
+mort <-  rep(mort, each = lgraz*lenv)
+
+replicates = 2
+init <- rep(init, each = length(mort) )
+env <- rep(env, times = replicates)
+graz <- rep(graz, times = replicates)
+mort <- rep(mort, times = replicates)
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1:length(graz), 
+	starting = init,
+	g = graz,		#grazing
+	m = mort, 		# intrinsic mortality
+	b = env, 		# beta*eps 
+	d = 0.1,		# degradation
+	c_ = 0.2, 		# beta*g  
+	del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+	r = 0.01, 	# regeneration rate
+	f = 0.9		# local fascilitation
+)
+
+
+# defining parameter set
+parameters_ini = list(
+	low = list(
+		m = 0.1, 		# intrinsic mortality
+		b = 0.25, 		# beta*eps 
+		d = 0.1,		# degradation
+		c_ = 0.2, 		# beta*g  
+		del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+		r = 0.01, 	# regeneration rate
+		f = 0.9, 		# local fascilitation
[…]

### 2013-11-21 data/old_runs/corrected_grazing/sim3/simulation_grazing_corrected.r

@@ -1,0 +1,624 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+rm(list=ls())
+
+setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\corrected_grazing")
+
+env <- 	seq(0.2,1,.025)
+graz <- seq(0,.5, 0.025)
+mort <- 0.1 # seq(0.1,.3, 0.1)
+init <- c("low", "high")
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+
+env <-	rep(rep(env, times = lgraz), times = lmort)
+graz <-	 rep(rep(graz, each = lenv), times = lmort)
+mort <-  rep(mort, each = lgraz*lenv)
+
+replicates = 2
+init <- rep(init, each = length(mort) )
+env <- rep(env, times = replicates)
+graz <- rep(graz, times = replicates)
+mort <- rep(mort, times = replicates)
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1:length(graz), 
+	starting = init,
+	g = graz,		#grazing
+	m = mort, 		# intrinsic mortality
+	b = env, 		# beta*eps 
+	d = 0.1,		# degradation
+	c_ = 0.2, 		# beta*g  
+	del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+	r = 0.01, 	# regeneration rate
+	f = 0.9		# local fascilitation
+)
+
+
+# defining parameter set
+parameters_ini = list(
+	low = list(
+		m = 0.1, 		# intrinsic mortality
+		b = 0.25, 		# beta*eps 
+		d = 0.1,		# degradation
+		c_ = 0.2, 		# beta*g  
+		del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+		r = 0.01, 	# regeneration rate
+		f = 0.9, 		# local fascilitation
[…]

### 2013-11-21 data/old_runs/corrected_grazing/simulation_grazing_corrected.r

@@ -1,0 +1,627 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+rm(list=ls())
+
+setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\corrected_grazing")
+setwd("C:\\Users\\SCHNEIDER\\Documents\\CASCADE\\2013_grazingmodel\\corrected_grazing")
+env <- 	seq(0.2,1,.025)
+graz <- seq(0,.5, 0.025)
+mort <- 0.1 # seq(0.1,.3, 0.1)
+init <- c("low", "high")
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+
+env <-	rep(rep(env, times = lgraz), times = lmort)
+graz <-	 rep(rep(graz, each = lenv), times = lmort)
+mort <-  rep(mort, each = lgraz*lenv)
+
+replicates = 2
+init <- rep(init, each = length(mort) )
+env <- rep(env, times = replicates)
+graz <- rep(graz, times = replicates)
+mort <- rep(mort, times = replicates)
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1:length(graz), 
+	starting = init,
+	g = graz,		#grazing
+	m = mort, 		# intrinsic mortality
+	b = env, 		# beta*eps 
+	d = 0.1,		# degradation
+	c_ = 0.2, 		# beta*g  
+	del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+	r = 0.01, 	# regeneration rate
+	f = 0.9		# local fascilitation
+)
+
+
+# defining parameter set
+parameters_ini = list(
+	low = list(
+		m = 0.1, 		# intrinsic mortality
+		b = 0.25, 		# beta*eps 
+		d = 0.1,		# degradation
+		c_ = 0.2, 		# beta*g  
+		del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
[…]

### 2013-11-21 data/old_runs/corrected_grazing/simulation_grazing_corrected_benchmark.r

@@ -1,0 +1,60 @@
+library(rbenchmark)
+
+benchmark(
+check1 = {		if(i == addgrazing/delta ) parms_temp <- as.list(parameters[iteration,]) },
+replace1 = {		x_new <- x_old 	},
+getrho = {		parms_temp$rho <- result$rho[[1]][i-1] },
+countQplus = { 		parms_temp$Q_plus <- count(x_old, "+")/4   },
+countQempty = {		parms_temp$Q_unveg <- ( count(x_old, c("0", "-")))/4 },
+		#parms_temp$vul <- sum(parms_temp$Q_unveg[x_old$cells == "+"]*4)/(width*height)
+getrandomnum = {		rnum <- runif(width*height) },
+	
+calcprobs = {	
+		recolonisation <- with(parms_temp, (del*rho+(1-del)*Q_plus)*(b-c_*rho)*delta)
+		degradation <- with(parms_temp, (d *delta))
+		death <- with(parms_temp, (m+g*Q_unveg)*delta)
+		regeneration <- with(parms_temp, (r + f*Q_plus)*delta)
+		},
+		
+checkprobs = {		if(any(c(recolonisation+degradation, death, regeneration) > .95 )) warning(paste("a set probability is exceeding 1 in run", iteration, "time step", i, "! decrease delta!!!")) 
+		if(any(c(recolonisation, degradation, death, regeneration) < 0)) warning(paste("a set probability falls below 0 in run", iteration, "in time step", i, "! balance parameters!!!")) 
+		},
+applyrules = {		
+		x_new$cells[which(x_old$cells == "0" & rnum <= recolonisation)] <- "+"
+		x_new$cells[which(x_old$cells == "0" & rnum > recolonisation & rnum <= recolonisation+degradation)] <- "-"
+		x_new$cells[which(x_old$cells == "+" & rnum <= death)] <- "0"
+		x_new$cells[which(x_old$cells == "-" & rnum <= regeneration)] <- "0"
+},
+
+savemort = {
+		result$mortality[i] <- mean(death[x_old$cells == "+"], na.rm = TRUE)/delta
+		},
+		
+savebordermort = {		result$mortality_border[i] <- mean(death[x_old$cells == "+" & parms_temp$Q_unveg > 0], na.rm = TRUE)/delta
+		},
+
+saverho = {		
+		for(j in 1:length(states)) {
[…]

### 2013-11-21 data/old_runs/corrected_grazing_vul/fitting_psd.r

@@ -1,0 +1,115 @@
+
+#setwd("/home/schneider/herbivory/")
+setwd("C:\\Users\\SCHNEIDER\\Documents\\CASCADE\\2013_grazingmodel\\corrected_grazing_vul\\sim4")
+
+load(paste("results\\result", 638, sep = "_"))
+
+sort(unlist(result$patches))
+
+result$cumpatch$size
+plot(mean~size, data = result$cumpatch, log= "xy", pch = 20)
+
+lmean <- mean(log(sort(unlist(result$patches))))
+lstd  <- sd(log(sort(unlist(result$patches))))
+
+logncdf <- function(x, lmean, lstd) 1/(std*sqrt(2*pi)*sum(exp(-(log())) seq())) 
+P <- ecdf(sort(unlist(result$patches)))
+
+qlnorm(sort(unlist(result$patches)))
+
+
+
+	bins <- sort(unique(unlist(result$patches[-1])))
+	bins <- seq(1,501, 10)
+	bins <- 10^seq(log10(1),log10(10000), .1)
+	plot(NA,NA, xlim = c(1,10000), ylim = c(1,1000), log = "xy")
+	for(i in 3:22) {
+	freq <- sapply(bins, function(j) length(which(result$patches[[i]] >= j)) )
+	
+	points(freq~bins, pch = 20, type = "b")
+}
+
+	with(result$cumpatch2, lines(mean~size, lwd = 3))
+	with(result$cumpatch2, lines(c(I(mean+sd),rev(I(mean-sd)))~c(size, rev(size)), lty = 2))
+
+	bins <- sort(unique(unlist(result$patches[-1])))
+	plot(NA,NA, log = "xy", xlim = c(1,1000), ylim = c(1,1000))
+	for(i in 3:22) {
+	freq <- sapply(bins, function(j) length(which(result$patches[[i]] == j)) )
+	
+	points(freq~bins, pch = 20)
+}
+
+
+
+# time and resolution of simulation
+timesteps = 3000
+addgrazing = 1000
+delta = 1/5
+
+snapshots <- c(1, c(addgrazing, timesteps-20:1*50, timesteps)/delta+1)
+
+plot(mean~size, type = "p", pch = 20, data = result$cumpatch, log = "xy")
+
+
+dd1 <- data.frame(size = NA, n = NA )[-1,]
+
+	for(i in as.character((snapshots-1)*delta)[-(1:2)] ) {
[…]

### 2013-11-21 data/old_runs/corrected_grazing_vul/sim2/simulation_grazing_corrected_vul.r

@@ -1,0 +1,512 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+rm(list=ls())
+
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+#setwd("/home/schneider/herbivory/")
+#setwd("C:\\Users\\SCHNEIDER\\Documents\\CASCADE\\2013_grazingmodel\\corrected_grazing_vul")
+
+env <- 	seq(0.2,1,.04)#seq(0.2,1,.025)
+graz <- seq(0,.1, 0.005)#seq(0,.9, 0.075)
+mort <- 0.1 #seq(0.1,.3, 0.1)
+init <- c("low", "high")
+global <- c(TRUE, FALSE)
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+linit <- 2
+lglobal <- 2
+
+
+#global <- rep(global, each = lgraz*lenv*lmort*linit)
+#init <- rep(rep(init, each = lgraz*lenv*lmort), times = lglobal)
+#mort <-	rep(rep(mort, each = lgraz*lenv), times = lglobal*linit)
+#graz <-	 rep(rep(graz, each = lenv), times = lmort*lglobal*linit)
+#env <-  rep(env, times = lgraz*lmort*lglobal*linit)
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1:(lgraz*lmort*lglobal*linit*lenv), 
+	global = rep(global, each = lgraz*lenv*lmort*linit),
+	starting = rep(rep(init, each = lgraz*lenv*lmort), times = lglobal),
+	g = rep(rep(graz, each = lenv), times = lmort*lglobal*linit),		#grazing
+	m = rep(rep(mort, each = lgraz*lenv), times = lglobal*linit), 		# intrinsic mortality
+	b = rep(env, times = lgraz*lmort*lglobal*linit), 		# beta*eps 
+	d = 0.1,		# degradation
+	c_ = 0.2, 		# beta*g  
[…]

### 2013-11-21 data/old_runs/corrected_grazing_vul/sim3/simulation_grazing_corrected_vul.r

@@ -1,0 +1,649 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+rm(list=ls())
+
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+#setwd("/home/schneider/herbivory/")
+setwd("C:\\Users\\SCHNEIDER\\Documents\\CASCADE\\2013_grazingmodel\\corrected_grazing_vul")
+
+env <- 	seq(0.2,1,.04)#seq(0.2,1,.025)
+graz <- seq(0,.1, 0.005)#seq(0,.9, 0.075)
+mort <- 0.1 #seq(0.1,.3, 0.1)
+init <- c("low", "high")
+global <- c(TRUE, FALSE)
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+linit <- 2
+lglobal <- 2
+
+
+#global <- rep(global, each = lgraz*lenv*lmort*linit)
+#init <- rep(rep(init, each = lgraz*lenv*lmort), times = lglobal)
+#mort <-	rep(rep(mort, each = lgraz*lenv), times = lglobal*linit)
+#graz <-	 rep(rep(graz, each = lenv), times = lmort*lglobal*linit)
+#env <-  rep(env, times = lgraz*lmort*lglobal*linit)
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1:(lgraz*lmort*lglobal*linit*lenv), 
+	global = rep(global, each = lgraz*lenv*lmort*linit),
+	starting = rep(rep(init, each = lgraz*lenv*lmort), times = lglobal),
+	g = rep(rep(graz, each = lenv), times = lmort*lglobal*linit),		#grazing
+	m = rep(rep(mort, each = lgraz*lenv), times = lglobal*linit), 		# intrinsic mortality
+	b = rep(env, times = lgraz*lmort*lglobal*linit), 		# beta*eps 
+	d = 0.1,		# degradation
+	c_ = 0.2, 		# beta*g  
[…]

### 2013-11-21 data/old_runs/corrected_grazing_vul/simulation_grazing_corrected_vul.r

@@ -1,0 +1,676 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+rm(list=ls())
+
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+#setwd("/home/schneider/herbivory/")
+#setwd("C:\\Users\\SCHNEIDER\\Documents\\CASCADE\\2013_grazingmodel\\corrected_grazing_vul")
+
+env <- 	sort(unique(c(seq(0.2,1,.04), seq(0.4,0.6, 0.02))) )#seq(0.2,1,.025)
+graz <- seq(0,.1, 0.005)#seq(0,.9, 0.075)
+mort <- 0.1 #seq(0.1,.3, 0.1)
+init <- c("low", "high")
+global <- c(TRUE, FALSE)
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+linit <- 2
+lglobal <- 2
+
+
+#global <- rep(global, each = lgraz*lenv*lmort*linit)
+#init <- rep(rep(init, each = lgraz*lenv*lmort), times = lglobal)
+#mort <-	rep(rep(mort, each = lgraz*lenv), times = lglobal*linit)
+#graz <-	 rep(rep(graz, each = lenv), times = lmort*lglobal*linit)
+#env <-  rep(env, times = lgraz*lmort*lglobal*linit)
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1:(lgraz*lmort*lglobal*linit*lenv), 
+	global = rep(global, each = lgraz*lenv*lmort*linit),
+	starting = rep(rep(init, each = lgraz*lenv*lmort), times = lglobal),
+	g = rep(rep(graz, each = lenv), times = lmort*lglobal*linit),		#grazing
+	m = rep(rep(mort, each = lgraz*lenv), times = lglobal*linit), 		# intrinsic mortality
[…]

### 2013-11-21 data/old_runs/corrected_grazing_vul/simulation_grazing_corrected_vul_analyses.r

@@ -1,0 +1,703 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+rm(list=ls())
+
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+#setwd("/home/schneider/herbivory/")
+setwd("C:\\Users\\SCHNEIDER\\Documents\\CASCADE\\2013_grazingmodel\\corrected_grazing_vul")
+
+env <- 	seq(0.2,1,.04)#seq(0.2,1,.025)
+graz <- seq(0,.1, 0.005)#seq(0,.9, 0.075)
+mort <- 0.1 #seq(0.1,.3, 0.1)
+init <- c("low", "high")
+global <- c(TRUE, FALSE)
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+linit <- 2
+lglobal <- 2
+
+
+#global <- rep(global, each = lgraz*lenv*lmort*linit)
+#init <- rep(rep(init, each = lgraz*lenv*lmort), times = lglobal)
+#mort <-	rep(rep(mort, each = lgraz*lenv), times = lglobal*linit)
+#graz <-	 rep(rep(graz, each = lenv), times = lmort*lglobal*linit)
+#env <-  rep(env, times = lgraz*lmort*lglobal*linit)
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1:(lgraz*lmort*lglobal*linit*lenv), 
+	global = rep(global, each = lgraz*lenv*lmort*linit),
+	starting = rep(rep(init, each = lgraz*lenv*lmort), times = lglobal),
+	g = rep(rep(graz, each = lenv), times = lmort*lglobal*linit),		#grazing
+	m = rep(rep(mort, each = lgraz*lenv), times = lglobal*linit), 		# intrinsic mortality
+	b = rep(env, times = lgraz*lmort*lglobal*linit), 		# beta*eps 
+	d = 0.1,		# degradation
+	c_ = 0.2, 		# beta*g  
[…]

### 2013-11-21 data/old_runs/corrected_grazing_vul/simulation_grazing_corrected_vul_linux.r

@@ -1,0 +1,650 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+rm(list=ls())
+
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+#setwd("/home/schneider/herbivory/")
+setwd("C:\\Users\\SCHNEIDER\\Documents\\CASCADE\\2013_grazingmodel\\corrected_grazing_vul\\sim4")
+
+env <- 	sort(unique(c(seq(0.2,1,.04), seq(0.4,0.6, 0.02))) )#seq(0.2,1,.025)
+graz <- seq(0,.1, 0.005)#seq(0,.9, 0.075)
+mort <- 0.1 #seq(0.1,.3, 0.1)
+init <- c("low", "high")
+global <- c(TRUE, FALSE)
+stock <- c(TRUE, FALSE)
+
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+linit <- 2
+lglobal <- 2
+lstock <- 2
+
+
+#global <- rep(global, each = lgraz*lenv*lmort*linit)
+#init <- rep(rep(init, each = lgraz*lenv*lmort), times = lglobal)
+#mort <-	rep(rep(mort, each = lgraz*lenv), times = lglobal*linit)
+#graz <-	 rep(rep(graz, each = lenv), times = lmort*lglobal*linit)
+#env <-  rep(env, times = lgraz*lmort*lglobal*linit)
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1:(lgraz*lmort*lglobal*linit*lenv), 
+	global = rep(global, each = lgraz*lenv*lmort*linit),
+	starting = rep(rep(init, each = lgraz*lenv*lmort), times = lglobal),
+	g = rep(rep(graz, each = lenv), times = lmort*lglobal*linit),		#grazing
[…]

### 2013-11-21 data/old_runs/corrected_grazing_vul/update_patchdist.r

@@ -1,0 +1,257 @@
+setwd("C:\\Users\\SCHNEIDER\\Documents\\CASCADE\\2013_grazingmodel\\corrected_grazing_vul")
+
+env <- 	seq(0.2,1,.04)#seq(0.2,1,.025)
+graz <- seq(0,.1, 0.005)#seq(0,.9, 0.075)
+mort <- 0.1 #seq(0.1,.3, 0.1)
+init <- c("low", "high")
+global <- c(TRUE, FALSE)
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+linit <- 2
+lglobal <- 2
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1:(lgraz*lmort*lglobal*linit*lenv), 
+	global = rep(global, each = lgraz*lenv*lmort*linit),
+	starting = rep(rep(init, each = lgraz*lenv*lmort), times = lglobal),
+	g = rep(rep(graz, each = lenv), times = lmort*lglobal*linit),		#grazing
+	m = rep(rep(mort, each = lgraz*lenv), times = lglobal*linit), 		# intrinsic mortality
+	b = rep(env, times = lgraz*lmort*lglobal*linit), 		# beta*eps 
+	d = 0.1,		# degradation
+	c_ = 0.2, 		# beta*g  
+	del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+	r = 0.01, 	# regeneration rate
+	f = 0.9		# local fascilitation
+)
+
+
+# specify lattice
+width = 100
+height = 100
+
+# initial cell states
+states = c("+","0","-")
+prob = c(9/10,.9/10,0.1/10)
+color <- c("black","grey80", "white") # define colors for the cell state levels
+
+
+highlight <- function(x, colrange = c("black", "red2"), steps = NULL, range = "auto"){
+
+if(is.null(steps)) steps = length(unique(x))
+if(range == "auto") {
+	min_val = min(x, na.rm = TRUE)
+	max_val = max(x, na.rm = TRUE)
+	} else {
+	min_val = range[1]
+	max_val = range[2]
+	}
+
+colorscale <- colorRampPalette(colrange, space = "rgb")
+cols <- colorscale( steps)
+
+cols[as.integer((x-min_val)/(max_val-min_val)*(steps-1)+1)]
+
+}
+
+count  <- function(x, neighbor, state = NULL) {
[…]

### 2013-11-21 data/old_runs/example_fig.r

@@ -1,0 +1,344 @@
+
+### example
+width = 20
+height = 20
+
+ex1 <- list(  
+	dim = c(5, 5),  # first element contains the dimensions of the landscape 
+	cells = sample(factor(1:length(states)), 25, replace = T, prob = c(0.3,0.3,0.4) ) #second element contains a random row-wise, factorial vector to fill the grid 
+	)
+levels(ex1$cells) <- states  #assign cell states 
+class(ex1) <- c("list","landscape") # set class of object (required for plotting)
+
+pdf("E:\\SkyDrive\\Uni\\talks\\ex_grids.pdf")
+{
+par(mfrow =c(1,1), mar = c(4,4,4,4))
+
+grd <- ex1 
+	vul <- .66
+	feeding <- rep((0.05 + 0.033/vul), times = length(which(grd$cells == "+")))  
+
+		breaks <- seq(.01,.25, length = 31) #10^seq(  -1, log10(0.11 ),length = 11) #
+
+		grd$cells <- as.numeric(grd$cells)+31
+		grd$cells[grd$cells == 32] <- findInterval(feeding, breaks)+1
+		grd$cells <- as.factor(grd$cells)
+		levels(grd$cells) <- c(levels(grd$cells), as.character(which(! as.character(1:34) %in% levels(grd$cells))))
+
+		plot(grd, grid = FALSE, cols = grazingcol[as.numeric(levels(grd$cells))], ani = FALSE)
+		
+
+grd <- ex1 
+	Q <- count(grd, "+" )/4
+	vul <- 0.165
+	feeding <- (0.05 + 0.033/vul*(1-Q)[grd$cells == "+"] )
+
+
+		breaks <- seq(.01,.25, length = 31) #10^seq(  -1, log10(0.11 ),length = 11) #
+
+		grd$cells <- as.numeric(grd$cells)+31
+		grd$cells[grd$cells == 32] <- findInterval(feeding, breaks)+1
+		grd$cells <- as.factor(grd$cells)
+		levels(grd$cells) <- c(levels(grd$cells), as.character(which(! as.character(1:34) %in% levels(grd$cells))))
+
+		plot(grd, grid = FALSE, cols = grazingcol[as.numeric(levels(grd$cells))], ani = FALSE)
+	
+
+
+grd <- ex1 
+	vul <- length(which(grd$cells == "+"))
[…]

### 2013-11-21 data/old_runs/grazing.r

@@ -1,0 +1,154 @@
+#################################################
+#
+#		model: random walk
+#
+#
+#author: Flo
+#date: 19.02.2013
+#
+#################################################
+rm(list=ls())
+#setwd("/home/flo/Desktop/cellularautomata")
+#setwd("E:\\Eigene Dateien\\Dropbox\\cellular automata\\")
+source("E:\\Dropbox\\SharedWithMarina\\code\\cellularautomata5.r")
+source("C:\\Users\\SCHNEIDER\\Dropbox\\SharedWithMarina\\code\\cellularautomata5.r")
+
+
+initial <- result$timeseries[[2501]]
+#initial <- landscape(100,100, levels = c("+","-","0"), prob = c(2/8,3/8))
+par(mar = c(0,0,0,0))
+plot(initial, grid = FALSE, col = c("black", "white", "grey80"))
+#plotcounts(initial, setcountparms(initial), "+")
+
+# time and resolution of simulation
+timesteps = 100
+delta = 1/10
+
+# defining transition rules
+rm(rules)
+setrule("0", "+", "recolonise", rule = "(del*rho+(1-del)*Q_plus)*(b-c_*rho)", ruleset = "rules")
+setrule("+", "0", "mortality", rule = "m", ruleset = "rules")
+setrule("+", "0", "grazing", rule = "g*(Q_minus+Q_zero)*rho_G", ruleset = "rules")
+
+setrule("0", "-", "degrade", rule = "d", ruleset = "rules")
+setrule("-", "0", "regenerate", rule = "r + f*Q_plus", ruleset = "rules")
+
+rules
+
+
+# defining parameter set
+parameters = list(
+	n = 10, 		# number of goats
+	g = 0.05,		# increased mortality due to individual grazing per year. goats eat 5% of body weight per day, 30 kg ~ 1.5 kg / day.
+	m = 0.02, 		# intrinsic mortality
+	b = 0.4, 		# beta*eps 
+	d = 0.1,		# degradation
+	c_ = 0.2, 		# beta*g  
+	del = 0.15,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+	r = 0.0001, 	# regeneration rate
+	f = 0.9 		# local fascilitation
+)
+
+countpar <- setcountparms(initial)
+
+
+
+# initialising result list 
+result <- list()   # create an output list and
+			result$dim <- initial$dim  # writes dimensions from initial object
[…]

### 2013-11-21 data/old_runs/grazing_density_smiple.r

@@ -1,0 +1,55 @@
+
+# specify lattice
+width = 100
+height = 100
+
+states = c("+","0","-")
+color <- c("black","grey80", "white") # define colors for the cell state levels
+
+parms_temp <- list(
+	m = 0.1,
+	g = .1
+
+)
+
+
+res = 200
+death <- data.frame(dens = seq(0,1, length=res))
+
+for(i in 1:res) {
+
+j = i/res 
+prob = c( i/res ,1- i/res ,0)
+
+#### first run with grazing
+# sampling the initial random grid into a list object
+initial <- list(  
+	dim = c(as.integer(width), as.integer(height)),  # first element contains the dimensions of the landscape 
+	cells = sample(factor(1:length(states)), width*height, replace = T, prob = prob ) #second element contains a random row-wise, factorial vector to fill the grid 
+	)
+levels(initial$cells) <- states  #assign cell states 
+class(initial) <- c("list","landscape") # set class of object (required for plotting)
+
+
+parms_temp$rho  <-  sum(initial$cells == "+")/(width*height)
+parms_temp$Q_plus <- count(initial, "+")/4   
+parms_temp$vul <- sum((1-parms_temp$Q_plus)[initial$cells == "+"])/(width*height)
+
+
+#if(parms_temp$vul == 0) parms_temp$vul = 1/(width*height)
+
+death$parent[i] <- with(parms_temp, (m+g/0.75))
+death$stock[i] <- with(parms_temp, (m+g/rho))
+death$assoc[i] <- mean(with(parms_temp, (m+g*(1-Q_plus)/0.1888)))
+death$stockassoc[i] <- mean(with(parms_temp, (m+g*(1-Q_plus)/vul)))
+
+
+}
+
+plot(NA,NA, xlim = c(0,1), ylim = c(0,1), ylab = "risk of death", xlab = "vegetation cover")
+
+lines(parent*dens ~ dens, data = death, col = "black", lwd = 2)
+
+lines(stock*dens ~ dens, data = death, col = "red3", lwd = 3)
+lines(assoc*dens ~ dens, data = death, col = "blue", lwd = 2)
+lines(stockassoc*dens ~ dens, data = death, col = "green2", lwd = 1)

### 2013-11-21 data/old_runs/grazing_gradient/transitions_baseline3.r

@@ -1,0 +1,342 @@
+################################################
+#
+#		model: iterator for cellular automata
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#
+#################################################
+
+
+
+rm(list=ls())
+#source("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
+			# cumulate the result for each of the four neighboring cells (for(k in interact))
+			} else neighbors <- 0
+			
+			
+			mean(neighbors/4, na.rm = TRUE)	
+			# devide by 4 to get density
[…]

### 2013-11-21 data/old_runs/grazing_gradient/transitions_baseline4.r

@@ -1,0 +1,393 @@
+################################################
+#
+#		model: iterator for cellular automata
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#
+#################################################
+
+
+
+rm(list=ls())
+#source("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
+			# cumulate the result for each of the four neighboring cells (for(k in interact))
+			} else neighbors <- 0
+			
+			
+			mean(neighbors/4, na.rm = TRUE)	
+			# devide by 4 to get density
[…]

### 2013-11-21 data/old_runs/grazing_gradient/transitions_corrected.r

@@ -1,0 +1,397 @@
+################################################
+#
+#		model: iterator for cellular automata
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#
+#################################################
+
+
+
+rm(list=ls())
+#source("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
+			# cumulate the result for each of the four neighboring cells (for(k in interact))
+			} else neighbors <- 0
+			
+			
+			mean(neighbors/4, na.rm = TRUE)	
+			# devide by 4 to get density
[…]

### 2013-11-21 data/old_runs/herbivorymodel.r

@@ -1,0 +1,265 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+
+rm(list=ls())
+source("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+source("E:\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+setwd("E:\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+# specify lattice
+width = 100
+height = 100
+
+# initial cell states
+states = c("+","0","-")
+prob = c(1/10,1/10,8/10)
+
+# time and resolution of simulation
+timesteps = 1000
+addgrazing = 500
+delta = 1/10
+
+
+# derive helper vectors for counting: 
+# transformation vector for evaluation at the border of the grid
+	# set evaluation matrix 
+	X <- matrix(as.integer(1:(width*height)), ncol = width, byrow =TRUE)
+	# setting the border of the evaluation matrix X
+	X <- cbind(X[,width], X, X[,1] )  
+	X <- rbind(X[height,], X, X[1,] ) 
+	# transformation vector which adds the border to the lattice:
+	x_with_border <- as.integer(t(X))
+	
+	# from the matrix X (with copied border cells), which cells are the actual cells (reverse transformation vector of the previous lines) 
+	x_to_evaluate <- sort(matrix(1:prod(dim(X)), ncol = dim(X)[2], byrow =TRUE)[-c(1, dim(X)[1]), -c(1,dim(X)[2])]	)		
+
+# defining the neighborhood which is to be evaluated	
+	# set interaction matrix
+	I <- matrix(c(0,1,0,1,NA,1,0,1,0), ncol = 3, byrow = TRUE)	
+	# coordinates of neighbours in Interaction matrix I: 
+	neighbours_in_I <- which(is.finite(abs(I)/abs(I)), arr.in = TRUE)
[…]

### 2013-11-21 data/old_runs/herbivorymodel2_baseline_high.r

@@ -1,0 +1,293 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+
+rm(list=ls())
+source("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+source("E:\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+setwd("E:\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+
+# specify lattice
+width = 50
+height = 50
+
+# initial cell states
+states = c("+","0","-")
+prob = c(1/10,1/10,8/10)
+
+# time and resolution of simulation
+timesteps = 200
+addgrazing = 200
+delta = 1/10
+
+
+# derive helper vectors for counting: 
+# transformation vector for evaluation at the border of the grid
+	# set evaluation matrix 
+	X <- matrix(as.integer(1:(width*height)), ncol = width, byrow =TRUE)
+	# setting the border of the evaluation matrix X
+	X <- cbind(X[,width], X, X[,1] )  
+	X <- rbind(X[height,], X, X[1,] ) 
+	# transformation vector which adds the border to the lattice:
+	x_with_border <- as.integer(t(X))
+	
+	# from the matrix X (with copied border cells), which cells are the actual cells (reverse transformation vector of the previous lines) 
+	x_to_evaluate <- sort(matrix(1:prod(dim(X)), ncol = dim(X)[2], byrow =TRUE)[-c(1, dim(X)[1]), -c(1,dim(X)[2])]	)		
+
+# defining the neighborhood which is to be evaluated	
+	# set interaction matrix
+	I <- matrix(c(0,1,0,1,NA,1,0,1,0), ncol = 3, byrow = TRUE)	
[…]

### 2013-11-21 data/old_runs/randomwalk.r

@@ -1,0 +1,57 @@
+x_new 
+
+
+grazerlayer <- function(x, n, state = c("0", "-"), start = "random") {
+
+starting <- sample( which(x$cells %in% state & count(x_new,  c("0", "-")) > 3 ), n) 
+Y <- matrix(FALSE, ncol = x$dim[1], nrow = x$dim[2]) 
+t(Y)[starting] <- TRUE
+out <- list()
+out$pos <- starting
+out$arr.ind <- which(Y, arr.ind = TRUE)
+out$cumulative <- Y
+out$orientation <- numeric(length = n)
+out$saturation <- numeric(length = n)
+out$state <- state
+class(out) <- c("list", "grazerlist")
+return(out)
+}
+
+out <- grazerlayer(x_new, 50)
+
+par(mar =c(0,0,0,0))
+plot(x_new, cols = color)
+points(out$arr.ind, pch = 21, bg = "red3")
+
+move <- function(x, steps, landscape) {
+n = length(x$pos)
+opencells_logical_with_border <- (landscape$cells %in% x$state)[x_with_border]
+
+for(i in 1:steps) {
+# for each x$pos get  a vector of type allow <- c(TRUE, TRUE, FALSE, TRUE ) for allowance to walk east,  north, west, south
+allow <- sapply(interact, function(k) opencells_logical_with_border[x_to_evaluate[x$pos]+k])
+
+# for each x$pos sample a random value from  c(1,2,3,4) with probability allow 
+goto <- sapply(1:n , function(k) sample(c(1:4), 1, prob = allow[k,]) )
+
+# calculate new x$pos
+newind <- x$arr.ind + data.frame(row = c(-1,0,1,0)[goto], col = c(0,1,0,-1)[goto])
+newind$row[newind$row > landscape$dim[1]] <- newind$row[newind$row > landscape$dim[1]]- landscape$dim[1]
+newind$col[newind$col > landscape$dim[2]] <- newind$col[newind$col > landscape$dim[2]]- landscape$dim[2]
+
+Y <- matrix(FALSE, ncol = landscape$dim[1], nrow = landscape$dim[2]) 
+
+set <- function(x,y) Y[x, y] <- TRUE
+set(newind$row, newind$col)
+
+which(Y == TRUE)
+
+x$cumulative 
+
+points(grazerlist, pch = 21, bg = "red3")
+}
+
+}
+
+
+

### 2013-11-21 data/old_runs/simulation_grazing_corrected_vul2_linux.r

@@ -1,0 +1,666 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+rm(list=ls())
+
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+#setwd("/home/schneider/herbivory/")
+#setwd("C:\\Users\\SCHNEIDER\\Documents\\CASCADE\\2013_grazingmodel\\corrected_grazing_vul\\sim5")
+#setwd("E:\\Eigene Dokumente\\Uni\\projects\\2013 CASCADE grazing\\sim5")
+
+env <- 	unique(c(seq(0.2,1,.04)) )#seq(0.2,1,.025)
+graz <- c(.05, .15, .25)*0.66  #c(0.0375, 0.1125, 0.1875)
+#graz <- seq(0,.1, 0.005)#seq(0,.9, 0.075)
+mort <- 0.05 #seq(0.1,.3, 0.1)
+init <- c("low", "high")
+global <- c(TRUE, FALSE)
+stock <- c(TRUE, FALSE)
+
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+linit <- 2
+lglobal <- 2
+lstock <- 2
+
+
+#global <- rep(global, each = lgraz*lenv*lmort*linit)
+#init <- rep(rep(init, each = lgraz*lenv*lmort), times = lglobal)
+#mort <-	rep(rep(mort, each = lgraz*lenv), times = lglobal*linit)
+#graz <-	 rep(rep(graz, each = lenv), times = lmort*lglobal*linit)
+#env <-  rep(env, times = lgraz*lmort*lglobal*linit)
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1:(lgraz*lmort*lglobal*linit*lenv*lstock), 
+	global = rep(global, each = lgraz*lenv*lmort*linit*lstock),
+	stock =  rep(rep(stock, each = lgraz*lenv*lmort*linit), times = lstock),
[…]

### 2013-11-21 data/old_runs/transitions_baseline.r

@@ -1,0 +1,327 @@
+################################################
+#
+#		model: iterator for cellular automata
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#
+#################################################
+
+
+
+rm(list=ls())
+#source("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
+			# cumulate the result for each of the four neighboring cells (for(k in interact))
+
+			mean(neighbors/4)	
+			# devide by 4 to get density
+			# average over all cells
+}
+
+library(foreach)
+library(doSNOW)
+
[…]

### 2013-11-21 data/old_runs/transitions_baseline2.r

@@ -1,0 +1,322 @@
+################################################
+#
+#		model: iterator for cellular automata
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#
+#################################################
+
+
+
+rm(list=ls())
+#source("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
+			# cumulate the result for each of the four neighboring cells (for(k in interact))
+			} else neighbors <- 0
+			
+			
+			mean(neighbors/4, na.rm = TRUE)	
+			# devide by 4 to get density
[…]

### 2013-11-21 data/old_runs/weak_grazing/transitions_baseline.r

@@ -1,0 +1,327 @@
+################################################
+#
+#		model: iterator for cellular automata
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#
+#################################################
+
+
+
+rm(list=ls())
+#source("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
+			# cumulate the result for each of the four neighboring cells (for(k in interact))
+
+			mean(neighbors/4)	
+			# devide by 4 to get density
+			# average over all cells
+}
+
+library(foreach)
+library(doSNOW)
+
[…]

### 2013-11-21 data/old_runs/without_correction/transitions_herbivory.r

@@ -1,0 +1,303 @@
+################################################
+#
+#		model: iterator for cellular automata
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#
+#################################################
+
+
+
+rm(list=ls())
+#source("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\automatafunctions.r")
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
+			# cumulate the result for each of the four neighboring cells (for(k in interact))
+
+			mean(neighbors/4)	
+			# devide by 4 to get density
+			# average over all cells
+}
+
+library(foreach)
+library(doSNOW)
+
[…]

### 2013-11-21 data/sim4/simulation_grazing_corrected_vul_linux.r

@@ -1,0 +1,676 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+rm(list=ls())
+
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+#setwd("/home/schneider/herbivory/")
+#setwd("C:\\Users\\SCHNEIDER\\Documents\\CASCADE\\2013_grazingmodel\\corrected_grazing_vul")
+
+env <- 	sort(unique(c(seq(0.2,1,.04), seq(0.4,0.6, 0.02))) )#seq(0.2,1,.025)
+graz <- seq(0,.1, 0.005)#seq(0,.9, 0.075)
+mort <- 0.1 #seq(0.1,.3, 0.1)
+init <- c("low", "high")
+global <- c(TRUE, FALSE)
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+linit <- 2
+lglobal <- 2
+
+
+#global <- rep(global, each = lgraz*lenv*lmort*linit)
+#init <- rep(rep(init, each = lgraz*lenv*lmort), times = lglobal)
+#mort <-	rep(rep(mort, each = lgraz*lenv), times = lglobal*linit)
+#graz <-	 rep(rep(graz, each = lenv), times = lmort*lglobal*linit)
+#env <-  rep(env, times = lgraz*lmort*lglobal*linit)
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1:(lgraz*lmort*lglobal*linit*lenv), 
+	global = rep(global, each = lgraz*lenv*lmort*linit),
+	starting = rep(rep(init, each = lgraz*lenv*lmort), times = lglobal),
+	g = rep(rep(graz, each = lenv), times = lmort*lglobal*linit),		#grazing
+	m = rep(rep(mort, each = lgraz*lenv), times = lglobal*linit), 		# intrinsic mortality
[…]

### 2013-11-21 data/sim5/simulation_grazing_corrected_vul2_linux.r

@@ -1,0 +1,666 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+rm(list=ls())
+
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+#setwd("/home/schneider/herbivory/")
+#setwd("C:\\Users\\SCHNEIDER\\Documents\\CASCADE\\2013_grazingmodel\\corrected_grazing_vul\\sim5")
+#setwd("E:\\Eigene Dokumente\\Uni\\projects\\2013 CASCADE grazing\\sim5")
+
+env <- 	unique(c(seq(0.2,1,.02)) )#seq(0.2,1,.025)
+graz <- c(.05, .15, .25)*0.66  #c(0.0375, 0.1125, 0.1875)
+#graz <- seq(0,.1, 0.005)#seq(0,.9, 0.075)
+mort <- 0.05 #seq(0.1,.3, 0.1)
+init <- c("low", "high")
+global <- c(TRUE, FALSE)
+stock <- c(TRUE, FALSE)
+
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+linit <- 2
+lglobal <- 2
+lstock <- 2
+
+
+#global <- rep(global, each = lgraz*lenv*lmort*linit)
+#init <- rep(rep(init, each = lgraz*lenv*lmort), times = lglobal)
+#mort <-	rep(rep(mort, each = lgraz*lenv), times = lglobal*linit)
+#graz <-	 rep(rep(graz, each = lenv), times = lmort*lglobal*linit)
+#env <-  rep(env, times = lgraz*lmort*lglobal*linit)
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1:(lgraz*lmort*lglobal*linit*lenv*lstock), 
+	global = rep(global, each = lgraz*lenv*lmort*linit*lstock),
+	stock =  rep(rep(stock, each = lgraz*lenv*lmort*linit), times = lstock),
[…]

### 2013-11-21 data/sim5/simulation_grazing_corrected_vul2_linux_analyses.r

@@ -1,0 +1,759 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+rm(list=ls())
+
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+#setwd("/home/schneider/herbivory/")
+#setwd("C:\\Users\\SCHNEIDER\\Documents\\CASCADE\\2013_grazingmodel\\corrected_grazing_vul\\sim5")
+#setwd("E:\\Eigene Dokumente\\Uni\\projects\\2013 CASCADE grazing\\sim5")
+
+env <- 	unique(c(seq(0.2,1,.04)) )#seq(0.2,1,.025)
+graz <- c(.05, .15, .25)*0.66  #c(0.0375, 0.1125, 0.1875)
+#graz <- seq(0,.1, 0.005)#seq(0,.9, 0.075)
+mort <- 0.05 #seq(0.1,.3, 0.1)
+init <- c("low", "high")
+global <- c(TRUE, FALSE)
+stock <- c(TRUE, FALSE)
+
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+linit <- 2
+lglobal <- 2
+lstock <- 2
+
+
+#global <- rep(global, each = lgraz*lenv*lmort*linit)
+#init <- rep(rep(init, each = lgraz*lenv*lmort), times = lglobal)
+#mort <-	rep(rep(mort, each = lgraz*lenv), times = lglobal*linit)
+#graz <-	 rep(rep(graz, each = lenv), times = lmort*lglobal*linit)
+#env <-  rep(env, times = lgraz*lmort*lglobal*linit)
+
+# defining parameter set
+parameters = data.frame(
+	ID = 1:(lgraz*lmort*lglobal*linit*lenv*lstock), 
+	global = rep(global, each = lgraz*lenv*lmort*linit*lstock),
+	stock =  rep(rep(stock, each = lgraz*lenv*lmort*linit), times = lstock),
[…]

### 2013-11-21 data/sim6/simulation_grazing_corrected_vul3_linux.r

@@ -1,0 +1,686 @@
+#################################################
+#
+#		model: facilitation model (Kefi et al 2007)
+#
+#
+#author: Flo
+#date: 02.04.2013
+#
+#
+#	Tasks for Flo: 
+#		identify bottlenecks for speedy and parallel computing
+#		implement test for stable equilibrium
+#		
+#   Tasks for Marina:
+# 		adapt for two-species model (implement rules, find parameters)
+# 		adapt output
+#
+#################################################
+
+rm(list=ls())
+
+#setwd("C:\\Users\\SCHNEIDER\\Dropbox\\Projects\\CASCADE\\cellularautomata\\code v1.0\\thread herbivory\\")
+#setwd("/home/schneider/herbivory/")
+#setwd("C:\\Users\\SCHNEIDER\\Documents\\CASCADE\\2013_grazingmodel\\corrected_grazing_vul\\sim6")
+#setwd("E:\\Eigene Dokumente\\Uni\\projects\\2013 CASCADE grazing\\sim6")
+#setwd("C:\\Users\\SCHNEIDER\\Documents\\projects\\CASCADE\\2013 CASCADE grazing\\sim6")
+
+first_ID = 1
+env <- 	unique(c(seq(0.2,1,.04)) )#seq(0.2,1,.025)
+graz <- seq(.0, .2, 0.025)*0.66  #c(0.0375, 0.1125, 0.1875)
+#graz <- seq(0,.1, 0.005)#seq(0,.9, 0.075)
+mort <- 0.1 #seq(0.05,.15, 0.05)
+init <- c("low", "high")
+global <- c(TRUE, FALSE)
+stock <- c(TRUE, FALSE)
+
+
+lgraz <- length(graz)
+lenv <- length(env)
+lmort <- length(mort)
+linit <- 2
+lglobal <- 2
+lstock <- 2
+replication <- 5
+
+#global <- rep(global, each = lgraz*lenv*lmort*linit)
+#init <- rep(rep(init, each = lgraz*lenv*lmort), times = lglobal)
+#mort <-	rep(rep(mort, each = lgraz*lenv), times = lglobal*linit)
+#graz <-	 rep(rep(graz, each = lenv), times = lmort*lglobal*linit)
+#env <-  rep(env, times = lgraz*lmort*lglobal*linit)
+
+# defining parameter set
+parameters = data.frame(
+	ID = first_ID:(lgraz*lmort*lglobal*linit*lenv*lstock*replication+first_ID-1), 
[…]

### 2013-11-21 data/sim9/grazing_gradient_model_comp_v9.r

@@ -1,0 +1,482 @@
+#################################################
+#
+#		model: facilitation model plus grazing
+#
+#
+#author: Flo
+#date: 12.09.2013
+#
+#todo: funktionen auslagern
+#  		gradienten definieren
+#################################################
+
+rm(list=ls())
+
+########################################################################################
+	
+
+highlight <- function(x, colrange = c("black", "red2"), steps = NULL, range = "auto"){
+
+if(is.null(steps)) steps = length(unique(x))
+if(range == "auto") {
+	min_val = min(x, na.rm = TRUE)
+	max_val = max(x, na.rm = TRUE)
+	} else {
+	min_val = range[1]
+	max_val = range[2]
+	}
+
+colorscale <- colorRampPalette(colrange, space = "rgb")
+cols <- colorscale( steps)
+
+cols[as.integer((x-min_val)/(max_val-min_val)*(steps-1)+1)]
+
+}
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
[…]

### 2013-11-27 manuscript/first draft/ms_main.tex
updated figures, mortality estimate

@@
 Arid rangelands are at high risk of sudden desertification.
-This process of persistent vegetation loss under livestock grazing is hardly understood but evidently an issue of the system's bistability.
+Overgrazing is considered one of the main causes of desertification worldwide (MEA 2005), an effect which is aggravated by reduced local precipitation and increased temperatures due to global climate change.
+The process of persistent vegetation loss under livestock grazing is hardly understood but appear to be  an issue of the system's bistability.
 Under particular environmental conditions and grazing regimes, two ecosystem states are likewise resilient to perturbations: One state with a robust vegetation cover of shrubs and grasses as well as a desert with no significant vegetation cover (Scheffer 2001, Aguiar and Sala 1999).
@@
 
-For arid ecosystems, the major mechanism behind this bistability was found to be local facilitation of plant individuals.
-The presence of a plant provides its direct local environment with abilities of water retention, organic matter accumulation and protection against external stressors.
+For arid ecosystems, the major mechanism behind this bistability was found to be local facilitation of plant individuals (REF?).
+The presence of a plant provides its direct local environment with abilities of water retention, organic matter accumulation and protection against external stressors (REF?).
 As a consequence, the vegetation grows in patches (Aguiar and Sala 1999).
@@
 This parent model was used before to investigate the consequences of increased plant mortality.
-This simple formulation of additional plant stress can be interpreted as the effects of increased grazing intensity by large herbivores (K\'efi et al 2007a).
-However, the model is lacking some important features of grazing: the additional individual plant mortality (1) did not vary with vegetation density, and (2) it applied globally, affecting all plants likewise.
+It's simple formulation of additional plant stress can be interpreted as the effects of increased grazing intensity by large herbivores (K\'efi et al 2007a).
+However, the model has two features that are unrealistic: the individual plant mortality which is caused by grazing (1) did not vary with total vegetation cover, and (2) it applied globally, affecting all plants likewise (Fig.
+1).
 Both assumptions are justified by the need for model simplicity, but disallow a translation of the model parameters to real livestock rates.
@@
 %Grazing depends on vegetation cover
-The first extension adds a contrasting assumption on the plant density dependence of feeding to the parent model.
+The first extension adds a contrasting assumption on the dependence on the total vegetation cover to the parent model.
 The prior model defined grazing as a constant risk of additional plant death.
-On the population level, grazing mortality increased with vegetation cover (K\'efi et al.
-2007).
+On the population level, the number of plants dying because of grazing increased with vegetation cover (K\'efi et al.
+2007, Fig.
+1, black solid lines).
 However, grazers forage in the landscape and consume to fulfill their daily demands.
@@
 In reality, the total feeding of large herbivores in the landscape is not entirely independent of vegetation cover but follows a non-linear saturating curve (type-II functional response, Holling 1959, Spalinger and Hobbs 1992).
-Thus, particularly at low vegetation cover grazing is limited by the search success of the grazer, which is resource-density dependent.
-In rangeland systems, however, the functional response at low vegetation cover depends on factors on a larger scale: the animals move between landscapes motivated by internal decision patterns or by external management decisions (Senft 1989).
+Thus, particularly at low vegetation cover, grazing is limited by the search success of the grazer, which is resource-density dependent.
+In rangeland systems, however, the functional response at low vegetation cover depends on factors on a larger scale: the animals move between landscapes motivated by internal decision patterns or their density is modulated by human management decisions (Senft 1989).
 We assume therefore, that smooth type-II functional responses, as described for wild ungulates, are unlikely to happen in intensive rangelands within the spatial scale adressed in the present study.
 
-Because of these uncertainties and for reasons of simplicity we contrast the parent assumption of density independent individual risk with density independent global feeding.
-We expect constant global feeding to consolidate both stable states: the desert state will be more resilient against recolonisation of single plants while the vegetated state will provide better protection to the plant individual.
+Because of these uncertainties and for reasons of simplicity we contrast the parent assumption of individual plant mortality beeing independent of vegetation cover with a constant amount of global feeding, which results in individual plant mortality linearly decreasing with vegetation cover.
+We expect constant global feeding to consolidate both stable states: an almost desertified landscape will be more resilient against recolonisation of single plants while the vegetated state will provide better protection to the plant individual.
 As a consequence, the reversibility of catastrophic shifts is expected to deteriorate.
@@
 %Grazing is spatialy differentiated
-The second extension implements a local differentiation of grazing pressure.
+The second extension implements a small-scaled, local differentiation of grazing pressure.
 %For some time, the relationship between vegetation structure and grazing is acknowledged to be interdependent.
@@
 The authors showed in a theoretical framework similar to the present study, that positive local interactions are superior to environmental heterogeneity or heterogeneous stress in explaining the patterns of plant spatial heterogeneity (Adler et al 2001).
+
 To differentiate the grazing locally, we implement an associative protection of plant individuals growing next to each other.
-This mechanism is observed frequently among shrubs in dryland ecosystems and considered to be a major driver of spatial heterogeneity of stress on local scales.
-
+This mechanism is observed frequently among shrubs in dryland ecosystems and considered to be a major driver of spatial heterogeneity of stress on local scales (REF).
 %Bisigato et al 2005 Ecography
 %Milchunas and Noy-Meir 2002
-
 Here, the individual vulnerability against grazing is highest for plants growing isolated.

### 2013-11-27 code/grazing_density_v4.r

@@ -1,0 +1,694 @@
+
+
+rm(list=ls())
+
+
+# specify lattice
+width = 200
+height = 200
+
+states = c("+","0","-")
+color <- c("black","grey80", "white") # define colors for the cell state levels
+
+parms_temp <- list(
+	m = 0.05,
+	g = 0.03,
+		b = .5 ,		# beta*eps 
+		d = 0.1,		# degradation
+		c_ = 0.01, 		# beta*g  
+		del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+		r = 0.01, 	# regeneration rate
+		f = 0.9 		# local fascilitation
+)
+
+
+# initial cell states
+states = c("+","0","-")
+prob = c(6/10,2/10,2/10)
+color <- c("black","grey80", "white") # define colors for the cell state levels
+
+color <- c("black","grey80", "white") 
+# plotting function for objects of class "landscape".
+plot.landscape <- function(x, grid = FALSE, axis = FALSE, cols = "auto", add = FALSE, ani = FALSE, ...) {
+  lvls <- levels(x$cells) 
+  nlev <- length(lvls)
+  if(cols[1] == "auto") cols = c("black","grey80", "white")  # default color value
+  
+  if(ani & Sys.info()[['sysname']] == "Windows") adj = -0.5 else adj = 0 #this adjustment constant is added when producing a pixel accurate png or gif file, requires TRUE when the function is used to plot animated figures. 
+  
+  if(!add) plot(NA,NA, xlim = c(0.5+adj, x$dim[1]+0.5+adj), ylim = c( x$dim[2]+0.5+adj, 0+0.5+adj), bty = c("n", "o")[grid+1], xaxs = "i", yaxs = "i",xlab = "", ylab = "", xaxt = "n", yaxt = "n", ... ) 
+  
+  if(axis && !add) axis(3) 
+  if(axis && !add) axis(2)
+  
+  if(grid) border = "grey80" else border = cols[as.numeric(x$cells)]
+  
+  rect(rep(1:x$dim[1], times = x$dim[2])-.5, rep(1:x$dim[2], each = x$dim[1])-.5, rep(1:x$dim[1], times = x$dim[2])+.5, rep(1:x$dim[2], each = x$dim[1])+.5, col = cols[as.numeric(x$cells)], border = border)
+  
+  if(grid) box()
+ }
+
[…]

### 2013-11-27 code/grazing_gradient_model_analysis_v10_2.r

@@ -234,28 +234,8 @@ out_fits <- read.csv("output_fits.csv")
 unst_eq <- read.csv("unstable_eq.csv", sep = ";")
 output <- output[ output$stable == TRUE, ] # & output$ID > 27541
 
- par(mfrow = c(1,4), oma = c(0,0,2,0))
-for(i in sort(unique(as.character(output$g)))[-c(1,6)] ) {
-	
-	with(output[output$g == 0 & output$stable == TRUE ,], 
-		plot(b, rho_plus , ylim = c(0,1), xlim = c(0,1), pch = 20, col = "gray80", main = paste("g = ", i)))
-	
-	with(output[output$g == i  & output$globalgrazing == TRUE & output$stock == FALSE & output$stable == TRUE,], 
-		points(b, rho_plus , pch = 20, col = "gray20"))
-	
-	with(output[output$g == i & output$globalgrazing == TRUE & output$stock == TRUE & output$stable == TRUE,], 
-		points(b, rho_plus, pch = 20, col = "red2", cex = 1 ))
-	
-	with(output[output$g == i & output$globalgrazing == FALSE & output$stock == FALSE & output$stable == TRUE,], 
-		points(b, rho_plus, pch = 20, col = "darkred", cex = 1 ))
-	
-	with(output[output$g == i & output$globalgrazing == FALSE & output$stock == TRUE & output$stable == TRUE,], 
-		points(b, rho_plus, pch = 20, col = "red", cex = 1 ))
-	
-}
 
-
-pdf("C:\\Users\\SCHNEIDER\\SkyDrive\\Uni\\projects\\2013 Grazing models (CASCADE)\\figures\\stability.pdf", height = 7, width = 8, paper = "special")
+pdf("C:\\Users\\SCHNEIDER\\Documents\\projects\\CAS01_grazing\\figures\\stability.pdf", height = 7, width = 8, paper = "special")
 
  par(mfrow = c(2,2), mar = c(2,2,0,0), oma = c(3,3,2,1))
  
@@ -488,120 +468,33 @@ for(i in 1:4) {
 
 dev.off()	
 
-layout(matrix(c(1,0,2,3), ncol = 2, byrow = TRUE ) )
-
-par(mar = c(1,2,2,2))
[…]

### 2013-12-06 manuscript/first draft/ms_main.tex
new result figures for manuscript

@@
 \begin{figure}[h]%
-\includegraphics[width=\columnwidth]{figures/stability.pdf}%
-\caption{Stable states of the vegetation cover, $\rho_+$, over gradient of environmental quality, $b$, under different grazing models.
-a) parent model with static, global grazing mortality, $g_0/0.5$.
-b) livestock model with density dependend, global grazing mortality, $g_0 / \rho_+$.
-c) associative protection model with differentiated grazing at patch border, $g_0/0.5 \frac{v}{\hat{v}}$ .
-d) livestock \& associative protection model with density dependend, differentiated grazing at patch border,  $g_0/ \rho_+ \frac{v}{\hat{v}}$ .
-Gradient of grazing intensity from grey: $g_0 = 0$, black to red: $g_0 \in \left\{ 0.025, 0.05, 0.075, 0.1 \right\}$.}%
+\includegraphics[width=\columnwidth]{figures/fig3.pdf}%
+\caption{}%
 \label{}%
@@
 \begin{figure}[h]%
-\includegraphics[width=\columnwidth]{figures/patterns.pdf}%
-\caption{Stable states of the vegetation cover, $\rho_+$, over gradient of environmental quality, $b$, under different grazing models.
-a) parent model with static, global grazing mortality, $g_0/0.5$.
-b) livestock model with density dependend, global grazing mortality, $g_0 / \rho_+$.
-c) associative protection model with differentiated grazing at patch border, $g_0/0.5 \frac{v}{\hat{v}}$ .
-d) livestock \& associative protection model with density dependend, differentiated grazing at patch border,  $g_0/ \rho_+ \frac{v}{\hat{v}}$ .
-Gradient of grazing intensity from grey: $g_0 = 0$, black to red: $g_0 \in \left\{ 0.025, 0.05, 0.075, 0.1 \right\}$.}%
+\includegraphics[width=\columnwidth]{figures/fig4.pdf}%
+\caption{}%
+\label{}%
+\end{figure}
+
+\begin{figure}[h]%
+\includegraphics[width=\columnwidth]{figures/fig5.pdf}%
+\caption{}%
 \label{}%
@@
 Limited Power Law &  $n = b + a s^{- \alpha} $ & $ \log{n} = \log a - \alpha \log{s} + \log(1+\frac{b}{a s^{-\alpha} })$\\
-Truncated Power Law & $n = a s^{- \alpha} \mathrm{e}^{\frac{s}{S_x}}$  & $\log{n} = \log a  - \alpha  \log s  - \frac{s}{Sx}  $ \\
-Exponential &  $n = a \mathrm{e}^{-\varepsilon s} $ & $ \log{n} =  \log{a} - \varepsilon s $\\
+Truncated Power Law & $n = a s^{- \alpha} \mathrm{e}^{s \frac{1}{S_x}}$  & $\log{n} = \log a  - \alpha  \log s  - s \frac{1}{Sx}  $ \\
+%Exponential &  $n = a \mathrm{e}^{-\varepsilon s} $ & $ \log{n} =  \log{a} - \varepsilon s $\\
 	\bottomrule

### 2013-12-06 code/grazing_gradient_model_analysis_v10_2.r

@@ -522,7 +522,7 @@ layout(matrix(c(c(1,0,0,0),2:17), ncol = 4, byrow = TRUE))
 	with(temp, rect(b+pos1, 0,b+pos2,1, col =  modelcols[bestnum], border = NA)
 	)
 	}
-	
+		
 	temp <- unst_eq[unst_eq$g == grazing & unst_eq$stock == c(FALSE, FALSE, TRUE, TRUE)[i] & unst_eq$global == c(TRUE, FALSE, TRUE, FALSE)[i] & unst_eq$eq > 0 & unst_eq$eq < unst_eq$rho_plus,]
 	
 	if(length(temp$eq) > 0) {

### 2013-12-06 code/patchdist_update_v10_3.r

@@ -23,13 +23,13 @@ filenames <- list.files("results/")
 library(foreach)
 library(doSNOW)
 
-workstation <-  list(host = "schneider@kefi118",
+#workstation <-  list(host = "162.38.184.118", user = "schneider",
          rscript = "/usr/lib/R/bin/Rscript",
-		 snowlib = "/usr/lib/R/bin/library")
+		 snowlib = "/usr/lib/R/library/")
 		
 #workerlist <- rep(list(workstation), times = 23)
 
-workerlist <- rep("localhost", times = 11)
+workerlist <-  rep("localhost", times = 10)
 
 cl <- makeSOCKcluster(workerlist)
 

### 2013-12-06 code/test_foreach.r

@@ -6,21 +6,39 @@ library(foreach)
 library(doSNOW)
 
 
-dir.create("results")
+dir.create("results_test")
 
-workerlist <- rep("localhost", times = 23)
+winWorker <- list(host = "localhost", 
+				rscript = "C:\\R\\R-2.15.3\\bin\\x64\\Rscript.exe", 
+				snowlib = "C:\\R\\R-2.15.3\\library"
+	)
+winWorker <- list(host = "localhost", 
+				rscript = "C:/R/R-2.15.3/bin/x64/Rscript.exe", 
+				snowlib = "C:/R/R-2.15.3/library"
+	)
 
-cl <- makeSOCKcluster(workerlist)
+workstation <-  list(host = "162.38.184.118", user = "schneider",
+         rscript = "/usr/lib/R/bin/Rscript",
+		 snowlib = "/usr/lib/R/library/")
+		
+workerlist <- rep(list(workstation), times = 20)
+workerlist <- c( rep(list(winWorker), times = 10),rep(list(workstation), times = 23))
 
+#workerlist <- rep("localhost", times = 23)
+
+cl <- makeSOCKcluster(workerlist, master = "162.38.184.88")
+
+clusterCall(cl,function() Sys.info()[c("nodename","machine")])
+
+# /usr/lib/R/bin/Rscript /usr/lib/R/library//snow/RSOCKnode.R  MASTER=LAB PORT=10187 OUT=/dev/null SNOWLIB=/usr/lib/R/library/ 
 registerDoSNOW(cl)
 
 
-foreach(iteration = 1:46) %dopar% {
+foreach(iteration = 1:46, .combine = rbind) %dopar% {
 #
-temp <- rnorm(20)
-save(temp, file = paste("results/temp", iteration, sep = "_"))
-	
+temp <- rnorm(10)
 
+return(temp)
 gc() 
 }
 

### 2013-12-22 manuscript/AGU-LaTeX/ms_main.tex
v1 of figures, v0.2 of manuscript

--- 2013-12-06 manuscript/first draft/ms_main.tex
@@
 
+\begin{figure}[t]
+	\centering
+		\includegraphics{figures/fig1.pdf}
+	\caption{a) Vegetation in arid ecosystems grows in patches.
+The associated growth provides enhanced local growth conditions as well as protection against grazing.
+With decreasing environmental quality, vegetation declines and the landscape becomes more fragmented (mid column), leading to desertification (right column).
+b) Schematic representation of plants growing in patches.
+The direct neighborhood of plants provides shelter, retention of organic matter and water below and atop the soil surface.
+Spots remote from vegetation suffer erosion and evaporation.
+c) The cellular automata model used in this study distinguishes three states: occupied by vegetation (black), empty but fertile (grey) and degraded (white).
+d) Cumulative patch size distributions described by power law models: with decreasing environmental quality, the number of large patches declines, turning the power law distribution from up-bent (left) to down-bent (mid).
+Degraded landscapes have lost all the larger patches and are not to be described by a power law (right).
+}
+	\label{fig:fig1}
+\end{figure}
+
 For arid ecosystems, the major mechanism behind this bistability was found to be local facilitation of plant individuals (REF?).
@@
 1).
+
+\begin{figure}[tp]
+	\centering
+		\includegraphics{figures/fig2.pdf}
+	\caption{Grazing affects the mortality of individual plants and the patch structure emerging in the landscape.
+The four models compared in this study use different assumptions on the quantity and spatial differentiation of a particular grazing pressure.
+Visualized in the left column are the plants dying due to grazing (red).
+In each model, plants suffer different individual risk, sometimes depending on the number of direct local neighbors (estimates for plants with 4, 2, or 0 neighbors).
+a) The parent model as formulated in Kéfi et al 2008 assumes that grazing affects a proportion of living plants, i.e.
+individual risk is independent of global vegetation cover and shared evenly among all plants.
+b) The livestock model assumes a constant number of plants dying, regardless of the number of plants available.
+The individual risk is increasing when vegetation cover is low.
+c) Assuming asso­cia­tive protection leaves plants in the patch center unaffected by grazing (4), where­as plants at the patch border (2) or growing isolated (0) are suffering more from grazing.
+Particularly at high vegetation cover, the proportional grazing concentrates on those plants.
+d) The combined model is integrating both assumptions leading to highest individual risk at low and high vegetation cover for isolated plants.
+}
+	\label{fig:fig1}
+\end{figure}
+
 Both assumptions are justified by the need for model simplicity, but disallow a translation of the model parameters to real livestock rates.
@@
 2007, Fig.
-1, black solid lines).
+2a).
 However, grazers forage in the landscape and consume to fulfill their daily demands.

### 2013-12-22 manuscript/ms_main.tex
v1 of figures, v0.2 of manuscript

--- 2013-12-22 manuscript/AGU-LaTeX/ms_main.tex
@@
 
-\begin{figure}[t]
+\begin{figure}[!t]
 	\centering
@@
 With decreasing environmental quality, vegetation declines and the landscape becomes more fragmented (mid column), leading to desertification (right column).
-b) Schematic representation of plants growing in patches.
+b) Schematic representation of plant patches.
 The direct neighborhood of plants provides shelter, retention of organic matter and water below and atop the soil surface.
-Spots remote from vegetation suffer erosion and evaporation.
-c) The cellular automata model used in this study distinguishes three states: occupied by vegetation (black), empty but fertile (grey) and degraded (white).
+Spots remote from vegetation suffer from erosion and evaporation.
+c) The cellular automata model used in this study therefore distinguishes three states: occupied by vegetation (black), empty but fertile (grey) and degraded (white).
 d) Cumulative patch size distributions described by power law models: with decreasing environmental quality, the number of large patches declines, turning the power law distribution from up-bent (left) to down-bent (mid).
@@
 
-For arid ecosystems, the major mechanism behind this bistability was found to be local facilitation of plant individuals (REF?).
+For arid ecosystems, a major mechanism behind this bistability was found to be local facilitation of plant individuals (REF?).
 The presence of a plant provides its direct local environment with abilities of water retention, organic matter accumulation and protection against external stressors (REF?).
-As a consequence, the vegetation grows in patches (Aguiar and Sala 1999).
-When implemented in a simple spatially explicit model, this positive neighborhood effect provides true bistability on the landscape scale (K\'efi et al 2007a,b).
+Opposingly, spots remote from any vegetation suffer most from erosion and evaporation, leading to locally degraded soils.
+As a consequence, the vegetation grows in patches (Aguiar and Sala 1999; Fig 1b).
+When implemented in a simple spatially explicit model, this positive neighborhood effect provides true bistability on the landscape scale (K\'efi et al 2007a,b; Fig 1c).
+It therefore can be utilised to investigate the suitability of spatio-temporal metrics to indicate the proximity of a catastrophic shift, such as cumulative patch size distributions or critical slowing down.
 
-This parent model was used before to investigate the consequences of increased plant mortality.
-It's simple formulation of additional plant stress can be interpreted as the effects of increased grazing intensity by large herbivores (K\'efi et al 2007a).
-However, the model has two features that are unrealistic: the individual plant mortality which is caused by grazing (1) did not vary with total vegetation cover, and (2) it applied globally, affecting all plants likewise (Fig.
-1).
+This parent model was used before to investigate the consequences of increased plant mortality (K\'efi et al 2007a).
+It's simple formulation of additional plant stress can be interpreted as the effects of increased rangeland utilisation, corresponding to the grazing intensity by large herbivores, such as goat and sheep (K\'efi et al 2007a).
+However, the model has two features that must be considered unrealistic: the individual risk to be harmed by grazing (1) did not vary with total vegetation cover, and (2) it was defined globally and ignored the protective patch structure of the vegetation (Fig.
+2a).
 
-\begin{figure}[tp]
+\begin{figure}[!tp]
 	\centering
@@
 In each model, plants suffer different individual risk, sometimes depending on the number of direct local neighbors (estimates for plants with 4, 2, or 0 neighbors).
-a) The parent model as formulated in Kéfi et al 2008 assumes that grazing affects a proportion of living plants, i.e.
+a) The \textbf{parent model} as formulated in Kéfi et al 2008 assumes that grazing affects a proportion of living plants, i.e.
 individual risk is independent of global vegetation cover and shared evenly among all plants.
-b) The livestock model assumes a constant number of plants dying, regardless of the number of plants available.
+b) The \textbf{livestock model} assumes a constant number of plants dying, regardless of the number of plants available.
 The individual risk is increasing when vegetation cover is low.
-c) Assuming asso­cia­tive protection leaves plants in the patch center unaffected by grazing (4), where­as plants at the patch border (2) or growing isolated (0) are suffering more from grazing.
+c) Assuming \textbf{associative protection} leaves plants in the patch center unaffected by grazing (4), where­as plants at the patch border (2) or growing isolated (0) are suffering more from grazing.
 Particularly at high vegetation cover, the proportional grazing concentrates on those plants.
-d) The combined model is integrating both assumptions leading to highest individual risk at low and high vegetation cover for isolated plants.
+d) The \textbf{combined model} integrates both assumptions leading to highest individual risk at low and high vegetation cover for isolated plants.
 }
-	\label{fig:fig1}
+	\label{fig:fig2}
 \end{figure}
@@
 Both assumptions are justified by the need for model simplicity, but disallow a translation of the model parameters to real livestock rates.
+Increasing individual plant mortality by factor two does not correspond to twice the number of animals per hectar: The proportional character of global plant mortality causes the total pressure on vegetation to decrease when intensive grazing reduces total vegetation cover (Fig 1a).
+
 The relationship between livestock rates and the stability of vegetation cover is a crucial information for management decisions.
@@
 The first extension adds a contrasting assumption on the dependence on the total vegetation cover to the parent model.
-The prior model defined grazing as a constant risk of additional plant death.
-On the population level, the number of plants dying because of grazing increased with vegetation cover (K\'efi et al.
+The prior model defined grazing as a constant addition to the risk of a plant to die.
+On the population level, the number of plants dying because of grazing decreased with a loss in vegetation cover (K\'efi et al.
 2007, Fig.
@@
 However, grazers forage in the landscape and consume to fulfill their daily demands.
-In other words, the number of plants suffering from grazing is rather constant with vegetation cover.
-As a consequence, at low vegetation cover the grazing risk on individual plants will be high, whereas when vegetation cover is high, risk will be shared among many plants.
+In other words, the number of plants suffering from grazing should be rather constant with vegetation cover.
+As a consequence, at low vegetation cover the grazing risk on individual plants will be high, whereas when vegetation cover is high, risk will be shared among many plants (Fig 2b).
 In reality, the total feeding of large herbivores in the landscape is not entirely independent of vegetation cover but follows a non-linear saturating curve (type-II functional response, Holling 1959, Spalinger and Hobbs 1992).
@@
 
-Because of these uncertainties and for reasons of simplicity we contrast the parent assumption of individual plant mortality beeing independent of vegetation cover with a constant amount of global feeding, which results in individual plant mortality linearly decreasing with vegetation cover.
+Because of these uncertainties and for reasons of simplicity we contrast the parent assumption of individual plant mortality beeing independent of vegetation cover with a constant amount of global feeding, which results in individual plant mortality increasing with vegetation cover (Fig 2b, right panel).
 We expect constant global feeding to consolidate both stable states: an almost desertified landscape will be more resilient against recolonisation of single plants while the vegetated state will provide better protection to the plant individual.
@@
 
-\begin{table}[t!]
+\begin{table}[!th]
 \label{tab:models}
@@
 The applicable transition probabilities for each single cell given in equations 1, 2, 3 and 4 are compared against uniform random numbers between 0 and 1 to determine if a transition occurs or if the cell remains unchanged.
-To initialize the grid with a standardized patchy vegetation structure, each model simulation was run for 500 years without any grazing (g = 0) using an initial environmental value $b_ini \in \left\{ 0.25, 3, 0.8, 0.9 \right\}$ before grazing is taking action.
+To initialize the grid with a standardized patchy vegetation structure, each model simulation was run for 500 years without any grazing (g = 0) using an initial environmental value $ b_ini \in \left\{ 0.25, 3, 0.8, 0.9 \right\} $ before grazing is taking action.
 %(Note: I want to complete this to a highly resoluted gradient, to investigate the unstable equilibria).
-Starting from the preconditioned grids, the four grazing models (Table 1) with varying grazing intensity ($g \in \left{ 0, 0.025, 0.05, 0.075 , 0.10 \right}$) were run over a gradient of environmental quality ($b \in \left{ 0.2, ...
-, 1.0 \right}$) for 1.500 timesteps.
+Starting from the preconditioned grids, the four grazing models (Table 1) with varying grazing intensity ($ g \in \left\{ 0, 0.025, 0.05, 0.075, 0.10 \right\} $) were run over a gradient of environmental quality ($ b \in \left\{ 0.2, ...
+, 1.0 \right\} $) for 1.500 timesteps.
 
@@
 
-To investigate the bistability properties of the vegetation cover, we simulate a gradient in environmental quality b ranging from 0.2 to 1 with a step-width of 0.025.
+To investigate the bistability properties of the vegetation cover, we simulate a high resolution gradient in environmental quality $b$ ranging from 0.2 to 1 with varying step-width between 0.0025 and 0.025.
 The models of this study inherit all parameters from the parent model in K\'efi et al.
-2007a ($r = 0.01$, $f = 0.9$, $\delta = 0.1$, $c = 0.2$, $d = 0.1$), except the ones regarding mortality.
-We assume intrinsic mortality m0 for all models to be 0.5, re
-ecting an average individual lifespan of 20 years, if no additional mortality due to grazing is taking effect.
-We investigate, how the stability properties and patch size distributions change over a grazing gradient.
+2007a ($r = 0.01$, $f = 0.9$, $\delta = 0.1$, $c = 0.2$, $d = 0.1$), while substituting mortality with Eq.
+4 and Table 1.
+We assume intrinsic mortality $m_0$ for all models to be 0.05, reflecting an average individual lifespan of 20 years, if no additional mortality due to grazing is taking effect.
+Further, we run all models over a linear grazing gradient with $g$ taking 0, 0.025, 0.050, 0.075, 0.100.
+For the parent model, this produces average individual lifespans of 20.0, 13.3, 10.0, 8.0, 6.7 years, respectively.
+
+\subsection{Cumulative patch--size distributions}
+We investigate qualitatively, how the patch size distributions change with environmental quality at the different levels of grazing.
+Therefore, we assigned each simulation run to one of the following five cases: (1) fully vegetated, if the vegetation cover was significant (\> 2 \% ) and aggregated into one large spanning cluster (ignoring additional patches up to a size of 2 cells); (2) up-bent power law, if the cumulative patch--size distribution was best described by a power law with a lower limit defined by a single large spanning cluster; (3) power law, if the cumulative patch--size distribution  was best described by a power-law; (4) down-bent power law, if the cumulative patch--size distribution was best described by a truncated power law; (5) degraded, if the vegetation cover was insignificant (\< 2\%).
+
+We compare the parameter space occupied by those cases. Additionally, we compare the exponent of the power law for the cases 2--4.
 

### 2013-12-22 code/conceptualfigure.r

@@ -1,0 +1,635 @@
+# plotting function for objects of class "landscape".
+plot.landscape <- function(x, grid = FALSE, axis = FALSE, cols = "auto", add = FALSE, ani = FALSE, ...) {
+  lvls <- levels(x$cells) 
+  nlev <- length(lvls)
+  if(cols[1] == "auto") cols = c("black","grey80", "white")  # default color value
+  
+  if(ani & Sys.info()[['sysname']] == "Windows") adj = -0.5 else adj = 0 #this adjustment constant is added when producing a pixel accurate png or gif file, requires TRUE when the function is used to plot animated figures. 
+  
+  if(!add) plot(NA,NA, xlim = c(0.5+adj, x$dim[1]+0.5+adj), ylim = c( x$dim[2]+0.5+adj, 0+0.5+adj), bty = c("n", "o")[grid+1], xaxs = "i", yaxs = "i",xlab = "", ylab = "", xaxt = "n", yaxt = "n", ... ) 
+  
+  if(axis && !add) axis(3) 
+  if(axis && !add) axis(2)
+  
+  if(grid) border = "grey80" else border = cols[as.numeric(x$cells)]
+  
+  rect(rep(1:x$dim[1], times = x$dim[2])-.5, rep(1:x$dim[2], each = x$dim[1])-.5, rep(1:x$dim[1], times = x$dim[2])+.5, rep(1:x$dim[2], each = x$dim[1])+.5, col = cols[as.numeric(x$cells)], border = border)
+  
+  if(grid) box()
+ }
+
+setwd("E:\\Eigene Dokumente\\Uni\\projects\\CAS01_grazing\\figures\\conceptual\\")
+
+require(misc3d)
+
+makeshrub <- function(id, height = 0.5, red = FALSE) {
+filename <- paste("shrub",id,".png", sep = "")
+if(red) filename <-paste("shrub",id,"red.png", sep = "")
+png(filename, width = 325, height = 325, bg = "transparent", type =  "cairo-png" )
+
+#bmp(paste("shrub",id,".bmp", sep = ""), width = 300, height = 300, bg = "transparent")
+
+set.seed(id)
+res = 25
[…]

### 2013-12-22 code/grazing_gradient_model_analysis_v10_2.r

@@ -105,6 +105,18 @@ patches <- function(x, state, cumulative = TRUE) {
 	} 
 
 	
+excerpt <- function(x, extr = NULL) {
+  out <- list()
+  out$dim <- c(length(extr[1]:extr[3]),length(extr[2]:extr[4]))
+  temp  <- matrix(1:length(x$cells), ncol = x$dim[2], byrow = TRUE)
+  if(!is.null(extr[1]) ) temp <- temp[extr[1]:extr[3], extr[2]:extr[4], drop = FALSE]
+  out$cells <-  x$cells[as.vector(temp), drop = FALSE]
+  
+  class(out) <- c("list", "landscape")
+  
+  return(out)
+}
+
 
 # specify lattice
 width = 100
@@ -124,8 +136,8 @@ snapshots <-  (timesteps-20:0*50)/delta+1
 t_eval <- ((timesteps/delta)-100/delta):(timesteps/delta)+1
 	
 
-#setwd("E:\\Eigene Dokumente\\Uni\\projects\\2013 CASCADE grazing\\sim9\\")
-setwd("C:\\Users\\SCHNEIDER\\Documents\\projects\\CAS01_grazing\\data\\sim10\\")
+setwd("E:\\Eigene Dokumente\\Uni\\projects\\CAS01_grazing\\data\\sim10\\")
+#setwd("C:\\Users\\SCHNEIDER\\Documents\\projects\\CAS01_grazing\\data\\sim10\\")
 
 if(FALSE) {
 filenames <- list.files("results\\")
@@ -231,7 +243,8 @@ write.table(out_fits, "output_fits.csv", row.names = FALSE, col.names = TRUE, se
 
 output <- read.csv("output.csv")
 #out_fits <- read.csv("output_fits.csv")
-unst_eq <- read.csv("unstable_eq.csv", sep = ";")
+unst_eq <- read.csv("unstable_eq.csv", sep = ",")
+unst_eq <- unst_eq[unst_eq$eq >= 0.01,]
 output <- output[ output$stable == TRUE, ] # & output$ID > 27541
 
 	output$bestnum <- 0
@@ -245,10 +258,19 @@ output <- output[ output$stable == TRUE, ] # & output$ID > 27541
 		
 		
 # define colors for: desert, PL, TPL_up,  TPL_down, exp, covered
[…]

### 2013-12-22 code/tools/test_foreach.r

@@ -6,7 +6,7 @@ library(foreach)
 library(doSNOW)
 
 
-dir.create("results_test")
+#dir.create("results_test")
 
 winWorker <- list(host = "localhost", 
 				rscript = "C:\\R\\R-2.15.3\\bin\\x64\\Rscript.exe", 
@@ -21,26 +21,28 @@ workstation <-  list(host = "162.38.184.118", user = "schneider",
          rscript = "/usr/lib/R/bin/Rscript",
 		 snowlib = "/usr/lib/R/library/")
 		
-workerlist <- rep(list(workstation), times = 20)
-workerlist <- c( rep(list(winWorker), times = 10),rep(list(workstation), times = 23))
+#workerlist <- rep(list(workstation), times = 1)
+#workerlist <- c( rep(list(winWorker), times = 10),rep(list(workstation), times = 23))
 
-#workerlist <- rep("localhost", times = 23)
+workerlist <- rep("localhost", times = 23)
 
-cl <- makeSOCKcluster(workerlist, master = "162.38.184.88")
+cl <- makeSOCKcluster(workerlist, outfile='out_messages.txt')
 
-clusterCall(cl,function() Sys.info()[c("nodename","machine")])
+#clusterCall(cl,function() Sys.info()[c("nodename","machine")])
 
 # /usr/lib/R/bin/Rscript /usr/lib/R/library//snow/RSOCKnode.R  MASTER=LAB PORT=10187 OUT=/dev/null SNOWLIB=/usr/lib/R/library/ 
 registerDoSNOW(cl)
 
 
-foreach(iteration = 1:46, .combine = rbind) %dopar% {
-#
-temp <- rnorm(10)
+foreach(iteration = 1:46, .combine = rbind) %do% {
+
+load(paste("~/herbivory/results/result", iteration, sep = "_"))		
+
+temp <- result$out
 
 return(temp)
 gc() 
-}
+} -> out
 
 
 

### 2013-12-22 code/unstable_equilibria.r

@@ -76,12 +76,10 @@ height = 100
 	# relative position of the four direct neighbours of a cell
 	interact <- (relrow * dim(X)[2] + relcol)
 
-	
+#setwd("E:\\Eigene Dokumente\\Uni\\projects\\CAS01_grazing\\data\\sim10\\")
 
 output <- read.csv("output.csv")
 #output_fits <- read.csv("output_fits.csv")
-
-output <- output[ output$g != 1 & output$stable == TRUE, ] # & output$ID > 27541
 
 # initial cell states
 states = c("+","0","-")
@@ -95,14 +93,24 @@ delta = 1/5
 library(foreach)
 library(doSNOW)
 
-#workerlist <- rep(list(ubuWorker), times = 2)
-workerlist <- rep("localhost", times = 15)
+workstation <-  list(host = "162.38.184.118", user = "schneider",
+        rscript = "/usr/lib/R/bin/Rscript",
+		 snowlib = "/usr/lib/R/library/")
 
-cl <- makeSOCKcluster(workerlist)
+workstation <-  list(host = "162.38.184.118", user = "schneider",
+        rscript = "/usr/lib/R/bin/Rscript",
+		 snowlib = "/usr/lib/R/library/")
+		
+#workerlist <- rep(list(workstation), times = 23)
+#workerlist <- c( rep(list(winWorker), times = 10),rep(list(workstation), times = 23))
+
+workerlist <- rep("localhost", times = 23)
+
+cl <- makeSOCKcluster(workerlist, outfile='out_messages.txt') #, master = "162.38.106.207"
+
+#clusterCall(cl,function() Sys.info()[c("nodename","machine")])
 
 registerDoSNOW(cl)
-
-
 
 		out <- data.frame(	global =  NA,
 			stock =  NA,
@@ -111,7 +119,7 @@ registerDoSNOW(cl)
 			b = NA, 		# beta*eps 
 			d = 0.1,		# degradation
 			c_ = 0.2, 		# beta*g  
-			del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+			del = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhoods
 			r = 0.01, 	# regeneration rate
 			f = 0.9,		# local fascilitation
 			rho_plus = NA,
@@ -120,34 +128,20 @@ registerDoSNOW(cl)
 			)[-1,]
 			
[…]

### 2014-01-12 manuscript/ms_main.tex
Home office christmas 2012, figures finalised, ms edited

@@
 
-We applied these rules on a grid of 100 $\times$ 100 cells of 0.25m$^2$.
-Each timestep is defined as one year.
-The applicable transition probabilities for each single cell given in equations 1, 2, 3 and 4 are compared against uniform random numbers between 0 and 1 to determine if a transition occurs or if the cell remains unchanged.
-To initialize the grid with a standardized patchy vegetation structure, each model simulation was run for 500 years without any grazing (g = 0) using an initial environmental value $ b_ini \in \left\{ 0.25, 3, 0.8, 0.9 \right\} $ before grazing is taking action.
-%(Note: I want to complete this to a highly resoluted gradient, to investigate the unstable equilibria).
-Starting from the preconditioned grids, the four grazing models (Table 1) with varying grazing intensity ($ g \in \left\{ 0, 0.025, 0.05, 0.075, 0.10 \right\} $) were run over a gradient of environmental quality ($ b \in \left\{ 0.2, ...
-, 1.0 \right\} $) for 1.500 timesteps.
+The applicable transition probabilities for each single cell given in equations 1, 2, 3 and 4 are compared against uniform random numbers between 0 and 1 to determine if a transition occurs or if the cell remains unchanged (see appendix for additional detail).
+To initialize the grid with a standardized patchy vegetation structure, each model simulation was run for 500 years without any grazing ($g = 0$) using random initial environmental values (uniform; $ 0.25 <  b_\mathrm{ini} < 0.9 $) before grazing is taking action.
+%the lower limit of these causes the desert stable state beeing undetectable if the unstable equilibrium is below rho(b = 0.25) for a given parameter combination.
+A final simulation run with true random initial conditions, including very low vegetation cover (rho ~ 0.005) would be required.
 
-An average vegetation cover, $\rho_+$, as well as the average cell mortality,$\hat{m}$ , is determined as the average density of all grids of the final 500 timesteps.
+Starting from these preconditioned grids, the four grazing models (Table 1) were run over a gradient of environmental quality with varying grazing intensity for 2.000 timesteps (see below for exact parameter values).
 
-\subsection{Choice of parameters}
+\subsubsection{Choice of parameters}
+We applied the transition rules on a grid of 100 $\times$ 100 cells, assuming each grid cell to cover an area of 0.25m$^2$, and a timestep to include the dynamics occuring within one year.
 
@@
 Further, we run all models over a linear grazing gradient with $g$ taking 0, 0.025, 0.050, 0.075, 0.100.
-For the parent model, this produces average individual lifespans of 20.0, 13.3, 10.0, 8.0, 6.7 years, respectively.
+For the parent model, this produces average individual lifespans of 20.0, 13.3, 10.0, 8.0, 6.7 years, respectively. The other models generate more dynamic and locally differentiated lifespans.
 
-\subsection{Cumulative patch--size distributions}
+\subsubsection{Recorded output of the simulations}
+
+The average vegetation cover, $\rho_+$, % as well as the average cell mortality,$\hat{m}$,
+is determined as the average density of vegetated cells in all grids of the final 500 timesteps.
+
+The unstable equilibria are estimated by approximating the point of vegetation cover were the dynamics switch from increasing cover to decreasing cover (see appendix for more detail on the estimation).
+
+In the last 1000 timesteps of the simulation we recorded the largest occuring patch in intervals of 50 timesteps, to determine the average size of the largest patch ($n = 20$).
+
+%\subsection{Cumulative patch--size distributions}
 We investigate qualitatively, how the patch size distributions change with environmental quality at the different levels of grazing. 
-Therefore, we assigned each simulation run to one of the following five cases: (1) fully vegetated, if the vegetation cover was significant (\> 2 \% ) and aggregated into one large spanning cluster (ignoring additional patches up to a size of 2 cells); (2) up-bent power law, if the cumulative patch--size distribution was best described by a power law with a lower limit defined by a single large spanning cluster; (3) power law, if the cumulative patch--size distribution  was best described by a power-law; (4) down-bent power law, if the cumulative patch--size distribution was best described by a truncated power law; (5) degraded, if the vegetation cover was insignificant (\< 2\%).
+From 20 snapshots taken in 50 year intervals over the last 1000 years of each simulation we calculated the cumulative patch-size distributions (White et al 2008), i.e.
+the probabilities that an observation is less or equal to a particular patch size.
+We pooled the 20 patch size distributions to fit three alternative probability density functions for each simulation: a pure power law, an up-bent power law with a lower limit defined by a single large spanning cluster, and a down-bent power law, approximating to a largest possible patch (Table 2; Kéfi et al 2007).
+The models were fitted using nonlinear least-squares on the linearized (i.e. log-transformed) versions of the three models (Table 2).
 
-We compare the parameter space occupied by those cases. Additionally, we compare the exponent of the power law for the cases 2--4.
+\begin{table}[t!]
+\label{tab:fitmodels}
+\caption{ Kandidate models describing patch size distribution }
+\centering
+\begin{tabular}{lll}
+
+\toprule
+ & function & log-transform \\ \cmidrule(rl){2-2} \cmidrule(rl){3-3}
+Power Law &  $n = a s^{ \alpha}$ & $ \log{n} =\log{a} + \alpha \log s$\\
+Limited Power Law &  $n = b + a s^{ \alpha} $ & $ \log{n} = \log a + \alpha \log{s} + \log(1+\frac{b}{a s^{\alpha} })$\\
+Truncated Power Law & $n = a s^{ \alpha} \mathrm{e}^{s \frac{1}{S_x}}$  & $\log{n} = \log a  + \alpha  \log s  - s \frac{1}{Sx}  $ \\
+%Exponential &  $n = a \mathrm{e}^{-\varepsilon s} $ & $ \log{n} =  \log{a} - \varepsilon s $\\
+	\bottomrule
+\multicolumn{3}{p{12.5cm}}{\footnotesize  }
+	\end{tabular}
+\end{table}
+
+We assigned each simulation replicate to one of the following five cases: (1) fully vegetated, if the vegetation cover was significant ($ > 2 \%$) and aggregated into one large spanning cluster (ignoring patches of size 1 or 2 cells); (2) up-bent power law, if the cumulative patch--size distribution was best described by a power law with a lower limit of one large patch spanning over the whole landscape; (3) power law, if the cumulative patch--size distribution  was best described by a power-law; (4) down-bent power law, if the cumulative patch--size distribution was best described by a truncated power law; (5) degraded, if the vegetation cover was insignificant ($< 2\%$).
+
+We compare the parameter space occupied by these five cases. Additionally, we compare the exponent $\alpha$ of the power law for the cases 2--4.
 
@@
 
-\newpage
+For each model, we investigate the response of the landscape structure to increasing pressure, which might be a decrease in environmental quality or an increase in grazing intensity.
+Landscape structure is assessed by four different metrics: the vegetation cover of the landscape, the character of the cumulative patch size distribution, the exponent of the most suitable power-law model, and the size of the largest patch.
 
-\begin{figure}[h]%
+The four models differ strongly regarding the stability of and the transition between the vegetated and the desert state with increasing pressure (Fig.
+3).
+The most apparent difference of the livestock model compared to the parent model is the larger basin of attraction of the desert state: The critical threshold of vegetation cover (white line in Fig.
+3a) below which the remaining vegetation will be lost as well, is much more tense in the livestock model.
+At the same time, the vegetated state becomes more relaxed, reflected by the higher values of vegetation cover and a small shift of the borders between the vegetation categories towards worse environmental conditions (dark, green and red background areas).
+
+\begin{figure}[t]%
 \includegraphics[width=\columnwidth]{figures/fig3.pdf}%
@@
 
-\begin{figure}[h]%
+The associative protection model has an opposite effect on the stability of the vegetated state, shifting the vegetation categories profoundly towards better environmental conditions and diminishing the average vegetation cover.
+The desert state, however, becomes more fragile as well.
+The parameter space, where coexistence of both stable states is detectable by the simulations becomes narrower.
+Remarkably, both models accellerate the breakdown in vegetation cover with decreasing environmental quality and increasing grazing intensity compared to the parent model.
+The critical threshold of catastrophic shift with increasing pressure becomes more sudden.
+The combined model seems to add the effects of the both models seamlessly.
+The shift of the vegetated categories is equaled out and similar to the parent model.
+The critical threshold of sudden vegetation loss is reached even earlier with increasing pressure.
+\par
+
+For those cases where the landscape's patch size distribution was best described by one of the power law models we compare the exponents of the power law, $\alpha$ with increasing pressure (Fig.~\ref{fig:fig4}).
+In the parent model, with incresing pressure  $\alpha$ increases from around -3 to a climax of VALUE .
+
+\begin{figure}[t]%
 \includegraphics[width=\columnwidth]{figures/fig4.pdf}%
-\caption{}%
-\label{}%
+\caption{For the cases where a power law description of the cumulative patch size distribution is appropriate, the exponent $\alpha$ of the power law is plotted as a function of environmental quality, $b$.
+Green symbols are up-bent power laws, red symbols are straight power laws and brown symbols are down-bent power laws.
+Dark and light grey areas are fully vegetated and desert landscapes, respectively.
+}%
+\label{fig:fig4}%
 \end{figure}
 
-\begin{figure}[h]%
+\begin{figure}[tb]%
 \includegraphics[width=\columnwidth]{figures/fig5.pdf}%
 \caption{}%
-\label{}%
+\label{fig:fig5}%
 \end{figure}
@@
 
+\newpage
+
 \section{APPENDIX}
@@
 The grid was defined to have periodic borders, connecting the east border to the west, the north to the south.
-Prerun for 100 years. 
+Prerun for 100 years
 during pre-run, environmental conditions $b_ini \in \left\{ 0.250, 0.274, 0.300, 0.329, 0.360, 0.395, 0.433, 0.474, 0.520, 0.570, 0.624, 0.684, 0.749, 0.821, 0.900 \right\} $ emerge a patch structure ranging from low to high vegetation cover.
@@
 
-\begin{table}[t!]
-\label{tab:fitmodels}
-\caption{ Models describing patch size distribution }
-\centering
-\begin{tabular}{lll}
-
-\toprule
- & function & log-transform \\ \cmidrule(rl){2-2} \cmidrule(rl){3-3}
-Power Law &  $n = a s^{- \alpha}$ & $ \log{n} =\log{a} - \alpha \log s$\\
-Limited Power Law &  $n = b + a s^{- \alpha} $ & $ \log{n} = \log a - \alpha \log{s} + \log(1+\frac{b}{a s^{-\alpha} })$\\
-Truncated Power Law & $n = a s^{- \alpha} \mathrm{e}^{s \frac{1}{S_x}}$  & $\log{n} = \log a  - \alpha  \log s  - s \frac{1}{Sx}  $ \\
-%Exponential &  $n = a \mathrm{e}^{-\varepsilon s} $ & $ \log{n} =  \log{a} - \varepsilon s $\\
-	\bottomrule
-\multicolumn{3}{p{12.5cm}}{\footnotesize  }
-	\end{tabular}
-\end{table}
-
 table with log-transformed models

### 2014-01-12 code/conceptualfigure.r

@@ -552,84 +552,87 @@ dev.off()
 #####
 
 
-pdf("E:\\Eigene Dokumente\\Uni\\projects\\CAS01_grazing\\figures\\conceptual\\fig2_mortalities.pdf", width =3, height = 1.5, paper = "special", useDingbats = FALSE)
+pdf("E:\\Eigene Dokumente\\Uni\\projects\\CAS01_grazing\\figures\\conceptual\\fig2_mortalities.pdf", width =4.5, height = 1.2, paper = "special", useDingbats = FALSE)
 
 
 layout(matrix(c(1,2,1,3,1,4), ncol = 2, byrow = TRUE), width = c(5,2))
-par(mar = c(1,2,0,0), oma = c(3,1,1,3), bty = "n",  las = 1)
+layout(matrix(c(1,0,2,3,4), ncol = 5, byrow = TRUE), width = c(1,0.7,1,1,1))
+par(mar = c(1,1,0,0), oma = c(2,2,1,1), bty = "l",  las = 1, yaxs = "i" , xaxs = "r")
 
-plot(NA,NA, xlim = c(1,0), ylim = c(0,.12), ylab = "", xlab = "", xaxt = "n", yaxs = "i" , xaxs = "i", yaxp = c(0, 0.1, 2))
+plot(NA,NA, xlim = c(1,0), ylim = c(0,.12), ylab = "", xlab = "", xaxt = "n", yaxp = c(0, 0.1, 2))
 #lines(rep(parms_temp$m, times = length(n))*dens ~ dens, data = result, col = "grey80", lwd = 2)
 axis(1, at  = c(0,1), labels = c("", ""), tck = 0)
 ## intrinsic mortality 
 abline(a = 0, b = parms_temp$g/0.5, lty = 1, lwd = 3) # parent model
 
-plot(parent ~ dens, data = result, type = "l", lwd = 3, xlim = c(1,0), ylim = c(-0.03,.3 ), ylab = "", xlab = "vegetation cover", xaxt = "n", yaxt = "n", yaxs = "i" , xaxs = "i", yaxp = c(0, 0.2, 1))
-axis(4, at  = c(0,.2))
-axis(1, at  = c(0,1), labels = c("", ""), tck = 0.2)
[…]

### 2014-01-12 code/grazing_gradient_model_analysis_v10_2.r

@@ -514,7 +514,7 @@ dev.off()
 pdf("figures\\cpdf_fits.pdf", height = 7, width = 10, paper = "special", useDingbats = FALSE)
 
 layout(matrix(c(c(1,0,0,0),2:17), ncol = 4, byrow = TRUE)) 
- par( oma = c(3,3,2,1), mar = c(2,1,0,1))
+ par( oma = c(3,3,1,1), mar = c(2,1,1,1))
 
  
  temp <- output[output$g == 0 & output$rho_plus > 0,]
@@ -556,10 +556,11 @@ layout(matrix(c(c(1,0,0,0),2:17), ncol = 4, byrow = TRUE))
 		axis(1)	
 		box()
 
- par(mar = c(1,1,1,1))
+ par(mar = c(1,1,2,1))
  
-for(j in c(0.025, 0.05, 0.075, 0.1) ){
-	for(i in 1:4) {
+for(i in 1:4) { 
+	for(j in c(0.025, 0.05, 0.075, 0.1) ){
+
 	grazing = j
 	
 	temp <- output[output$g == grazing & 
@@ -601,8 +602,8 @@ for(j in c(0.025, 0.05, 0.075, 0.1) ){
 	temp <- output[output$g == grazing & output$stock == c(FALSE, TRUE, FALSE, TRUE)[i] & output$globalgrazing == c(TRUE, TRUE, FALSE, FALSE)[i],]
 	
 	points(temp$b, temp$rho_plus , pch = 20, col = "black", cex = 1.25)
-		if(i== 1 ) axis(2, at = c(0,0.5,1), las = 1)	
-		if(j == .1 ) axis(1)
+		if(j == 0.025 ) axis(2, at = c(0,0.5,1), las = 1)	
+		if(i == 4 ) axis(1)
 		box()
 	#text(0.3, .9, labels = paste("g = ", grazing))
 		}
@@ -806,9 +807,9 @@ dev.off()
  
 	
 	
-pdf("figures\\alpha.pdf", height = 5, width = 10, paper = "special", useDingbats = FALSE)
+pdf("figures\\alpha.pdf", height = 5.5, width = 10, paper = "special", useDingbats = FALSE)
 	
- par(mfcol= c(4,4), oma = c(3,3,2,1), mar = c(1,1,1,1))
+ par(mfrow= c(4,4), oma = c(3,3,2,1), mar = c(1,1,2,1))
 for(l in 1:4) {
[…]

### 2014-01-12 code/grazing_simulation.r

@@ -1,0 +1,846 @@
+
+rm(list=ls())
+
+########################################################################################
+	
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# function which counts the density of a state within neighboring cells of a particular state
+# requires a previous global definition of the transformation vectors "x_to_evaluate" and "x_with_border" and the neighborhood vector "interact"
+localdens <- function(x, neighbor, state) {
+			neighbors <- numeric(length = length(which(x$cells == state))) # preallocate memory
+			
+			# create a logical vector x(which cells are in state "state"?) 
+			# and transfer the vector x to a vector x_with_border: x[x_with_border]
+			x_logical_with_border <- (x$cells == neighbor)[x_with_border]
+			
+			#of those values which are original values (x_to_evaluate) and which are "state": x_to_evaluate[initial$cells == "state"]
+			x_state <- x_to_evaluate[x$cells == state]
+			
+			if(length(x_state) > 0) { 
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_state+k]
+			}
+			# derive a vector which contains the value to count (also "state") in one of the neighbor cells j
+			# cumulate the result for each of the four neighboring cells (for(k in interact))
+			} else neighbors <- 0
+			
+			
+			mean(neighbors/4, na.rm = TRUE)	
+			# devide by 4 to get density
+			# average over all cells
+}
+
+# plotting function for objects of class "landscape".
+plot.landscape <- function(x, grid = FALSE, axis = FALSE, cols = "auto", add = FALSE, ani = FALSE, ...) {
+  lvls <- levels(x$cells) 
+  nlev <- length(lvls)
[…]

### 2014-01-12 code/patchdist_update_v10_3.r

@@ -2,7 +2,7 @@
 
 rm(list=ls())
 
-#setwd("E:\\Eigene Dokumente\\Uni\\projects\\2013 CASCADE grazing\\sim9\\")
+#setwd("E:\\Eigene Dokumente\\Uni\\projects\\CAS01_grazing\\data\\sim10\\")
 setwd("C:\\Users\\SCHNEIDER\\Documents\\projects\\CAS01_grazing\\data\\sim10\\")
 
 

### 2014-01-12 code/simfunctions.r

@@ -1,0 +1,64 @@
+
+count  <- function(x, neighbor, state = NULL) {
+			if(is.null(state)) state = neighbor
+			
+			neighbors <- numeric(length = prod(x$dim))
+			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
+			for(k in interact) {
+				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
+			}
+			return(neighbors)	
+}
+
+# plotting function for objects of class "landscape".
+plot.landscape <- function(x, grid = FALSE, axis = FALSE, cols = "auto", add = FALSE, ani = FALSE, ...) {
+  lvls <- levels(x$cells) 
+  nlev <- length(lvls)
+  if(cols[1] == "auto") cols = c("black","grey80", "white")  # default color value
+  
+  if(ani & Sys.info()[['sysname']] == "Windows") adj = -0.5 else adj = 0 #this adjustment constant is added when producing a pixel accurate png or gif file, requires TRUE when the function is used to plot animated figures. 
+  
+  if(!add) plot(NA,NA, xlim = c(0.5+adj, x$dim[1]+0.5+adj), ylim = c( x$dim[2]+0.5+adj, 0+0.5+adj), bty = c("n", "o")[grid+1], xaxs = "i", yaxs = "i",xlab = "", ylab = "", xaxt = "n", yaxt = "n", ... ) 
+  
+  if(axis && !add) axis(3) 
+  if(axis && !add) axis(2)
+  
+  if(grid) border = "grey80" else border = cols[as.numeric(x$cells)]
+  
+  rect(rep(1:x$dim[1], times = x$dim[2])-.5, rep(1:x$dim[2], each = x$dim[1])-.5, rep(1:x$dim[1], times = x$dim[2])+.5, rep(1:x$dim[2], each = x$dim[1])+.5, col = cols[as.numeric(x$cells)], border = border)
+  
+  if(grid) box()
+ }
+
+# get patch size and patchsize distribution
+patches <- function(x, state, cumulative = TRUE) {
+	pattern <- x$cells
+	pattern <- pattern %in% state
+	map <- rep(NA, times = prod(x$dim))
+	old <- rep(99, times = prod(x$dim)) 
+	
+	while(!identical(old[pattern], map[pattern])) {
+		old <- map
+		count = as.integer(1)
+		for(i in which(pattern)) {
[…]

### 2014-11-13 documentation/model_formulation.Rmd
initialise model formulation

---
title: "Basic Model formulation"
author: "Florian D. Schneider"
date: "Tuesday, November 13, 2014"
output:
  pdf_document:
    fig_height: 3.5
    fig_width: 4
  html_document:
    keep_md: yes
    self_contained: no
geometry: margin=1.4in
fontsize: 12pt
---
<!--

bibliography: cas02.bib
csl: amnat.csl
-->

This is supposed to become an interactive model representation of the pair-approximation model.

The default form of the differential equation describing vegetation cover is

$$ \frac{dV}{dt} = G(V) - C(V) $$

with growth function $G$ and a function $C$ which represents consumption by herbivores. Formally, intrinsic mortality could be included in both terms. In the formulations of @noy-meir75 and @vandekoppel97 it is assumed to be covered by the logistic growth function.

It might become important in the pair-approximation model, where intrinsic plant mortality is included as a proportional loss of mature plants.

```{r, echo=FALSE}
# Default plotting function

defplot <- function(...) plot(..., 
            type = "l", 
            bty = "l", cex = 0.7, las = 1, 
            family = "Times",
            xlim = c(0,1), ylim = c(0,0.2), 
            xaxs = "i", yaxs = "i", 
            xaxp = c(0,1,2), yaxp = c(0,0.2,2), 
            xlab = "vegetation cover", ) 
```

## logistic growth

$$ G(V) = rV(1-V/K)$$

```{r, echo=FALSE}
G <- function(P, r = 0.5, K = 0.9) r*P*(1-P/K)
rho <- seq(0,1,length = 100)

defplot(rho, G(rho), ylab = "growth")

```

## functional response

$$ C(V) = c_\mathrm{max} \frac{V}{V+T_h} $$

```{r, echo=FALSE}
C <- function(P, c_max = 0.1, T_h = 0.05) c_max*P/(P+T_h)
rho <- seq(0,1,length = 100)

defplot(rho, C(rho), ylab = "plant mortality")

```

```{r, echo=FALSE}

defplot(rho, G(rho), ylab = "growth")
lines(rho, C(rho, c_max = 0.1, T_h =0.3))
lines(rho, C(rho, c_max = 0.1))
```

### 2014-11-18 documentation/model_formulation.Rmd
update model conception

@@
 author: "Florian D. Schneider"
-date: "Tuesday, November 13, 2014"
+date: "Tuesday, November 18, 2014"
 output:
@@
 fontsize: 12pt
+bibliography: ../manuscript/cas02.bib
+csl: ../manuscript/amnat.csl
 ---
-<!--
 
-bibliography: cas02.bib
-csl: amnat.csl
--->
-
-This is supposed to become an interactive model representation of the pair-approximation model.
+This document is supposed to become an interactive model representation of the pair-approximation model.
 
@@
             xaxp = c(0,1,2), yaxp = c(0,0.2,2), 
-            xlab = "vegetation cover", ) 
+            xlab = "vegetation cover") 
 ```
 
-## logistic growth
+## non-spatial models
+### logistic growth
 
-$$ G(V) = rV(1-V/K)$$
+If assuming phenomenological logistic growth, the gain in vegetation as a function of present vegetation cover is
+
+$$ G(V) = rV(1-V/K) \,,$$
+
+with the growth rate, $r$, and the carrying capacity, $K$, decelerating the exponential growth as vegetation cover increases and reducing it to zero if $V \geq K$.
+Thus, it is a decelerating effect (a negative feed-back) at high vegetation cover.
 
@@
 ```
+#### mechanistic growth assumptions
 
-## functional response
+#### decelerating growth at low cover
 
-$$ C(V) = c_\mathrm{max} \frac{V}{V+T_h} $$
+In arid ecosystems, growth would be decelerated at very low cover, because retention of water and organic matter in the site is fundamentally reduced.
+This postive feed-back (low cover reduces growth, leading to even lower cover) was discussed by @vandekoppel97 and @rietkerk97 to explain bistability in arid ecosystems.
+
+$$ G(V) = rV(1-V/K) \,,$$
+
+### functional response
+
+#### linear plant mortality
+
+In the cellular automata model and mean-field approximations of Sonia, individual plant mortality is a constant rate, which means it is increasing linearly with vegetation cover.
+
+Plant mortality due to grazing therefore would be described as
+
+$$ C(V) = mV \,, $$
+
+with the grazing mortality, $m$.
 
 ```{r, echo=FALSE}
-C <- function(P, c_max = 0.1, T_h = 0.05) c_max*P/(P+T_h)
+C <- function(V, m = 0.25) (m*V)
 rho <- seq(0,1,length = 100)
@@
 
+#### constant plant mortality
+
+In fact, when assuming a constant density of herbivores, grazing mortality will certainly saturate at some point, due to limitations of the herbivores handling times.
+A very parsimonious model would assume that grazing always is defined by this limitation of handling times, even at very low densities.
+
+Mortality due to grazing,
+
+$$ C(V) = m \,, $$
+
+thus would be defined as constant, $m$, regardless of the vegetation cover.
+
+```{r, echo=FALSE}
+C <- function(V, m = 0.1) m + V*0
+rho <- seq(0,1,length = 100)
+
+defplot(rho, C(rho), ylab = "plant mortality")
+
+```
+
+#### type 2 functional response
+
+In natural systems, however, consumption is defined by saturating functional responses, i.e.
+at low cover, density dependent mortality defined by search efficiency dominates, but at high cover handling times of the consumer will limit its intake rates.
+
+Therefore, plant mortality due to grazing should be
+
+$$ C(V) = \frac{aV}{1+ahV} \,,$$
+
+with the handling time, $h$ [time per resource unit], and the search efficiency, $a$ [area foraged in per time].
+
+```{r, echo=FALSE}
+C <- function(V, a = 0.5, h = 10) (a*V)/(1+a*h*V)
+rho <- seq(0,1,length = 100)
+
+defplot(rho, C(rho), ylab = "plant mortality")
+
+```
+
+#### type 3 functional response
+
+$$ C(V) = \frac{bV^{1+q}}{1+bhV^{1+q}} $$
+
+```{r, echo=FALSE}
+C <- function(V, a = 1, h = 10, q = 1) (a*V^(1+q))/(1+a*h*(V^(1+q)) ) 
+rho <- seq(0,1,length = 100)
+
+defplot(rho, C(rho), ylab = "plant mortality")
+
+```
+
+### bistability
+
+Plotting the growth rate against the different mortality rates shows how the definition of grazing mortality alters our prediction of the steady states of the system.
+
+Compared to the more realistic non-linear functions, the constant grazing mortality assumption seems to predict both the attractor of the vegetated state and the tipping point much better than the density dependent mortality used in Kéfi et al 2007 TPB.
+In fact, the linear density dependent mortality does not predict bistability at all, given a simple logistic growht function.
+
+The critical threshold that would lead to degradation is overestimated by the constant model.
+
+The type 3 functional response finally predicts the existance of a stable degraded state that is not a desert, i.e. not entirely unvegetated.
+
 ```{r, echo=FALSE}
 
-defplot(rho, G(rho), ylab = "growth")
-lines(rho, C(rho, c_max = 0.1, T_h =0.3))
-lines(rho, C(rho, c_max = 0.1))
+defplot(rho, G(rho), ylab = "growth", col = "green3")
+lines(rho, C(rho, a = 10, q = 1))
+lines(rho, rep(0.1, length(rho)))
+lines(rho, C(rho, a = 3, q = 0))
+lines(rho, C(rho, a = .2, h = 0.00001, q = 0))
+
 ```
+
+### open questions
+
+#### Intrinsic mortality
+
+The mortality of plants regardless of grazing cover must be considered to be already internalised into the logistic growth definition.
+
+Alternatively, it can be considered as an independent constant rate, $M(V) = mV$,
+
+$$ \frac{dV}{dt} = G(V) - M(V) - C(V) $$
+
+and thus would affect the alternative stable states of the ecosystem.
+
+```{r, echo=FALSE}
+
+defplot(rho, G(rho), ylab = "growth", col = "green3")
+m <- rho*0.05
+lines(rho, m, lty = 2)
+lines(rho, m+C(rho, a = 10, q = 1))
+lines(rho, m+rep(0.1, length(rho)))
+lines(rho, m+C(rho, a = 3, q = 0))
+lines(rho, m+C(rho, a = .2, h = 0.00001, q = 0))
+
+```
+
+## spatially-explicit models
+
+The models described above were discussed by @noy-meir75 for grazing systems where vegetation is assumed to be of homogeneous distribution, and no spatially-explicit effects are of relevance. 
+Refugia or fencing of areas, however, could be considered as spatial effects.
+
+In arid and semi-arid rangelands, spatial vegetation structure was considered to be of major importance for plant growth and mortality [@rietkerk97; @kefi07].
+It generates patterns that act as positive feed-backs on the local scale by providing refugia to grazing and overcome growth limitations at low vegetation cover.
+
+These mechanisms can be implemented in the model by using a pair-approximation approach.
+
+### growth
+
+### spatially-explicit grazing mortality
+
+# References

### 2014-11-19 documentation/model_formulation.Rmd
update documentation

@@
     keep_md: yes
-    self_contained: no
+fontsize: 12pt
 geometry: margin=1.4in
-fontsize: 12pt
+csl: ../manuscript/amnat.csl
 bibliography: ../manuscript/cas02.bib
-csl: ../manuscript/amnat.csl
 ---
@@
 
-$$ \frac{dV}{dt} = G(V) - C(V) $$
+$$ \frac{dV}{dt} = G(V) - C(V)H $$
 
-with growth function $G$ and a function $C$ which represents consumption by herbivores.
+with growth function $G$ and a function $C$ which represents consumption by a herbivore multiplied by the density of herbivores, $H$.
 Formally, intrinsic mortality could be included in both terms.
@@
             xlab = "vegetation cover")
+
+rho <- seq(0,1,length = 100)
+
 ```
@@
 ## non-spatial models
+
 ### logistic growth
@@
 G <- function(P, r = 0.5, K = 0.9) r*P*(1-P/K)
-rho <- seq(0,1,length = 100)
 
@@
 ```
+
 #### mechanistic growth assumptions
+
+The logistic growth model is phenomenological and does not provide a particular ecological explanation, why growth is decelerating to zero with increasing cover.
+It assumes just some form of global competition leading to a balance of gains and losses at the carrying capacity of the ecosystem.
+
+Mechanistic explanations can be the limitations by a globally diffuse resource, like water, or the occupation of space and the subsequent limitation of photosynthetic active surface.
+
+A multitude of such global feedback processes can define the precise shape of the growth function leading to asymetrical or truncated shapes, which could importantly affect the bistability properties of the system [@noy-meir75].
+He defines
+
+> The general shape of the growth function relating net growth to plant biomass, $G(V)$, is convex with a single maximum (Fig.
+1).
+The increase in the low-biomass range expresses the increase in photosynthetic capacity with increasing leaf area.
+The levelling
+off to a plateau and the decrease in the higher range expresses self-interference effects (shading, competition); at the point $V = V_\mathrm{max}$ 'maintenance' losses equal photosynthesis and $G = 0$.
+Vmax is the maximum (stable) biomass in ungrazed vegetation.
+[...] Of the explicit forms the best known is the logistic [...] but other functions are possible.
+
+```{r, echo=FALSE}
+G <- function(P, r = 0.5, K = 0.9) r*P*(1-P/K)
+
+defplot(rho, G(rho), ylab = "growth")
+
+```
+
+These are global feed-backs. But local feedbacks also could lead to this deceleration:  
+For instance, if plant cover would have a dominantly negative effect on it's surroundings, by depleting nutrients or light, the local neighborhood of a plant cannot be occupied as easy as more remote places.
+As cover increases and the interspace areas are closing in, the space available for rejuvenation approaches zero.
 
@@
 
-In the cellular automata model and mean-field approximations of Sonia, individual plant mortality is a constant rate, which means it is increasing linearly with vegetation cover.
+In the cellular automata model and mean-field approximations of @kefi07, individual plant mortality is a constant rate, which means it is increasing linearly with vegetation cover.
 
@@
 C <- function(V, m = 0.25) (m*V)
-rho <- seq(0,1,length = 100)
 
@@
 C <- function(V, m = 0.1) m + V*0
-rho <- seq(0,1,length = 100)
 
@@
 ```{r, echo=FALSE}
-C <- function(V, a = 0.5, h = 10) (a*V)/(1+a*h*V)
-rho <- seq(0,1,length = 100)
+C <- function(V, a = 8, h = 10) (a*V)/(1+a*h*V)
 
@@
 ```{r, echo=FALSE}
-C <- function(V, a = 1, h = 10, q = 1) (a*V^(1+q))/(1+a*h*(V^(1+q)) ) 
-rho <- seq(0,1,length = 100)
+C <- function(V, a = 8, h = 10, q = 1) (a*V^(1+q))/(1+a*h*(V^(1+q)) )
 
@@
 
+In the formalism of the pair-approximation model of @kefi07, growth, *i.e.* the total amount of empty cells transitioning into vegetated cells, becomes a complex function of the relative cover of the multiple pairs including a $+$.
+
+growth part of equation (28) from @kefi07:
+
+$$  \rho_{0}w_{0,+} = \left( \delta \rho_+ + (1-\delta) \frac{\rho_+ - \rho_{+-} - \rho_{++} }{1 - \rho_+ - \rho_-} \right) (b - c \rho_+)(1- \rho_+ - \rho_-)  $$
+
+The goal of this exercise would be to transform this equation into a function of $\rho_+$, assuming that the landscape is at a vegetated equilibrium.
+
+Therefore,
+
+$$ \frac{d\rho_+}{dt} = \frac{d\rho_{+,-}}{dt} = \frac{d\rho_-}{dt} =0$$
+
 ### spatially-explicit grazing mortality

### 2014-11-20 documentation/solving_ODEs.Rmd
initialise solving_ODEs.Rmd

---
title: " Solving ODEs in R"
author: "Florian Schneider"
date: "Thursday, November 20, 2014"
output:
  html_document:
    keep_md: yes
  pdf_document:
    fig_height: 3.5
    fig_width: 4
fontsize: 12pt
geometry: margin=1.4in
csl: ../manuscript/amnat.csl
bibliography: ../manuscript/cas02.bib
---

To evaluate a spatially explicit model, we can run the pair-approximation descibed by @kefi07 into equilibrium.

We translate the system of five equations given in @kefi07 for the pair-approximation model into `R` functions.

(25)
$$ \frac{d\rho_{+-}}{dt} = d(\rho_+ - \rho_{+-} - \rho_{++}) +  (\rho_- - \rho_{--} - \rho_{+-}) \times \left( \delta\rho_+ + \frac{z-1}{z}(1-\delta)\frac{\rho_+ - \rho_{+-}  \rho_{++}}{1 - \rho_{+}  \rho_{-}}  \right) \times (b-c\rho_+) - \rho_{+-}\left( r + \frac{f}{z}+\frac{z-1}{z}f\frac{\rho_{+-}}{\rho_-}+m \right) $$

where $z$ is the number of neighbors considered, i.e. $z = 4$.

```{r}
d_rho_pm <- function(rho, parms, z = 4) { 
  with(parms,                             
      d * (rho[4] - rho[1] - rho[2]) + 
        (rho[5] - rho[3] - rho[1]) * 
        (delta*rho[4] + (z-1)/z * (1 - delta) * 
           (rho[4] - rho[1] - rho[2])/(1 - rho[4] - rho[5]) ) *
        (b - c*rho[4]) - 
        rho[1]*( r + f/z + (z-1)/z * f * rho[1]/rho[5] + m)
   )
}

```

(26)

$$  \frac{\mathrm{d}\rho_{++}}{\mathrm{d}t} = 
  2(\rho_+ - \rho_{+-}  \rho_{++}) \times \left( \delta\rho_+ + \frac{1-\delta}{z} + \frac{z-1}{z}(1-\delta)\frac{\rho_+ - \rho_{+-}  \rho_{++}}{1 - \rho_{+}  \rho_{-}}  \right) \times
  (b-c\rho_+) - 2\rho_{++}m
  $$

```{r}
d_rho_pp <- function(rho, parms, z = 4) { 
  with(parms,                             
      2 * (rho[4] - rho[1] - rho[2]) * 
        (delta*rho[4] + (1-delta)/z +(z-1)/z * (1 - delta) * 
           (rho[4] - rho[1] - rho[2])/(1 - rho[4] - rho[5]) ) *
        (b - c*rho[4]) - 2*rho[2]*m
   )
}

```

(27)

$$  \frac{\mathrm{d}\rho_{--}}{\mathrm{d}t} = 
    2d(\rho_{-} - \rho_{--} - \rho_{+-}) - 2\rho_{--} 
    \left( r + \frac{z-1}{z} f \frac{\rho_{+-}}{\rho_{-}} \right) $$

```{r}
d_rho_mm <- function(rho, parms, z = 4) { 
  with(parms,                             
      2 * (rho[5] - rho[3] - rho[1]) - 
        2*rho[3]*(r+(z-1)/z*f*rho[1]/rho[5])
   )
}

```

(28)

$$  \frac{\mathrm{d}\rho_{+}}{\mathrm{d}t} = 
    \left( \delta\rho_+ + (1-\delta)\frac{\rho_+ - \rho_{+-} -  \rho_{++}}{1 - \rho_{+}  \rho_{-}}  \right) \times
    (b-c\rho_+)(1-\rho_+ - \rho_-) - m\rho_{+}
     $$

```{r}
d_rho_p <- function(rho, parms, z = 4) { 
  with(parms,                             
      (delta*rho[4] + (1 - delta) * 
           (rho[4] - rho[1] - rho[2])/(1 - rho[4] - rho[5]) ) *
        (b-c*rho[4])*(1-rho[4]-rho[5])-m*rho[4]
      )
}

```

(29)
$$  \frac{\mathrm{d}\rho_{-}}{\mathrm{d}t} = 
d (1-\rho_+ - \rho_-) - \left(r + f \frac{\rho_{+-}}{\rho_{-}}  \right)\rho_{-} $$

```{r}
d_rho_m <- function(rho, parms, z = 4) { 
  with(parms,                             
      d*(1-rho[4]-rho[5])-(r+f*rho[1]/rho[5])*rho[5]
      )
}

```

## solving the ODE-system

In R, the package 'deSolve' provides solvers for differential equations. 
```{r}
library("deSolve")
```

It requires a single function that calculates the current state of a system from a previous state of the system (entered as a vector, here `rho`) with a timestep `t`. The third parameter expects any parameters that are required inside the function.

I define a function `odesys()` that returns the results of the single equations concenated into a vector.

```{r}
odesys <- function (t, rho, parms = model_parms) {
  list(c(
    d_rho_pm(rho, parms),
    d_rho_pp(rho, parms),
    d_rho_mm(rho, parms),
    d_rho_p(rho, parms),
    d_rho_m(rho, parms)
    ))
}

```

We need to define the constants, as well as the initial vegetation structure which is unambigously given by the values $\rho_{+-}$, $\rho_{++}$, $\rho_{--}$, $\rho_{+}$ and $\rho_{-}$. We define the starting values as a static function of $\rho_{+}$ and store all of them in a single vector called `rho` which is expected by the functions defined above.

```{r}
# model parameters
model_parms <- list(
  delta = 0.1, # proportion of seeds dispersed globally
  b = 0.75, # environmental quality
  m = 0.15, # plant mortality
  f = 0.9, # effect of abiotic facilitation on regeneration
  r = 0.01, # intrinsic regeneration 
  d = 0.1, # rate of degradation
  c = 0.2 # global competition among plants 
  )

# initial vegetation cover
rho_ini <- 0.2

rho <- c(rho_pm = rho_ini*0.4, # rho[1]
         rho_pp = rho_ini*0.4 , # rho[2]
         rho_mm = (1-rho_ini)*0.6, # rho[3]
         rho_p = rho_ini,  # rho[4]
         rho_m = (1-rho_ini)*0.9   # rho[5]
        )
```

Finally, the model is run by calling the function `ode()`. The output is stored in object `out` and the missing $\rho$s are completed.

```{r}
# running the ode-solver
runmodel <- ode(y = rho, func = odesys, times = 1.1^(1:60), parms = model_parms)

# transfer into ouput and calculate missing rho values
out <- as.data.frame(runmodel)
out$rho_0 <- 1 - out$rho_p - out$rho_m
out$rho_p0 <- with(out, rho_p-rho_pm-rho_pp)
out$rho_m0 <- with(out, rho_m-rho_pm-rho_mm)
out$rho_00 <- with(out, rho_0-rho_m0-rho_p0)

# plot 
plot(rho_p ~ time, data = out, type = "l", lwd = 2, ylab = "rho_plus", main = "pair-approximation", ylim = c(0,1))
lines(rho_m ~ time, data = out, lty = 3)
lines(rho_0 ~ time, data = out, col = "grey50")
```

## simulating over given $\rho_+$
For the project, we need a prediction of growth for a given value of $\rho_+$. To achieve this, I reformulate the ODE system to be a function of $\rho_+$, which I add to the list of parameters.

```{r}

odesys <- function (t, rho, parms = model_parms) {
  list(c(
    d_rho_pm(rho, parms),
    d_rho_pp(rho, parms),
    d_rho_mm(rho, parms),
    model_parms$rho_p,
    d_rho_m(rho, parms)
    ))
}

# initial vegetation cover
rho_ini <- 0.6
model_parms$rho_p <- rho_ini

rho <- c(rho_pm = rho_ini*0.4, # rho[1]
         rho_pp = rho_ini*0.4 , # rho[2]
         rho_mm = (1-rho_ini)*0.6, # rho[3]
         rho_p = rho_ini,  # rho[4]
         rho_m = (1-rho_ini)*0.9   # rho[5]
        )

```

```{r}
# running the ode-solver
runmodel <- ode(y = rho, func = odesys, times = 1.1^(1:60), parms = model_parms)

# transfer into ouput and calculate missing rho values
out <- as.data.frame(runmodel)
out$rho_0 <- 1 - out$rho_p - out$rho_m
out$rho_p0 <- with(out, rho_p-rho_pm-rho_pp)
out$rho_m0 <- with(out, rho_m-rho_pm-rho_mm)
out$rho_00 <- with(out, rho_0-rho_m0-rho_p0)

# plot 
plot(rho_p ~ time, data = out, type = "l", lwd = 2, ylab = "rho_plus", main = "pair-approximation", ylim = c(0,1))
lines(rho_m ~ time, data = out, lty = 3)
lines(rho_0 ~ time, data = out, col = "grey50")
```

# References

### 2014-11-20 documentation/model_formulation.Rmd
model formulation: equilibrium?

@@
 output:
+  html_document:
+    keep_md: yes
   pdf_document:
@@
     fig_width: 4
-  html_document:
-    keep_md: yes
 fontsize: 12pt
@@
 
-Therefore,
+Therefore, we are looking for the general case of $G(\rho_+)$ where
 
-$$ \frac{d\rho_+}{dt} = \frac{d\rho_{+,-}}{dt} = \frac{d\rho_-}{dt} =0$$
+$$ \frac{d\rho_+}{dt} = \frac{d\rho_-}{dt} = \frac{d\rho_{+,-}}{dt}  = \frac{d\rho_{+,+}}{dt} = \frac{d\rho_{-,-}}{dt}   = 0$$
+
+Equilibrium has to fulfil
+
+$$ \rho_0w_{0,+} = - m \rho_+ \,.$$
+
+thus
+
+$$  - m \rho_+ = \left( \delta \rho_+ + (1-\delta) \frac{\rho_+ - \rho_{+-} - \rho_{++} }{1 - \rho_+ - \rho_-} \right) (b - c \rho_+)(1- \rho_+ - \rho_-)  $$
 

### 2014-11-20 code/ode_kefi.r
for interactive model: externalise odes in file

@@ -1,0 +1,49 @@
+# system of ODEs
+
+d_rho_pm <- function(rho, parms, z = 4) { 
+  with(parms,                             
+       d * (rho[4] - rho[1] - rho[2]) + 
+         (rho[5] - rho[3] - rho[1]) * 
+         (delta*rho[4] + (z-1)/z * (1 - delta) * 
+            (rho[4] - rho[1] - rho[2])/(1 - rho[4] - rho[5]) ) *
+         (b - c*rho[4]) - 
+         rho[1]*( r + f/z + (z-1)/z * f * rho[1]/rho[5] + m)
+  )
+}
+d_rho_pp <- function(rho, parms, z = 4) { 
+  with(parms,                             
+       2 * (rho[4] - rho[1] - rho[2]) * 
+         (delta*rho[4] + (1-delta)/z +(z-1)/z * (1 - delta) * 
+            (rho[4] - rho[1] - rho[2])/(1 - rho[4] - rho[5]) ) *
+         (b - c*rho[4]) - 2*rho[2]*m
+  )
+}
+d_rho_mm <- function(rho, parms, z = 4) { 
+  with(parms,                             
+       2 * (rho[5] - rho[3] - rho[1]) - 
+         2*rho[3]*(r+(z-1)/z*f*rho[1]/rho[5])
+  )
+}
+d_rho_p <- function(rho, parms, z = 4) { 
+  with(parms,                             
+       (delta*rho[4] + (1 - delta) * 
+          (rho[4] - rho[1] - rho[2])/(1 - rho[4] - rho[5]) ) *
+         (b-c*rho[4])*(1-rho[4]-rho[5])-m*rho[4]
+  )
+}
+
+d_rho_m <- function(rho, parms, z = 4) { 
+  with(parms,                             
+       d*(1-rho[4]-rho[5])-(r+f*rho[1]/rho[5])*rho[5]
+  )
+}
+
+odesys <- function (t, rho, parms = model_parms) {
+  list(c(
+    d_rho_pm(rho, parms),
+    d_rho_pp(rho, parms),
+    d_rho_mm(rho, parms),
+    d_rho_p(rho, parms),
+    d_rho_m(rho, parms)
+  ))
+}

### 2014-11-20 documentation/interactive_kefi_2007_bifurcation.Rmd
interactive bifurcation graph (parallel backend)

---
title: "Pair-approximation model"
author: "Florian Schneider"
date: "Thursday, November 20, 2014"
output: html_document
runtime: shiny
---

## original model (Kéfi et al 2007)

Running the pair approximation model of Kéfi et al 2007 TPB dynamically.

```{r, echo = FALSE}
# provides parallel backend
library(foreach)
library(doSNOW)

workstation <-  list(host = "162.38.184.118", user = "schneider",
                     rscript = "/usr/lib/R/bin/Rscript",
                     snowlib = "/usr/lib/R/library")

workerlist <- c(rep("localhost", times = 10))

cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')

registerDoSNOW(cl)

```

### plant and soil parameters
```{r, echo=FALSE}

inputPanel(
  sliderInput("m", label = "plant mortality:",
              min = 0.0, max = 1, value = 0.25, step = 0.01),
  sliderInput("c", label = "competition:",
              min = 0.0, max = 1, value = 0.2, step = 0.01),
  sliderInput("delta", label = "seed distpersal:",
              min = 0.0, max = 1, value = 0.9, step = 0.01),
  sliderInput("f", label = "facilitation:",
              min = 0.0, max = 1, value = 0.9, step = 0.01), 
  sliderInput("r", label = "regeneration rate:",
              min = 0.0, max = 1, value = 0.01, step = 0.01), 
  sliderInput("d", label = "degradation rate:",
              min = 0.0, max = 1, value = 0.1, step = 0.01)
)

```

### initial conditions
```{r, echo=FALSE}

renderPlot({

# sets parameters of iterations

global <- list(  
  m = as.numeric(input$m), #intrinsic mortality
  d = as.numeric(input$d),    # degradation
  c = as.numeric(input$c) ,     # beta*g  
  delta = as.numeric(input$delta) ,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
  r = as.numeric(input$r), 	# regeneration rate
  f = as.numeric(input$f)		# local fascilitation
)

parameters <- list(
  b = seq(0,1,.005) #seq(0,1,.002)+0.001
)

iterations <- expand.grid(parameters)
iterations <- cbind(ID = 1:dim(iterations)[1],iterations, global)

rho_ini <- 0.8
rho <- c(rho_pm = rho_ini*0.4, # rho[1]
         rho_pp = rho_ini*0.4 , # rho[2]
         rho_mm = (1-rho_ini)*0.6, # rho[3]
         rho_p = rho_ini,  # rho[4]
         rho_m = (1-rho_ini)*0.9   # rho[5]
        )

foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
  setwd("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock")
  source("code/ode_kefi.r")

  model_parms <- as.list(iterations[iteration,])

  # running the ode-solver
  runmodel <- ode(y = rho, func = odesys, times = c(1,10000), parms = model_parms)

  # transfer into ouput and calculate missing rho values
  out <- as.data.frame(runmodel)
  out$rho_0 <- 1 - out$rho_p - out$rho_m
  out$rho_p0 <- with(out, rho_p-rho_pm-rho_pp)
  out$rho_m0 <- with(out, rho_m-rho_pm-rho_mm)
  out$rho_00 <- with(out, rho_0-rho_m0-rho_p0)
  out$ID <- iteration

  return(out[dim(out)[1],])
  } -> output

plot(output$rho_p ~ iterations$b, type = "l", ylim = c(0,1))

rho_ini <- 0.01
rho <- c(rho_pm = rho_ini*0.4, # rho[1]
         rho_pp = rho_ini*0.4 , # rho[2]
         rho_mm = (1-rho_ini)*0.6, # rho[3]
         rho_p = rho_ini,  # rho[4]
         rho_m = (1-rho_ini)*0.9   # rho[5]
        )

foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
  setwd("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock")
  source("code/ode_kefi.r")

  model_parms <- as.list(iterations[iteration,])

  # running the ode-solver
  runmodel <- ode(y = rho, func = odesys, times = c(1,10000), parms = model_parms)

  # transfer into ouput and calculate missing rho values
  out <- as.data.frame(runmodel)
  out$rho_0 <- 1 - out$rho_p - out$rho_m
  out$rho_p0 <- with(out, rho_p-rho_pm-rho_pp)
  out$rho_m0 <- with(out, rho_m-rho_pm-rho_mm)
  out$rho_00 <- with(out, rho_0-rho_m0-rho_p0)
  out$ID <- iteration

  return(out[dim(out)[1],])
  } -> output

lines(output$rho_p ~ iterations$b)

} )

```

### 2014-11-20 documentation/solving_ODEs.Rmd
run simulation of ODE system of Kefi et al 2007

@@
 ## simulating over given $\rho_+$
-For the project, we need a prediction of growth for a given value of $\rho_+$. To achieve this, I reformulate the ODE system to be a function of $\rho_+$, which I add to the list of parameters.
+For the project, we need a prediction of growth for a given value of $\rho_+$. To achieve this, I am simulating a gradient of replicates with different environmental conditions.
+
+```{r, echo = FALSE}
+library(foreach)
+library(doSNOW)
+
+workstation <-  list(host = "162.38.184.118", user = "schneider",
+                     rscript = "/usr/lib/R/bin/Rscript",
+                     snowlib = "/usr/lib/R/library")
+
+workerlist <- c(rep("localhost", times = 10))
+
+cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')
+
+registerDoSNOW(cl)
+
+```
 
@@
 
-odesys <- function (t, rho, parms = model_parms) {
-  list(c(
-    d_rho_pm(rho, parms),
-    d_rho_pp(rho, parms),
-    d_rho_mm(rho, parms),
-    model_parms$rho_p,
-    d_rho_m(rho, parms)
-    ))
-}
+first_ID = 1  #
+
+global <- list(  
+  m = 0.1, #intrinsic mortality
+  d = 0.1,    # degradation
+  c = 0.2, 		# beta*g  
+  delta = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+  r = 0.01, 	# regeneration rate
+  f = 0.9		# local fascilitation
+)
+
+parameters <- list(
+  b = seq(0,1,.01) #seq(0,1,.002)+0.001
+)
+
+iterations <- expand.grid(parameters)
+iterations <- cbind(ID = 1:dim(iterations)[1],iterations, global)
+str(iterations)
 
 # initial vegetation cover
-rho_ini <- 0.6
-model_parms$rho_p <- rho_ini
+rho_ini <- 0.2
 
@@
 ```{r}
+
+foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
+ model_parms <- as.list(iterations[iteration,])
+
   # running the ode-solver
-runmodel <- ode(y = rho, func = odesys, times = 1.1^(1:60), parms = model_parms)
+  runmodel <- ode(y = rho, func = odesys, times = c(1,60), parms = model_parms)
 
@@
   out$rho_00 <- with(out, rho_0-rho_m0-rho_p0)
+  out$ID <- iteration
 
-# plot 
-plot(rho_p ~ time, data = out, type = "l", lwd = 2, ylab = "rho_plus", main = "pair-approximation", ylim = c(0,1))
-lines(rho_m ~ time, data = out, lty = 3)
-lines(rho_0 ~ time, data = out, col = "grey50")
+  return(out[dim(out)[1],])
+  } -> output
+
+plot(output$rho_p ~ iterations$b, type = "l", ylim = c(0,1))
+
+stopCluster(cl)
+
 ```

### 2014-11-20 documentation/solving_ODEs.Rmd
solving ODEs: add simulation of gradient, add references

@@
 ---
-title: " Solving ODEs in R"
+title: ' Solving ODEs in R'
 author: "Florian Schneider"
@@
   pdf_document:
-    fig_height: 3.5
-    fig_width: 4
-fontsize: 12pt
-geometry: margin=1.4in
+    fig_height: 5
+    fig_width: 7
+fontsize: 10pt
+geometry: margin=1.2in
 csl: ../manuscript/amnat.csl
@@
 
-In R, the package 'deSolve' provides solvers for differential equations. 
+In R, the package 'deSolve' provides solvers for differential equations [@soetaert10].
+
 ```{r}
@@
 
-We need to define the constants, as well as the initial vegetation structure which is unambigously given by the values $\rho_{+-}$, $\rho_{++}$, $\rho_{--}$, $\rho_{+}$ and $\rho_{-}$.
-We define the starting values as a static function of $\rho_{+}$ and store all of them in a single vector called `rho` which is expected by the functions defined above.
+**Important:** *Equations of ODE and function for ODE solver are saved in file `code/ode_kefi.r` for further use in simulations and applications.*
+
+We need to define the constants, which I take from the example of @kefi07, Fig.
+3 to evaluate this implementation (looks very similar, differences might originate from the starting conditions, which are not given precisely in the paper).
 
@@
 model_parms <- list(
-  delta = 0.1, # proportion of seeds dispersed globally
-  b = 0.75, # environmental quality
-  m = 0.15, # plant mortality
-  f = 0.9, # effect of abiotic facilitation on regeneration
-  r = 0.01, # intrinsic regeneration 
-  d = 0.1, # rate of degradation
-  c = 0.2 # global competition among plants 
+  delta = 0, # proportion of seeds dispersed globally
+  b = 0.4, # environmental quality
+  m = 0.1, # plant mortality
+  f = 0.8, # effect of abiotic facilitation on regeneration
+  r = 0.0, # intrinsic regeneration 
+  d = 0.2, # rate of degradation
+  c = 0.3 # global competition among plants 
   )
+```
 
+And the initial vegetation structure which is unambigously given by the values $\rho_{+-}$, $\rho_{++}$, $\rho_{--}$, $\rho_{+}$ and $\rho_{-}$.
+We define the starting values as a static function of $\rho_{+}$ and store all of them in a single vector called `rho` which is expected by the functions defined above.
+
+```{r}
 # initial vegetation cover
-rho_ini <- 0.2
+rho_ini <- 0.1
 
@@
 
+Note that I am determining the timesteps exponentially, to have a decline in resolution as the dynamics steady out.
+
 ```{r}
 # running the ode-solver
-runmodel <- ode(y = rho, func = odesys, times = 1.1^(1:60), parms = model_parms)
+runmodel <- ode(y = rho, func = odesys, times = 1.1^(1:53), parms = model_parms)
 
@@
 out$rho_00 <- with(out, rho_0-rho_m0-rho_p0)
-
+```
+```{r, "fig1"}
 # plot
-plot(rho_p ~ time, data = out, type = "l", lwd = 2, ylab = "rho_plus", main = "pair-approximation", ylim = c(0,1))
+plot(rho_p ~ time, data = out, type = "l", lwd = 2, ylab = "vegetation cover", ylim = c(0,1))
 lines(rho_m ~ time, data = out, lty = 3)
 lines(rho_0 ~ time, data = out, col = "grey50")
+legend(60,1, c("rho_+", "rho_-", "rho_0"), col = c("black", "black", "grey50"), lty = c(1,3,1), lwd = c(2,1,1), bty = "n", horiz = TRUE)
 ```
 
-## simulating over given $\rho_+$
+## simulating gradients
+
 For the project, we need a prediction of growth for a given value of $\rho_+$. To achieve this, I am simulating a gradient of replicates with different environmental conditions.
 
-```{r, echo = FALSE}
+First, we need to generate a parallel backend for the use of the `foreach()` function.
+
+```{r, }
 library(foreach)
@@
 
+Then, build a database of parameters, using a global dataset (valid for all iterations) plus a gradient over $b$, stored in `parameters` (I use `expand.grid()` to blow up the parameter space if varying multiple parameters at a time).
+
 ```{r}
@@
 global <- list(  
-  m = 0.1, #intrinsic mortality
+  m = 0.2, #intrinsic mortality
   d = 0.1,    # degradation
   c = 0.2, 		# beta*g  
-  delta = 0.1,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
+  delta = 0.9,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
   r = 0.01, 	# regeneration rate
@@
 parameters <- list(
-  b = seq(0,1,.01) #seq(0,1,.002)+0.001
+  b = seq(0,1,.0025) #seq(0,1,.002)+0.001
 )
@@
 iterations <- cbind(ID = 1:dim(iterations)[1],iterations, global)
-str(iterations)
 
+```
+
+In a first simulation run, the initial vegetation cover should be high.
+
+```{r}
 # initial vegetation cover
-rho_ini <- 0.2
+rho_ini <- 0.8
 
@@
         )
-
 ```
 
+Then, call `foreach()` to run all iterations in parallel.
+
 ```{r}
-
 foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
@@
   # running the ode-solver
-  runmodel <- ode(y = rho, func = odesys, times = c(1,60), parms = model_parms)
+  runmodel <- ode(y = rho, func = odesys, times = c(1,10000), parms = model_parms)
 
@@
   return(out[dim(out)[1],])
-  } -> output
-
-plot(output$rho_p ~ iterations$b, type = "l", ylim = c(0,1))
-
-stopCluster(cl)
+  } -> output_high
 
@@
 
+In a second run, we initialise the same set of parameters, but starting from low cover:
+
+```{r}
+# initial vegetation cover
+rho_ini <- 0.01
+
+rho <- c(rho_pm = rho_ini*0.4, # rho[1]
+         rho_pp = rho_ini*0.4 , # rho[2]
+         rho_mm = (1-rho_ini)*0.6, # rho[3]
+         rho_p = rho_ini,  # rho[4]
+         rho_m = (1-rho_ini)*0.9   # rho[5]
+        )
+
+foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
+ model_parms <- as.list(iterations[iteration,])
+
+  # running the ode-solver
+  runmodel <- ode(y = rho, func = odesys, times = c(1,10000), parms = model_parms)
+
+  # transfer into ouput and calculate missing rho values
+  out <- as.data.frame(runmodel)
+  out$rho_0 <- 1 - out$rho_p - out$rho_m
+  out$rho_p0 <- with(out, rho_p-rho_pm-rho_pp)
+  out$rho_m0 <- with(out, rho_m-rho_pm-rho_mm)
+  out$rho_00 <- with(out, rho_0-rho_m0-rho_p0)
+  out$ID <- iteration
+
+  return(out[dim(out)[1],])
+  } -> output_low
+```
+
+```{r, "fig2"}
+plot(output_high$rho_p ~ iterations$b, type = "p", pch = 20, cex =0.75, ylim = c(0,1), xlab = "environmental quality, b", ylab = "vegetation cover, rho_+")
+points(output_low$rho_p ~ iterations$b, pch = 20, cex =0.75)
+```
+
+Don't forget to shut down the parallel backend.
+```{r}
+stopCluster(cl)
+```
+
 # References

### 2014-12-19 documentation/interactive_kefi_2007_bifurcation.Rmd
update model exploration

@@
               min = 0.0, max = 1, value = 0.2, step = 0.01),
-  sliderInput("delta", label = "seed distpersal:",
+  sliderInput("delta", label = "seed dispersal:",
               min = 0.0, max = 1, value = 0.9, step = 0.01),
@@
 
-plot(output$rho_p ~ iterations$b, type = "l", ylim = c(0,1))
+plot(output$rho_p ~ iterations$b, type = "p",pch  =20, ylim = c(0,1))
 
@@
 
-lines(output$rho_p ~ iterations$b)
+points(output$rho_p ~ iterations$b, pch = 20)
 

### 2014-12-19 documentation/solving_ODEs.Rmd
update model exploration

@@
 
+## Mortality vs Growth (simulated)
+
+It would be great to solve the equation system into a function of vegetation cover, $\rho_+$, if this is possible.
+
+```{r}
+output_high$M <- output_high$rho_p * model_parms$m
+output_high$G <- with(output_high, (rho_p * model_parms$delta + (1 - model_parms$delta ) * (rho_p - rho_pm - rho_pp)/(1-rho_p*rho_m))*(model_parms$b - model_parms$c * rho_p)*(1-rho_p-rho_m)  )
+
+plot( output_high$G ~ output_high$rho_p,
+      type = "l",
+      xlim = c(0,1) )
+
+lines(  output_high$M ~ output_high$rho_p, type = "l")
+
+```
+
 Don't forget to shut down the parallel backend.

### 2015-01-23 documentation/model_formulation.Rmd
model development, accelerating global growth

@@
 
-This document is supposed to become an interactive model representation of the pair-approximation model.
+This document is supposed to develop a model representation of the global mean field as well as pair-approximations of local feed back mechanisms.
 
@@
 
-$$ \frac{dV}{dt} = G(V) - C(V)H $$
+$$ \frac{dV}{dt} = G(V) - C(V)g $$
 
-with growth function $G$ and a function $C$ which represents consumption by a herbivore multiplied by the density of herbivores, $H$.
+with growth function $G$ and a function $C$ which represents consumption by a herbivore multiplied by the density of herbivores, the livestock rate, $g$.
 Formally, intrinsic mortality could be included in both terms.
@@
             family = "Times",
-            xlim = c(0,1), ylim = c(0,0.2), 
+            xlim = c(0,1), ylim = c(0,1), 
             xaxs = "i", yaxs = "i", 
-            xaxp = c(0,1,2), yaxp = c(0,0.2,2), 
+            xaxp = c(0,1,2), yaxp = c(0,1,2), 
             xlab = "vegetation cover")
@@
 ```{r, echo=FALSE}
-G <- function(P, r = 0.5, K = 0.9) r*P*(1-P/K)
+
+G <- function(P, r = 2, K = 0.9) r*P*(1-P/K)
 
@@
 ```{r, echo=FALSE}
-G <- function(P, r = 0.5, K = 0.9) r*P*(1-P/K)
+G <- function(P, r = 2, K = 0.9) r*P*(1-P/K)
 
@@
 
-#### decelerating growth at low cover
+#### accelerating growth
 
-In arid ecosystems, growth would be decelerated at very low cover, because retention of water and organic matter in the site is fundamentally reduced.
+In arid ecosystems, growth would be low at very low cover and accelerate as cover increases, because retention of water and organic matter in the site is fundamentally reduced.
 This postive feed-back (low cover reduces growth, leading to even lower cover) was discussed by @vandekoppel97 and @rietkerk97 to explain bistability in arid ecosystems.
 
-$$ G(V) = rV(1-V/K) \,,$$
+Starting out from the logistic growth term, we now additionally assume that $r$ is accelerated with cover,
+
+$$ r = r_0 * V^a $$
+
+$$ G(V) = r_0 V^{a+1} (1-V/K) \,,$$
+
+```{r, echo=FALSE}
+
+r <- function(P, r_0 = 2, a = 2) r_0*P^a
+
+defplot(rho, r(rho), ylab = "growth")
+
+G <- function(P, r_0 = 2, a = 2, K = 0.9) r(P, r_0, a)*P*(1-P/K)
+
+defplot(rho, G(rho), ylab = "growth")
+
+```
 

### 2015-01-27 documentation/model_formulation.Rmd
update model formulation

@@
             bty = "l", cex = 0.7, las = 1,
-            family = "Times",
             xlim = c(0,1), ylim = c(0,1), 
@@
 
-$$ G(V) = rV(1-V/K) \,,$$
+$$ G(V) = rV(1-\frac{V}{K}) \,,$$
 
@@
 ```{r, echo=FALSE}
-G <- function(P, r = 2, K = 0.9) r*P*(1-P/K)
+G <- function(P, r = 2.25, K = 0.9) r*P*(1-P/K)
 
@@
 ```
+
+**This model is going to be the starting point for the model development**
 
@@
 
-$$ G(V) = r_0 V^{a+1} (1-V/K) \,,$$
+$$ G(V) = r_0 V^{a+1} (1-\frac{V}{K}) \,,$$
 
@@
 
-G <- function(P, r_0 = 2, a = 2, K = 0.9) r(P, r_0, a)*P*(1-P/K)
+G_sig <- function(P, r_0 = 2, a = 2, K = 0.9) r(P, r_0, a)*P*(1-P/K)
 
-defplot(rho, G(rho), ylab = "growth")
+defplot(rho, G_sig(rho), ylab = "growth")
 
@@
 
-$$ C(V) = mV \,, $$
+$$ C(V) = gVL \,, $$
 
-with the grazing mortality, $m$.
+with the grazing mortality, $g$, per grazer individual multiplied by livestock density, $L$.
+That is to say, per individual, a constant percentage of cover is eaten.
+That is, ecologically speaking, not realistic.
 
 ```{r, echo=FALSE}
-C <- function(V, m = 0.25) (m*V)
+C <- function(V, L = 10, g = 0.01) (g*V*L)
 
@@
 
-$$ C(V) = m \,, $$
+$$ C(V) = gL \,, $$
 
-thus would be defined as constant, $m$, regardless of the vegetation cover.
+thus would be defined as constant, $g$, per grazer individual regardless of the vegetation cover.
 
 ```{r, echo=FALSE}
-C <- function(V, m = 0.1) m + V*0
+C <- function(V, L = 10, g = 0.01) g * L + V*0
 
@@
 ```
+
+Note that grazing mortality can exceed 1, meaning that grazers would exploit more biomass than provided at full cover.
+Thus stability can be only reached if growth is sufficiently higher and the landscape regenerates itself multiple times per year.
 
@@
 
-$$ C(V) = \frac{aV}{1+ahV} \,,$$
+$$ C(V) = \frac{aVL}{1+ahV} \,,$$
 
-with the handling time, $h$ [time per resource unit], and the search efficiency, $a$ [area foraged in per time].
+with the handling time, $h$ [time per landscape unit], and the search efficiency, $a$ [percent of landscape foraged in per time].
 
 ```{r, echo=FALSE}
-C <- function(V, a = 8, h = 10) (a*V)/(1+a*h*V)
+C <- function(V, L = 10, a = 0.6, h = 100) (a*V*L)/(1+a*h*V)
 
@@
 
+Since we are calculating in fractions of landscape, and not in cells, the handling time defines the time a grazer individual would need to handle the entire landscape as readily available fodder.
+The search time is the percentage of the landscape foraged in over the course of one year.
+
 #### type 3 functional response
+
+The
 
@@
 ```{r, echo=FALSE}
-C <- function(V, a = 8, h = 10, q = 1) (a*V^(1+q))/(1+a*h*(V^(1+q)) )
+C <- function(V, L = 10, a = 0.6, h = 100, q = 1) (a*V^(1+q)*L)/(1+a*h*(V^(1+q)) )
 
@@
 ```
+
+#### null model
+
+We assume the classic type II functional response plus an intrinsic mortality rate,
+
+$$ C(V) = m*V + \frac{aVL}{1+ahV} \,,$$
+
+```{r, echo=FALSE}
+
+C <- function(V, m_0 = 0.05, L = 10, a = 0.6, h = 100, q = 0) (m_0 * V) +(a*V^(1+q)*L)/(1+a*h*(V^(1+q)) )
+
+defplot(rho, C(rho), ylab = "plant mortality")
+abline(a = 0, b = 0.05, lty = 2)
+```
+
+**This model is going to be the starting point for our model development**
 
@@
 Compared to the more realistic non-linear functions, the constant grazing mortality assumption seems to predict both the attractor of the vegetated state and the tipping point much better than the density dependent mortality used in Kéfi et al 2007 TPB.
-In fact, the linear density dependent mortality does not predict bistability at all, given a simple logistic growht function.
+In fact, the linear density dependent mortality does not predict bistability at all, given a simple logistic growth function.
 
-The critical threshold that would lead to degradation is overestimated by the constant model.
+```{r, echo=FALSE}
+
+defplot(rho, G(rho), ylab = "growth", col = "green3")
+lines(rho, C(rho, L = 10, a = 0.01, h = 0.00001))
+lines(rho, C(rho, L = 25, a = 0.01, h = 0.00001))
+lines(rho, C(rho, L = 50, a = 0.01, h = 0.00001))
+abline( h = 0.1 )
+abline( h = 0.25 )
+abline( h = 0.5 )
+```
+
+When assuming a type II functional response, the system gains true bistability.
+
+```{r, echo=FALSE}
+
+defplot(rho, G(rho), ylab = "growth", col = "green3")
+lines(rho, C(rho, L = 10))
+lines(rho, C(rho, L = 25))
+lines(rho, C(rho, L = 50))
+
+```
 
@@
 defplot(rho, G(rho), ylab = "growth", col = "green3")
-lines(rho, C(rho, a = 10, q = 1))
-lines(rho, rep(0.1, length(rho)))
-lines(rho, C(rho, a = 3, q = 0))
-lines(rho, C(rho, a = .2, h = 0.00001, q = 0))
-
-```
-
-### open questions
-
-#### Intrinsic mortality
-
-The mortality of plants regardless of grazing cover must be considered to be already internalised into the logistic growth definition.
-
-Alternatively, it can be considered as an independent constant rate, $M(V) = mV$,
-
-$$ \frac{dV}{dt} = G(V) - M(V) - C(V) $$
-
-and thus would affect the alternative stable states of the ecosystem.
-
-```{r, echo=FALSE}
-
-defplot(rho, G(rho), ylab = "growth", col = "green3")
-m <- rho*0.05
-lines(rho, m, lty = 2)
-lines(rho, m+C(rho, a = 10, q = 1))
-lines(rho, m+rep(0.1, length(rho)))
-lines(rho, m+C(rho, a = 3, q = 0))
-lines(rho, m+C(rho, a = .2, h = 0.00001, q = 0))
+lines(rho, C(rho, L = 10, q = 1))
+lines(rho, C(rho, L = 25, q = 1))
+lines(rho, C(rho, L = 50, q = 1))
 
@@
 These mechanisms can be implemented in the model by using a pair-approximation approach. 
-
-### growth
+In numerical simulations $G$ and $C$ can be implemented as the transition rates in a two-state spatial model.
 
@@
 
-growth part of equation (28) from @kefi07:
+In a model with only global effects, the transition probability of indiviual plant death would be
 
-$$  \rho_{0}w_{0,+} = \left( \delta \rho_+ + (1-\delta) \frac{\rho_+ - \rho_{+-} - \rho_{++} }{1 - \rho_+ - \rho_-} \right) (b - c \rho_+)(1- \rho_+ - \rho_-)  $$
+$$ w_{1,0} = C(V)/V = m + \frac{aL}{1+ahV} $$
 
-The goal of this exercise would be to transform this equation into a function of $\rho_+$, assuming that the landscape is at a vegetated equilibrium.
+and the transition probability of individual plant growth
 
-Therefore, we are looking for the general case of $G(\rho_+)$ where
+$$ w_{0,1} = G(V)/V = r (1-\frac{V}{K}) \,.$$
 
-$$ \frac{d\rho_+}{dt} = \frac{d\rho_-}{dt} = \frac{d\rho_{+,-}}{dt}  = \frac{d\rho_{+,+}}{dt} = \frac{d\rho_{-,-}}{dt}   = 0$$
+### spatially-explicit growth
 
-Equilibrium has to fulfil
+We implement the accelerating and decelerating features on a local level.
 
-$$ \rho_0w_{0,+} = - m \rho_+ \,.$$
+#### accelerating growth
 
-thus
+This is assuming that growth rate is enhanced by a local facilitation term similar to the implementation in Kéfi et al 2007. The growth rate becomes a function of local vegetation cover $\nu$:
 
-$$  - m \rho_+ = \left( \delta \rho_+ + (1-\delta) \frac{\rho_+ - \rho_{+-} - \rho_{++} }{1 - \rho_+ - \rho_-} \right) (b - c \rho_+)(1- \rho_+ - \rho_-)  $$
+$$ r = r_0 + f \nu \,,$$
 
-### spatially-explicit grazing mortality
+with an intrinsic growth rate $r_0$ which determines the probability of growth in absence of local vegetation and the additive facilitation $f$ which gradually determines the enhancement of the growth with an increasing local vegetation cover (maximizing at $\nu = 1$ if the cell has 4 neighbors).
+
+#### decelerating growth
+
+This is assuming that growth is diminished locally by competition, $c$. This affects carrying capacity rather than growth rate itself.
+
+$$ K = K_0 (1 - c \nu)$$
+
+### spatially-explicit mortality
+
+On the mortality side, we can implement spatially explicit effects on the parameters of the functional response.
+
+#### accelerating mortality
+
+Mortality becomes enhanced in sparse environments by patches that are attracting grazers. We assume this attractivity, $v$, reduces search times locally.
+
+$$ a = a_0 * (1 - v * \nu) $$
+
+#### decelerating mortality
+
+Plants in grazed habitats develop protective traits, such as thorns or cushion growth.
+They thereby provide protected habitat to their direct neighborhood, or share the investments in those traits with their neighbors.
+Overall grazing mortality thus is reduced through an increase in handling time by associational resistance, $p$
+
+$$ h = h_0 * p * \nu $$
+
+$$ $$
+
+#### decelerating mortality
 

### 2015-01-29 code/simfunctions.r
include runCA and animateCA functions

@@ -1,20 +1,217 @@
+########################################################
+# The MIT License (MIT)
+#
+# Copyright (c) 2014 Florian D. Schneider & Sonia Kéfi
+# 
+# Permission is hereby granted, free of charge, to any person obtaining a copy
+# of this software and associated documentation files (the "Software"), to deal
+# in the Software without restriction, including without limitation the rights
+# to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+# copies of the Software, and to permit persons to whom the Software is
+# furnished to do so, subject to the following conditions:
+#   
+# The above copyright notice and this permission notice shall be included in all
+# copies or substantial portions of the Software.
+# 
+# THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+# IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+# FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+# AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+# LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+# OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+# SOFTWARE.
+#
+########################################################
 
-count  <- function(x, neighbor, state = NULL) {
-			if(is.null(state)) state = neighbor
-			
-			neighbors <- numeric(length = prod(x$dim))
-			x_logical_with_border <- (x$cells %in% neighbor)[x_with_border]
-			for(k in interact) {
-				neighbors <- neighbors + x_logical_with_border[x_to_evaluate+k]
-			}
-			return(neighbors)	
+
+######################
+## mapping function ##
+######################
+
+# required to vectorise the counting and plotting. 
+# returns a map of the landscape to translate it into a vector with boundaries and another one to back-translate it to a vector without boundaries into the global environment. Needs to be called only once for the dimensions of the lattice. 
+
+
[…]

### 2015-01-29 documentation/model_formulation.Rmd
update model formulation, include ODE and CA models

@@
 
-G <- function(P, r = 2, K = 0.9) r*P*(1-P/K)
+G <- function(P, r = 2, K = 1) r*P*(1-P/K)
 
@@
 
-The logistic growth model is phenomenological and does not provide a particular ecological explanation, why growth is decelerating to zero with increasing cover.
+In the first place the logistic growth model is phenomenological and does not provide a particular ecological explanation, why growth is decelerating to zero with increasing cover.
 It assumes just some form of global competition leading to a balance of gains and losses at the carrying capacity of the ecosystem.
@@
 The increase in the low-biomass range expresses the increase in photosynthetic capacity with increasing leaf area.
-The levelling
-off to a plateau and the decrease in the higher range expresses self-interference effects (shading, competition); at the point $V = V_\mathrm{max}$ 'maintenance' losses equal photosynthesis and $G = 0$.
+The levelling off to a plateau and the decrease in the higher range expresses self-interference effects (shading, competition); at the point $V = V_\mathrm{max}$ 'maintenance' losses equal photosynthesis and $G = 0$.
 Vmax is the maximum (stable) biomass in ungrazed vegetation.
 [...] Of the explicit forms the best known is the logistic [...] but other functions are possible.
+
+We consider these multiple competitive effects being of minor relevance for now and start with a single limitation by the factor space.
+As the landscape becomes more and more filled with vegetation, the empty space become scarce.
+Thus, growth declines until each potentially habitable spot is occupied.
+We implement a generally valid carrying capacity, $K$.
 
@@
 
-These are global feed-backs. But local feedbacks also could lead to this deceleration:  
-For instance, if plant cover would have a dominantly negative effect on it's surroundings, by depleting nutrients or light, the local neighborhood of a plant cannot be occupied as easy as more remote places.
-As cover increases and the interspace areas are closing in, the space available for rejuvenation approaches zero.
+We assume that growth rate is linearly related to environmental quality, $b$, which is inverse to aridity.
+Since drylands are water limited in the first place, this factor defines the reproduction rate (i.e.
+number of seeds produced per plant, number of seedlings surviving the harshness of the dry soils).
+We assume that
+
+$$ r = r_0 * b $$
+
+to have a gradient of environmental harshness, $b$, reaching from 0 to 1. 
+The model thus starts out with the assumption of decelerating growth due to competition for space.
 
@@
 
-In arid ecosystems, growth would be low at very low cover and accelerate as cover increases, because retention of water and organic matter in the site is fundamentally reduced.
+In arid ecosystems, growth rates would be lowest at very low cover and accelerate as cover increases, because retention of water and organic matter in the site is fundamentally reduced and water runoff is high.
 This postive feed-back (low cover reduces growth, leading to even lower cover) was discussed by @vandekoppel97 and @rietkerk97 to explain bistability in arid ecosystems.
@@
 
-$$ r = r_0 * V^a $$
+$$ r = r_0 b V^a $$
 
-$$ G(V) = r_0 V^{a+1} (1-\frac{V}{K}) \,,$$
+$$ G(V) = r_0 b V^{a+1} (1-\frac{V}{K}) \,,$$
 
@@
 
-r <- function(P, r_0 = 2, a = 2) r_0*P^a
+r <- function(P, r_0 =  2.25, a = 1) r_0 *P^a
 
-defplot(rho, r(rho), ylab = "growth")
+#defplot(rho, r(rho), ylab = "growth")
 
-G_sig <- function(P, r_0 = 2, a = 2, K = 0.9) r(P, r_0, a)*P*(1-P/K)
+G_sig <- function(P, r_0 = 2.25, a = 1,  K = 0.9) r(P, r_0, a)*P*(1-P/K)
 
 defplot(rho, G_sig(rho), ylab = "growth")
-
+lines(rho, G_sig(rho, a = 0))
+lines(rho, G_sig(rho,r_0 = 2.25, a = 1))
+lines(rho, G_sig(rho,r_0 = 2.25, a = 0.2))
 ```
 
-### functional response
+### plant mortality
 
@@
 In the cellular automata model and mean-field approximations of @kefi07, individual plant mortality is a constant rate, which means it is increasing linearly with vegetation cover.
-
-Plant mortality due to grazing therefore would be described as
+Total plant mortality due to grazing therefore would be described as
 
@@
 
-thus would be defined as constant, $g$, per grazer individual regardless of the vegetation cover.
+thus could be defined as constant, $g$, per grazer individual regardless of the vegetation cover.
 
@@
 at low cover, density dependent mortality defined by search efficiency dominates, but at high cover handling times of the consumer will limit its intake rates.
-
 Therefore, plant mortality due to grazing should be
@@
 
-Since we are calculating in fractions of landscape, and not in cells, the handling time defines the time a grazer individual would need to handle the entire landscape as readily available fodder.
+Since we are calculating in fractions of landscape, i.e.
+vegetation cover, and not in cells, the handling time defines the time a grazer individual would need to handle the entire landscape as readily available fodder.
 The search time is the percentage of the landscape foraged in over the course of one year.
@@
 
+### population dynamics
+
+The differential equation describing population cover above with $V$ and $C$ substituted would then become
+
+$$ \frac{\mathrm{d}\rho_1}{\mathrm{d}t} = r b V (1-\frac{V}{K}) - m V - \frac{aVL}{1+ahV} $$
+
+Given a parameter set  
+```{r, echo=FALSE}
+
+parms <- list(
+    m = 0.05,
+    r = 1,
+    b = 0.9, 
+    K = 0.9, 
+    a = 0.6,
+    h = 100,
+    L = 10
+  )
+
+ as.data.frame(parms)
+```
+we receive a projection of the intersection of growth and mortality.
+
+```{r, echo=FALSE}
+plot(rho,  with(parms,  m * rho + (a * rho * L)/(1 + a * h * rho)), ylab = "growth", type = "l", ylim = c(0,0.5))
+lines(rho, with(parms, r* b* rho * (1 - rho/K) ), col = "green3" )
+
+```
+
+And solving the ODE, starting with values larger than the unstable equilibrium point on the left,  returns a timeseries
+
+```{r, echo=FALSE}
+d_rho <- function(rho, parms, z = 4) { 
+  with(parms, r* b* rho * (1 - rho/K) - m * rho - (a * rho * L)/(1 + a * h * rho) )  
+}
+
+odesys <- function (t, rho, parms = model_parms) {
+  list( d_rho(rho, parms)  )
+}
+
+library(deSolve)
+# running the ode-solver
+runmodel <- ode(y = 0.90, func = odesys, times = 1.1^(1:53), parms = parms)
+
+# transfer into ouput and calculate missing rho values
+ODEnullmodel <- as.data.frame(runmodel)
+
+plot(ODEnullmodel, type = "l", ylim = c(0,1))
+
+```
+
+### Cellular Automata
+
+The differential equations are formulated as changes in vegetation cover. Growth are positive changes, and deaths are negative changes.
+
+Now, the CA uses transition probabilities. That is, a) the individual risk of death of vegetated cells and b) the individual chance of becoming vegetated for an empty cell.
+To translate into the CA context, we need to find rules that apply to a) the vegetated cells, and b) the empty cells.
+Thus, it would be a) the reduction of term of the population of vegetated cells, divided by cover, to get the individual risk of death
+
+$$ w_{1,0} = m + \frac{aL}{1+ah\rho_1} $$
+
+This would be b) the growth term for the population of empty cells, divided by cover of empty cells, to get the individual chance of growth.
+
+$$ w_{0,1} = \frac{r \rho_1 b (1-\frac{\rho_1}{K})}{1-\rho_1} $$
+
+From an individual cell perspective, the probabilities would change along with cover like this
+
+```{r, echo=FALSE}
+par(mar = c(4,4,2,4))
+plot(rho,  with(parms,  r * b *  ( 1 - (rho / K)) / (1 - rho) ) , ylab = "individual growth on empty cells", type = "l", ylim = c(0,2.5), las = 1, yaxs = "i")
+
+lines(rho, with(parms, m + ( a * L )/( 1 + a * h * rho ) ), col = "red3" )
+axis(4, las = 1)
+mtext("individual risk of death of vegetated cells", 4, 3)
+```
+
+Running this in a timeseries of 150 years yields
+
+```{r, echo = FALSE}
+
+source("../code/simfunctions.r")
+
+set.seed(584232) # setting seed for random number generation
+
+nullmodel <- runCA(0.9, parms, delta = 0.2, t_max = 150, t_min = 100, t_eval = 50, isstable = 0.0001, saveeach = 5)
+
+plot(nullmodel$rho_one ~ nullmodel$time, 
+     ylim = c(0,1), ylab = expression( rho["+"]), 
+     xlab = "time [yr]", type = "l")
+
+lines(ODEnullmodel, lty = 3)
+
+#animateCA(nullmodel, "ca_nullmodel.gif")
+
+```
+
 ## spatially-explicit models
@@
 
-We implement the accelerating and decelerating features on a local level.
+We implement accelerating and decelerating features on growth and mortality on the local level.
 
@@
 
-This is assuming that growth rate is enhanced by a local facilitation term similar to the implementation in Kéfi et al 2007. The growth rate becomes a function of local vegetation cover $\nu$:
+This is assuming that the local environmental suitability is enhanced by a local facilitation term similar to the implementation in Kéfi et al 2007.
+Thus, the growth rate becomes a function of local vegetation cover $\nu$:
 
@@
 
+For instance, if plant cover would have a dominantly negative effect on it's surroundings, by depleting nutrients or light, the local neighborhood of a plant cannot be occupied as easy as more remote places.
+As cover increases and the interspace areas are closing in, the space available for rejuvenation approaches zero.
 This is assuming that growth is diminished locally by competition, $c$. This affects carrying capacity rather than growth rate itself.
@@
 
-$$ $$
+### implementation in a cellular automaton
 
-#### decelerating mortality
+It is straigthforward to implement the model in a cellular automata model.
+It is defined on a defined grid of cells where each cell can take the values '1' if it is vegetated and '0' if it is empty.
+The transition rules between those states depend either on global vegetation cover $\rho_1$ or on the local cover in the direct neighborhood of the cell $\nu_1$ (closest 4 cells, von-Neumann-neighborhood).
+
+Thus, two transition probabilities need to be defined: the probability of **growth**, $w_{0,1}$, turning an empty cell into a vegetated, and the probability of **death**, $w_{0,1}$, removing vegetation from the cell.
+
+The null model, i.e. the model with only global effects on mortality and growth, with the basic assumption of a type II functional response and logistic growth, would thus be defined by
+
+$$ w_{0,1} = r (1-\frac{\rho_1}{K})$$
+
+$$ w_{1,0} =  m + \frac{aL}{1+ah\rho_1} $$
+
+### implementation in a pair-approximation model
+
+The pair-approximation framework (Matsuda et al 1992) is defined as a system of ordinary differential equations, describing the populations of *pairs* of cells.
+In a two state system, three pairs can be defined with population densities $\rho_{1,1}$ , $\rho_{0,0}$, and $\rho_{0,1}$.
+Besides, there are the population densities of singletons, $\rho_0$ and $\rho_1$.
+
+There are conservation equations which allow to calculate population densities of one pair out of the other pairs and singletons.
+
+$$ \rho_1 + \rho_0 = 1$$
+
+$$ \rho_{1,1} + \rho_{0,1} = \rho_1$$
+
+$$ \rho_{0,1} + \rho_{0,0} = \rho_0$$
+
+This means, the system can be fully described by two differential equations:
+
+$$ \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = $$
+
+$$ \frac{\mathrm{d}\rho_0}{\mathrm{d}t} = $$
 

### 2015-01-29 code/simulation.r
a simple CA simulation framework

@@ -1,0 +1,59 @@
+########################################################
+# The MIT License (MIT)
+#
+# Copyright (c) 2014 Florian D. Schneider & Sonia Kéfi
+# 
+# Permission is hereby granted, free of charge, to any person obtaining a copy
+# of this software and associated documentation files (the "Software"), to deal
+# in the Software without restriction, including without limitation the rights
+# to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+# copies of the Software, and to permit persons to whom the Software is
+# furnished to do so, subject to the following conditions:
+#   
+# The above copyright notice and this permission notice shall be included in all
+# copies or substantial portions of the Software.
+# 
+# THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+# IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+# FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+# AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+# LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+# OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+# SOFTWARE.
+#
+########################################################
+
+rm(list=ls())
+
+########################################################################################
+source("code/simfunctions.r")
+
+set.seed(584232) # setting seed for random number generation
+
+# parameter settings:
+parameters <- list(  
+  m = 0.05, #intrinsic mortality
+  r = 0.5, # intrinsic growth rate
+  b = 1,   # environmental quality / 1 - aridity
+  K = 0.9, # carrying capacity 
+  a = 0.6, # search efficiency
+  h = 100, # handling time
+  L = 10 # livestock units / grazing intensity
+)
+
+
+nullmodel <- runCA(runif(1, 0.8,0.9), parameters, delta = 0.2, t_max = 150, t_min = 100, t_eval = 50, stability = 0.0001, saveeach = 5)
+
+
+par(mfrow = c(2,1))
[…]

### 2015-01-29 code/simfunctions.r
fix bug in runCA function; works now !!!

@@ -234,7 +234,9 @@ animateCA <- function(result, filename) {
   library(animation)
   if(Sys.info()[['sysname']] == "Linux") X11.options(antialias = "none") #for Linux Systems to enable pixel-wise plotting in (animated) gif-files. 
   if(Sys.info()[['sysname']] == "Windows") windows.options(antialias = "none") #for Windows Systems to enable pixel-wise plotting in (animated) gif-files. 
-  
+  width = result$timeseries[[1]]$dim[1]
+  height = result$timeseries[[1]]$dim[2]
+    
   saveGIF( 
     for(i in 1:length(result$timeseries) ) {
       par(mar = c(0,0,0,0))
@@ -289,6 +291,7 @@ runCA <- function(init, parms, width = 100, height = 100, delta = 0.1, t_max = 1
   
   # initialise simulation variables: 
   x_old <- initial  # ghost matrix at t_i
+  x_new <- x_old
   stability <- 1  # check value for stability
   i = 0  # iterator for simulation timesteps
   
@@ -296,8 +299,7 @@ runCA <- function(init, parms, width = 100, height = 100, delta = 0.1, t_max = 1
   while(stability > isstable & i <= t_max/delta & parms_temp$rho_one > 0) {
     
     i <- i +1  # increase iterator
-    x_new <- x_old   	# copy x_old into an object x_new to allocate memory	
-    
+   
     # model specific part:
     # 1 - setting time-step parameters
     #parms_temp$local <- count(x_old, "1")/4  # count local density of occupied fields for each cell:
@@ -308,7 +310,7 @@ runCA <- function(init, parms, width = 100, height = 100, delta = 0.1, t_max = 1
     # 3 - setting transition probabilities
     growth <- with(parms_temp, r * b * rho_one * ( 1 - (rho_one / K)) / (1 - rho_one)  *delta)  # recolonisation rates of all cells 
     
-    death <- with(parms_temp, m + ( a * L )/( 1 + a * h * rho_one ) *delta)   # set probability of death for each cell
+    death <- with(parms_temp, (m + ( a * L )/( 1 + a * h * rho_one )) *delta)   # set probability of death for each cell
     
     # check for sum of probabilities to be inferior 1 and superior 0
     if(any(c(growth, death) > 1 )) warning(paste("a set probability is exceeding 1 in time step", i, "! decrease delta!!!")) 

### 2015-01-29 documentation/model_formulation.Rmd
develop nullmodel for cellular automata and ODE

@@
 par(mar = c(4,4,2,4))
-plot(rho,  with(parms,  r * b *  ( 1 - (rho / K)) / (1 - rho) ) , ylab = "individual growth on empty cells", type = "l", ylim = c(0,2.5), las = 1, yaxs = "i")
+plot(rho,  with(parms,  r * b * rho * ( 1 - (rho / K)) / (1 - rho) ) , ylab = "individual growth on empty cells", type = "l", ylim = c(0,2.5), las = 1, yaxs = "i")
 
@@
 
-Running this in a timeseries of 150 years yields
+Running this on a 100 x 100 grid of cells in a timeseries of 150 years yields (dashed line: ODE prediction as reference)
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, warning = FALSE}
 
@@
 
-nullmodel <- runCA(0.9, parms, delta = 0.2, t_max = 150, t_min = 100, t_eval = 50, isstable = 0.0001, saveeach = 5)
+nullmodel <- runCA(0.9, parms, delta = 0.2, t_max = 150, t_min = 100, t_eval = 50, isstable = 0.0001, saveeach = 1)
 
@@
 lines(ODEnullmodel, lty = 3)
+```
 
-#animateCA(nullmodel, "ca_nullmodel.gif")
-
+And the animated lattice would look like this:
+```{r, eval = FALSE, echo = FALSE, warning = FALSE}
+animateCA(nullmodel, "ca_nullmodel.gif")
 ```
+![](ca_nullmodel.gif)
 

### 2015-01-29 documentation/interactive_nullmodel.Rmd
add interactive Version of Nullmodel in R Shiny

---
title: "null-model ode"
author: "Florian Schneider"
date: "Thursday, November 20, 2014"
output: html_document
runtime: shiny
---

## original model (Kéfi et al 2007)

Running the pair approximation model of Kéfi et al 2007 TPB dynamically.

```{r, echo=FALSE}

library("deSolve")

rho <- seq(0,1,length = 100)

```

### plant and soil parameters
```{r, echo=FALSE}

inputPanel(
  sliderInput("m", label = "intrinsic mortality:",
              min = 0.0, max = 1, value = 0.05, step = 0.01),
  sliderInput("r", label = "max. growth rate:",
              min = 0.0, max = 3, value = 1, step = 0.01),
  sliderInput("b", label = "environmental quality:",
              min = 0.0, max = 1, value = 0.9, step = 0.01),
  sliderInput("K", label = "carrying capacity:",
              min = 0.0, max = 1, value = 0.9, step = 0.01),
  sliderInput("a", label = "search efficiency:",
              min = 0.0, max = 1, value = 0.6, step = 0.01), 
  sliderInput("h", label = "handling time:",
              min = 0.0, max = 1000, value = 100, step = 1), 
  sliderInput("L", label = "livestock units:",
              min = 0.0, max = 100, value = 10, step = 1)
)

```

### initial conditions
```{r, echo=FALSE}

inputPanel(
  sliderInput("rho_1", label = "initial plant cover:",
              min = 0.0, max = 1, value = 0.9, step = 0.01)
  )

renderPlot({

 parms <- list()
  parms$m <- as.numeric(input$m)
  parms$r <- as.numeric(input$r)
  parms$b <- as.numeric(input$b) 
  parms$K <- as.numeric(input$K) 
  parms$a <- as.numeric(input$a) 
  parms$h <- as.numeric(input$h) 
  parms$L <- as.numeric(input$L)

  plot(rho,  with(parms,  m * rho + (a * rho * L)/(1 + a * h * rho)), ylab = "growth", type = "l", ylim = c(0,0.5))
  lines(rho, with(parms, r* b* rho * (1 - rho/K) ), col = "green3" )

  })

renderPlot({

d_rho <- function(rho, parms, z = 4) { 
  with(parms, r* b* rho * (1 - rho/K) - m * rho - (a * rho * L)/(1 + a * h * rho) )  
}

odesys <- function (t, rho, parms = model_parms) {
  list( d_rho(rho, parms)  )
}

 parms <- list()
  parms$m <- as.numeric(input$m)
  parms$r <- as.numeric(input$r)
  parms$b <- as.numeric(input$b) 
  parms$K <- as.numeric(input$K) 
  parms$a <- as.numeric(input$a) 
  parms$h <- as.numeric(input$h) 
  parms$L <- as.numeric(input$L)

# running the ode-solver
runmodel <- ode(y = input$rho_1, func = odesys, times = 1.1^(1:53), parms = parms)

# transfer into ouput and calculate missing rho values
out <- as.data.frame(runmodel)

plot(out, type = "l", ylim = c(0,1))
axis(2, at = round(tail(out,1)[["1"]],2), las = 1 )

} )
```

### 2015-01-30 documentation/interactive_nullmodel.Rmd
labelling nullmodel document
ODE of global model, without spatially-explicit terms (Shiny)

@@
 ---
-title: "null-model ode"
+title: "global-model (ODE)"
 author: "Florian Schneider"
-date: "Thursday, November 20, 2014"
+date: "Thursday, January 30, 2015"
 output: html_document
@@
 
-## original model (Kéfi et al 2007)
+## developping the null model
 
-Running the pair approximation model of Kéfi et al 2007 TPB dynamically.
+Non-spatial global model solved as ordinary differential equations. It includes global accelerating and decelerating feedbacks on mortality and grazing.
 
@@
 
-### plant and soil parameters
+### growth parameters
+
+The basic assumption is logistic plant growth, i.e. growth goes to zero as the carrying capacity is approached (decelerating effect).
+
+As an accelerating feedback we add water runoff, which reduces growth at low cover.
+
 ```{r, echo=FALSE}
@@
 inputPanel(
-  sliderInput("m", label = "intrinsic mortality:",
-              min = 0.0, max = 1, value = 0.05, step = 0.01),
   sliderInput("r", label = "max. growth rate:",
@@
               min = 0.0, max = 1, value = 0.9, step = 0.01),
+  sliderInput("alpha", label = "water runoff:",
+              min = 0.0, max = 1, value = 0, step = 0.01)
+)
+
+```
+
+### mortality parameters
+
+The basic assumption is a constant intrinsic mortality of plants and a type II functional response, i.e. grazing mortality saturates with cover (decelerating effect).
+
+As an additional accelerating effect we assume the search efficiency to be lowest at low cover and increase with cover, turning the function into a type III functional response.
+
+```{r, echo=FALSE}
+
+inputPanel(
+  sliderInput("m", label = "intrinsic mortality:",
+              min = 0.0, max = 1, value = 0.05, step = 0.01),
   sliderInput("a", label = "search efficiency:",
@@
   sliderInput("L", label = "livestock units:",
-              min = 0.0, max = 100, value = 10, step = 1)
+              min = 0.0, max = 100, value = 10, step = 1),
+  sliderInput("q", label = "hill coefficient:",
+              min = 0.0, max = 3, value = 1, step = 0.01)
 )
@@
   parms$L <- as.numeric(input$L) 
+  parms$alpha <- as.numeric(input$alpha) 
+  parms$q <- as.numeric(input$q)
 
-  plot(rho,  with(parms,  m * rho + (a * rho * L)/(1 + a * h * rho)), ylab = "growth", type = "l", ylim = c(0,0.5))
-  lines(rho, with(parms, r* b* rho * (1 - rho/K) ), col = "green3" )
+  plot(rho,  with(parms,  m * rho + (a * rho^( 1 + q) * L)/(1 + a * h * rho^( 1 + q))), ylab = "growth", type = "l", ylim = c(0,0.5))
+  lines(rho, with(parms, r* b* rho^( 1 + alpha) * (1 - rho/K) ), col = "green3" )
 
@@
 d_rho <- function(rho, parms, z = 4) { 
-  with(parms, r* b* rho * (1 - rho/K) - m * rho - (a * rho * L)/(1 + a * h * rho) )  
+  with(parms, r* b* rho^( 1 + alpha) * (1 - rho/K) - m * rho - (a * rho^( 1 + q) * L)/(1 + a * h * rho^( 1 + q)) )  
 }
@@
   parms$L <- as.numeric(input$L) 
+  parms$alpha <- as.numeric(input$alpha) 
+  parms$q <- as.numeric(input$q)
 
 # running the ode-solver
-runmodel <- ode(y = input$rho_1, func = odesys, times = 1.1^(1:53), parms = parms)
+runmodel <- ode(y = input$rho_1, func = odesys, times = 1.1^(1:50), parms = parms)
 

### 2015-01-30 documentation/model_formulation.Rmd
update model formulation. spatially-explicit model development

@@
 output:
+  pdf_document:
+    fig_height: 4
+    fig_width: 6
   html_document:
     keep_md: yes
-  pdf_document:
-    fig_height: 3.5
-    fig_width: 4
 fontsize: 12pt
@@
 
+**This model is going to be the starting point for the model development**
+
 #### mechanistic growth assumptions
@@
 
-```{r, echo=FALSE}
-G <- function(P, r = 2.25, K = 0.9) r*P*(1-P/K)
-
-defplot(rho, G(rho), ylab = "growth")
-
-```
-
-**This model is going to be the starting point for the model development**
-
 We assume that growth rate is linearly related to environmental quality, $b$, which is inverse to aridity.
@@
 
-$$ r = r_0 * b $$
+$$ r = r_0 b $$
 
@@
 
-The
+As an accelerating assumption on plant mortality we add the formulation of density dependent search efficiencies.
+
+$$ b = b_0 N^q $$
+
+Which turns the consumption term into
 
 $$ C(V) = \frac{bV^{1+q}}{1+bhV^{1+q}} $$
+
+Thus, plant mortality is accelerating with increasing cover, protecting plants at low cover, due to inefficient search behaviour, and then saturates at the inverse of handling time, $h$.
 
@@
 
-#### null model
+#### consumption null model
 
@@
 
-$$ C(V) = m*V + \frac{aVL}{1+ahV} \,,$$
+$$ C(V) = mV + \frac{aVL}{1+ahV} \,,$$
 
@@
 
-$$ \frac{\mathrm{d}\rho_1}{\mathrm{d}t} = r b V (1-\frac{V}{K}) - m V - \frac{aVL}{1+ahV} $$
+$$ \frac{\mathrm{d}V}{\mathrm{d}t} = r b V (1-\frac{V}{K}) - m V - \frac{aVL}{1+ahV} $$
 
@@
 plot(ODEnullmodel, type = "l", ylim = c(0,1))
+axis(4, at = round(tail(ODEnullmodel,1)[["1"]],2),cex.axis = 0.8, las = 1 )
+
+```
+
+#### global model
+
+The ODE that includes the additional global feed backs of water runoff and  reduced search efficience at low cover would read:
+
+$$ \frac{\mathrm{d}V}{\mathrm{d}t} = r b V^{1 + \alpha} (1-\frac{V}{K}) - m V - \frac{aV^{1+q}L}{1+ahV^{1+q}} $$
+
+We extend the parameter set by
+
+```{r, echo=FALSE}
+
+parms <- list(
+    m = 0.05,
+    r = 1,
+    b = 0.9, 
+    K = 0.9, 
+    a = 0.6,
+    h = 100,
+    L = 16, 
+    alpha = 0.1,
+    q = 1
+  )
+
+ as.data.frame(parms)
+```
+
+which gives us the prediction of complex bistability.
+
+```{r, echo=FALSE}
+plot(rho,  with(parms,  m * rho + (a * rho^( 1 + q) * L)/(1 + a * h * rho^( 1 + q))), ylab = "growth", type = "l", ylim = c(0,0.5))
+  lines(rho, with(parms, r* b* rho^( 1 + alpha) * (1 - rho/K) ), col = "green3" )
+
+d_rho <- function(rho, parms, z = 4) { 
+  with(parms, r* b* rho^( 1 + alpha) * (1 - rho/K) - m * rho - (a * rho^( 1 + q) * L)/(1 + a * h * rho^( 1 + q)) )  
+  }
+
+# running the ode-solver
+ODEglobal_high <- ode(y = 0.9, func = odesys, times = 1.1^(1:50), parms = parms)
+
+# transfer into ouput and calculate missing rho values
+ODEglobal_high <- as.data.frame(ODEglobal_high)
+
+plot(ODEglobal_high, type = "l", ylim = c(0,1))
+axis(4, at = round(tail(ODEglobal_high,1)[["1"]],2),cex.axis = 0.75, las = 1 )
+
+# running the ode-solver
+ODEglobal_low <- ode(y = 0.1, func = odesys, times = 1.1^(1:50), parms = parms)
+
+# transfer into ouput and calculate missing rho values
+ODEglobal_low <- as.data.frame(ODEglobal_low)
+axis(4, at = round(tail(ODEglobal_low,1)[["1"]],2),cex.axis = 0.75, las = 1 )
+
+lines(ODEglobal_low)
 
@@
 
-The differential equations are formulated as changes in vegetation cover. Growth are positive changes, and deaths are negative changes.
+It is straigthforward to implement the model in a cellular automata model.
+It is defined on a defined grid of cells where each cell can take the values '1' if it is vegetated and '0' if it is empty.
+The transition rules between those states depend on global vegetation cover, $\rho_1$.
 
-Now, the CA uses transition probabilities. That is, a) the individual risk of death of vegetated cells and b) the individual chance of becoming vegetated for an empty cell.
+While the differential equations are formulated as changes in vegetation cover(growth are positive changes, and deaths are negative changes), the CA uses transition probabilities between the two possible cell states.
+Thus, two transition probabilities need to be defined: the probability of **growth**, $w_{0,1}$, turning an empty cell into a vegetated, and the probability of **death**, $w_{0,1}$, removing vegetation from the cell.
+
 To translate into the CA context, we need to find rules that apply to a) the vegetated cells, and b) the empty cells.
@@
 ```{r, echo=FALSE}
+
+parms <- list(
+    m = 0.05,
+    r = 1,
+    b = 0.9, 
+    K = 0.9, 
+    a = 0.6,
+    h = 100,
+    L = 10, 
+    alpha = 0,
+    q = 0
+  )
+
 par(mar = c(4,4,2,4))
@@
 animateCA(nullmodel, "ca_nullmodel.gif")
+
+![](ca_nullmodel.gif)
+
 ```
-![](ca_nullmodel.gif)
+
+#### additional feedbacks
+
+The next step is to include the additional global feedback terms into the model. This is straightforward
+
+$$ w_{1,0} = m + \frac{aL\rho_1^{q}}{1+ah\rho_1^{1+q}} $$
+
+$$ w_{0,1} = \frac{r \rho_1^{1+\alpha} b (1-\frac{\rho_1}{K})}{1-\rho_1} $$
+
+```{r, echo=FALSE}
+
+parms <- list(
+    m = 0.05,
+    r = 1,
+    b = 0.9, 
+    K = 0.9, 
+    a = 0.6,
+    h = 100,
+    L = 16, 
+    alpha = 0.1,
+    q = 1
+  )
+
+set.seed(5535732) # setting seed for random number generation
+
+globalmodel <- runCA(0.9, parms, delta = 0.2, t_max = 150, t_min = 100, t_eval = 50, isstable = 0.0001, saveeach = 1)
+
+plot(globalmodel$rho_one ~ globalmodel$time, 
+     ylim = c(0,1), ylab = expression( rho["+"]), 
+     xlab = "time [yr]", type = "l")
+
+globalmodel <- runCA(0.1, parms, delta = 0.2, t_max = 150, t_min = 100, t_eval = 50, isstable = 0.0001, saveeach = 1)
+lines(globalmodel$rho_one ~ globalmodel$time)
+lines(ODEglobal_low, lty = 3)
+lines(ODEglobal_high, lty = 3)
+```
+
+\newpage
 
@@
 
-These mechanisms can be implemented in the model by using a pair-approximation approach. 
-In numerical simulations $G$ and $C$ can be implemented as the transition rates in a two-state spatial model.
-
-In the formalism of the pair-approximation model of @kefi07, growth, *i.e.* the total amount of empty cells transitioning into vegetated cells, becomes a complex function of the relative cover of the multiple pairs including a $+$.
-
-In a model with only global effects, the transition probability of indiviual plant death would be
-
-$$ w_{1,0} = C(V)/V = m + \frac{aL}{1+ahV} $$
-
-and the transition probability of individual plant growth
-
-$$ w_{0,1} = G(V)/V = r (1-\frac{V}{K}) \,.$$
+These mechanisms can be implemented in the model by using a pair-approximation approach and numerical simulations, where $G$ and $C$ can be implemented as the transition rates in a two-state spatial model depending on global and local vegetation cover.
 
@@
 This is assuming that the local environmental suitability is enhanced by a local facilitation term similar to the implementation in Kéfi et al 2007.
-Thus, the growth rate becomes a function of local vegetation cover $\nu$:
+Thus, the reduction in growth rate, $r$, by aridity, $a$, is compensated by a function of $q_{0|1}$, the local density of cells in state 1 (i.e.
+vegetation) given that the focal cell is in state 0 (i.e.
+empty), and maximizes to one if the cell has four neighbors:
 
-$$ r = r_0 + f \nu \,,$$
+$$ a = a_0 + (f - a_0) q_{0|1} \,,$$
 
-with an intrinsic growth rate $r_0$ which determines the probability of growth in absence of local vegetation and the additive facilitation $f$ which gradually determines the enhancement of the growth with an increasing local vegetation cover (maximizing at $\nu = 1$ if the cell has 4 neighbors).
+with the effect of aridity, $a_0$ in absence of local vegetation being increased to the value of $f$ if additive facilitation due to the presence of neighboring plants occurs.
+This term gradually determines the enhancement of the growth with an increasing local vegetation cover (maximizing at $q_{0|1} = 1$ if the cell has 4 neighbors).
 
@@
 
-$$ K = K_0 (1 - c \nu)$$
+$$ K = K_0 (1 - c q_{0|1})$$
+
+That is, if a cell has a fully vegetated local neighborhood, it's carrying capacity will be reduced by $c$. Thus, $c \geq K$ otherwise carrying capacity will fall below $0$ at high cover.
 
@@
 
-Mortality becomes enhanced in sparse environments by patches that are attracting grazers. We assume this attractivity, $v$, reduces search times locally.
+Mortality becomes enhanced in sparse environments by patches that are attracting grazers. We assume this attractivity enhances search efficiency locally by the value $v$
 
-$$ a = a_0 * (1 - v * \nu) $$
+$$ b = b_0  + v q_{1|1} $$
+
+Thus, at low local cover feeding will be high on plants with neighbors, whereas at high cover the term has no effect since handling time is limiting consumption.
+An enhanced search efficiency will raise the critical thresholds for a collapse.
 
@@
 They thereby provide protected habitat to their direct neighborhood, or share the investments in those traits with their neighbors.
-Overall grazing mortality thus is reduced through an increase in handling time by associational resistance, $p$
+Overall grazing mortality thus is reduced through an increase in handling time by associational resistance, $p$,
 
-$$ h = h_0 * p * \nu $$
+$$ h = h_0  / (1 - p  q_{1|1}) $$
 
-### implementation in a cellular automaton
-
-It is straigthforward to implement the model in a cellular automata model.
-It is defined on a defined grid of cells where each cell can take the values '1' if it is vegetated and '0' if it is empty.
-The transition rules between those states depend either on global vegetation cover $\rho_1$ or on the local cover in the direct neighborhood of the cell $\nu_1$ (closest 4 cells, von-Neumann-neighborhood).
-
-Thus, two transition probabilities need to be defined: the probability of **growth**, $w_{0,1}$, turning an empty cell into a vegetated, and the probability of **death**, $w_{0,1}$, removing vegetation from the cell.
-
-The null model, i.e. the model with only global effects on mortality and growth, with the basic assumption of a type II functional response and logistic growth, would thus be defined by
-
-$$ w_{0,1} = r (1-\frac{\rho_1}{K})$$
-
-$$ w_{1,0} =  m + \frac{aL}{1+ah\rho_1} $$
+An alternative formulation should be found. The term is highly non-linear and provokes relevant changes only at values above 0.9. It is defined within $(0,1($
 
@@
 
-The pair-approximation framework (Matsuda et al 1992) is defined as a system of ordinary differential equations, describing the populations of *pairs* of cells.
+The pair-approximation framework [@matsuda92] is defined as a system of ordinary differential equations, describing the populations of *pairs* of cells.
 In a two state system, three pairs can be defined with population densities $\rho_{1,1}$ , $\rho_{0,0}$, and $\rho_{0,1}$.
@@
 
-This means, the system can be fully described by two differential equations:
+Since formally $q_{i|j}$ is a conditional probability (given that one neighbor is in state $i$, what is the probability to find the other in state $j$) we can express it in terms of densities of pairs and ‘singletons,’ as
 
-$$ \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = $$
+$$ q_{i|j} = \frac{\rho_{ij}}{\rho_i} $$
 
-$$ \frac{\mathrm{d}\rho_0}{\mathrm{d}t} = $$
+The vegetation dynamics of the system can be fully described by two differential equations (omitting indexed zeros of equations for easier reading):
+
+$$ \frac{\mathrm{d}\rho_1}{\mathrm{d}t} = r ( a + (f - a) q_{0|1} ) \rho_1^{1 + \alpha} (1-\frac{\rho_1}{K (1 - c q_{0|1})}) - m \rho_1 - \frac{(b  + v q_{1|1} )\rho_1^{1+q}L}{1+( b  + v q_{1|1} ) ( h  / (p  q_{1|1}))\rho_1^{1+q}} $$
+
+$$ \frac{\mathrm{d}\rho_0}{\mathrm{d}t} =  m \rho_1 + \frac{(b  + v q_{1|1} )\rho_1^{1+q}L}{1+( b  + v q_{1|1} ) ( h  / (p  q_{1|1}))\rho_1^{1+q}} - r ( a + (f - a) q_{0|1} ) \rho_1^{1 + \alpha} (1-\frac{\rho_1}{K (1 - c q_{0|1})})$$
+
+Next steps:
+
+- formulate equations for $\frac{\mathrm{d}\rho_{0,1}}{\mathrm{d}t}$, $\frac{\mathrm{d}\rho_{0,0}}{\mathrm{d}t}$, and $\frac{\mathrm{d}\rho_{1,1}}{\mathrm{d}t}$
+
+### implementation in CA
+
+It is quite straightforward to implement the mechanisms described above in the syntax of a cellular automata model, where the probabilities of transition can not only depend on the global cover $\rho_1$ but also on the local cover in the direct neighborhood of the cell $\nu_1$ (closest 4 cells, von-Neumann-neighborhood).
+
+see implementation in NetLogo: `documentation/livestockmodel.nlogo`
 

### 2015-01-30 code/simfunctions.r
updated simCA function to include global terms (R)

@@ -308,9 +308,9 @@ runCA <- function(init, parms, width = 100, height = 100, delta = 0.1, t_max = 1
     rnum <- runif(width*height) # one random number between 0 and 1 for each cell
     
     # 3 - setting transition probabilities
-    growth <- with(parms_temp, r * b * rho_one * ( 1 - (rho_one / K)) / (1 - rho_one)  *delta)  # recolonisation rates of all cells 
+    growth <- with(parms_temp, (r * b * rho_one^(1 + alpha) * ( 1 - (rho_one / K)) / (1 - rho_one))  *delta)  # recolonisation rates of all cells 
     
-    death <- with(parms_temp, (m + ( a * L )/( 1 + a * h * rho_one )) *delta)   # set probability of death for each cell
+    death <- with(parms_temp, (m + ( a * L * rho_one^(q) )/( 1 + a * h * rho_one^(1+q) )) *delta)   # set probability of death for each cell
     
     # check for sum of probabilities to be inferior 1 and superior 0
     if(any(c(growth, death) > 1 )) warning(paste("a set probability is exceeding 1 in time step", i, "! decrease delta!!!")) 

### 2015-02-02 documentation/model_formulation.Rmd
develop pair-approximation ODE system, solve ODE

@@
 This is assuming that the local environmental suitability is enhanced by a local facilitation term similar to the implementation in Kéfi et al 2007.
-Thus, the reduction in growth rate, $r$, by aridity, $a$, is compensated by a function of $q_{0|1}$, the local density of cells in state 1 (i.e.
+Thus, the reduction in growth rate, $r$, by aridity, i.e.
+the inverse of environmental quality $b$, is compensated by a function of $q_{0|1}$, the local density of cells in state 1 (i.e.
 vegetation) given that the focal cell is in state 0 (i.e.
@@
 
-$$ a = a_0 + (f - a_0) q_{0|1} \,,$$
+$$ b = b^* + (f - b^*) q_{0|1} \,,$$
 
@@
 
-$$ K = K_0 (1 - c q_{0|1})$$
+$$ K = K^* (1 - c q_{0|1})$$
 
@@
 
-Mortality becomes enhanced in sparse environments by patches that are attracting grazers. We assume this attractivity enhances search efficiency locally by the value $v$
+Mortality becomes enhanced in sparse environments by patches that are attracting grazers. We assume this attractivity enhances search efficiency, $a$, locally by the value $v$
 
-$$ b = b_0  + v q_{1|1} $$
+$$ a = a^*  + v q_{1|1} $$
 
@@
 
-$$ h = h_0  / (1 - p  q_{1|1}) $$
+$$ h = h^*  / (1 - p  q_{1|1}) $$
 
 An alternative formulation should be found. The term is highly non-linear and provokes relevant changes only at values above 0.9. It is defined within $(0,1($
+
+### implementation in CA
+
+It is quite straightforward to implement the mechanisms described above in the syntax of a cellular automata model, where the probabilities of transition can not only depend on the global cover $\rho_1$ but also on the local cover in the direct neighborhood of the cell $\nu_1$ (closest 4 cells, von-Neumann-neighborhood).
+
+the transition probabilities therefore become
+
+$$ w_{1,0} = m + ( a^*  + v q_{1|1})L\rho_1^{q}{1+( a^*  + v q_{1|1})(h^*/ (1 - p  q_{1|1}))\rho_1^{1+q}} $$
+
+$$ w_{0,1} = \frac{r \rho_1^{1+\alpha} (b^* + (f - b^*) q_{0|1}) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})})}{1-\rho_1} $$
+
+see implementation in NetLogo: `documentation/livestockmodel.nlogo`
 
@@
 
+We can use the transition probabilities of cells  to describe the full system of differential equations.
+
+$$  \frac{\mathrm{d}\rho_{11}}{\mathrm{d}t} = 2 \rho_{10}w_{0,1} - 2 \rho_{11} w_{1,0}  $$
+
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = \rho_{0}w_{0,1} - \rho_{1}w_{1,0}   $$
+
+$$  \frac{\mathrm{d}\rho_{10}}{\mathrm{d}t} = 2 \rho_{00}w_{0,1} + 2 \rho_{11}w_{1,0} - \rho_{10}w_{0,1}    $$
+
+$$  \frac{\mathrm{d}\rho_{0}}{\mathrm{d}t} = \rho_{1}w_{1,0} - \rho_{0}w_{0,1} $$
+
+$$  \frac{\mathrm{d}\rho_{00}}{\mathrm{d}t} = 2 \rho_{10}w_{1,0} - 2 \rho_{00}w_{0,1}  $$
+
+Due to the conservation equations, two of these are sufficient to fully describe the dynamics of the system.
+
+We chose to go with the first two and substitute
+
+$$  \frac{\mathrm{d}\rho_11}{\mathrm{d}t} = 2 \rho_{10} \frac{r \rho_1^{1+\alpha} (b^* + (f - b^*) q_{0|1}) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a^*  + v q_{1|1})L\rho_1^{q}}{1+(a^*  + v q_{1|1})(h^*  / (1 - p  q_{1|1}))\rho_1^{1+q}}  $$
+
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = \rho_{0}\frac{r \rho_1^{1+\alpha} (b^* + (f - b^*) q_{0|1}) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})})}{1-\rho_1} - \rho_{1}m - \frac{(a^*  + v q_{1|1})L\rho_1^{1+q}}{1+(a^*  + v q_{1|1})(h^*  / (1 - p  q_{1|1}))\rho_1^{1+q}}   $$
+
+they can be simplified to
+
+$$  \frac{\mathrm{d}\rho_11}{\mathrm{d}t} = 2 \rho_{10} \frac{r \rho_1^{1+\alpha} (b^* + (f - b^*) q_{0|1}) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a^*  + v q_{1|1})L\rho_1^{q}}{1+(a^*  + v q_{1|1})(h^*  / (1 - p  q_{1|1}))\rho_1^{1+q}}  $$
+
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b^* + (f - b^*) q_{0|1}) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})}) - \rho_{1}m - \frac{(a^*  + v q_{1|1})L\rho_1^{1+q}}{1+(a^*  + v q_{1|1})(h^*  / (1 - p  q_{1|1}))\rho_1^{1+q}}   $$
+
+as a matter of fact the equation for $\frac{\mathrm{d}\rho_{1}}{\mathrm{d}t}$ will re-convert into the regular population equation.
+
 Since formally $q_{i|j}$ is a conditional probability (given that one neighbor is in state $i$, what is the probability to find the other in state $j$) we can express it in terms of densities of pairs and ‘singletons,’ as
@@
 
-The vegetation dynamics of the system can be fully described by two differential equations (omitting indexed zeros of equations for easier reading):
+thus,
 
-$$ \frac{\mathrm{d}\rho_1}{\mathrm{d}t} = r ( a + (f - a) q_{0|1} ) \rho_1^{1 + \alpha} (1-\frac{\rho_1}{K (1 - c q_{0|1})}) - m \rho_1 - \frac{(b  + v q_{1|1} )\rho_1^{1+q}L}{1+( b  + v q_{1|1} ) ( h  / (p  q_{1|1}))\rho_1^{1+q}} $$
+$$ q_{0|1} = \frac{\rho_{10}}{\rho_0} =  \frac{\rho_{1} - \rho_{11}}{1 - \rho_1} $$
 
-$$ \frac{\mathrm{d}\rho_0}{\mathrm{d}t} =  m \rho_1 + \frac{(b  + v q_{1|1} )\rho_1^{1+q}L}{1+( b  + v q_{1|1} ) ( h  / (p  q_{1|1}))\rho_1^{1+q}} - r ( a + (f - a) q_{0|1} ) \rho_1^{1 + \alpha} (1-\frac{\rho_1}{K (1 - c q_{0|1})})$$
+$$ q_{1|1} = \frac{\rho_{11}}{\rho_1} $$
 
-Next steps:
+and substituting yields
 
-- formulate equations for $\frac{\mathrm{d}\rho_{0,1}}{\mathrm{d}t}$, $\frac{\mathrm{d}\rho_{0,0}}{\mathrm{d}t}$, and $\frac{\mathrm{d}\rho_{1,1}}{\mathrm{d}t}$
+$$  \frac{\mathrm{d}\rho_{11} }{\mathrm{d}t} = 2 (\rho_1 - \rho_{11}) \frac{r \rho_1^{1+\alpha} (b^* + (f - b^*) \frac{\rho_{1} - \rho_{11}}{1 - \rho_1}) (1-\frac{\rho_1}{K^* (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a^*  + v \frac{\rho_{11}}{\rho_1})L\rho_1^{q}}{1+(a^*  + v \frac{\rho_{11}}{\rho_1})(h^*  / (1 - p  \frac{\rho_{11}}{\rho_1}))\rho_1^{1+q}}  $$
 
-### implementation in CA
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b^* + (f - b^*) \frac{\rho_{1} - \rho_{11}}{1 - \rho_1}) (1-\frac{\rho_1}{K^* (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})}) - \rho_{1}m -  \frac{(a^*  + v \frac{\rho_{11}}{\rho_1})L\rho_1^{1+q}}{1+(a^*  + v \frac{\rho_{11}}{\rho_1})(h^*  / (1 - p  \frac{\rho_{11}}{\rho_1}))\rho_1^{1+q}}   $$
 
-It is quite straightforward to implement the mechanisms described above in the syntax of a cellular automata model, where the probabilities of transition can not only depend on the global cover $\rho_1$ but also on the local cover in the direct neighborhood of the cell $\nu_1$ (closest 4 cells, von-Neumann-neighborhood).
+Thus, we can solve the ODE for a given set of parameters
 
-see implementation in NetLogo: `documentation/livestockmodel.nlogo`
+```{r}
+
+parms <- list(
+    m = 0.05,  # intrinsic mortality of plants 
+    r = 1,   # growth rate
+    f = 0.7, #  local facilitation
+    b = 0.3,  # environmental quality inverse of aridity
+    K = 0.9,  # carrying capacity
+    alpha = 0.1, # water runoff
+    c = 0.2, # local competition
+    a = 0.6, # search efficiency
+    v = 0.1, # attraction effect
+    h = 50, # handling time (one individual on landscape unit)
+    p = 0.2, # protection effect
+    L = 4, # livestock units per landscape unit 
+    q = 1 # search ineffiency at low cover
+  )
+
+```
+
+we get the following dynamics
+
+```{r, echo=FALSE}
+
+rho <- c(rho_1 = 0.9, rho_11 = 0.8, rho_10 = 0.1, rho_00 = 0, rho_0 = 0.1)
+
+d_rho_1 <- function(rho, parms) { 
+  with(parms,
+       r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) - m * rho[1] - ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q))
+       )  
+  }
+
+d_rho_11 <- function(rho,  parms) { 
+  with(parms,
+       2* (rho[1] - rho[2]) * r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) / (1-rho[1]) - 2 * rho[2] * m  - 2 * rho[2] * ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q))
+       )  
+  }
+
+odesys_spex <- function(t, rho, parms = model_parms) {
+  rho_1 <- d_rho_1(rho, parms)
+  rho_11 <- d_rho_11(rho, parms)
+  list(c(rho_1, rho_11, rho_1-rho_11, 1-rho_1-(rho_1-rho_11),1-rho_1  
+               ) )
+}
+
+# running the ode-solver
+ODE_spex <- ode(y = rho, func = odesys_spex, times = 1.1^(1:50), parms = parms)
+
+# transfer into ouput and calculate missing rho values
+ODE_spex <- as.data.frame(ODE_spex )
+
+plot(rho_1 ~ time , data = ODE_spex , type = "l", ylim = c(0,1))
+axis(4, at = round(tail(ODE_spex,1)$rho_1,2),cex.axis = 0.7, las = 1 )
+
+```
+
+Now, we can also plot the clustering coefficient of this landscape, being
+
+$$ c_11 = q_{1|1}/\rho_1 = \frac{\rho_11}{\rho_1^2}$$
+
+```{r, echo=FALSE}
+plot(I(rho_11/rho_1^2) ~ time , data = ODE_spex , type = "l", ylim = c(0,2))
+
+```
 

### 2015-02-02 documentation/interactive_spexmodel.Rmd
interactive implementation of pair-approximation ODE model

---
title: "global-model (ODE)"
author: "Florian Schneider"
date: "Thursday, January 30, 2015"
output: html_document
runtime: shiny
---

## developping the null model

Non-spatial global model solved as ordinary differential equations. It includes global accelerating and decelerating feedbacks on mortality and grazing.

```{r, echo=FALSE}

library("deSolve")

  ini_rho <- function(rho_1) {
    c(
    rho_1 = rho_1,
    rho_11 = rho_1*rho_1,
    rho_10 = rho_1*(1-rho_1)*2,
    rho_00 = (1-rho_1)*(1-rho_1),
    rho_0 = 1-rho_1
    )

  }

d_rho_1 <- function(rho, parms) { 
  with(parms,
       r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) - m * rho[1] - ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q)) 
       )  
  }

d_rho_11 <- function(rho,  parms) { 
  with(parms,
       2* (rho[1] - rho[2]) * r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) / (1-rho[1]) - 2 * rho[2] * m  - 2 * rho[2] * ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q)) 
       )  
  }

odesys_spex <- function(t, rho, parms = model_parms) {
  rho_1 <- d_rho_1(rho, parms)
  rho_11 <- d_rho_11(rho, parms)
  list(c(rho_1, rho_11, rho_1-rho_11, 1-rho_1-(rho_1-rho_11),1-rho_1  
               ) )
}

```

### growth parameters

The basic assumption is logistic plant growth, i.e. growth goes to zero as the carrying capacity is approached (decelerating effect).

As an accelerating feedback we add water runoff, which reduces growth at low cover.

```{r, echo=FALSE}

inputPanel(
  sliderInput("r", label = "max. growth rate:",
              min = 0.0, max = 3, value = 1, step = 0.01),
  sliderInput("b", label = "environmental quality:",
              min = 0.0, max = 1, value = 0.1, step = 0.01),
  sliderInput("K", label = "carrying capacity:",
              min = 0.0, max = 1, value = 0.9, step = 0.01),
  sliderInput("alpha", label = "water runoff:",
              min = 0.0, max = 1, value = 0, step = 0.01),
  sliderInput("f", label = "local facilitation:",
              min = 0.0, max = 1, value = 0.9, step = 0.01),
  sliderInput("c", label = "local competition:",
              min = 0.0, max = 1, value = 0.1, step = 0.01)
)

```

### mortality parameters

The basic assumption is a constant intrinsic mortality of plants and a type II functional response, i.e. grazing mortality saturates with cover (decelerating effect).

As an additional accelerating effect we assume the search efficiency to be lowest at low cover and increase with cover, turning the function into a type III functional response.

```{r, echo=FALSE}

inputPanel(
  sliderInput("m", label = "intrinsic mortality:",
              min = 0.0, max = 1, value = 0.05, step = 0.01),
  sliderInput("a", label = "search efficiency:",
              min = 0.0, max = 1, value = 0.6, step = 0.01), 
  sliderInput("h", label = "handling time:",
              min = 0.0, max = 200, value = 50, step = 1), 
  sliderInput("L", label = "livestock units:",
              min = 0.0, max = 100, value = 6, step = 1),
  sliderInput("q", label = "hill coefficient:",
              min = 0.0, max = 1, value = 0, step = 0.01),
  sliderInput("v", label = "local attractivity:",
              min = 0.0, max = 1, value = 0.0, step = 0.01),
  sliderInput("p", label = "local protection:",
              min = 0.0, max = 1, value = 0.8, step = 0.01)
)

```

### initial conditions
```{r, echo=FALSE}

inputPanel(
  sliderInput("rho_1", label = "initial plant cover:",
              min = 0.0, max = 1, value = 0.9, step = 0.01)
  )

renderPlot({

 parms <- list(
  m = as.numeric(input$m),
  r = as.numeric(input$r),
  b = as.numeric(input$b),
  K = as.numeric(input$K),
  a = as.numeric(input$a), 
  h = as.numeric(input$h),
  L = as.numeric(input$L),
  alpha = as.numeric(input$alpha), 
  q = as.numeric(input$q),
  f = as.numeric(input$f),
  c = as.numeric(input$c),
  v = as.numeric(input$v),
  p = as.numeric(input$p) 
  )

# running the ode-solver
ODE_spex <- ode(y = ini_rho(input$rho_1), func = odesys_spex, times = 1.1^(1:50), parms = parms)

# transfer into ouput and calculate missing rho values
ODE_spex <- as.data.frame(ODE_spex )

par(mar = c(4,4,1,3)+0.1)
plot(rho_1 ~ time , data = ODE_spex , type = "l", ylim = c(0,1))
axis(4, at = round(tail(ODE_spex,1)$rho_1,2),cex.axis = 0.7, las = 1 )

} )
```

### 2015-02-02 documentation/interactive_spexmodel_bifurcations.Rmd
dynamic bifurcation diagram of full spatially-explicit model (Shiny)

--- 2015-02-02 documentation/interactive_spexmodel.Rmd
@@
 ```{r, echo=FALSE}
+
+# provides parallel backend
+library(foreach)
+library(doSNOW)
+
+workstation <-  list(host = "162.38.184.118", user = "schneider",
+                     rscript = "/usr/lib/R/bin/Rscript",
+                     snowlib = "/usr/lib/R/library")
+
+workerlist <- c(rep("localhost", times = 10))
+
+cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')
+
+registerDoSNOW(cl)
 
@@
 
-inputPanel(
-  sliderInput("rho_1", label = "initial plant cover:",
-              min = 0.0, max = 1, value = 0.9, step = 0.01)
-  )
-
 renderPlot({
@@
 
+parms$b <- seq(0,1,.005) #seq(0,1,.002)+0.001
+parms$rho_ini <- c(0.9, 0.01)
+
+iterations <- expand.grid(parms)
+iterations <- cbind(ID = 1:dim(iterations)[1],iterations)
+
+iterations$f[iterations$f < iterations$b] <- iterations$b[iterations$f < iterations$b]
+
+foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
+  source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")
+
+  model_parms <- as.list(iterations[iteration,])
+
   # running the ode-solver
-ODE_spex <- ode(y = ini_rho(input$rho_1), func = odesys_spex, times = 1.1^(1:50), parms = parms)
+  runmodel <- ode(y = ini_rho(model_parms$rho_ini), func = odesys_spex, times = c(1,1000), parms = model_parms)
 
   # transfer into ouput and calculate missing rho values
-ODE_spex <- as.data.frame(ODE_spex )
+  return(as.data.frame(runmodel)[dim(runmodel)[1],])
+  } -> output
 
-par(mar = c(4,4,1,3)+0.1)
-plot(rho_1 ~ time , data = ODE_spex , type = "l", ylim = c(0,1))
-axis(4, at = round(tail(ODE_spex,1)$rho_1,2),cex.axis = 0.7, las = 1 )
+output <- cbind(output, iterations)
+
+plot(rho_1 ~ b, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
 

### 2015-02-02 documentation/model_formulation.Rmd
including bifurcation diagrams

@@
     r = 1,   # growth rate
-    f = 0.7, #  local facilitation
-    b = 0.3,  # environmental quality inverse of aridity
+    f = 0.8, #  local facilitation
+    b = 0.4,  # environmental quality inverse of aridity
     K = 0.9,  # carrying capacity
@@
     h = 50, # handling time (one individual on landscape unit)
-    p = 0.2, # protection effect
-    L = 4, # livestock units per landscape unit 
+    p = 0.8, # protection effect
+    L = 6, # livestock units per landscape unit 
     q = 1 # search ineffiency at low cover
@@
 
+### Bifurcation diagrams
+
+```{r, echo=FALSE}
+
+parms$b <- seq(0,1,.005) #seq(0,1,.002)+0.001
+parms$rho_ini <- c(0.9, 0.01)
+
+iterations <- expand.grid(parms)
+iterations <- cbind(ID = 1:dim(iterations)[1],iterations)
+
+iterations$f[iterations$f < iterations$b] <- iterations$b[iterations$f < iterations$b]
+
+# provides parallel backend
+library(foreach)
+library(doSNOW)
+
+workstation <-  list(host = "162.38.184.118", user = "schneider",
+                     rscript = "/usr/lib/R/bin/Rscript",
+                     snowlib = "/usr/lib/R/library")
+
+workerlist <- c(rep("localhost", times = 10))
+
+cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')
+
+registerDoSNOW(cl)
+
+foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
+  source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")
+
+  model_parms <- as.list(iterations[iteration,])
+
+  # running the ode-solver
+  runmodel <- ode(y = ini_rho(model_parms$rho_ini), func = odesys_spex, times = c(1,1000), parms = model_parms)
+
+  # transfer into ouput and calculate missing rho values
+  return(as.data.frame(runmodel)[dim(runmodel)[1],])
+  } -> output
+
+output <- cbind(output, iterations)
+
+plot(rho_1 ~ b, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
+
+```
+
+```{r, echo=FALSE}
+
+parms$L <- seq(0,20,0.2) #seq(0,1,.002)+0.001
+parms$rho_ini <- c(0.9, 0.01)
+
+iterations <- expand.grid(parms)
+iterations <- cbind(ID = 1:dim(iterations)[1],iterations)
+
+iterations$f[iterations$f < iterations$b] <- iterations$b[iterations$f < iterations$b]
+
+foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
+  source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")
+
+  model_parms <- as.list(iterations[iteration,])
+
+  # running the ode-solver
+  runmodel <- ode(y = ini_rho(model_parms$rho_ini), func = odesys_spex, times = c(1,1000), parms = model_parms)
+
+  # transfer into ouput and calculate missing rho values
+  return(as.data.frame(runmodel)[dim(runmodel)[1],])
+  } -> output
+
+output <- cbind(output, iterations)
+
+plot(rho_1 ~ L, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
+
+stopCluster(cl)
+
+```
+
 # References

### 2015-02-02 code/simfunctions.r
update simfunctions.r ; include ODE equations

@@ -354,3 +354,35 @@ runCA <- function(init, parms, width = 100, height = 100, delta = 0.1, t_max = 1
 }
 
 
+ini_rho <- function(rho_1) {
+  c(
+    rho_1 = rho_1,
+    rho_11 = rho_1*rho_1,
+    rho_10 = rho_1*(1-rho_1)*2,
+    rho_00 = (1-rho_1)*(1-rho_1),
+    rho_0 = 1-rho_1
+  )
+  
+}
+
+d_rho_1 <- function(rho, parms) { 
+  with(parms,
+       r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) - m * rho[1] - ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q)) 
+  )  
+}
+
+
+d_rho_11 <- function(rho,  parms) { 
+  with(parms,
+       2* (rho[1] - rho[2]) * r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) / (1-rho[1]) - 2 * rho[2] * m  - 2 * rho[2] * ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q)) 
+  )  
+}
+
+
+odesys_spex <- function(t, rho, parms = model_parms) {
+  rho_1 <- d_rho_1(rho, parms)
+  rho_11 <- d_rho_11(rho, parms)
+  list(c(rho_1, rho_11, rho_1-rho_11, 1-rho_1-(rho_1-rho_11),1-rho_1  
+  ) )
+}
+

### 2015-02-03 documentation/interactive_spexmodel_bifurcations.Rmd
fix model definitions, local feedbacks

@@
   with(parms,
-       r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) - m * rho[1] - ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q))
+       r * (b + (1 - b) * f * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) - m * rho[1] - ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L * (1 - p * rho[2]/rho[1]))/(1 +(a + v*rho[2]/rho[1]) * h  * rho[1]^( 1 + q))
   )  
@@
   with(parms,
-       2* (rho[1] - rho[2]) * r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) / (1-rho[1]) - 2 * rho[2] * m  - 2 * rho[2] * ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q))
+       2* (rho[1] - rho[2]) * r * (b + (1 - b) * f * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) / (1-rho[1]) - 2 * rho[2] * m  - 2 * rho[2] * ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L * (1 - p * rho[2]/rho[1]))/(1 +(a + v*rho[2]/rho[1]) * h  * rho[1]^( 1 + q))
   )  
@@
 
+inputPanel(
+  selectInput("gradient", label = "select gradient:",             choices = c("b", "L"), selected = "b")
+  )
+
 renderPlot({
@@
 
-parms$b <- seq(0,1,.005) #seq(0,1,.002)+0.001
+ if(input$gradient == "b") parms$b <- seq(0,1,.005) 
+ if(input$gradient == "L") parms$L <- seq(0,20,.1)
+
 parms$rho_ini <- c(0.9, 0.01)
@@
 
-plot(rho_1 ~ b, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
+ if(input$gradient == "b") plot(rho_1 ~ b, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
+ if(input$gradient == "L") plot(rho_1 ~ L, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
 

### 2015-02-03 documentation/model_formulation.Rmd
bugfix bifurcation over L
fix model definitions, local feedbacks

@@
 
-$$ \frac{dV}{dt} = G(V) - C(V)g $$
+$$ \frac{dV}{dt} = G(V) - C(V)L $$
 
-with growth function $G$ and a function $C$ which represents consumption by a herbivore multiplied by the density of herbivores, the livestock rate, $g$.
+with growth function $G$ and a function $C$ which represents consumption by a herbivore multiplied by the density of herbivores, the livestock rate, $L$.
 Formally, intrinsic mortality could be included in both terms.
@@
             bty = "l", cex = 0.7, las = 1,
-            xlim = c(0,1), ylim = c(0,1), 
+            xlim = c(0,1), ylim = c(0,0.4), 
             xaxs = "i", yaxs = "i", 
@@
 
-G <- function(P, r = 2, K = 1) r*P*(1-P/K)
+G <- function(P, r = 1, K = 1) r*P*(1-P/K)
 
@@
 
-r <- function(P, r_0 =  2.25, a = 1) r_0 *P^a
+r <- function(P, r_0 =  1, a = 1) r_0 *P^a
 
@@
 
-G_sig <- function(P, r_0 = 2.25, a = 1,  K = 0.9) r(P, r_0, a)*P*(1-P/K)
+G_sig <- function(P, r_0 = 1, a = 1,  K = 0.9) r(P, r_0, a)*P*(1-P/K)
 
@@
 lines(rho, G_sig(rho, a = 0))
-lines(rho, G_sig(rho,r_0 = 2.25, a = 1))
-lines(rho, G_sig(rho,r_0 = 2.25, a = 0.2))
+lines(rho, G_sig(rho, a = 1))
+lines(rho, G_sig(rho, a = 0.2))
 ```
@@
 ```{r, echo=FALSE}
-C <- function(V, L = 10, g = 0.01) (g*V*L)
+C <- function(V, L = 5, g = 0.01) (g*V*L)
 
@@
 ```{r, echo=FALSE}
-C <- function(V, L = 10, g = 0.01) g * L + V*0
+C <- function(V, L = 5, g = 0.01) g * L + V*0
 
@@
 ```{r, echo=FALSE}
-C <- function(V, L = 10, a = 0.6, h = 100) (a*V*L)/(1+a*h*V)
+C <- function(V, L = 5, a = 0.6, h = 100) (a*V*L)/(1+a*h*V)
 
@@
 ```{r, echo=FALSE}
-C <- function(V, L = 10, a = 0.6, h = 100, q = 1) (a*V^(1+q)*L)/(1+a*h*(V^(1+q)) )
+C <- function(V, L = 5, a = 0.6, h = 100, q = 1) (a*V^(1+q)*L)/(1+a*h*(V^(1+q)) )
 
@@
 
-C <- function(V, m_0 = 0.05, L = 10, a = 0.6, h = 100, q = 0) (m_0 * V) +(a*V^(1+q)*L)/(1+a*h*(V^(1+q)) )
+C <- function(V, m_0 = 0.05, L = 5, a = 0.6, h = 100, q = 0) (m_0 * V) +(a*V^(1+q)*L)/(1+a*h*(V^(1+q)) )
 
@@
 abline( h = 0.25 )
-abline( h = 0.5 )
 ```
@@
 defplot(rho, G(rho), ylab = "growth", col = "green3")
+lines(rho, C(rho, L = 5))
 lines(rho, C(rho, L = 10))
 lines(rho, C(rho, L = 25))
-lines(rho, C(rho, L = 50))
 
@@
 defplot(rho, G(rho), ylab = "growth", col = "green3")
+lines(rho, C(rho, L = 5, q = 1))
 lines(rho, C(rho, L = 10, q = 1))
 lines(rho, C(rho, L = 25, q = 1))
-lines(rho, C(rho, L = 50, q = 1))
 
@@
     h = 100,
-    L = 10
+    L = 5
   )
@@
     h = 100,
-    L = 10, 
+    L = 5, 
     alpha = 0,
-    q = 0
+    q = 0,
+    c = 0,
+    f = 0,
+    v = 0,
+    p = 0
   )
@@
     alpha = 0.1,
-    q = 1
+    q = 1,
+    c = 0,
+    f = 0,
+    v = 0,
+    p = 0
+
   )
@@
 
-$$ b = b^* + (f - b^*) q_{0|1} \,,$$
+$$ b = b^* + (1 - b^*)  f q_{0|1} \,,$$
 
@@
 This term gradually determines the enhancement of the growth with an increasing local vegetation cover (maximizing at $q_{0|1} = 1$ if the cell has 4 neighbors).
+
+```{r, echo=FALSE}
+
+G <- function(V, q_01 = "auto", r_0 = 1, b = 0.1, f = 0, alpha = 0,  K = 0.9, c = 0) {
+  if(q_01 == "auto") q_01 <- 1
+  r_0*(b + (1-b)*f*q_01)*V^(1+alpha)*(1-V/(K * (1-c * q_01) ) )
+
+}
+
+defplot(rho, G(rho, 1, f = 0.8), ylab = "growth", col = "green3")
+lines(rho,G(rho, 0.75, f = 0.8), col = "green3")
+lines(rho,G(rho, 0.5, f = 0.8), col = "green3")
+lines(rho,G(rho, 0.25, f = 0.8), col = "green3")
+lines(rho,G(rho, 0, f = 0.8), col = "green3")
+
+lines(rho, C(rho, L = 10))
+
+```
 
@@
 That is, if a cell has a fully vegetated local neighborhood, it's carrying capacity will be reduced by $c$. Thus, $c \geq K$ otherwise carrying capacity will fall below $0$ at high cover.
+
+```{r, echo=FALSE}
+
+defplot(rho, G(rho, 1, c = 0.3, b = 0.9), ylab = "growth", col = "green3")
+lines(rho,G(rho, 0.75, c = 0.3, b = 0.9), col = "green3")
+lines(rho,G(rho, 0.5, c = 0.3, b = 0.9), col = "green3")
+lines(rho,G(rho, 0.25, c = 0.3, b = 0.9), col = "green3")
+lines(rho,G(rho, 0, c = 0.3, b = 0.9), col = "green3")
+
+lines(rho, C(rho, L = 10))
+
+```
 
@@
 
+```{r, echo=FALSE}
+
+C <- function(V, q_01 = 1, m_0 = 0.05, L = 5, a = 0.1, h = 100, q = 0, p = 0, v = 0) (m_0 * V) +( (a+v*q_01  ) *V^(1+q)*L*(1-p*q_01))/(1+(a+v*q_01)*h*(V^(1+q)) )
+
+defplot(rho, C(rho, 1, v = 0.5), ylab = "mortality")
+lines(rho,C(rho, 0.75, v = 0.5))
+lines(rho,C(rho, 0.5, v = 0.5))
+lines(rho,C(rho, 0.25, v = 0.5))
+lines(rho,C(rho, 0, v = 0.5))
+
+lines(rho, G(rho, b = 0.6), col = "green3")
+
+```
+
 #### decelerating mortality
@@
 An alternative formulation should be found. The term is highly non-linear and provokes relevant changes only at values above 0.9. It is defined within $(0,1($ 
+Instead, we could define associational protection as a reduced impact due to a reduced local density of grazers.
+
+$$ L = L^* (1 - p  q_{1|1}) $$
+
+```{r, echo=FALSE}
+
+defplot(rho, C(rho, 1, p = 0.8, L = 10), ylab = "mortality")
+lines(rho,C(rho, 0.75, p = 0.8, L = 10))
+lines(rho,C(rho, 0.5, p = 0.8, L = 10))
+lines(rho,C(rho, 0.25, p = 0.8, L = 10))
+lines(rho,C(rho, 0, p = 0.8, L = 10))
+
+lines(rho, G(rho, b = 0.6), col = "green3")
+
+```
+
+#### multiple effects
+
+If those effects coincide, potentially even with the global feedbacks, the intersections are difficult to estimate.
+
+```{r, echo=FALSE}
+
+defplot(rho, C(rho, 1, p = 0.8, v = 0.6, L = 10, q = .5), ylab = "mortality")
+lines(rho,C(rho, 0.75, p = 0.8, v = 0.6, L = 10, q = .5))
+lines(rho,C(rho, 0.5, p = 0.8,v = 0.6,  L = 10, q = .5))
+lines(rho,C(rho, 0.25, p = 0.8,v = 0.6,  L = 10, q = .5))
+lines(rho,C(rho, 0, p = 0.8,v = 0.6,  L = 10, q = .5))
+
+lines(rho, G(rho, 1, c = 0.3, f = 0.9, b = 0.3, alpha = 0.5), col = "green3")
+lines(rho,G(rho, 0.75, c = 0.3,f = 0.9,  b = 0.3, alpha = 0.5), col = "green3")
+lines(rho,G(rho, 0.5, c = 0.3,f = 0.9,  b = 0.3, alpha = 0.5), col = "green3")
+lines(rho,G(rho, 0.25, c = 0.3, f = 0.9, b = 0.3, alpha = 0.5), col = "green3")
+lines(rho,G(rho, 0, c = 0.3,f = 0.9,  b = 0.3, alpha = 0.5), col = "green3")
+
+```
 
@@
 
-$$ w_{1,0} = m + ( a^*  + v q_{1|1})L\rho_1^{q}{1+( a^*  + v q_{1|1})(h^*/ (1 - p  q_{1|1}))\rho_1^{1+q}} $$
+$$ w_{1,0} = m + ( a^* + v q_{1|1})L^* (1 - p  q_{1|1})\rho_1^{q}{1+( a^*  + v q_{1|1})h\rho_1^{1+q}} $$
 
-$$ w_{0,1} = \frac{r \rho_1^{1+\alpha} (b^* + (f - b^*) q_{0|1}) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})})}{1-\rho_1} $$
+$$ w_{0,1} = \frac{r \rho_1^{1+\alpha} (b^* + (1 - b^*) f q_{0|1}) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})})}{1-\rho_1} $$
 
 see implementation in NetLogo: `documentation/livestockmodel.nlogo`
+
+Thus, we can run the CA for a given set of parameters
+
+```{r}
+
+parms <- list(
+    m = 0.05,  # intrinsic mortality of plants 
+    r = 1,   # growth rate
+    f = 0.8, #  local facilitation
+    b = 0.4,  # environmental quality inverse of aridity
+    K = 0.9,  # carrying capacity
+    alpha = 0.1, # water runoff
+    c = 0.2, # local competition
+    a = 0.6, # search efficiency
+    v = 0.1, # attraction effect
+    h = 50, # handling time (one individual on landscape unit)
+    p = 0.8, # protection effect
+    L = 6, # livestock units per landscape unit 
+    q = 1 # search ineffiency at low cover
+  )
+
+```
+
+```{r, echo = FALSE, warning = FALSE}
+
+set.seed(565432) # setting seed for random number generation
+
+spexmodel <- runCA(0.9, parms, delta = 0.2, t_max = 150, t_min = 100, t_eval = 50, isstable = 0.0001, saveeach = 1)
+
+plot(spexmodel$rho_one ~ spexmodel$time, 
+     ylim = c(0,1), ylab = expression( rho["+"]), 
+     xlab = "time [yr]", type = "l")
+
+```
+
+And the animated lattice would look like this:
+```{r, eval = FALSE, echo = FALSE, warning = FALSE}
+animateCA(nullmodel, "ca_nullmodel.gif")
+
+![](ca_nullmodel.gif)
+
+```
 
@@
 
-$$  \frac{\mathrm{d}\rho_11}{\mathrm{d}t} = 2 \rho_{10} \frac{r \rho_1^{1+\alpha} (b^* + (f - b^*) q_{0|1}) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a^*  + v q_{1|1})L\rho_1^{q}}{1+(a^*  + v q_{1|1})(h^*  / (1 - p  q_{1|1}))\rho_1^{1+q}}  $$
+$$  \frac{\mathrm{d}\rho_{11}}{\mathrm{d}t} = 2 \rho_{10} \frac{r \rho_1^{1+\alpha} (b^* + (1 - b) f q_{0|1})) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a^*  + v q_{1|1})L^* (1 - p  q_{1|1})\rho_1^{q}}{1+(a^*  + v q_{1|1})h\rho_1^{1+q}}  $$
 
-$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = \rho_{0}\frac{r \rho_1^{1+\alpha} (b^* + (f - b^*) q_{0|1}) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})})}{1-\rho_1} - \rho_{1}m - \frac{(a^*  + v q_{1|1})L\rho_1^{1+q}}{1+(a^*  + v q_{1|1})(h^*  / (1 - p  q_{1|1}))\rho_1^{1+q}}   $$
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = \rho_{0}\frac{r \rho_1^{1+\alpha} (b^* + (1 - b)  f q_{0|1})) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})})}{1-\rho_1} - \rho_{1}m - \frac{(a^*  + v q_{1|1})L^* (1 - p  q_{1|1})\rho_1^{1+q}}{1+(a^*  + v q_{1|1})h\rho_1^{1+q}}   $$
 
@@
 
-$$  \frac{\mathrm{d}\rho_11}{\mathrm{d}t} = 2 \rho_{10} \frac{r \rho_1^{1+\alpha} (b^* + (f - b^*) q_{0|1}) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a^*  + v q_{1|1})L\rho_1^{q}}{1+(a^*  + v q_{1|1})(h^*  / (1 - p  q_{1|1}))\rho_1^{1+q}}  $$
+$$  \frac{\mathrm{d}\rho_{11}}{\mathrm{d}t} = 2 \rho_{10} \frac{r \rho_1^{1+\alpha} (b^* + (1 - b)  f q_{0|1})) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a^*  + v q_{1|1})L^* (1 - p  q_{1|1}) \rho_1^{q}}{1+(a^*  + v q_{1|1})h)\rho_1^{1+q}}  $$
 
-$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b^* + (f - b^*) q_{0|1}) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})}) - \rho_{1}m - \frac{(a^*  + v q_{1|1})L\rho_1^{1+q}}{1+(a^*  + v q_{1|1})(h^*  / (1 - p  q_{1|1}))\rho_1^{1+q}}   $$
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b^* + (1 - b)  f q_{0|1})) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})}) - \rho_{1}m - \frac{(a^*  + v q_{1|1})L^* (1 - p  q_{1|1})\rho_1^{1+q}}{1+(a^*  + v q_{1|1})h\rho_1^{1+q}}   $$
 
@@
 
-$$  \frac{\mathrm{d}\rho_{11} }{\mathrm{d}t} = 2 (\rho_1 - \rho_{11}) \frac{r \rho_1^{1+\alpha} (b^* + (f - b^*) \frac{\rho_{1} - \rho_{11}}{1 - \rho_1}) (1-\frac{\rho_1}{K^* (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a^*  + v \frac{\rho_{11}}{\rho_1})L\rho_1^{q}}{1+(a^*  + v \frac{\rho_{11}}{\rho_1})(h^*  / (1 - p  \frac{\rho_{11}}{\rho_1}))\rho_1^{1+q}}  $$
+$$  \frac{\mathrm{d}\rho_{11} }{\mathrm{d}t} = 2 (\rho_1 - \rho_{11}) \frac{r \rho_1^{1+\alpha} (b^* + (1 - b)  f \frac{\rho_{1} - \rho_{11}}{1 - \rho_1} )) (1-\frac{\rho_1}{K^* (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a^*  + v \frac{\rho_{11}}{\rho_1})L^* (1 - p \frac{\rho_{11}}{\rho_1})\rho_1^{q}}{1+(a^*  + v \frac{\rho_{11}}{\rho_1})h\rho_1^{1+q}}  $$
 
-$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b^* + (f - b^*) \frac{\rho_{1} - \rho_{11}}{1 - \rho_1}) (1-\frac{\rho_1}{K^* (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})}) - \rho_{1}m -  \frac{(a^*  + v \frac{\rho_{11}}{\rho_1})L\rho_1^{1+q}}{1+(a^*  + v \frac{\rho_{11}}{\rho_1})(h^*  / (1 - p  \frac{\rho_{11}}{\rho_1}))\rho_1^{1+q}}   $$
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b^* + (1 - b) f  \frac{\rho_{1} - \rho_{11}}{1 - \rho_1}) (1-\frac{\rho_1}{K^* (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})}) - \rho_{1}m -  \frac{(a^*  + v \frac{\rho_{11}}{\rho_1})L^* (1 - p \frac{\rho_{11}}{\rho_1})\rho_1^{1+q}}{1+(a^*  + v \frac{\rho_{11}}{\rho_1})h\rho_1^{1+q}}   $$
 
-Thus, we can solve the ODE for a given set of parameters
-
-```{r}
-
-parms <- list(
-    m = 0.05,  # intrinsic mortality of plants 
-    r = 1,   # growth rate
-    f = 0.8, #  local facilitation
-    b = 0.4,  # environmental quality inverse of aridity
-    K = 0.9,  # carrying capacity
-    alpha = 0.1, # water runoff
-    c = 0.2, # local competition
-    a = 0.6, # search efficiency
-    v = 0.1, # attraction effect
-    h = 50, # handling time (one individual on landscape unit)
-    p = 0.8, # protection effect
-    L = 6, # livestock units per landscape unit 
-    q = 1 # search ineffiency at low cover
-  )
-
-```
-
-we get the following dynamics
+for the set of parameters given above, we get the following dynamics
 
@@
   with(parms,
-       r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) - m * rho[1] - ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q))
+       r * (b + (1 - b) * f * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) - m * rho[1] - ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L * (1 - p * rho[2]/rho[1]))/(1 +(a + v*rho[2]/rho[1]) * h  * rho[1]^( 1 + q))
        )  
@@
   with(parms,
-       2* (rho[1] - rho[2]) * r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) / (1-rho[1]) - 2 * rho[2] * m  - 2 * rho[2] * ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q))
+       2* (rho[1] - rho[2]) * r * ( b + (1 - b) * f * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) / (1-rho[1]) - 2 * rho[2] * m  - 2 * rho[2] * ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L * (1 - p * rho[2]/rho[1]))/(1 +(a + v*rho[2]/rho[1]) * h  * rho[1]^( 1 + q))
        )  
@@
 axis(4, at = round(tail(ODE_spex,1)$rho_1,2),cex.axis = 0.7, las = 1 )
+
+lines(spexmodel$rho_one ~ spexmodel$time, col = "grey50")
 
@@
 parms$L <- seq(0,20,0.2) #seq(0,1,.002)+0.001
+parms$b <- 0.4
 parms$rho_ini <- c(0.9, 0.01)
@@
 
+## analytical solutions
+
 # References

### 2015-02-03 code/simfunctions.r

@@ -302,15 +302,15 @@ runCA <- function(init, parms, width = 100, height = 100, delta = 0.1, t_max = 1
    
     # model specific part:
     # 1 - setting time-step parameters
-    #parms_temp$local <- count(x_old, "1")/4  # count local density of occupied fields for each cell:
+    parms_temp$q_one_one <- count(x_old, "1")/4  # count local density of occupied fields for each cell:
     
     # 2 - drawing random numbers
     rnum <- runif(width*height) # one random number between 0 and 1 for each cell
     
     # 3 - setting transition probabilities
-    growth <- with(parms_temp, (r * b * rho_one^(1 + alpha) * ( 1 - (rho_one / K)) / (1 - rho_one))  *delta)  # recolonisation rates of all cells 
+    growth <- with(parms_temp, (r * (b + (1-b)*f*q_one_one) * rho_one^(1 + alpha) * ( 1 - (rho_one / (K * (1-c*q_one_one) ))) / (1 - rho_one))  *delta)  # recolonisation rates of all cells 
     
-    death <- with(parms_temp, (m + ( a * L * rho_one^(q) )/( 1 + a * h * rho_one^(1+q) )) *delta)   # set probability of death for each cell
+    death <- with(parms_temp, (m + ( (a + v*q_one_one) * L * (1-p*q_one_one) * rho_one^(q) )/( 1 + (a + v*q_one_one) * h * rho_one^(1+q) )) *delta)   # set probability of death for each cell
     
     # check for sum of probabilities to be inferior 1 and superior 0
     if(any(c(growth, death) > 1 )) warning(paste("a set probability is exceeding 1 in time step", i, "! decrease delta!!!")) 
@@ -367,14 +367,13 @@ ini_rho <- function(rho_1) {
 
 d_rho_1 <- function(rho, parms) { 
   with(parms,
[…]

### 2015-02-03 documentation/analytical.Rmd
search analytical solution (crazy)

---
title: "Untitled"
author: "Florian Schneider"
date: "Tuesday, February 03, 2015"
output: html_document
---

Consider the equation system:

$$  \frac{\mathrm{d}\rho_{11} }{\mathrm{d}t} = 2 (\rho_1 - \rho_{11}) \frac{r \rho_1^{1+\alpha} (b + (1 - b)  f \frac{\rho_{1} - \rho_{11}}{1 - \rho_1} )) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a  + v \frac{\rho_{11}}{\rho_1})L (1 - p \frac{\rho_{11}}{\rho_1})\rho_1^{q}}{1+(a  + v \frac{\rho_{11}}{\rho_1})h\rho_1^{1+q}}  $$

$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b + (1 - b) f  \frac{\rho_{1} - \rho_{11}}{1 - \rho_1}) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})}) - \rho_{1}m -  \frac{(a  + v \frac{\rho_{11}}{\rho_1})L (1 - p \frac{\rho_{11}}{\rho_1})\rho_1^{1+q}}{1+(a  + v \frac{\rho_{11}}{\rho_1})h\rho_1^{1+q}}   $$

### 1.

assume:

$$  \frac{\mathrm{d}\rho_{11} }{\mathrm{d}t} = 0 $$

then,

$$  0 = 2 (\rho_1 - \rho_{11}) \frac{r \rho_1^{1+\alpha} (b + (1 - b)  f \frac{\rho_{1} - \rho_{11}}{1 - \rho_1} )) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a  + v \frac{\rho_{11}}{\rho_1})L (1 - p \frac{\rho_{11}}{\rho_1})\rho_1^{q}}{1+(a  + v \frac{\rho_{11}}{\rho_1})h\rho_1^{1+q}}  $$

$$  \rho_{11} m +  \rho_{11} \frac{(a  + v \rho_{11}\rho_1^{-1})L (1 - p \rho_{11}\rho_1^{-1})\rho_1^{q}}{1+(a  + v\rho_{11}\rho_1^{-1})h\rho_1^{1+q}} =  (\rho_1 - \rho_{11}) r \rho_1^{1+\alpha} (b + (1 - b)  f \frac{\rho_{1} - \rho_{11}}{1 - \rho_1} )) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})}) \frac{1}{1-\rho_1}   $$

$$  \rho_{11} m +  \frac{(a\rho_1^{q} \rho_{11}  + v \rho_{11}^2\rho_1^{q-1}) (L - L p \rho_{11}\rho_1^{-1})}{1+(a  + v\rho_{11}\rho_1^{-1})h\rho_1^{1+q}} =  (\rho_1 - \rho_{11}) r \rho_1^{1+\alpha} (b + (f - bf) (\rho_{1} - \rho_{11}) \frac{1}{1 - \rho_1} )) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})}) \frac{1}{1-\rho_1}   $$

$$  \rho_{11} m +  \frac{aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2}  }{1+(a  + v\rho_{11}\rho_1^{-1})h\rho_1^{1+q}} =  (\rho_1 - \rho_{11}) r \rho_1^{1+\alpha} (b + (f - bf) (\rho_{1} - \rho_{11}) \frac{1}{1 - \rho_1} )) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})}) \frac{1}{1-\rho_1}   $$

$$  \rho_{11} m +  \frac{aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2}  }{1+ah\rho_1^{1+q}  + vh\rho_1^{q}\rho_{11}} =  (\rho_1 - \rho_{11}) r \rho_1^{1+\alpha} (b + (f - bf) (\rho_{1} - \rho_{11}) \frac{1}{1 - \rho_1} )) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})}) \frac{1}{1-\rho_1}   $$

$$  \rho_{11} m +  \frac{aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2}  }{1+ah\rho_1^{1+q}  + vh\rho_1^{q}\rho_{11}} =  (\rho_1 r \rho_1^{1+\alpha}  - \rho_{11} r \rho_1^{1+\alpha} )(b + (f\rho_{1} - f\rho_{11} - bf\rho_{1} + bf \rho_{11}) \frac{1}{1 - \rho_1} )) (1-\rho_1K^{-1} (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})^{-1}) \frac{1}{1-\rho_1}   $$

$$  \rho_{11} m +  \frac{aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2}  }{1+ah\rho_1^{1+q}  + vh\rho_1^{q}\rho_{11}} =  (\rho_1 r \rho_1^{1+\alpha}  - \rho_{11} r \rho_1^{1+\alpha} )(b + f\rho_{1}(1 - \rho_1)^{-1} - f\rho_{11}(1 - \rho_1)^{-1} - bf\rho_{1}(1 - \rho_1)^{-1} + bf \rho_{11}(1 - \rho_1)^{-1}  ) (1-\rho_1K^{-1} (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})^{-1}) (1 - \rho_1)^{-1}    $$

$$  \rho_{11} m +  \frac{aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2}  }{1+ah\rho_1^{1+q}  + vh\rho_1^{q}\rho_{11}} =  (\rho_1 r \rho_1^{1+\alpha}  - \rho_{11} r \rho_1^{1+\alpha} )(b  (1 - \rho_1)^{-1}  + f\rho_{1}(1 - \rho_1)^{-2} - f\rho_{11}(1 - \rho_1)^{-2} - bf\rho_{1}(1 - \rho_1)^{-2} + bf \rho_{11}(1 - \rho_1)^{-2}  ) (1-\rho_1K^{-1} (1 - c (\rho_{1} - \rho_{11})\frac{1}{1 - \rho_1})^{-1})  $$

$$   m\rho_{11} + amh\rho_1^{1+q}\rho_{11} + vmh\rho_1^{q}\rho_{11}^2  + aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2}  =  (\rho_1 r \rho_1^{1+\alpha}  - \rho_{11} r \rho_1^{1+\alpha} )(b  (1 - \rho_1)^{-1}  + f\rho_{1}(1 - \rho_1)^{-2} - f\rho_{11}(1 - \rho_1)^{-2} - bf\rho_{1}(1 - \rho_1)^{-2} + bf \rho_{11}(1 - \rho_1)^{-2}  ) (1-\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) (1+ah\rho_1^{1+q}  + vh\rho_1^{q}\rho_{11}) $$

$$ m\rho_{11} + amh\rho_1^{1+q}\rho_{11} + vmh\rho_1^{q}\rho_{11}^2  + aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2}  = (1 - \rho_1)^{-1}(b    + f\rho_{1}(1 - \rho_1)^{-1} - f\rho_{11}(1 - \rho_1)^{-1} - bf\rho_{1}(1 - \rho_1)^{-1} + bf \rho_{11}(1 - \rho_1)^{-1}  ) (1-\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) (r \rho_1^{2+\alpha}  - r \rho_1^{1+\alpha} \rho_{11} +  arh\rho_1^{3+q+\alpha} - rah\rho_1^{2+q+\alpha}\rho_{11}  + vrh\rho_1^{2+q+\alpha}\rho_{11} - vrh\rho_1^{1+q+\alpha}\rho_{11}^2   )  $$

$$ m\rho_{11} + amh\rho_1^{1+q}\rho_{11} + vmh\rho_1^{q}\rho_{11}^2  + aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2}  = (1 - \rho_1)^{-2}(b (1 - \rho_1)^{1}   + f\rho_{1} - f\rho_{11} - bf\rho_{1} + bf \rho_{11} ) (1-\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) (r \rho_1^{2+\alpha}  - r \rho_1^{1+\alpha} \rho_{11} +  arh\rho_1^{3+q+\alpha} - rah\rho_1^{2+q+\alpha}\rho_{11}  + vrh\rho_1^{2+q+\alpha}\rho_{11} - vrh\rho_1^{1+q+\alpha}\rho_{11}^2   )  $$

$$ ( m\rho_{11} + amh\rho_1^{1+q}\rho_{11} + vmh\rho_1^{q}\rho_{11}^2  + aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2} )(1 - \rho_1)^{2} = (b - b\rho_1 + f\rho_{1} - f\rho_{11} - bf\rho_{1} + bf \rho_{11} ) (1-\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) (r \rho_1^{2+\alpha}  - r \rho_1^{1+\alpha} \rho_{11} +  arh\rho_1^{3+q+\alpha} - rah\rho_1^{2+q+\alpha}\rho_{11}  + vrh\rho_1^{2+q+\alpha}\rho_{11} - vrh\rho_1^{1+q+\alpha}\rho_{11}^2   ) $$

$$ ( m\rho_{11} + amh\rho_1^{1+q}\rho_{11} + vmh\rho_1^{q}\rho_{11}^2  + aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2} )(1 - \rho_1)^{2} = (b - (b+f-bf) \rho_1 - (f +bf)\rho_{11} ) (1 -\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) (r \rho_1^{2+\alpha}  - r \rho_1^{1+\alpha}\rho_{11} + arh\rho_1^{3+q+\alpha} - rah\rho_1^{2+q+\alpha}\rho_{11}  + vrh\rho_1^{2+q+\alpha}\rho_{11} - vrh\rho_1^{1+q+\alpha}\rho_{11}^2   ) $$

to simplify, I assume $q = 0$ and $\alpha = 0$:

$$ ( m\rho_{11} + amh\rho_1\rho_{11} + aL\rho_1\rho_{11} + vmh\rho_1\rho_{11}^2   - aLp \rho_1^{-1}\rho_{11}^2 + v L\rho_1^{-1}  \rho_{11}^2 - v L p \rho_1^{-2}\rho_{11}^3 )(1 - \rho_1)^{2} = (b - (b+f-bf) \rho_1 - (f +bf)\rho_{11} ) (1 -\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) (  r \rho_1^{2} + arh\rho_1^{3} - r \rho_1^{1}\rho_{11}  - rah\rho_1^{2}\rho_{11}  + vrh\rho_1^{2}\rho_{11} - vrh\rho_1\rho_{11}^2   ) $$

$$ ( m\rho_{11} + amh\rho_1\rho_{11} + aL\rho_1\rho_{11} + vmh\rho_1\rho_{11}^2   - aLp \rho_1^{-1}\rho_{11}^2 + v L\rho_1^{-1}  \rho_{11}^2 - v L p \rho_1^{-2}\rho_{11}^3 )(1 - \rho_1)^{2} =  (1 -\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) ( br \rho_1^{2} - (br+fr-brf) \rho_1^{3} + barh\rho_1^{3}- (barh+farh-bfarh) \rho_1^{4} - (fr +brf) \rho_1^{2}\rho_{11} - (farh +bfarh)\rho_1^{3}\rho_{11}  - br \rho_1^{1}\rho_{11} + (br+fr-bfr) \rho_1^{2}\rho_{11} + (fr +brf) \rho_1^{1}\rho_{11}^2  - brah\rho_1^{2}\rho_{11} + (brah+frah-bfrah) \rho_1^{3}\rho_{11} + (rahf +bfrah)\rho_1^{2}\rho_{11}^2 + bvrh\rho_1^{2}\rho_{11} - (bvrh+fvrh-bfvrh) \rho_1^{3}\rho_{11} - (fvrh +bfvrh)\rho_1^{2}\rho_{11}^2  - bvrh\rho_1\rho_{11}^2 + (bvrh+fvrh-bvrhf) \rho_1^2\rho_{11}^2 + (fvrh +bvrhf)\rho_1\rho_{11}^3 ) $$

$$ ( m\rho_{11} + amh\rho_1\rho_{11} + aL\rho_1\rho_{11} + vmh\rho_1\rho_{11}^2   - aLp \rho_1^{-1}\rho_{11}^2 + v L\rho_1^{-1}  \rho_{11}^2 - v L p \rho_1^{-2}\rho_{11}^3 )(1 - \rho_1)^{2} =  (1 -\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) ( br \rho_1^{2} - (br+fr-brf+ barh) \rho_1^{3} - (barh+farh-bfarh) \rho_1^{4} - br \rho_1\rho_{11}- (2fr+br-brah+ bvrh) \rho_1^{2}\rho_{11} - (2farh + brah-bvrh-fvrh+bfvrh)\rho_1^{3}\rho_{11}+ (fr +brf- bvrh) \rho_1\rho_{11}^2  + (rahf +bfrah+ bvrh)\rho_1^{2}\rho_{11}^2 + (fvrh +bvrhf)\rho_1\rho_{11}^3 ) $$

$$  m\rho_{11} - (2-amh-aL)\rho_1\rho_{11} + (m- 2amh- 2aL)\rho_1^{2}\rho_{11}+ amh\rho_1^{3}\rho_{11} + aL\rho_1^{3}\rho_{11}+ (vL- aLp) \rho_1^{-1}\rho_{11}^2 + (vmh- aLp+ vL)\rho_1\rho_{11}^2- 2vmh\rho_1^2\rho_{11}^2 + vmh\rho_1^{3}\rho_{11}^2+ 2aLp\rho_{11}^2 - 2vL\rho_{11}^2 - vLp\rho_1^{-2}\rho_{11}^3 + 2vLp\rho_1^{-1}\rho_{11}^3 - vLp  \rho_{11}^3 =  (1 -\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) ( br \rho_1^{2} - (br+fr-brf+ barh) \rho_1^{3} - (barh+farh-bfarh) \rho_1^{4} - br \rho_1\rho_{11}- (2fr+br-brah+ bvrh) \rho_1^{2}\rho_{11} - (2farh + brah-bvrh-fvrh+bfvrh)\rho_1^{3}\rho_{11}+ (fr +brf- bvrh) \rho_1\rho_{11}^2  + (rahf +bfrah+ bvrh)\rho_1^{2}\rho_{11}^2 + (fvrh +bvrhf)\rho_1\rho_{11}^3 ) $$

I substitute the left hand side

$$ x =  m\rho_{11} - (2-amh-aL)\rho_1\rho_{11} + (m- 2amh- 2aL)\rho_1^{2}\rho_{11}+ amh\rho_1^{3}\rho_{11} + aL\rho_1^{3}\rho_{11}+ (vL- aLp) \rho_1^{-1}\rho_{11}^2 + (vmh- aLp+ vL)\rho_1\rho_{11}^2- 2vmh\rho_1^2\rho_{11}^2 + vmh\rho_1^{3}\rho_{11}^2+ 2aLp\rho_{11}^2 - 2vL\rho_{11}^2 - vLp\rho_1^{-2}\rho_{11}^3 + 2vLp\rho_1^{-1}\rho_{11}^3 - vLp  \rho_{11}^3$$

and the right hand part

$$ y =  ( br \rho_1^{2} - (br+fr-brf+ barh) \rho_1^{3} - (barh+farh-bfarh) \rho_1^{4} - br \rho_1\rho_{11}- (2fr+br-brah+ bvrh) \rho_1^{2}\rho_{11} - (2farh + brah-bvrh-fvrh+bfvrh)\rho_1^{3}\rho_{11}+ (fr +brf- bvrh) \rho_1\rho_{11}^2  + (rahf +bfrah+ bvrh)\rho_1^{2}\rho_{11}^2 + (fvrh +bvrhf)\rho_1\rho_{11}^3 )$$

to get the simplified

$$ x =  (1 -\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) y$$

$$ x =  y - y\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1} $$

$$ x - y = - y\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1} $$

$$ x - y (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  ) = - y\rho_1K^{-1} $$

$$ (x - y) - (x - y)(c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1}) = - y\rho_1K^{-1} $$

$$ (x - y) - (x - y)(c \rho_{1} - c \rho_{11})(1 - \rho_1)^{-1} = - y\rho_1K^{-1} $$

$$ x - y + y\rho_1K^{-1} =  (x - y)(c \rho_{1} - c \rho_{11})(1 - \rho_1)^{-1}$$

$$( xK - yK + y\rho_1 )(1 - \rho_1) =  (xK - yK)(c \rho_{1} - c \rho_{11})$$

$$xK - yK + y\rho_1 - xK\rho_1 + yK\rho_1 - y\rho_1^2 =  (xK - yK)(c \rho_{1} - c \rho_{11})$$

$$xK - yK + y\rho_1 - xK\rho_1 + yK\rho_1 - y\rho_1^2 =  xKc\rho_{1} - yKc\rho_{1} -xK c \rho_{11} + yK c \rho_{11}$$

Now I resubstitute $x$ and $y$:

$$xK - yK + y\rho_1 - xK\rho_1 + yK\rho_1 - y\rho_1^2 =  xKc\rho_{1} - yKc\rho_{1} - xKc\rho_{11} + yKc \rho_{11}$$

$$mK \rho_{11} - (2-amh-aL)K \rho_1\rho_{11} + (m- 2amh- 2aL)K \rho_1^{2}\rho_{11}+ amhK \rho_1^{3}\rho_{11} + aLK \rho_1^{3}\rho_{11}+ (vL- aLp)K  \rho_1^{-1}\rho_{11}^2 + (vmh- aLp+ vL)K \rho_1\rho_{11}^2- 2vmhK \rho_1^2\rho_{11}^2 + vmhK \rho_1^{3}\rho_{11}^2+ 2aLpK \rho_{11}^2 - 2vLK \rho_{11}^2 - vLpK \rho_1^{-2}\rho_{11}^3 + 2vLpK \rho_1^{-1}\rho_{11}^3 - vLpK \rho_{11}^3
- yK
+ y\rho_1 
- xK\rho_1 
+ yK\rho_1 
- y\rho_1^2 
+ yKc\rho_{1}
+ xKc\rho_{11}
- yKc\rho_{11}
- xKc\rho_{1}
= 0
$$

### 2015-02-04 documentation/analytical.Rmd
work on analytical solution

@@
 $$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b + (1 - b) f  \frac{\rho_{1} - \rho_{11}}{1 - \rho_1}) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})}) - \rho_{1}m -  \frac{(a  + v \frac{\rho_{11}}{\rho_1})L (1 - p \frac{\rho_{11}}{\rho_1})\rho_1^{1+q}}{1+(a  + v \frac{\rho_{11}}{\rho_1})h\rho_1^{1+q}}   $$
-
-### 1.
 
@@
 
-$$mK \rho_{11} - (2-amh-aL)K \rho_1\rho_{11} + (m- 2amh- 2aL)K \rho_1^{2}\rho_{11}+ amhK \rho_1^{3}\rho_{11} + aLK \rho_1^{3}\rho_{11}+ (vL- aLp)K  \rho_1^{-1}\rho_{11}^2 + (vmh- aLp+ vL)K \rho_1\rho_{11}^2- 2vmhK \rho_1^2\rho_{11}^2 + vmhK \rho_1^{3}\rho_{11}^2+ 2aLpK \rho_{11}^2 - 2vLK \rho_{11}^2 - vLpK \rho_1^{-2}\rho_{11}^3 + 2vLpK \rho_1^{-1}\rho_{11}^3 - vLpK \rho_{11}^3
-- yK
-+ y\rho_1 
-- xK\rho_1 
-+ yK\rho_1 
-- y\rho_1^2 
-+ yKc\rho_{1}
-+ xKc\rho_{11}
-- yKc\rho_{11}
-- xKc\rho_{1}
-= 0
-$$
+Which gives the messy:
+
+$$mK \rho_{11} - (2-amh-aL)K \rho_1\rho_{11} + (m- 2amh- 2aL)K \rho_1^{2}\rho_{11}+ amhK \rho_1^{3}\rho_{11} + aLK \rho_1^{3}\rho_{11}+ (vL- aLp)K  \rho_1^{-1}\rho_{11}^2 + (vmh- aLp+ vL)K \rho_1\rho_{11}^2- 2vmhK \rho_1^2\rho_{11}^2 + vmhK \rho_1^{3}\rho_{11}^2+ 2aLpK \rho_{11}^2 - 2vLK \rho_{11}^2 - vLpK \rho_1^{-2}\rho_{11}^3 + 2vLpK \rho_1^{-1}\rho_{11}^3 - vLpK \rho_{11}^3    - brK \rho_1^{2} + (br+fr-brf+ barh)K \rho_1^{3} + (barh+farh-bfarh)K \rho_1^{4} + brK \rho_1\rho_{11} + (2fr+br-brah+ bvrh)K \rho_1^{2}\rho_{11} + (2farh + brah-bvrh-fvrh+bfvrh)K\rho_1^{3}\rho_{11} - (fr +brf- bvrh)K \rho_1\rho_{11}^2 - (rahf +bfrah+ bvrh)K\rho_1^{2}\rho_{11}^2 - (fvrh +bvrhf)K\rho_1\rho_{11}^3     + br \rho_1^{3} - (br+fr-brf+ barh) \rho_1^{4} - (barh+farh-bfarh) \rho_1^{5} - br \rho_1^2\rho_{11}- (2fr+br-brah+ bvrh) \rho_1^{3}\rho_{11} - (2farh + brah-bvrh-fvrh+bfvrh)\rho_1^{4}\rho_{11}+ (fr +brf- bvrh) \rho_1^2\rho_{11}^2  + (rahf +bfrah+ bvrh)\rho_1^{3}\rho_{11}^2 + (fvrh +bvrhf)\rho_1^2\rho_{11}^3     - mK \rho_1\rho_{11} + (2-amh-aL)K \rho_1^2\rho_{11} - (m- 2amh- 2aL)K \rho_1^{3}\rho_{11}- amhK \rho_1^{4}\rho_{11} - aLK \rho_1^{4}\rho_{11}- (vL- aLp)K  \rho_{11}^2 - (vmh- aLp+ vL)K \rho_1^2\rho_{11}^2+ 2vmhK \rho_1^3\rho_{11}^2 - vmhK \rho_1^{4}\rho_{11}^2- 2aLpK \rho_1 \rho_{11}^2 + 2vLK \rho_1\rho_{11}^2 + vLpK \rho_1^{-1}\rho_{11}^3 - 2vLpK \rho_{11}^3 + vLpK \rho_1\rho_{11}^3    + brK \rho_1^{3} - (br+fr-brf+ barh)K \rho_1^{4} - (barh+farh-bfarh)K \rho_1^{5} - brK \rho_1^2\rho_{11}- (2fr+br-brah+ bvrh)K \rho_1^{3}\rho_{11} - (2farh + brah-bvrh-fvrh+bfvrh)K\rho_1^{4}\rho_{11}+ (fr +brf- bvrh)K \rho_1^2\rho_{11}^2  + (rahf +bfrah+ bvrh)K\rho_1^{3}\rho_{11}^2 + (fvrh +bvrhf)K\rho_1^2\rho_{11}^3     - br \rho_1^{4} + (br+fr-brf+ barh) \rho_1^{5} + (barh+farh-bfarh) \rho_1^{6} + br \rho_1^3\rho_{11}+ (2fr+br-brah+ bvrh) \rho_1^{4}\rho_{11} + (2farh + brah-bvrh-fvrh+bfvrh)\rho_1^{5}\rho_{11}- (fr +brf- bvrh) \rho_1^3\rho_{11}^2  - (rahf +bfrah+ bvrh)\rho_1^{4}\rho_{11}^2 - (fvrh +bvrhf)\rho_1^3\rho_{11}^3     + brKc \rho_1^{3} - (br+fr-brf+ barh)Kc \rho_1^{4} - (barh+farh-bfarh)Kc \rho_1^{5} - brKc \rho_1^2\rho_{11}- (2fr+br-brah+ bvrh)Kc \rho_1^{3}\rho_{11} - (2farh + brah-bvrh-fvrh+bfvrh)Kc\rho_1^{4}\rho_{11}+ (fr +brf- bvrh)Kc \rho_1^2\rho_{11}^2  + (rahf +bfrah+ bvrh)Kc\rho_1^{3}\rho_{11}^2 + (fvrh +bvrhf)Kc\rho_1^2\rho_{11}^3     - mKc \rho_1\rho_{11} + (2-amh-aL)Kc \rho_1^2\rho_{11} - (m- 2amh- 2aL)Kc \rho_1^{3}\rho_{11}- amhKc \rho_1^{4}\rho_{11} - aLKc \rho_1^{4}\rho_{11}- (vL- aLp)Kc \rho_{11}^2 - (vmh- aLp+ vL)Kc \rho_1^2\rho_{11}^2+ 2vmhKc \rho_1^3\rho_{11}^2 - vmhKc \rho_1^{4}\rho_{11}^2- 2aLpKc \rho_1 \rho_{11}^2 + 2vLKc \rho_1\rho_{11}^2 + vLpKc \rho_1^{-1}\rho_{11}^3 - 2vLpKc \rho_{11}^3 + vLpKc \rho_1\rho_{11}^3    + mKc \rho_{11}^2 - (2-amh-aL)Kc \rho_1\rho_{11}^2 + (m- 2amh- 2aL)Kc \rho_1^{2}\rho_{11}^2+ amhKc \rho_1^{3}\rho_{11}^2 + aLKc \rho_1^{3}\rho_{11}^2+ (vL- aLp)Kc  \rho_1^{-1}\rho_{11}^3 + (vmh- aLp+ vL)Kc \rho_1\rho_{11}^3- 2vmhKc \rho_1^2\rho_{11}^3 + vmhKc \rho_1^{3}\rho_{11}^3+ 2aLpKc \rho_{11}^3 - 2vLKc \rho_{11}^3 - vLpKc \rho_1^{-2}\rho_{11}^4 + 2vLpKc \rho_1^{-1}\rho_{11}^4 - vLpKc \rho_{11}^4    - brKc \rho_1^{2}\rho_{11} + (br+fr-brf+ barh)Kc \rho_1^{3}\rho_{11} + (barh+farh-bfarh)Kc \rho_1^{4}\rho_{11} + brKc \rho_1\rho_{11}^2 + (2fr+br-brah+ bvrh)Kc \rho_1^{2}\rho_{11}^2 + (2farh + brah-bvrh-fvrh+bfvrh)Kc\rho_1^{3}\rho_{11}^2 - (fr +brf- bvrh)Kc \rho_1\rho_{11}^3 - (rahf +bfrah+ bvrh)Kc\rho_1^{2}\rho_{11}^3 - (fvrh +bvrhf)Kc\rho_1\rho_{11}^4 = 0 $$
+
+I think I stop here, since there are fourth order polinomials that will not resolve.
+
+<!--
+
+m*K*x-(2-a*m*h-a*L)*K*y*x + (m-2*a*m*h-2*a*L)*K*y^2*x+ a*m*h*K*y^3*x + a*L*K*y^3*x+ (v*L-a*L*p)*K*y^-1*x^2 + (v*m*h-a*L*p+v*L)*K*y*x^2- 2*v*m*h*K*y^2*x^2 + v*m*h*K*y^3*x^2+ 2*a*L*p*K*x^2 - 2*v*L*K*x^2 - v*L*p*K*y^(-2)*x^3 + 2*v*L*p*K*y^(-1)*x^3 - v*L*p*K*x^3    - b*r*K*y^2  + (b*r+f*r-b*r*f+b*a*r*h)*K*y^3  + (b*a*r*h+f*a*r*h-b*f*a*r*h)*K y^4  + b*r*K*y*x + (2*f*r+b*r-b*r*a*h+ b*v*r*h)*K*y^2*x + (2*f*a*r*h + b*r*a*h-b*v*r*h-f*v*r*h+b*f*v*r*h)*K*y^3*x - (f*r +b*r*f- b*v*r*h)*K*y*x^2 - (r*a*h*f +b*f*r*a*h+ b*v*r*h)*K*y^2*x^2 - (f*v*r*h +b*v*r*h*f)*K*y*x^3     + b*r*y^3  - (b*r+f*r-b*r*f+ b*a*r*h)*y^4  - (b*a*r*h+f*a*r*h-b*f*a*r*h)*y^5  - b*r*y^2*x- (2*f*r+b*r-b*r*a*h+ b*v*r*h)*y^3*x - (2*f*a*r*h + b*r*a*h-b*v*r*h-f*v*r*h+b*f*v*r*h)*y^4*x+ (f*r +b*r*f- b*v*r*h)*y^2*x^2 + (r*a*h*f +b*f*r*a*h+ b*v*r*h)*y^3*x^2 + (f*v*r*h +b*v*r*h*f)*y^2*x^3     - m*K*y*x + (2-a*m*h-a*L)*K*y^2*x - (m- 2*a*m*h- 2*a*L)*K *y^3*x- a*m*h*K*y^4*x - a*L*K*y^4*x- (v*L- a*L*p)*K*x^2 - (v*m*h- a*L*p+ v*L)*K*y^2*x^2+ 2*v*m*h*K*y^3*x^2 - v*m*h*K*y^4*x^2- 2*a*L*p*K*y*x^2 + 2*v*L*K*y*x^2 + v*L*p*K*y^(-1)*x^3 - 2*v*L*p*K*x^3 + v*L*p*K*y*x^3   + b*r*K*y^3  - (b*r+f*r-b*r*f+ b*a*r*h)*K*y^4  - (b*a*r*h+f*a*r*h-b*f*a*r*h)*K*y^5  - b*r*K*y^2*x- (2*f*r+b*r-b*r*a*h+ b*v*r*h)*K*y^3*x - (2*f*a*r*h + b*r*a*h-b*v*r*h-f*v*r*h+b*f*v*r*h)*K*y^4*x+ (f*r +b*r*f- b*v*r*h)*K*y^2*x^2  + (r*a*h*f +b*f*r*a*h+ b*v*r*h)*K*y^3*x^2 + (f*v*r*h +b*v*r*h*f)*K*y^2*x^3     - b*r*y^4  + (b*r+f*r-b*r*f+ b*a*r*h)*y^5  + (b*a*r*h+f*a*r*h-b*f*a*r*h)*y^6  + b*r*y^3*x+ (2*f*r+b*r-b*r*a*h+ b*v*r*h)*y^4*x + (2*f*a*r*h+b*r*a*h-b*v*r*h-f*v*r*h+b*f*v*r*h)*y^5*x- (f*r +b*r*f- b*v*r*h)*y^3*x^2  - (r*a*h*f+b*f*r*a*h+b*v*r*h)*y^4*x^2 - (f*v*r*h +b*v*r*h*f)*y^3*x^3     + b*r*K*c*y^3  - (b*r+f*r-b*r*f+ b*a*r*h)*K*c*y^4  - (b*a*r*h+f*a*r*h-b*f*a*r*h)*K*c*y^5  - b*r*K*c* y^2*x- (2*f*r+b*r-b*r*a*h+ b*v*r*h)*K*c*y^3*x - (2*f*a*r*h + b*r*a*h-b*v*r*h-f*v*r*h+b*f*v*r*h)*K*c*y^4*x+ (f*r +b*r*f- b*v*r*h)*K*c*y^2*x^2  + (r*a*h*f +b*f*r*a*h+ b*v*r*h)*K*c*y^3*x^2 + (f*v*r*h+b*v*r*h*f)*K*c*y^2*x^3     - m*K*c*y*x + (2-a*m*h-a*L)*K*c*y^2*x - (m- 2*a*m*h- 2*a*L)*K*c*y^3*x - a*m*h*K*c*y^4*x - a*L*K*c*y^4*x- (v*L- a*L*p)*K*c*x^2 - (v*m*h-a*L*p+v*L)*K*c*y^2*x^2+ 2*v*m*h*K*c* y^3*x^2 - v*m*h*K*c*y^4*x^2- 2*a*L*p*K*c*y*x^2 + 2*v*L*K*c*y*x^2 + v*L*p*K*c*y^(-1)*x^3 - 2*v*L*p*K*c*x^3 + v*L*p*K*c*y*x^3    + m*K*c*x^2 - (2-a*m*h-a*L)*K*c*y*x^2 + (m- 2*a*m*h-2*a*L)*K*c*y^2*x^2+ a*m*h*K*c*y^3*x^2 + a*L*K*c*y^3*x^2+ (v*L- a*L*p)*K*c*y^(-1)*x^3 + (v*m*h-a*L*p+v*L)*K*c*y*x^3 - 2*v*m*h*K*c*y^2*x^3 + v*m*h*K*c*y^3*x^3 + 2*a*L*p*K*c*x^3 - 2*v*L*K*c*x^3 - v*L*p*K*c*y^(-2)*x^4 + 2*v*L*p*K*c*y^(-1)*x^4 - v*L*p*K*c*x^4    - b*r*K*c*y^2*x + (b*r+f*r-b*r*f+b*a*r*h)*K*c*y^3*x + (b*a*r*h+f*a*r*h-b*f*a*r*h)*K*c*y^4*x + b*r*K*c*y*x^2 + (2*f*r+b*r-b*r*a*h+ b*v*r*h)*K*c*y^2*x^2 + (2*f*a*r*h+b*r*a*h-b*v*r*h-f*v*r*h+b*f*v*r*h)*K*c*y^3*x^2 - (f*r+b*r*f-b*v*r*h)*K*c*y*x^3 - (r*a*h*f+b*f+r*a*h+b*v*r*h)*K*c*y^2*x^3 - (f*v*r*h+b*v*r*h*f)*K*c*y*x^4 = 0
+
+-->

### 2015-02-04 documentation/model_formulation.Rmd
update model formulation, produce pdf

@@
 author: "Florian D. Schneider"
-date: "Tuesday, November 18, 2014"
+date: "Tuesday, February 4th, 2014"
 output:
@@
 
-And the animated lattice would look like this:
 ```{r, eval = FALSE, echo = FALSE, warning = FALSE}
-animateCA(nullmodel, "ca_nullmodel.gif")
+animateCA(spexmodel, "ca_spexmodel.gif")
 
@@
 
-for the set of parameters given above, we get the following dynamics
+for the set of parameters given above, we can project the following dynamics from solving the ODE (CA model simulation as grey line).
 
@@
 
+The following shows bifurcation diagrams over an environmental gradient.
+
 ```{r, echo=FALSE}
@@
 
+The following shows bifurcation diagrams over a livestock density gradient.
+
 ```{r, echo=FALSE}
@@
 
+A dynamic Shiny implementation of that can be found under 'documentation/interactive_spexmodel_bifurcations.Rmd' ( run in RStudio, specify parallel backend.).
+
+The simulation backend is now all set up. As a final step, we want to check out if the model can be solved analytically.
+
 ## analytical solutions
 
+I tried to solve the equilibria of the equation system by assuming
+
+$$  \frac{\mathrm{d}\rho_{11} }{\mathrm{d}t} = 0 $$
+
+thus,
+
+$$  0 = 2 (\rho_1 - \rho_{11}) \frac{r \rho_1^{1+\alpha} (b + (1 - b)  f \frac{\rho_{1} - \rho_{11}}{1 - \rho_1} )) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a  + v \frac{\rho_{11}}{\rho_1})L (1 - p \frac{\rho_{11}}{\rho_1})\rho_1^{q}}{1+(a  + v \frac{\rho_{11}}{\rho_1})h\rho_1^{1+q}}  $$
+
+and then solving towards $\rho_{11}$. The simplification process can be found in the file 'analytical.Rmd'. It  leads to 4th order polynomials, which makes it impossible to solve. 
+However, solutions might be possible for certain cases, e.g. with only facilitation, or only associational protection, and all other effects set to zero.
+
+For instance, assuming competition $c = 0$ and attraction $v = 0$ yields a much simpler equation of second order
+
+$$0 = - brK \rho_1^{2}  + (br+2brK+rfK-brfK+barhK) \rho_1^{3}- (-2br-brK-frK+brfK+farhK-bfarhK-fr+brf-barh) \rho_1^{4} + (br+fr-brf-barhK-farhK+bfarhK-farh+bfarh) \rho_1^{5}+ (barh+farh-bfarh) \rho_1^{6}        + (mK + (brK-mK-2K+amhK+aLK) \rho_1 + (mK- br- 2amKh- 2aLK+2frK+brK-brahK+2K-amhK-aLK- brK) \rho_1^{2}   + (-mK+aLK-br+3amhK+2farhK+2brahK-2fr+brah-2frK+2aLK) \rho_1^{3}-  (2farh-amhK-aLK-2farhK-brahK+2fr+br)\rho_1^{4} + (2farh + brah)\rho_1^{5}  )     \rho_{11}  + (aLpK\rho_1^{-1}+ 2aLpK  + aLpK + (-frK-brfK-3aLpK+2vLK) \rho_1  + (frbrf+frK+brfK-rahfK-bfrahK+aLpK) \rho_1^{2} + (rahf +bfrah+rahfK +bfrahK-fr -brf)\rho_1^{3} + (-rahf-bfrah) \rho_1^{4}    )   \rho_{11}^2   $$
+
 # References

### 2015-02-04 documentation/analytical.Rmd
work on analytical solution, simplifications

@@
 
-<!--
+$$0 =  (br+fr-brf+ barh)K \rho_1^{3} - brK \rho_1^{2} + (barh+farh-bfarh)K \rho_1^{4}   + br \rho_1^{3} - (br+fr-brf+ barh) \rho_1^{4} - (barh+farh-bfarh) \rho_1^{5} + brK \rho_1^{3}   - (br+fr-brf+ barh)K \rho_1^{4}- (barh+farh-bfarh)K \rho_1^{5}  - br \rho_1^{4} + (br+fr-brf+ barh) \rho_1^{5} + (barh+farh-bfarh) \rho_1^{6}  + brKc \rho_1^{3} - (br+fr-brf+ barh)Kc \rho_1^{4} - (barh+farh-bfarh)Kc \rho_1^{5}+ ( mK + (m- 2amh- 2aL)K \rho_1^{2} - (2-amh-aL)K \rho_1 + amhK \rho_1^{3} + aLK \rho_1^{3} + brK \rho_1 + (2fr+br-brah+ bvrh)K \rho_1^{2} + (2farh + brah-bvrh-fvrh+bfvrh)K\rho_1^{3} - br \rho_1^2- (2fr+br-brah+ bvrh) \rho_1^{3}  - (2farh + brah-bvrh-fvrh+bfvrh)\rho_1^{4} - mK \rho_1 + (2-amh-aL)K \rho_1^2  - (m- 2amh- 2aL)K \rho_1^{3}- amhK \rho_1^{4} - aLK \rho_1^{4}- brK \rho_1^2- (2fr+br-brah+ bvrh)K \rho_1^{3} - (2farh + brah-bvrh-fvrh+bfvrh)K\rho_1^{4}+ br \rho_1^3+ (2fr+br-brah+ bvrh) \rho_1^{4} + (2farh + brah-bvrh-fvrh+bfvrh)\rho_1^{5} - brKc \rho_1^2- (2fr+br-brah+ bvrh)Kc \rho_1^{3} - (2farh + brah-bvrh-fvrh+bfvrh)Kc\rho_1^{4} - mKc \rho_1 + (2-amh-aL)Kc \rho_1^2 - (m- 2amh- 2aL)Kc \rho_1^{3}- amhKc \rho_1^{4} - aLKc \rho_1^{4} - brKc \rho_1^{2} + (br+fr-brf+ barh)Kc \rho_1^{3} + (barh+farh-bfarh)Kc \rho_1^{4} ) \rho_{11}+ ((vL- aLp)K  \rho_1^{-1} + (vmh- aLp+ vL)K \rho_1- 2vmhK \rho_1^2 + vmhK \rho_1^{3}+ 2aLpK  - 2vLK   - (fr +brf- bvrh)K \rho_1 - (rahf +bfrah+ bvrh)K\rho_1^{2} + (fr +brf- bvrh) \rho_1^2  + (rahf +bfrah+ bvrh)\rho_1^{3}- (vL- aLp)K  - (vmh- aLp+ vL)K \rho_1^2 + 2vmhK \rho_1^3  - vmhK \rho_1^{4}- 2aLpK \rho_1  + 2vLK \rho_1   + (fr +brf- bvrh)K \rho_1^2  + (rahf +bfrah+ bvrh)K\rho_1^{3}   - (fr +brf- bvrh) \rho_1^3  - (rahf +bfrah+ bvrh)\rho_1^{4}  + (fr +brf- bvrh)Kc \rho_1^2  + (rahf +bfrah+ bvrh)Kc\rho_1^{3} - (vL- aLp)Kc  - (vmh- aLp+ vL)Kc \rho_1^2+ 2vmhKc \rho_1^3 - vmhKc \rho_1^{4}- 2aLpKc \rho_1  + 2vLKc \rho_1 + mKc  - (2-amh-aL)Kc \rho_1 + (m- 2amh- 2aL)Kc \rho_1^{2}+ amhKc \rho_1^{3} + aLKc \rho_1^{3}+ brKc \rho_1 + (2fr+br-brah+ bvrh)Kc \rho_1^{2} + (2farh + brah-bvrh-fvrh+bfvrh)Kc\rho_1^{3})\rho_{11}^2- (vLpK \rho_1^{-2} + 2vLpK \rho_1^{-1} - vLpK   - (fvrh +bvrhf)K\rho_1+ (fvrh +bvrhf)\rho_1^2  + vLpK \rho_1^{-1}    - 2vLpK  + vLpK \rho_1 + (fvrh +bvrhf)K\rho_1^2  - (fvrh +bvrhf)\rho_1^3+ (fvrh +bvrhf)Kc\rho_1^2 + vLpKc \rho_1^{-1} - 2vLpKc  + vLpKc \rho_1  + (vL- aLp)Kc  \rho_1^{-1} + (vmh- aLp+ vL)Kc \rho_1- 2vmhKc \rho_1^2 + vmhKc \rho_1^{3}+ 2aLpKc  - 2vLKc  - (fr +brf- bvrh)Kc \rho_1 - (rahf +bfrah+ bvrh)Kc\rho_1^{2} ) \rho_{11}^3- (vLpKc \rho_1^{-2} + 2vLpKc \rho_1^{-1} - vLpKc  - (fvrh +bvrhf)Kc\rho_1 ) \rho_{11}^4$$
 
-m*K*x-(2-a*m*h-a*L)*K*y*x + (m-2*a*m*h-2*a*L)*K*y^2*x+ a*m*h*K*y^3*x + a*L*K*y^3*x+ (v*L-a*L*p)*K*y^-1*x^2 + (v*m*h-a*L*p+v*L)*K*y*x^2- 2*v*m*h*K*y^2*x^2 + v*m*h*K*y^3*x^2+ 2*a*L*p*K*x^2 - 2*v*L*K*x^2 - v*L*p*K*y^(-2)*x^3 + 2*v*L*p*K*y^(-1)*x^3 - v*L*p*K*x^3    - b*r*K*y^2  + (b*r+f*r-b*r*f+b*a*r*h)*K*y^3  + (b*a*r*h+f*a*r*h-b*f*a*r*h)*K y^4  + b*r*K*y*x + (2*f*r+b*r-b*r*a*h+ b*v*r*h)*K*y^2*x + (2*f*a*r*h + b*r*a*h-b*v*r*h-f*v*r*h+b*f*v*r*h)*K*y^3*x - (f*r +b*r*f- b*v*r*h)*K*y*x^2 - (r*a*h*f +b*f*r*a*h+ b*v*r*h)*K*y^2*x^2 - (f*v*r*h +b*v*r*h*f)*K*y*x^3     + b*r*y^3  - (b*r+f*r-b*r*f+ b*a*r*h)*y^4  - (b*a*r*h+f*a*r*h-b*f*a*r*h)*y^5  - b*r*y^2*x- (2*f*r+b*r-b*r*a*h+ b*v*r*h)*y^3*x - (2*f*a*r*h + b*r*a*h-b*v*r*h-f*v*r*h+b*f*v*r*h)*y^4*x+ (f*r +b*r*f- b*v*r*h)*y^2*x^2 + (r*a*h*f +b*f*r*a*h+ b*v*r*h)*y^3*x^2 + (f*v*r*h +b*v*r*h*f)*y^2*x^3     - m*K*y*x + (2-a*m*h-a*L)*K*y^2*x - (m- 2*a*m*h- 2*a*L)*K *y^3*x- a*m*h*K*y^4*x - a*L*K*y^4*x- (v*L- a*L*p)*K*x^2 - (v*m*h- a*L*p+ v*L)*K*y^2*x^2+ 2*v*m*h*K*y^3*x^2 - v*m*h*K*y^4*x^2- 2*a*L*p*K*y*x^2 + 2*v*L*K*y*x^2 + v*L*p*K*y^(-1)*x^3 - 2*v*L*p*K*x^3 + v*L*p*K*y*x^3   + b*r*K*y^3  - (b*r+f*r-b*r*f+ b*a*r*h)*K*y^4  - (b*a*r*h+f*a*r*h-b*f*a*r*h)*K*y^5  - b*r*K*y^2*x- (2*f*r+b*r-b*r*a*h+ b*v*r*h)*K*y^3*x - (2*f*a*r*h + b*r*a*h-b*v*r*h-f*v*r*h+b*f*v*r*h)*K*y^4*x+ (f*r +b*r*f- b*v*r*h)*K*y^2*x^2  + (r*a*h*f +b*f*r*a*h+ b*v*r*h)*K*y^3*x^2 + (f*v*r*h +b*v*r*h*f)*K*y^2*x^3     - b*r*y^4  + (b*r+f*r-b*r*f+ b*a*r*h)*y^5  + (b*a*r*h+f*a*r*h-b*f*a*r*h)*y^6  + b*r*y^3*x+ (2*f*r+b*r-b*r*a*h+ b*v*r*h)*y^4*x + (2*f*a*r*h+b*r*a*h-b*v*r*h-f*v*r*h+b*f*v*r*h)*y^5*x- (f*r +b*r*f- b*v*r*h)*y^3*x^2  - (r*a*h*f+b*f*r*a*h+b*v*r*h)*y^4*x^2 - (f*v*r*h +b*v*r*h*f)*y^3*x^3     + b*r*K*c*y^3  - (b*r+f*r-b*r*f+ b*a*r*h)*K*c*y^4  - (b*a*r*h+f*a*r*h-b*f*a*r*h)*K*c*y^5  - b*r*K*c* y^2*x- (2*f*r+b*r-b*r*a*h+ b*v*r*h)*K*c*y^3*x - (2*f*a*r*h + b*r*a*h-b*v*r*h-f*v*r*h+b*f*v*r*h)*K*c*y^4*x+ (f*r +b*r*f- b*v*r*h)*K*c*y^2*x^2  + (r*a*h*f +b*f*r*a*h+ b*v*r*h)*K*c*y^3*x^2 + (f*v*r*h+b*v*r*h*f)*K*c*y^2*x^3     - m*K*c*y*x + (2-a*m*h-a*L)*K*c*y^2*x - (m- 2*a*m*h- 2*a*L)*K*c*y^3*x - a*m*h*K*c*y^4*x - a*L*K*c*y^4*x- (v*L- a*L*p)*K*c*x^2 - (v*m*h-a*L*p+v*L)*K*c*y^2*x^2+ 2*v*m*h*K*c* y^3*x^2 - v*m*h*K*c*y^4*x^2- 2*a*L*p*K*c*y*x^2 + 2*v*L*K*c*y*x^2 + v*L*p*K*c*y^(-1)*x^3 - 2*v*L*p*K*c*x^3 + v*L*p*K*c*y*x^3    + m*K*c*x^2 - (2-a*m*h-a*L)*K*c*y*x^2 + (m- 2*a*m*h-2*a*L)*K*c*y^2*x^2+ a*m*h*K*c*y^3*x^2 + a*L*K*c*y^3*x^2+ (v*L- a*L*p)*K*c*y^(-1)*x^3 + (v*m*h-a*L*p+v*L)*K*c*y*x^3 - 2*v*m*h*K*c*y^2*x^3 + v*m*h*K*c*y^3*x^3 + 2*a*L*p*K*c*x^3 - 2*v*L*K*c*x^3 - v*L*p*K*c*y^(-2)*x^4 + 2*v*L*p*K*c*y^(-1)*x^4 - v*L*p*K*c*x^4    - b*r*K*c*y^2*x + (b*r+f*r-b*r*f+b*a*r*h)*K*c*y^3*x + (b*a*r*h+f*a*r*h-b*f*a*r*h)*K*c*y^4*x + b*r*K*c*y*x^2 + (2*f*r+b*r-b*r*a*h+ b*v*r*h)*K*c*y^2*x^2 + (2*f*a*r*h+b*r*a*h-b*v*r*h-f*v*r*h+b*f*v*r*h)*K*c*y^3*x^2 - (f*r+b*r*f-b*v*r*h)*K*c*y*x^3 - (r*a*h*f+b*f+r*a*h+b*v*r*h)*K*c*y^2*x^3 - (f*v*r*h+b*v*r*h*f)*K*c*y*x^4 = 0
+$$0 = - brK \rho_1^{2}  + (br+2brK+rfK-brfK+barhK+brKc) \rho_1^{3}- (brKc-2br+frKc-brfKc+barhKc-brK-frK+brfK+farhK-bfarhK-fr+brf-barh) \rho_1^{4} + (br+fr-brf-barhK-farhK+bfarhK-farh+bfarh-barhKc-farhKc+bfarhKc) \rho_1^{5}+ (barh+farh-bfarh) \rho_1^{6}        + (mK + (brK-mK-mKc-2K+amhK+aLK) \rho_1 + (mK- br- 2amKh- 2aLK+2frK+brK-brahK+ bvrhK+2K-amhK-aLK- brK- 2brKc+ 2Kc-amhKc-aLKc) \rho_1^{2}   + (-mK-mKc+aLK-br++3amhK2farhK+2brahK-2bvrhK-fvrhK+bfvrhK-2fr+brah-bvrh-2frK+2aLK-2frKc+brahKc-bvrhKc+2amhKc+2aLKc+frKc-brfKc+barhKc) \rho_1^{3}-  (2farh-fvrh+bfvrh-amhK-aLK-2farhK-brahK+bvrhK+fvrhK-bfvrhK+2fr+br-amhKc-aLKc-bfarhKc-farhKc +bvrhKc+fvrhKc-bfvrhKc )\rho_1^{4} + (2farh + brah-bvrh-fvrh+bfvrh)\rho_1^{5}  )     \rho_{11}
 
--->
++ ((vLK- aLpK)\rho_1^{-1}+ 2aLpK  - 2vLK - vLK+ aLpK - vLKc+ aLpKc + mKc+ (vmhK+vLK-frK-brfK+bvrhK-3aLpK+2vLK-2aLpKc+2vLKc-2Kc+amhKc+aLKc+brKc) \rho_1  + (2frKc+frbrf-bvrh+brKc-brahKc+ bvrhKc- 2vmhK+frK+brfK-bvrhK- rahfK -bfrahK-bvrhK- vmhK+ aLpK- vLK + frKc+brfKc- bvrhKc +mKc- 2amhKc- 2aLKc-vmhKc+aLpKc- vLKc) \rho_1^{2} + (3farhKc + brahKc-fvrhKc+bfvrhKc+ vmhK+rahf +bfrah+ bvrh+ 2vmhK+rahfK +bfrahK+ bvrhK-fr -brf+bvrh +bfrahKc+ 2vmhKc+ amhKc + aLKc )\rho_1^{3} + (-vmhK-rahf-bfrah-bvrh- vmhKc) \rho_1^{4}    )   \rho_{11}^2
+
+- (+vLpK \rho_1^{-2} + (2vLpK + vLpK  + vLpKc+ vLKc- aLpKc)\rho_1^{-1} - vLpK - 2vLpK + 2aLpKc  - 2vLKc  - 2vLpKc - (fvrhK +bvrhfK+ vLpK+ vLpKc + vmhKc- aLpKc+ vLKc-frKc  -brfKc + bvrhKc )\rho_1 + (fvrh +bvrhf+fvrhK +bvrhfK+fvrhKc +bvrhfKc- 2vmhKc-rahfKc -bfrahKc- bvrhKc )\rho_1^2 - (fvrh +bvrhf+ vmhKc)\rho_1^3 )   \rho_{11}^3  - (vLpKc \rho_1^{-2} + 2vLpKc \rho_1^{-1} - vLpKc  - (fvrhKc +bvrhfKc)\rho_1 )     \rho_{11}^4    $$
+
+assuming competition $c = 0$ and attraction $v = 0$ yields a much simpler model
+
+$$0 = - brK \rho_1^{2}  + (br+2brK+rfK-brfK+barhK) \rho_1^{3}- (-2br-brK-frK+brfK+farhK-bfarhK-fr+brf-barh) \rho_1^{4} + (br+fr-brf-barhK-farhK+bfarhK-farh+bfarh) \rho_1^{5}+ (barh+farh-bfarh) \rho_1^{6}        + (mK + (brK-mK-2K+amhK+aLK) \rho_1 + (mK- br- 2amKh- 2aLK+2frK+brK-brahK+2K-amhK-aLK- brK) \rho_1^{2}   + (-mK+aLK-br+3amhK+2farhK+2brahK-2fr+brah-2frK+2aLK) \rho_1^{3}-  (2farh-amhK-aLK-2farhK-brahK+2fr+br)\rho_1^{4} + (2farh + brah)\rho_1^{5}  )     \rho_{11}  + (aLpK\rho_1^{-1}+ 2aLpK  + aLpK + (-frK-brfK-3aLpK+2vLK) \rho_1  + (frbrf+frK+brfK-rahfK-bfrahK+aLpK) \rho_1^{2} + (rahf +bfrah+rahfK +bfrahK-fr -brf)\rho_1^{3} + (-rahf-bfrah) \rho_1^{4}    )   \rho_{11}^2   $$

### 2015-02-20 documentation/model_formulation.Rmd
try to visualise attractor of pair-approximation model

@@
 
+### Attractor
+
+Visualising the attractor of the system is not as straightforward as for the models that are determined by global cover only, where the intersections of growth and mortality defined the equilibria.
+In the spatially-explict context the attractor can be visualised if we assume that clustering in the initial state is random, i.e.
+$q_{11} = \rho_{1}$.
+
+```{r, echo=FALSE}
+
+parms$rho_ini <- seq(0, 1, 0.1)
+parms$m_ini <- seq(0, 0.2, length = 11)
+
+iterations <- expand.grid(parms)
+iterations <- cbind(ID = 1:dim(iterations)[1],iterations)
+
+rho_one_one_ini <- function(prms) {
+  out <- sapply(1:nrow(prms), function(x) {
+  i <- with(prms[x,], v * L * p * rho_ini^(q-1) )
+  ii <- with(prms[x,], ( v*m*h*rho_ini - v*h*m_ini + p*a*L - v*L )*rho_ini^q )
+  iii <- with(prms[x,], -m_ini + m*rho_ini + a*h*m*rho_ini^(2+q) - a*h*m_ini*rho_ini^(1+q) - a*L*rho_ini^(1+q) )
+
+  solutions <- data.frame( 
+    s1 = (- ii + sqrt( ii^2 - 4*i*iii ) ) / 2 * i,
+    s2 = (- ii - sqrt( ii^2 - 4*i*iii ) ) / 2 * i)
+  return(solutions)
+  })
+  return(as.data.frame(t(out)))
+}
+
+iterations$rho_11_ini <- unlist(rho_one_one_ini(iterations)$s1)
+
+ini_rho <- function(rho_1, rho_11 = NULL) {
+    if(is.null(rho_11)) out <- {
+      c(
+    rho_1 = rho_1,
+    rho_11 = rho_1*rho_1,
+    rho_10 = rho_1*(1-rho_1)*2,
+    rho_00 = (1-rho_1)*(1-rho_1),
+    rho_0 = 1-rho_1
+    ) 
+    } else {
+      out <- c(
+    rho_1 = rho_1,
+    rho_11 = rho_11,
+    rho_10 = rho_1-rho_11,
+    rho_00 = 1-2*rho_1+rho_11,
+    rho_0 = 1-rho_1
+    )
+    if(any(out < 0)) out <- c(
+    rho_1 = NA,
+    rho_11 = NA,
+    rho_10 = NA,
+    rho_00 = NA,
+    rho_0 = NA
+    )
+    }
+
+  return(out)
+}
+
+starting <- sapply(1:nrow(iterations), function(x)  ini_rho(iterations[x, "rho_ini"], iterations[x, "rho_11_ini"])  )
+
+iterations$f[iterations$f < iterations$b] <- iterations$b[iterations$f < iterations$b]
+
+defplot(rho, C(rho, 1), ylab = "mortality")
+lines(rho,C(rho, 0.75))
+lines(rho,C(rho, 0.5))
+lines(rho,C(rho, 0.25))
+lines(rho,C(rho, 0))
+
+lines(rho, G(rho, 1), col = "green3")
+lines(rho,G(rho, 0.75), col = "green3")
+lines(rho,G(rho, 0.5), col = "green3")
+lines(rho,G(rho, 0.25), col = "green3")
+lines(rho,G(rho, 0), col = "green3")
+
+points(iterations$rho_ini, iterations$m_ini, pch = 20, col = is.na(starting[1,])+1)
+
+# provides parallel backend
+library(foreach)
+library(doSNOW)
+
+workstation <-  list(host = "162.38.184.118", user = "schneider",
+                     rscript = "/usr/lib/R/bin/Rscript",
+                     snowlib = "/usr/lib/R/library")
+
+workerlist <- c(rep("localhost", times = 10))
+
+cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')
+
+registerDoSNOW(cl)
+
+foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
+  source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")
+
+  model_parms <- as.list(iterations[iteration,])
+  rho_starting <- ini_rho(iterations[iteration, "rho_ini"], iterations[iteration, "rho_11_ini"])
+  if(!any(is.na(rho_starting)) ) {
+
+  # running the ode-solver
+  runmodel <- ode(y = rho_starting, func = odesys_spex, times = c(1,2), parms = model_parms)
+
+  out <-as.data.frame(runmodel)[dim(runmodel)[1],]
+  }  else { 
+    out <- c(time = 100,rho_starting)  
+  }
+  # transfer into ouput and calculate missing rho values
+  return(out)
+  } -> output
+
+output <- cbind(output, iterations)
+
+output$rho_11_ini <- iterations$rho_11_ini
+
+G <- function(rho_1, q_01 = "auto", prms = parms) {
+  if(q_01 == "auto") q_01 <- 1
+  with(prms, r*(b + (1-b)*f*q_01)*rho_1^(1+alpha)*(1-rho_1/(K * (1-c * q_01) ) ))
+
+}
+
+C <- function(rho_1, q_01 = 1, prms = parms) with(prms, (m * rho_1) +( (a+v*q_01  ) *rho_1^(1+q)*L*(1-p*q_01))/(1+(a+v*q_01)*h*(rho_1^(1+q)) ) )
+
+rho <- seq(0,1,length = 100)
+
+arrows(output$rho_ini, C(output$rho_ini, output$rho_11_ini),output$rho_1, C(output$rho_1, output$rho_11), length = 0.05)
+
+stopCluster(cl)
+
+```
+
 ### Bifurcation diagrams

### 2015-02-23 documentation/model_formulation.Rmd
add visual attractor analysis

--- 2015-02-04 documentation/model_formulation.Rmd
@@
             bty = "l", cex = 0.7, las = 1,
-            xlim = c(0,1), ylim = c(0,0.4), 
+            xlim = c(0,1), ylim = c(0,0.25), 
             xaxs = "i", yaxs = "i", 
-            xaxp = c(0,1,2), yaxp = c(0,1,2), 
+            xaxp = c(0,1,2), yaxp = c(0,0.25,2), 
             xlab = "vegetation cover")
@@
 
-odesys_spex <- function(t, rho, parms = model_parms) {
+odesys_spex <- function(t, rho, parms) {
   rho_1 <- d_rho_1(rho, parms)
@@
 plot(I(rho_11/rho_1^2) ~ time , data = ODE_spex , type = "l", ylim = c(0,2))
+
+```
+
+### Attractor
+
+Visualising the attractor of the system is not as straightforward as for the models that are determined by global cover only, where the intersections of growth and mortality defined the equilibria.
+In the spatially-explict pair-approximation model the attractor can be visualised if we assume that clustering in the initial state is non-random, and we assign starting values for $\rho_1$ and global mortality of the population $m_1$.
+To do this, we need to solve the mortality function towards $\rho{1\,,1}$ to obtain initial covers of all the different pairs and singletons.
+
+Thus,
+
+$$ m_1 = \rho_1 m_0 + \frac{(a+v\frac{\rho_{11}}{\rho_1})L(1-p\frac{\rho_{11}}{\rho_1}\rho_1^{1+q})}{1+(a+v\frac{\rho_{11}}{\rho_1})h\rho_1^{1+q}} $$
+
+rearranging yields:
+
+$$ 0 = m_1 - m_0\rho_1+m_qah\rho_1^{1+q}-aL\rho_1^{1+q}-m_0ah\rho_1^{1+q} + (m_1v\rho_1^{q}+aLp\rho_1^{q}-vL\rho_1^{q}-m_0v\rho_1^{1+q})\rho_{11} + vLp\rho_1^{q-1}\rho_{11}^2$$
+Polynomials of second order of the form $0 =  c+ bx +ax^2$ can be solved by the quadratic formula
+
+$$x = \frac{-b\pm \sqrt{b^2-4ac} }{2a}$$
+
+```{r, echo=FALSE}
+
+ini_rho <- function(rho_1, rho_11 = NULL) {
+    if(is.null(rho_11)) out <- {
+      c(
+    rho_1 = rho_1,
+    rho_11 = rho_1*rho_1,
+    rho_10 = rho_1*(1-rho_1)*2,
+    rho_00 = (1-rho_1)*(1-rho_1),
+    rho_0 = 1-rho_1
+    ) 
+    } else {
+      out <- c(
+    rho_1 = rho_1,
+    rho_11 = rho_11,
+    rho_10 = rho_1-rho_11,
+    rho_00 = 1-2*rho_1+rho_11,
+    rho_0 = 1-rho_1
+    )
+    if(any(out < 0)) out <- c(
+    rho_1 = NA,
+    rho_11 = NA,
+    rho_10 = NA,
+    rho_00 = NA,
+    rho_0 = NA
+    )
+    }
+
+  return(out)
+}
+
+G <- function(rho_1, q_01 = "auto", prms = parms) {
+  if(q_01 == "auto") q_01 <- 1
+  with(prms, r*(b + (1-b)*f*q_01)*rho_1^(1+alpha)*(1-rho_1/(K * (1-c * q_01) ) ))
+
+}
+
+C <- function(rho_1, q_11 = 1, prms = parms) with(prms, (m * rho_1) +( (a+v*q_11  ) *rho_1^(1+q)*L*(1-p*q_11))/(1+(a+v*q_11)*h*(rho_1^(1+q)) ) )
+
+attractor <- function(parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini = seq(0,1, length = 11)) {
+
+ini <- list(
+  rho_1 = rho_1_ini,
+  rho_11 = rho_11_ini,
+  rho_10 = NA,
+  rho_00 = NA,
+  rho_0 = NA
+  )
+
+ini <- expand.grid(ini)
+
+for(x in 1:nrow(ini)) {
+      temp <- ini_rho(ini[x,]$rho_1, ini[x,]$rho_11)
+      ini[x,]$rho_1 <- temp[1]
+      ini[x,]$rho_11 <- temp[2]
+      ini[x,]$rho_10 <- temp[3]
+      ini[x,]$rho_00 <- temp[4]
+      ini[x,]$rho_0 <- temp[5]
+      }
+
+ini$m_ini <- C(ini$rho_1, ini$rho_11/ini$rho_1)
+ini$g_ini <- G(ini$rho_1, (ini$rho_1-ini$rho_11)/(1-ini$rho_1))
+
+ini <- subset(ini, !is.na(m_ini))
+ini <- cbind(ID = 1:nrow(ini),ini)
+
+foreach(iteration = ini$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
+  source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")
+
+   rho_starting <- ini[iteration, 2:6]
+
+  # running the ode-solver
+  runmodel <- ode(y = as.numeric(rho_starting), func = odesys_spex, times = seq(1,2), parms = parms)
+
+  out <-as.data.frame(runmodel)[dim(runmodel)[1],]
+
+  # transfer into ouput and calculate missing rho values
+  return(out)
+  } -> output
+
+names(output) <- c("time", "rho_1", "rho_11", "rho_10", "rho_00", "rho_0")
+
+rho <- seq(0,1,length = 100)
+
+defplot(rho, C(rho, 1), ylab = "mortality", col= "#00000020")
+lines(rho,C(rho, 0.75), col= "#00000040")
+lines(rho,C(rho, 0.5), col= "#00000080")
+lines(rho,C(rho, 0.25), col= "#00000040")
+lines(rho,C(rho, 0), col= "#00000020")
+
+lines(rho, G(rho, 1), col = "#00993320")
+lines(rho,G(rho, 0.75), col = "#00993340")
+lines(rho,G(rho, 0.5), col = "#00993380")
+lines(rho,G(rho, 0.25), col = "#00993340")
+lines(rho,G(rho, 0), col = "#00993320")
+
+arrows(ini$rho_1, C(ini$rho_1, ini$rho_11/ini$rho_1),output$rho_1, C(output$rho_1, output$rho_11/output$rho_1), length = 0.02)
+
+arrows(ini$rho_1, G(ini$rho_1, (ini$rho_1-ini$rho_11)/(1-ini$rho_1)),output$rho_1, G(output$rho_1, (output$rho_1-output$rho_11)/(1-output$rho_1) ), length = 0.02, col = "#009933" )
+
+high_equ <- as.data.frame(ode(y = ini_rho(0.8), func = odesys_spex, times = c(1,1000), parms = parms) )[2,]
+
+points(high_equ$rho_1, C(high_equ$rho_1, high_equ$rho_11/high_equ$rho_1), pch = 20)
+
+low_equ <- as.data.frame(ode(y = ini_rho(0.0001), func = odesys_spex, times = c(1,1000), parms = parms) )[2,]
+
+points(low_equ$rho_1, C(low_equ$rho_1,low_equ$rho_11/low_equ$rho_1), pch = 20)
+
+}
+
+# provides parallel backend
+library(foreach)
+library(doSNOW)
+
+workstation <-  list(host = "162.38.184.118", user = "schneider",
+                     rscript = "/usr/lib/R/bin/Rscript",
+                     snowlib = "/usr/lib/R/library")
+
+workerlist <- c(rep("localhost", times = 10))
+
+cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')
+
+registerDoSNOW(cl)
+
+attractor(parms)
+
+```
+
+```{r, echo=FALSE}
+
+parms <- list(
+    m = 0.05,  # intrinsic mortality of plants 
+    r = 1,   # growth rate
+    f = 0.8, #  local facilitation
+    b = 0.25,  # environmental quality inverse of aridity
+    K = 0.9,  # carrying capacity
+    alpha = 0.1, # water runoff
+    c = 0.2, # local competition
+    a = 0.6, # search efficiency
+    v = 0.1, # attraction effect
+    h = 50, # handling time (one individual on landscape unit)
+    p = 0.8, # protection effect
+    L = 6, # livestock units per landscape unit 
+    q = 1 # search ineffiency at low cover
+  )
+
+attractor(parms)
+
+stopCluster(cl)
 

### 2015-02-23 documentation/interactive_spexmodel_attractor.Rmd
interactive simulation of attractor (Shiny)

--- 2015-02-03 documentation/interactive_spexmodel_bifurcations.Rmd
@@
 date: "Thursday, January 30, 2015"
-output: html_document
+output:
+  html_document:
+    fig_height: 6.5
 runtime: shiny
@@
                ) )
+}
+
+defplot <- function(...) plot(..., 
+            type = "l", 
+            bty = "l", cex = 0.7, las = 1,
+            xlim = c(0,1), ylim = c(0,0.25), 
+            xaxs = "i", yaxs = "i", 
+            xaxp = c(0,1,5), yaxp = c(0,0.25,5), 
+            xlab = "vegetation cover")
+
+rho <- seq(0,1,length = 100)
+
+defaultparms <- list(
+    m = 0.05,  # intrinsic mortality of plants 
+    r = 1,   # growth rate
+    f = 0.8, #  local facilitation
+    b = 0.25,  # environmental quality inverse of aridity
+    K = 0.9,  # carrying capacity
+    alpha = 0.1, # water runoff
+    c = 0.2, # local competition
+    a = 0.6, # search efficiency
+    v = 0.1, # attraction effect
+    h = 50, # handling time (one individual on landscape unit)
+    p = 0.8, # protection effect
+    L = 6, # livestock units per landscape unit 
+    q = 1 # search ineffiency at low cover
+  )
+
+ini_rho <- function(rho_1, rho_11 = NULL) {
+    if(is.null(rho_11)) out <- {
+      c(
+    rho_1 = rho_1,
+    rho_11 = rho_1*rho_1,
+    rho_10 = rho_1*(1-rho_1)*2,
+    rho_00 = (1-rho_1)*(1-rho_1),
+    rho_0 = 1-rho_1
+    ) 
+    } else {
+      out <- c(
+    rho_1 = rho_1,
+    rho_11 = rho_11,
+    rho_10 = rho_1-rho_11,
+    rho_00 = 1-2*rho_1+rho_11,
+    rho_0 = 1-rho_1
+    )
+    if(any(out < 0)) out <- c(
+    rho_1 = NA,
+    rho_11 = NA,
+    rho_10 = NA,
+    rho_00 = NA,
+    rho_0 = NA
+    )
+    }
+
+  return(out)
+}
+
+G <- function(rho_1, q_01 = "auto", prms ) {
+  if(q_01 == "auto") q_01 <- 1
+  with(prms, r*(b + (1-b)*f*q_01)*rho_1^(1+alpha)*(1-rho_1/(K * (1-c * q_01) ) ))
+
+}
+
+C <- function(rho_1, q_11 = 1, prms ) with(prms, (m * rho_1) +( (a+v*q_11  ) *rho_1^(1+q)*L*(1-p*q_11))/(1+(a+v*q_11)*h*(rho_1^(1+q)) ) )
+
+attractor <- function(parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini = seq(0,1, length = 11)) {
+
+ini <- list(
+  rho_1 = rho_1_ini,
+  rho_11 = rho_11_ini,
+  rho_10 = NA,
+  rho_00 = NA,
+  rho_0 = NA
+  )
+
+ini <- expand.grid(ini)
+
+for(x in 1:nrow(ini)) {
+      temp <- ini_rho(ini[x,]$rho_1, ini[x,]$rho_11)
+      ini[x,]$rho_1 <- temp[1]
+      ini[x,]$rho_11 <- temp[2]
+      ini[x,]$rho_10 <- temp[3]
+      ini[x,]$rho_00 <- temp[4]
+      ini[x,]$rho_0 <- temp[5]
+      }
+
+ini$m_ini <- C(ini$rho_1, ini$rho_11/ini$rho_1, parms)
+ini$g_ini <- G(ini$rho_1, (ini$rho_1-ini$rho_11)/(1-ini$rho_1), parms)
+
+ini <- subset(ini, !is.na(rho_1))
+ini <- cbind(ID = 1:nrow(ini),ini)
+
+foreach(iteration = ini$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
+  source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")
+
+   rho_starting <- ini[iteration, 2:6]
+
+  # running the ode-solver
+  runmodel <- ode(y = as.numeric(rho_starting), func = odesys_spex, times = seq(1,2), parms = parms)
+
+  out <-as.data.frame(runmodel)[dim(runmodel)[1],]
+
+  # transfer into ouput and calculate missing rho values
+  return(out)
+  } -> output
+
+names(output) <- c("time", "rho_1", "rho_11", "rho_10", "rho_00", "rho_0")
+
+rho <- seq(0,1,length = 100)
+
+defplot(rho, C(rho, 1, parms), ylab = "mortality", col= "#00000020")
+lines(rho,C(rho, 0.75, parms), col= "#00000040")
+lines(rho,C(rho, 0.5, parms), col= "#00000080")
+lines(rho,C(rho, 0.25, parms), col= "#00000040")
+lines(rho,C(rho, 0, parms), col= "#00000020")
+
+lines(rho, G(rho, 1, parms), col = "#00993320")
+lines(rho,G(rho, 0.75, parms), col = "#00993340")
+lines(rho,G(rho, 0.5, parms), col = "#00993380")
+lines(rho,G(rho, 0.25, parms), col = "#00993340")
+lines(rho,G(rho, 0, parms), col = "#00993320")
+
+arrows(ini$rho_1, C(ini$rho_1, ini$rho_11/ini$rho_1, parms),output$rho_1, C(output$rho_1, output$rho_11/output$rho_1, parms), length = 0.04)
+
+arrows(ini$rho_1, G(ini$rho_1, (ini$rho_1-ini$rho_11)/(1-ini$rho_1), parms),output$rho_1, G(output$rho_1, (output$rho_1-output$rho_11)/(1-output$rho_1), parms ), length = 0.02, col = "#009933" )
+
+high_equ <- as.data.frame(ode(y = ini_rho(0.9), func = odesys_spex, times = c(1,1000), parms = parms) )[2,]
+
+points(high_equ$rho_1, C(high_equ$rho_1, high_equ$rho_11/high_equ$rho_1, parms), cex = 2, pch = 20, col = 2)
+
+low_equ <- as.data.frame(ode(y = ini_rho(0.0001), func = odesys_spex, times = c(1,1000), parms = parms) )[2,]
+
+points(low_equ$rho_1, C(low_equ$rho_1,low_equ$rho_11/low_equ$rho_1, parms), pch = 20, cex = 2, col = 2)
+
 }
@@
 ### initial conditions
-```{r, echo=FALSE}
-
-inputPanel(
-  selectInput("gradient", label = "select gradient:",             choices = c("b", "L"), selected = "b")
-  )
+```{r, echo=FALSE, fig.height=7.5}
 
@@
 
- if(input$gradient == "b") parms$b <- seq(0,1,.005) 
- if(input$gradient == "L") parms$L <- seq(0,20,.1)
+# if(input$gradient == "b") parms$b <- seq(0,1,.005) 
+# if(input$gradient == "L") parms$L <- seq(0,20,.1)
 
-parms$rho_ini <- c(0.9, 0.01)
+#iterations$f[iterations$f < iterations$b] <- iterations$b[iterations$f < iterations$b]
 
-iterations <- expand.grid(parms)
-iterations <- cbind(ID = 1:dim(iterations)[1],iterations)
+attractor(parms)
 
-iterations$f[iterations$f < iterations$b] <- iterations$b[iterations$f < iterations$b]
-
-foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
-  source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")
-
-  model_parms <- as.list(iterations[iteration,])
-
-  # running the ode-solver
-  runmodel <- ode(y = ini_rho(model_parms$rho_ini), func = odesys_spex, times = c(1,1000), parms = model_parms)
-
-  # transfer into ouput and calculate missing rho values
-  return(as.data.frame(runmodel)[dim(runmodel)[1],])
-  } -> output
-
-output <- cbind(output, iterations)
-
- if(input$gradient == "b") plot(rho_1 ~ b, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
- if(input$gradient == "L") plot(rho_1 ~ L, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
+# if(input$gradient == "b") plot(rho_1 ~ b, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
+# if(input$gradient == "L") plot(rho_1 ~ L, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
 

### 2015-02-26 documentation/model_formulation.Rmd
update model formulation

@@
 
-Mechanistic explanations can be the limitations by a globally diffuse resource, like water, or the occupation of space and the subsequent limitation of photosynthetic active surface.
+Mechanistic explanations can be the limitations by a globally diffuse resource, like water. 
+(We do not discuss the occupation of space as a global resource and the subsequent limitation of photosynthetic active surface, since this is implemented by adding explicit space to the model.)
 
@@
 
-$$ c_11 = q_{1|1}/\rho_1 = \frac{\rho_11}{\rho_1^2}$$
+$$ c_{11} = q_{1|1}/\rho_1 = \frac{\rho_{11}}{\rho_1^2}$$
 
@@
 $$x = \frac{-b\pm \sqrt{b^2-4ac} }{2a}$$
+
+**apparently this solution is wrong!! Instead, I draw  initial $rho_11$ values systematically and elimitate those that produce invalid $m_ini$**
+
+This works quite well, but produces unequal distributions of arrows along the y-axis.
 
@@
 
-The following shows bifurcation diagrams over an environmental gradient.
-
-```{r, echo=FALSE}
+The following shows bifurcation diagrams over an environmental gradient, $b$, as well as over a gradient of livestock density, $L$.
+
+```{r, echo=FALSE, fig.height = 3}
 
@@
 
+par(mfrow= c(1,2), bty = "n", mar = c(4,4,1,0))
+
 foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
@@
 
-plot(rho_1 ~ b, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
-
-```
-
-The following shows bifurcation diagrams over a livestock density gradient.
-
-```{r, echo=FALSE}
+plot(rho_1 ~ b, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1), cex = 0.5)
 
@@
 
-plot(rho_1 ~ L, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
+plot(rho_1 ~ L, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1), cex = 0.5)
 

### 2015-02-26 code/simfunctions.r
complete simfunctions.r collection

@@ -354,16 +354,6 @@ runCA <- function(init, parms, width = 100, height = 100, delta = 0.1, t_max = 1
 }
 
 
-ini_rho <- function(rho_1) {
-  c(
-    rho_1 = rho_1,
-    rho_11 = rho_1*rho_1,
-    rho_10 = rho_1*(1-rho_1)*2,
-    rho_00 = (1-rho_1)*(1-rho_1),
-    rho_0 = 1-rho_1
-  )
-  
-}
 
 d_rho_1 <- function(rho, parms) { 
   with(parms,
@@ -385,3 +375,238 @@ odesys_spex <- function(t, rho, parms = model_parms) {
   ) )
 }
 
+# function to get initial rho values for pair-approximation
+
+ini_rho <- function(rho_1, rho_11 = NULL) {
+  if(is.null(rho_11[1])) out <- {
+    c(
+      rho_1 = rho_1,
+      rho_11 = rho_1*rho_1,
+      rho_10 = rho_1*(1-rho_1)*2,
+      rho_00 = (1-rho_1)*(1-rho_1),
+      rho_0 = 1-rho_1
+    ) 
+  } else {
+    out <- c(
+      rho_1 = rho_1,
+      rho_11 = rho_11,
+      rho_10 = rho_1-rho_11,
+      rho_00 = 1-2*rho_1+rho_11,
+      rho_0 = 1-rho_1
+    )
+    if(any(out < 0)) out <- c(
+      rho_1 = NA,
+      rho_11 = NA,
+      rho_10 = NA,
+      rho_00 = NA,
+      rho_0 = NA
+    )
+  }
+  
+  return(out)
+}
+
+
+# function for local growth
+# takes global cover, rho_1 and local cover, q_01, as parameters, plus a list object with model parameters. 
+
+G <- function(rho_1, q_01 = "auto", prms = parms, set = list(NA)) {
+  
+  if(q_01[1] == "auto") q_01 <- 1
+  
+  
+  prms.names <- names(prms)
+  set.names <- names(set)
+  m.names <- sort(unique(c(prms.names, set.names)))
+  
+  prms_temp <- sapply(m.names, function(i) {
+    if (i %in% set.names) set[[i]]
+    else prms[[i]]
+  }, simplify = FALSE)
+  
+  with(prms_temp, r*(b + (1-b)*f*q_01)*rho_1^(1+alpha)*(1-rho_1/(K * (1-c * q_01) ) ))
+  
+} 
+
+
+# function for local grazing mortality
[…]

### 2015-02-26 documentation/model_exploration.Rmd
init model exploration document

---
title: "Model exploration"
author: "Florian D. Schneider"
date: "Tuesday, February 25th, 2014"
output:
  pdf_document:
    fig_height: 3
    fig_width: 7
  html_document:
    keep_md: yess
fontsize: 12pt
geometry: margin=1.4in
csl: ../manuscript/amnat.csl
bibliography: ../manuscript/cas02.bib
---

```{r, echo=FALSE}
# ODE solving
library("deSolve")

source("../code/simfunctions.r")

parallel = TRUE
if(parallel) {
# provides parallel backend
library(foreach)
library(doSNOW)

workstation <-  list(host = "162.38.184.118", user = "schneider",
                     rscript = "/usr/lib/R/bin/Rscript",
                     snowlib = "/usr/lib/R/library")

workerlist <- c(rep("localhost", times = 10))

cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')

registerDoSNOW(cl)

}

# default parameters, global model without spatially-explicit effects

parms <- list(
    m = 0.05,
    r = 1,
    b = 0.9, 
    K = 0.9, 
    a = 0.6,
    h = 50,
    L = 4, 
    alpha = 0,
    q = 0,
    c = 0,
    f = 0,
    v = 0,
    p = 0
  )

```

# Introduction

The paper departs from a general perception of grazing systems as non-equilibrium systems that are subject to 'dynamic regimes' or 'alternative stable states'.
This means, they can shift their attractor between a vegetated state and a degraded state, if they experience a sufficient perturbation.
The question is how resilience thinking in managed rangelands can be improved [@folke06; @walker04; @walker12; @folke10]. How can the ecosystem be put in a state that is far from the tipping point or the critical thresholds [@may75; @westoby89; @walker81]?

It aims at an integration of spatial grazing mortality into a graphical population-dynamics framework as developed by @noy-meir75 and extended by @vandekoppel97 and @rietkerk97.
The model is of conceptual character and identifies how certain management practices or characteristics of vegetation affect the thresholds of catastrophic shifts. This kind of model can be used to inform state-and-transition models that are of high relevance in decisionmaking in the management of rangelands [@westoby89; @briske05; @bestelmeyer03; @briske08].

# alternative stable states due to feedbacks

Ecological systems with non-linear stability are common in nature [@carpenter03; @scheffer01; @gunderson02; @berkes03] and this has enormous impacts on how we analyse and manage those systems. 
As early as 40 years ago, rangelands were identified as a system with non-linear stability, i.e. with alternative stable states [@noy-meir75], from graphical analysis as a simple consumer-resource system.

```{r, echo=FALSE, fig.height = 6}
colpal <- list(mort = c("#000000","#00000060","#00000040","#00000020"), grow = c("#009933","#00993360","#00993340","#00993320") )
rho <- seq(0,1,length = 100)

par(mfrow = c(2,2), bty ="n", las = 1 ) 
defplot(rho, G(rho), ylab = "plant mortality/growth", col = colpal$grow[1])
lines(rho, C(rho, set = list(L = 2)), col = colpal$mort[1] )
lines(rho, C(rho, set = list(L = 5)), col = colpal$mort[1] )
lines(rho, C(rho, set = list(L = 8)), col = colpal$mort[1] )

bifurcation(parms, "L", c(0,12))

defplot(rho, C(rho), ylab = "plant mortality/growth", col = colpal$mort[1])
lines(rho, G(rho, set = list(b = 0.2) ), col = colpal$grow[1] )
lines(rho, G(rho, set = list(b = 0.5)), col = colpal$grow[1] )
lines(rho, G(rho, set = list(b = 0.8)), col = colpal$grow[1] )

bifurcation(parms, "b", c(0,1))

```

## spatially-explicit mechanisms

We now add a spatially explicit feedback to the model.

### facilitation.

```{r, echo = FALSE}

par(mfrow = c(1,2), bty ="n", las = 1 )

parms$f = 0.9
parms$b = 0.4

attractor(parms)

bifurcation(parms, "b", c(0,1))

```

### local competition

```{r, echo = FALSE}

par(mfrow = c(1,2), bty ="n", las = 1 )

parms$f = 0
parms$c = 0.3
parms$b = 0.7

attractor(parms)

bifurcation(parms, "b", c(0,1))

```

### associational resistance

```{r, echo = FALSE}

par(mfrow = c(1,2), bty ="n", las = 1 ) 
parms$c = 0
parms$v = 0
parms$L = 8
parms$p = 0.8
parms$b = 0.7

attractor(parms)

bifurcation(parms, "L", c(0,12))

```

### attractant-decoy

```{r, echo = FALSE}

par(mfrow = c(1,2), bty ="n", las = 1 )

parms$p = 0
parms$L = 4
parms$q = 0.4
parms$v = 0.9

attractor(parms)

bifurcation(parms, "L", c(0,12))

```

```{r, echo=FALSE}
if(parallel) {
stopCluster(cl)
}
```

# References

### 2015-02-27 documentation/model_formulation.Rmd
update model formulation, rely on simfunctions.r; add mean field model

@@
 
-parms <- list(
+defparms <- list(
     m = 0.05,
@@
   )
-
+parms <- defparms
 par(mar = c(4,4,2,4))
@@
   )
+parms <- defparms
 
@@
 
-G <- function(V, q_01 = "auto", r_0 = 1, b = 0.1, f = 0, alpha = 0,  K = 0.9, c = 0) {
-  if(q_01 == "auto") q_01 <- 1
-  r_0*(b + (1-b)*f*q_01)*V^(1+alpha)*(1-V/(K * (1-c * q_01) ) )
-
-}
-
-defplot(rho, G(rho, 1, f = 0.8), ylab = "growth", col = "green3")
-lines(rho,G(rho, 0.75, f = 0.8), col = "green3")
-lines(rho,G(rho, 0.5, f = 0.8), col = "green3")
-lines(rho,G(rho, 0.25, f = 0.8), col = "green3")
-lines(rho,G(rho, 0, f = 0.8), col = "green3")
-
-lines(rho, C(rho, L = 10))
+defplot(rho, G(rho, 1, set = list( f = 0.8)), ylab = "growth", col = "green3")
+lines(rho,G(rho, 0.75, set = list( f = 0.8)), col = "green3")
+lines(rho,G(rho, 0.5, set = list( f = 0.8)), col = "green3")
+lines(rho,G(rho, 0.25, set = list( f = 0.8)), col = "green3")
+lines(rho,G(rho, 0, set = list( f = 0.8)), col = "green3")
+
+lines(rho, C(rho, set = list( L = 10)))
 
@@
 
-defplot(rho, G(rho, 1, c = 0.3, b = 0.9), ylab = "growth", col = "green3")
-lines(rho,G(rho, 0.75, c = 0.3, b = 0.9), col = "green3")
-lines(rho,G(rho, 0.5, c = 0.3, b = 0.9), col = "green3")
-lines(rho,G(rho, 0.25, c = 0.3, b = 0.9), col = "green3")
-lines(rho,G(rho, 0, c = 0.3, b = 0.9), col = "green3")
-
-lines(rho, C(rho, L = 10))
+defplot(rho, G(rho, 1, set= list(c = 0.3, b = 0.9)), ylab = "growth", col = "green3")
+lines(rho,G(rho, 0.75, set= list( c = 0.3, b = 0.9)), col = "green3")
+lines(rho,G(rho, 0.5, set= list( c = 0.3, b = 0.9)), col = "green3")
+lines(rho,G(rho, 0.25, set= list(c = 0.3, b = 0.9)), col = "green3")
+lines(rho,G(rho, 0, set= list( c = 0.3, b = 0.9)), col = "green3")
+
+lines(rho, C(rho, set= list( L = 10)))
 
@@
 
-C <- function(V, q_01 = 1, m_0 = 0.05, L = 5, a = 0.1, h = 100, q = 0, p = 0, v = 0) (m_0 * V) +( (a+v*q_01  ) *V^(1+q)*L*(1-p*q_01))/(1+(a+v*q_01)*h*(V^(1+q)) )
-
-defplot(rho, C(rho, 1, v = 0.5), ylab = "mortality")
-lines(rho,C(rho, 0.75, v = 0.5))
-lines(rho,C(rho, 0.5, v = 0.5))
-lines(rho,C(rho, 0.25, v = 0.5))
-lines(rho,C(rho, 0, v = 0.5))
-
-lines(rho, G(rho, b = 0.6), col = "green3")
+defplot(rho, C(rho, 1, set = list( v = 0.5)), ylab = "mortality")
+lines(rho,C(rho, 0.75, set = list( v = 0.5)))
+lines(rho,C(rho, 0.5, set = list( v = 0.5)))
+lines(rho,C(rho, 0.25, set = list( v = 0.5)))
+lines(rho,C(rho, 0, set = list( v = 0.5)))
+
+lines(rho, G(rho, set = list( b = 0.6)), col = "green3")
 
@@
 
-defplot(rho, C(rho, 1, p = 0.8, L = 10), ylab = "mortality")
-lines(rho,C(rho, 0.75, p = 0.8, L = 10))
-lines(rho,C(rho, 0.5, p = 0.8, L = 10))
-lines(rho,C(rho, 0.25, p = 0.8, L = 10))
-lines(rho,C(rho, 0, p = 0.8, L = 10))
-
-lines(rho, G(rho, b = 0.6), col = "green3")
+defplot(rho, C(rho, 1, set= list(p = 0.8, L = 10)), ylab = "mortality")
+lines(rho,C(rho, 0.75, set= list(p = 0.8, L = 10)))
+lines(rho,C(rho, 0.5, set= list(p = 0.8, L = 10)))
+lines(rho,C(rho, 0.25, set= list(p = 0.8, L = 10)))
+lines(rho,C(rho, 0, set= list(p = 0.8, L = 10)))
+
+lines(rho, G(rho, set= list(b = 0.6)), col = "green3")
 
@@
 
-defplot(rho, C(rho, 1, p = 0.8, v = 0.6, L = 10, q = .5), ylab = "mortality")
-lines(rho,C(rho, 0.75, p = 0.8, v = 0.6, L = 10, q = .5))
-lines(rho,C(rho, 0.5, p = 0.8,v = 0.6,  L = 10, q = .5))
-lines(rho,C(rho, 0.25, p = 0.8,v = 0.6,  L = 10, q = .5))
-lines(rho,C(rho, 0, p = 0.8,v = 0.6,  L = 10, q = .5))
-
-lines(rho, G(rho, 1, c = 0.3, f = 0.9, b = 0.3, alpha = 0.5), col = "green3")
-lines(rho,G(rho, 0.75, c = 0.3,f = 0.9,  b = 0.3, alpha = 0.5), col = "green3")
-lines(rho,G(rho, 0.5, c = 0.3,f = 0.9,  b = 0.3, alpha = 0.5), col = "green3")
-lines(rho,G(rho, 0.25, c = 0.3, f = 0.9, b = 0.3, alpha = 0.5), col = "green3")
-lines(rho,G(rho, 0, c = 0.3,f = 0.9,  b = 0.3, alpha = 0.5), col = "green3")
+defplot(rho, C(rho, 1, set = list( p = 0.8, v = 0.6, L = 10, q = .5)), ylab = "mortality")
+lines(rho,C(rho, 0.75, set = list( p = 0.8, v = 0.6, L = 10, q = .5)))
+lines(rho,C(rho, 0.5, set = list( p = 0.8,v = 0.6,  L = 10, q = .5)))
+lines(rho,C(rho, 0.25, set = list( p = 0.8,v = 0.6,  L = 10, q = .5)))
+lines(rho,C(rho, 0, set = list( p = 0.8,v = 0.6,  L = 10, q = .5)))
+
+lines(rho, G(rho, 1, set = list(c = 0.3, f = 0.9, b = 0.3, alpha = 0.5)), col = "green3")
+lines(rho,G(rho, 0.75, set = list( c = 0.3,f = 0.9,  b = 0.3, alpha = 0.5)), col = "green3")
+lines(rho,G(rho, 0.5, set = list(c = 0.3,f = 0.9,  b = 0.3, alpha = 0.5)), col = "green3")
+lines(rho,G(rho, 0.25, set = list( c = 0.3, f = 0.9, b = 0.3, alpha = 0.5)), col = "green3")
+lines(rho,G(rho, 0, set = list(c = 0.3,f = 0.9,  b = 0.3, alpha = 0.5)), col = "green3")
 
@@
 
+### mean field model
+
+As a simple approximation of the model serves a mean field model, assuming that the conditional local vegetation cover of empty cells $q_{0|1}$ and of vegetated cells $q_{1|1}$ both are well represented by the global vegetation cover $\rho_q$.
+Of course $\rho_0 = 1-\rho_1$.
+The change in cover derives from the growth and mortality terms
+
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = \rho_{0}w_{0,1} - \rho_{1}w_{1,0}   $$
+
+Thus, we can substitute $w_{1,0}$ and $w_{0,1}$ and obtain
+
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b + (1 - b)  f \rho_1)) (1-\frac{\rho_1}{K(1 - c \rho_1}) - \rho_{1}m - \frac{(a  + v \rho_1)L (1 - p \rho_1)\rho_1^{1+q}}{1+(a  + v \rho_1)h\rho_1^{1+q}}  $$
+
+Plotting the meanfield growth and mortality over vegetation cover is straigthforward now.
+
+$$ G = r \rho_1^{1+\alpha} (b + (1 - b)  f \rho_1)) (1-\frac{\rho_1}{K(1 - c \rho_1}) $$
+
+$$ C = \rho_{1}m + \frac{(a  + v \rho_1)L (1 - p \rho_1)\rho_1^{1+q}}{1+(a  + v \rho_1)h\rho_1^{1+q}}$$
+
+The steady states can be simulated by solving the ODE.
+
+```{r, echo=FALSE}
+
+parms <- defparms
+parms$f = 0.9
+parms$b = 0.4
+parms$L = 5
+
+attractor(parms, pairapprox = FALSE, localvals = TRUE)
+
+```
+
+But the steady state also can be solved analytically.
+
+Therefore, we solve for the equilibrium case
+
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = 0 $$
+
+$$  0 = r \rho_1^{1+\alpha} (b + (1 - b)  f \rho_1)) (1-\frac{\rho_1}{K(1 - c \rho_1}) - \rho_{1}m - \frac{(a  + v \rho_1)L (1 - p \rho_1)\rho_1^{1+q}}{1+(a  + v \rho_1)h\rho_1^{1+q}}  $$
+
 ### implementation in a pair-approximation model
@@
 ```{r, echo=FALSE}
-
-ini_rho <- function(rho_1, rho_11 = NULL) {
-    if(is.null(rho_11)) out <- {
-      c(
-    rho_1 = rho_1,
-    rho_11 = rho_1*rho_1,
-    rho_10 = rho_1*(1-rho_1)*2,
-    rho_00 = (1-rho_1)*(1-rho_1),
-    rho_0 = 1-rho_1
-    ) 
-    } else {
-      out <- c(
-    rho_1 = rho_1,
-    rho_11 = rho_11,
-    rho_10 = rho_1-rho_11,
-    rho_00 = 1-2*rho_1+rho_11,
-    rho_0 = 1-rho_1
-    )
-    if(any(out < 0)) out <- c(
-    rho_1 = NA,
-    rho_11 = NA,
-    rho_10 = NA,
-    rho_00 = NA,
-    rho_0 = NA
-    )
-    }
-
-  return(out)
-}
-
-G <- function(rho_1, q_01 = "auto", prms = parms) {
-  if(q_01 == "auto") q_01 <- 1
-  with(prms, r*(b + (1-b)*f*q_01)*rho_1^(1+alpha)*(1-rho_1/(K * (1-c * q_01) ) ))
-
-}
-
-C <- function(rho_1, q_11 = 1, prms = parms) with(prms, (m * rho_1) +( (a+v*q_11  ) *rho_1^(1+q)*L*(1-p*q_11))/(1+(a+v*q_11)*h*(rho_1^(1+q)) ) )
-
-attractor <- function(parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini = seq(0,1, length = 11)) {
-
-ini <- list(
-  rho_1 = rho_1_ini,
-  rho_11 = rho_11_ini,
-  rho_10 = NA,
-  rho_00 = NA,
-  rho_0 = NA
-  )
-
-ini <- expand.grid(ini)
-
-for(x in 1:nrow(ini)) {
-      temp <- ini_rho(ini[x,]$rho_1, ini[x,]$rho_11)
-      ini[x,]$rho_1 <- temp[1]
-      ini[x,]$rho_11 <- temp[2]
-      ini[x,]$rho_10 <- temp[3]
-      ini[x,]$rho_00 <- temp[4]
-      ini[x,]$rho_0 <- temp[5]
-      }
-
-ini$m_ini <- C(ini$rho_1, ini$rho_11/ini$rho_1)
-ini$g_ini <- G(ini$rho_1, (ini$rho_1-ini$rho_11)/(1-ini$rho_1))
-
-ini <- subset(ini, !is.na(m_ini))
-ini <- cbind(ID = 1:nrow(ini),ini)
-
-foreach(iteration = ini$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
-  source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")
-
-   rho_starting <- ini[iteration, 2:6]
-
-  # running the ode-solver
-  runmodel <- ode(y = as.numeric(rho_starting), func = odesys_spex, times = seq(1,2), parms = parms)
-
-  out <-as.data.frame(runmodel)[dim(runmodel)[1],]
-
-  # transfer into ouput and calculate missing rho values
-  return(out)
-  } -> output
-
-names(output) <- c("time", "rho_1", "rho_11", "rho_10", "rho_00", "rho_0")
-
-rho <- seq(0,1,length = 100)
-
-defplot(rho, C(rho, 1), ylab = "mortality", col= "#00000020")
-lines(rho,C(rho, 0.75), col= "#00000040")
-lines(rho,C(rho, 0.5), col= "#00000080")
-lines(rho,C(rho, 0.25), col= "#00000040")
-lines(rho,C(rho, 0), col= "#00000020")
-
-lines(rho, G(rho, 1), col = "#00993320")
-lines(rho,G(rho, 0.75), col = "#00993340")
-lines(rho,G(rho, 0.5), col = "#00993380")
-lines(rho,G(rho, 0.25), col = "#00993340")
-lines(rho,G(rho, 0), col = "#00993320")
-
-arrows(ini$rho_1, C(ini$rho_1, ini$rho_11/ini$rho_1),output$rho_1, C(output$rho_1, output$rho_11/output$rho_1), length = 0.02)
-
-arrows(ini$rho_1, G(ini$rho_1, (ini$rho_1-ini$rho_11)/(1-ini$rho_1)),output$rho_1, G(output$rho_1, (output$rho_1-output$rho_11)/(1-output$rho_1) ), length = 0.02, col = "#009933" )
-
-high_equ <- as.data.frame(ode(y = ini_rho(0.8), func = odesys_spex, times = c(1,1000), parms = parms) )[2,]
-
-points(high_equ$rho_1, C(high_equ$rho_1, high_equ$rho_11/high_equ$rho_1), pch = 20)
-
-low_equ <- as.data.frame(ode(y = ini_rho(0.0001), func = odesys_spex, times = c(1,1000), parms = parms) )[2,]
-
-points(low_equ$rho_1, C(low_equ$rho_1,low_equ$rho_11/low_equ$rho_1), pch = 20)
-
-}
 

### 2015-02-27 documentation/model_exploration.Rmd
update model exploration. Add meanfield comparison, rely on simfunctions.r

@@
     fig_width: 7
-  html_document:
-    keep_md: yess
 fontsize: 12pt
@@
 
-parms <- list(
+defparms <- list(
     m = 0.05,
     r = 1,
-    b = 0.9, 
+    b = 0.7, 
     K = 0.9, 
@@
 ```{r, echo=FALSE, fig.height = 6}
-colpal <- list(mort = c("#000000","#00000060","#00000040","#00000020"), grow = c("#009933","#00993360","#00993340","#00993320") )
+colpal <- list(mort = c("#000000","#00000040","#00000030","#00000020"), grow = c("#009933","#00993340","#00993330","#00993320") )
 rho <- seq(0,1,length = 100)
+parms <- defparms
 
@@
 
+The model is neutral in terms of species and assumes only one relevant type of vegetation and one population of grazer.
+Noy-meir notes that the model can apply also if there is one plant species that dominates the ecological dynamics of the entire system.
+[@noy-meir75]
+We stick to this approach of neglecting species-specific differences.
+
 ## spatially-explicit mechanisms
 
-We now add a spatially explicit feedback to the model.
+The spatial extend at which ecological feedback mechanisms act often differ between small local scales and the entire landscape.
+Previous models have shown that feedbacks acting simultaneously but on different spatial scales emerge and alter the stability properties of the ecosystem [@kefi07a].
+Therefore, we are going to implement different kinds of local feedback, complementary to the global feedbacks described above, into the model framework.
 
-### facilitation.
+The graphical approach pursued by @noy-meir75 and @vandekoppel97 is limited by it's dimensionality.
+The functions of vegetation growth and mortality are well defined by only one parameter, vegetation cover.
+Especially in arid and semi-arid ecosystems vegetation cover turned out to be insufficient to describe the state of the ecosystem.
+
+```{r, echo=FALSE}
+
+par(mfrow = c(1,2), bty ="n", las = 1 )
+
+p1 <- defparms
+p1$f = 0.9
+p1$b = 0.4
+
+attractor(p1, pairapprox = FALSE, localvals = TRUE)
+
+attractor(p1)
+
+```
+
+### facilitation
+
+The mechanism of facilitation, i.e.
+the positive effect of one individual or species onto another, is well described in terms of species interaction networks [@bruno09; @kefi12].
+Such mechanisms can be species-pair specific but they also play a general abiotic role as the cause of local amelioration of the environmental conditions of the neighborhood of a 'nurse' plant.
 
@@
 par(mfrow = c(1,2), bty ="n", las = 1 ) 
-
+parms <- defparms
 parms$f = 0.9
@@
 par(mfrow = c(1,2), bty ="n", las = 1 ) 
+parms <- defparms
 
-parms$f = 0
 parms$c = 0.3
-parms$b = 0.7
 
@@
 par(mfrow = c(1,2), bty ="n", las = 1 ) 
-parms$c = 0
-parms$v = 0
+parms <- defparms
+
 parms$L = 8
 parms$p = 0.8
-parms$b = 0.7
 
@@
 
-parms$p = 0
-parms$L = 4
-parms$q = 0.4
+parms <- defparms
+
+parms$L = 6
+parms$q = 0
 parms$v = 0.9
@@
 
+##
+
 ```{r, echo=FALSE}

### 2015-02-27 code/simfunctions.r
update simfunctions.r

@@ -355,6 +355,35 @@ runCA <- function(init, parms, width = 100, height = 100, delta = 0.1, t_max = 1
 
 
 
+### Ordinary differential equations
+
+## global model
+
+d_rho <- function(rho, parms, z = 4) { 
+  with(parms, r* b* rho * (1 - rho/K) - m * rho - (a * rho * L)/(1 + a * h * rho) )  
+}
+
+
+odesys <- function (t, rho, parms = model_parms) {
+  list( d_rho(rho, parms)  )
+}
+
+## mean-field
+
+
+d_rho_mean <- function(rho, parms) { 
+  with(parms, r *  rho^( 1 + alpha) * (b + (1-b) * f * rho)  * (1 - rho/(K * (1-c*rho) ) )  -  m * rho - ((a+v*rho) * rho^( 1 + q) * L * (1-p*rho))/(1 + (a+v*rho) * h * rho^( 1 + q)) )  
+}
+
+
+odesys_mean <- function (t, rho, parms) {
+  list( d_rho_mean(rho, parms )  )
+}
+
+
+## pair-approximation
+
+
 d_rho_1 <- function(rho, parms) { 
   with(parms,
        r * (b + (1 - b) * f * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) - m * rho[1] - ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L * (1 - p * rho[2]/rho[1]))/(1 +(a + v*rho[2]/rho[1]) * h  * rho[1]^( 1 + q)) 
@@ -410,21 +439,21 @@ ini_rho <- function(rho_1, rho_11 = NULL) {
 # function for local growth
 # takes global cover, rho_1 and local cover, q_01, as parameters, plus a list object with model parameters. 
 
-G <- function(rho_1, q_01 = "auto", prms = parms, set = list(NA)) {
+G <- function(rho_1, q_01 = "auto", parms = defparms, set = list(NA)) {
   
   if(q_01[1] == "auto") q_01 <- 1
   
   
-  prms.names <- names(prms)
+  parms.names <- names(parms)
   set.names <- names(set)
-  m.names <- sort(unique(c(prms.names, set.names)))
+  m.names <- sort(unique(c(parms.names, set.names)))
   
[…]

### 2015-03-03 code/simfunctions.r
allow starting conditions < 0; BUG #1 : flawed conservation equations

@@ -423,13 +423,7 @@ ini_rho <- function(rho_1, rho_11 = NULL) {
       rho_00 = 1-2*rho_1+rho_11,
       rho_0 = 1-rho_1
     )
-    if(any(out < 0)) out <- c(
-      rho_1 = NA,
-      rho_11 = NA,
-      rho_10 = NA,
-      rho_00 = NA,
-      rho_0 = NA
-    )
+    
   }
   
   return(out)
@@ -606,13 +600,12 @@ attractor <- function(model_parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini
   
 }
 
-
 bifurcation <- function(parms, over, xrange, res = 201, times = c(1,1000), ini = c(0.99, 0.0001)  ) {
-  
+    
   require(foreach)
   
   parms[[over]] <- seq(xrange[1],xrange[2],length = res)
-  
+ 
   parms$rho_ini <- ini
   
   iterations <- expand.grid(parms)
@@ -639,6 +632,8 @@ bifurcation <- function(parms, over, xrange, res = 201, times = c(1,1000), ini =
   foreach(i = upper[,over], .combine = rbind, .packages = c("deSolve") ) %dopar% {
     
     source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")
+    
+    
     
     model_parms <- upper[upper[, over] == i,]
     

### 2015-03-03 documentation/interactive_spexmodel_bifurcations.Rmd
interactive bifurcation diagram (Shiny): refer to simfunctions.r

--- 2015-02-02 documentation/interactive_spexmodel.Rmd
@@
 
+# provides parallel backend
+library(foreach)
+library(doSNOW)
+
+workstation <-  list(host = "162.38.184.118", user = "schneider",
+                     rscript = "/usr/lib/R/bin/Rscript",
+                     snowlib = "/usr/lib/R/library")
+
+workerlist <- c(rep("localhost", times = 10))
+
+cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')
+
+registerDoSNOW(cl)
+
 library("deSolve")
 
-  ini_rho <- function(rho_1) {
-    c(
-    rho_1 = rho_1,
-    rho_11 = rho_1*rho_1,
-    rho_10 = rho_1*(1-rho_1)*2,
-    rho_00 = (1-rho_1)*(1-rho_1),
-    rho_0 = 1-rho_1
+defparms <- list(
+    m = 0.05,
+    r = 1,
+    b = 0.7, 
+    K = 0.9, 
+    a = 0.6,
+    h = 50,
+    L = 4, 
+    alpha = 0,
+    q = 0,
+    c = 0,
+    f = 0,
+    v = 0,
+    p = 0
   )
-
-  }
-
-d_rho_1 <- function(rho, parms) { 
-  with(parms,
-       r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) - m * rho[1] - ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q))
-       )  
-  }
-
-d_rho_11 <- function(rho,  parms) { 
-  with(parms,
-       2* (rho[1] - rho[2]) * r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) / (1-rho[1]) - 2 * rho[2] * m  - 2 * rho[2] * ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q))
-       )  
-  }
-
-odesys_spex <- function(t, rho, parms = model_parms) {
-  rho_1 <- d_rho_1(rho, parms)
-  rho_11 <- d_rho_11(rho, parms)
-  list(c(rho_1, rho_11, rho_1-rho_11, 1-rho_1-(rho_1-rho_11),1-rho_1  
-               ) )
-}
-
 ```
@@
   sliderInput("r", label = "max. growth rate:",
-              min = 0.0, max = 3, value = 1, step = 0.01),
+              min = 0.0, max = 3, value = defparms$r, step = 0.01),
   sliderInput("b", label = "environmental quality:",
-              min = 0.0, max = 1, value = 0.1, step = 0.01),
+              min = 0.0, max = 1, value =   defparms$b, step = 0.01),
   sliderInput("K", label = "carrying capacity:",
-              min = 0.0, max = 1, value = 0.9, step = 0.01),
+              min = 0.0, max = 1, value =  defparms$K, step = 0.01),
   sliderInput("alpha", label = "water runoff:",
-              min = 0.0, max = 1, value = 0, step = 0.01),
+              min = 0.0, max = 1, value =  defparms$alpha, step = 0.01),
   sliderInput("f", label = "local facilitation:",
-              min = 0.0, max = 1, value = 0.9, step = 0.01),
+              min = 0.0, max = 1, value =  defparms$f, step = 0.01),
   sliderInput("c", label = "local competition:",
-              min = 0.0, max = 1, value = 0.1, step = 0.01)
+              min = 0.0, max = 1, value =  defparms$c, step = 0.01)
 )
@@
   sliderInput("m", label = "intrinsic mortality:",
-              min = 0.0, max = 1, value = 0.05, step = 0.01),
+              min = 0.0, max = 1, value =  defparms$m, step = 0.01),
   sliderInput("a", label = "search efficiency:",
-              min = 0.0, max = 1, value = 0.6, step = 0.01), 
+              min = 0.0, max = 1, value =  defparms$a, step = 0.01), 
   sliderInput("h", label = "handling time:",
-              min = 0.0, max = 200, value = 50, step = 1), 
+              min = 0.0, max = 200, value =  defparms$h, step = 1), 
   sliderInput("L", label = "livestock units:",
-              min = 0.0, max = 100, value = 6, step = 1),
+              min = 0.0, max = 100, value =  defparms$L, step = 1),
   sliderInput("q", label = "hill coefficient:",
-              min = 0.0, max = 1, value = 0, step = 0.01),
+              min = 0.0, max = 1, value =  defparms$q, step = 0.01),
   sliderInput("v", label = "local attractivity:",
-              min = 0.0, max = 1, value = 0.0, step = 0.01),
+              min = 0.0, max = 1, value =  defparms$v, step = 0.01),
   sliderInput("p", label = "local protection:",
-              min = 0.0, max = 1, value = 0.8, step = 0.01)
+              min = 0.0, max = 5, value =  defparms$p, step = 0.01)
 )
@@
 inputPanel(
-  sliderInput("rho_1", label = "initial plant cover:",
-              min = 0.0, max = 1, value = 0.9, step = 0.01)
+  selectInput("gradient", label = "select gradient:",             choices = c("b", "L"), selected = "b")
   )
@@
 renderPlot({
+
+source("../code/simfunctions.r")
+
+  rho <- seq(0,1,length = 100)
 
@@
 
-# running the ode-solver
-ODE_spex <- ode(y = ini_rho(input$rho_1), func = odesys_spex, times = 1.1^(1:50), parms = parms)
+ par(mfrow = c(1,2))
 
-# transfer into ouput and calculate missing rho values
-ODE_spex <- as.data.frame(ODE_spex )
+  attractor(parms, localvals = TRUE, rho_1_ini = seq(0,1, length = 21), rho_11_ini = seq(0,1, length = 11))
 
-par(mar = c(4,4,1,3)+0.1)
-plot(rho_1 ~ time , data = ODE_spex , type = "l", ylim = c(0,1))
-axis(4, at = round(tail(ODE_spex,1)$rho_1,2),cex.axis = 0.7, las = 1 )
+ if(input$gradient == "b") bifurcation(parms, "b", c(0,1))
+ if(input$gradient == "L") bifurcation(parms, "L", c(0,12))
 

### 2015-03-03 code/simfunctions.r
add projection of alternative stable states to meanfield model in attractor()

@@ -507,17 +507,29 @@ attractor <- function(model_parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini
     lines(rho,G(rho, rho, model_parms), col = colpal$grow[1], lwd = 2)
     
     
-    # running the ode-solver to get steady states  
-    ODEmeanfield_hi <-as.data.frame(ode(y = 0.9, func = odesys_mean, times = c(1,10000), parms = model_parms))
+    runmodel_high <- as.data.frame(ode(y = 0.99, func = odesys_mean, times = c(1,1000), parms = model_parms))
     
-    points(ODEmeanfield_hi[2,2], C(ODEmeanfield_hi[2,2],ODEmeanfield_hi[2,2], model_parms),  pch = 20)
+    points(runmodel_high[2,2],G(runmodel_high[2,2], runmodel_high[2,2], model_parms), xpd = TRUE, pch = 20, cex = 2)
     
-    ODEmeanfield_lo <-as.data.frame(ode(y = 0.0001, func = odesys_mean, times = c(1,10000), parms = model_parms))
+    runmodel_low <- as.data.frame(ode(y = 0.0001, func = odesys_mean, times = c(1,1000), parms = model_parms))
     
-    points(ODEmeanfield_lo[2,2], C(ODEmeanfield_lo[2,2],ODEmeanfield_lo[2,2], model_parms),  pch = 20, xpd = TRUE)
+    points(runmodel_low[2,2],G(runmodel_low[2,2],runmodel_low[2,2],model_parms), xpd = TRUE, pch = 20, cex = 2)
+    
+    lo <- runmodel_low[2,2]
+    hi <- runmodel_high[2,2]
+    
+    for(i in 1:10) {
+      
+      mid <- (lo+hi)/2 
+      runmodel_mid <- as.data.frame(ode(mid, func = odesys_mean, times = c(1,1.01), parms = model_parms))
+      
+      if(runmodel_mid[2,2] > mid) { hi <- mid } else { lo <- mid}
+      
+    }
     
     
-    
+    points(mid,G(mid,mid,model_parms), xpd = TRUE, pch = 21, cex = 1.5, bg = "white")
+        
     
   }
   
@@ -607,7 +619,7 @@ attractor <- function(model_parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini
 }
 
 
[…]

### 2015-03-03 documentation/model_formulation.Rmd
clean out and homogenize document model_formulation.Rmd

@@
 
-This document is supposed to develop a model representation of the global mean field as well as pair-approximations of local feed back mechanisms.
+```{r, echo=FALSE}
 
-The default form of the differential equation describing vegetation cover is
+# provides parallel backend
+library(foreach)
+library(doSNOW)
 
-$$ \frac{dV}{dt} = G(V) - C(V)L $$
+workstation <-  list(host = "162.38.184.118", user = "schneider",
+                     rscript = "/usr/lib/R/bin/Rscript",
+                     snowlib = "/usr/lib/R/library")
 
-with growth function $G$ and a function $C$ which represents consumption by a herbivore multiplied by the density of herbivores, the livestock rate, $L$.
-Formally, intrinsic mortality could be included in both terms.
-In the formulations of @noy-meir75 and @vandekoppel97 it is assumed to be covered by the logistic growth function.
+workerlist <- c(rep("localhost", times = 10))
 
-It might become important in the pair-approximation model, where intrinsic plant mortality is included as a proportional loss of mature plants.
+cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')
 
-```{r, echo=FALSE}
-# Default plotting function
+registerDoSNOW(cl)
+
+# default plot for graphs over G or C
 
@@
 
-rho <- seq(0,1,length = 100)
+```
 
-```
+This document is supposed to develop a model representation of the global mean field as well as pair-approximations of local feed back mechanisms.
+
+As our starting point for model development, the default form of the differential equation describing the change of vegetation cover, $\rho$, is
+
+$$ \frac{d\rho}{dt} = G(\rho) - C(\rho)L $$
+
+with growth function $G$ and a function $C$ which represents consumption by a herbivore multiplied by the density of herbivores, the livestock rate, $L$.
+Formally, intrinsic mortality could be included in both terms.
+In the formulations of @noy-meir75 and @vandekoppel97 it is assumed to be covered by the logistic growth function.
+
+It might become important in the pair-approximation model, where intrinsic plant mortality is included as a proportional loss of mature plants.
 
@@
 
-$$ G(V) = rV(1-\frac{V}{K}) \,,$$
+$$ G(\rho) = r\rho(1-\frac{\rho}{K}) \,,$$
 
-with the growth rate, $r$, and the carrying capacity, $K$, decelerating the exponential growth as vegetation cover increases and reducing it to zero if $V \geq K$.
+with the growth rate, $r$, and the carrying capacity, $K$, decelerating the exponential growth as vegetation cover increases and reducing it to zero if $\rho \geq K$.
 Thus, it is a decelerating effect (a negative feed-back) at high vegetation cover.
@@
 
-G <- function(P, r = 1, K = 1) r*P*(1-P/K)
+G <- function(rho, r = 1, K = 1) r*rho*(1-rho/K)
+rho <- seq(0,1,length = 100)
 
@@
 The levelling off to a plateau and the decrease in the higher range expresses self-interference effects (shading, competition); at the point $V = V_\mathrm{max}$ 'maintenance' losses equal photosynthesis and $G = 0$.
-Vmax is the maximum (stable) biomass in ungrazed vegetation.
+$V_\mathrm{max}$ is the maximum (stable) biomass in ungrazed vegetation.
 [...] Of the explicit forms the best known is the logistic [...] but other functions are possible.
@@
 
-$$ r = r_0 b V^a $$
+$$ r = r_0 b \rho^a $$
 
-$$ G(V) = r_0 b V^{a+1} (1-\frac{V}{K}) \,,$$
+$$ G(\rho) = r_0 b \rho^{a+1} (1-\frac{\rho}{K}) \,,$$
 
@@
 
-r <- function(P, r_0 =  1, a = 1) r_0 *P^a
+r <- function(rho, r_0 =  1, b = 1, a = 1) r_0*b *rho^a
+
+G <- function(rho, r = 1, b =1, K = 1) r*b*rho*(1-rho/K)
 
@@
 
-G_sig <- function(P, r_0 = 1, a = 1,  K = 0.9) r(P, r_0, a)*P*(1-P/K)
+G_sig <- function(rho, r_0 = 1, b = 1, a = 1,  K = 0.9) r(rho, r_0, b, a)*rho*(1-rho/K)
 
@@
 
-$$ C(V) = gVL \,, $$
+$$ C( \rho ) = g \rho L \,, $$
 
@@
 ```{r, echo=FALSE}
-C <- function(V, L = 5, g = 0.01) (g*V*L)
+C <- function(rho, L = 5, g = 0.01) (g*rho*L)
 
@@
 
-$$ C(V) = gL \,, $$
+$$ C( \rho ) = gL \,, $$
 
@@
 ```{r, echo=FALSE}
-C <- function(V, L = 5, g = 0.01) g * L + V*0
+C <- function(rho, L = 5, g = 0.01) g * L + rho*0
 
@@
 
-$$ C(V) = \frac{aVL}{1+ahV} \,,$$
+$$ C( \rho ) = \frac{a \rho L}{1+ah \rho } \,,$$
 
@@
 ```{r, echo=FALSE}
-C <- function(V, L = 5, a = 0.6, h = 100) (a*V*L)/(1+a*h*V)
+C <- function(rho, L = 5, a = 0.6, h = 100) (a*rho*L)/(1+a*h*rho)
 
@@
 
-$$ C(V) = \frac{bV^{1+q}}{1+bhV^{1+q}} $$
+$$ C( \rho ) = \frac{b \rho^{1+q}}{1+bh \rho^{1+q}} $$
 
@@
 ```{r, echo=FALSE}
-C <- function(V, L = 5, a = 0.6, h = 100, q = 1) (a*V^(1+q)*L)/(1+a*h*(V^(1+q)) )
+C <- function(rho, L = 5, a = 0.6, h = 100, q = 1) (a*rho^(1+q)*L)/(1+a*h*(rho^(1+q)) )
 
@@
 
-$$ C(V) = mV + \frac{aVL}{1+ahV} \,,$$
+$$ C(\rho) = m \rho + \frac{a \rho L}{1+a h \rho} \,,$$
 
@@
 
-C <- function(V, m_0 = 0.05, L = 5, a = 0.6, h = 100, q = 0) (m_0 * V) +(a*V^(1+q)*L)/(1+a*h*(V^(1+q)) )
+C <- function(rho, m_0 = 0.05, L = 5, a = 0.6, h = 100, q = 0) (m_0 * rho) +(a*rho^(1+q)*L)/(1+a*h*(rho^(1+q)) )
 
@@
 
-The differential equation describing population cover above with $V$ and $C$ substituted would then become
+The differential equation describing population cover above with $G$ and $C$ substituted would then become
 
-$$ \frac{\mathrm{d}V}{\mathrm{d}t} = r b V (1-\frac{V}{K}) - m V - \frac{aVL}{1+ahV} $$
+$$ \frac{\mathrm{d}\rho}{\mathrm{d}t} = r b \rho (1-\frac{\rho}{K}) - m \rho - \frac{a\rho L}{1+ah\rho} $$
 
@@
 ```{r, echo=FALSE}
-plot(rho,  with(parms,  m * rho + (a * rho * L)/(1 + a * h * rho)), ylab = "growth", type = "l", ylim = c(0,0.5))
-lines(rho, with(parms, r* b* rho * (1 - rho/K) ), col = "green3" )
+defplot(rho,  C(rho, m = parms$m, L = parms$L, h = parms$h, a = parms$a ) )
+
+lines(rho, G(rho, r=parms$r, K = parms$K, b=parms$b), col = "green3" )
 
@@
 
-odesys <- function (t, rho, parms = model_parms) {
+odesys <- function (t, rho, parms = parms) {
   list( d_rho(rho, parms)  )
@@
 
+If we run this for low and high starting conditions, we can get the stable equilibria. Further, running an iterative series of intermediate starting conditions gives us the unstable equilibrium.
+
+```{r, echo=FALSE}
+
+defplot(rho,  C(rho, m = parms$m, L = parms$L, h = parms$h, a = parms$a ) )
+
+lines(rho, G(rho, r=parms$r, K = parms$K, b=parms$b), col = "green3" )
+
+runmodel_high <- as.data.frame(ode(y = 0.99, func = odesys, times = c(1,1000), parms = parms))
+
+points(runmodel_high[2,2],G(runmodel_high[2,2], r=parms$r, K = parms$K, b=parms$b), xpd = TRUE, pch = 20, cex = 2)
+
+runmodel_low <- as.data.frame(ode(y = 0.0001, func = odesys, times = c(1,1000), parms = parms))
+
+points(runmodel_low[2,2],G(runmodel_low[2,2], r=parms$r, K = parms$K, b=parms$b), xpd = TRUE, pch = 20, cex = 2)
+
+lo <- runmodel_low[2,2]
+hi <- runmodel_high[2,2]
+
+for(i in 1:10) {
+
+  mid <- (lo+hi)/2 
+  runmodel_mid <- as.data.frame(ode(mid, func = odesys, times = c(1,1.01), parms = parms))
+
+  if(runmodel_mid[2,2] > mid) { hi <- mid } else { lo <- mid}
+
+}
+
+points(mid,G(mid, r=parms$r, K = parms$K, b=parms$b), xpd = TRUE, pch = 21, cex = 1.5, bg = "white")
+
+```
+
 #### global model
@@
 
-$$ \frac{\mathrm{d}V}{\mathrm{d}t} = r b V^{1 + \alpha} (1-\frac{V}{K}) - m V - \frac{aV^{1+q}L}{1+ahV^{1+q}} $$
+$$ \frac{\mathrm{d}\rho}{\mathrm{d}t} = r b \rho^{1 + \alpha} (1-\frac{\rho}{K}) - m \rho - \frac{a\rho^{1+q}L}{1+ah\rho^{1+q}} $$
 
@@
 ```{r, echo=FALSE}
+par(mar = c(5,4,2,3)+0.1)
 plot(rho,  with(parms,  m * rho + (a * rho^( 1 + q) * L)/(1 + a * h * rho^( 1 + q))), ylab = "growth", type = "l", ylim = c(0,0.5))
@@
   )
+
 parms <- defparms
@@
 plot(nullmodel$rho_one ~ nullmodel$time, 
-     ylim = c(0,1), ylab = expression( rho["+"]), 
+     ylim = c(0,1), ylab = expression( rho["1"]), 
      xlab = "time [yr]", type = "l")
@@
 And the animated lattice would look like this:
+
 ```{r, eval = FALSE, echo = FALSE, warning = FALSE}
@@
     a = 0.6,
-    h = 100,
-    L = 16, 
-    alpha = 0.1,
-    q = 1,
+    h = 50,
+    L = 6, 
+    alpha = 0,
+    q = 0,
     c = 0,
@@
 plot(globalmodel$rho_one ~ globalmodel$time, 
-     ylim = c(0,1), ylab = expression( rho["+"]), 
+     ylim = c(0,1), ylab = expression( rho["1"]), 
      xlab = "time [yr]", type = "l")
@@
 
-## spatially-explicit models
+## local feedbacks
 
@@
 It generates patterns that act as positive feed-backs on the local scale by providing refugia to grazing and overcome growth limitations at low vegetation cover.
-
-These mechanisms can be implemented in the model by using a pair-approximation approach and numerical simulations, where $G$ and $C$ can be implemented as the transition rates in a two-state spatial model depending on global and local vegetation cover.
 
@@
 
-defplot(rho, G(rho, 1, set = list( f = 0.8)), ylab = "growth", col = "green3")
-lines(rho,G(rho, 0.75, set = list( f = 0.8)), col = "green3")
-lines(rho,G(rho, 0.5, set = list( f = 0.8)), col = "green3")
-lines(rho,G(rho, 0.25, set = list( f = 0.8)), col = "green3")
-lines(rho,G(rho, 0, set = list( f = 0.8)), col = "green3")
+parms <- defparms
+parms$f = 0.8
+parms$b = 0.3
 
-lines(rho, C(rho, set = list( L = 10)))
+attractor(parms, pairapprox = FALSE, meanfield = FALSE, localvals = TRUE)
 
@@
 
-defplot(rho, G(rho, 1, set= list(c = 0.3, b = 0.9)), ylab = "growth", col = "green3")
-lines(rho,G(rho, 0.75, set= list( c = 0.3, b = 0.9)), col = "green3")
-lines(rho,G(rho, 0.5, set= list( c = 0.3, b = 0.9)), col = "green3")
-lines(rho,G(rho, 0.25, set= list(c = 0.3, b = 0.9)), col = "green3")
-lines(rho,G(rho, 0, set= list( c = 0.3, b = 0.9)), col = "green3")
+parms <- defparms
+parms$c = 0.3
+parms$b = 0.9
 
-lines(rho, C(rho, set= list( L = 10)))
+attractor(parms, pairapprox = FALSE, meanfield = FALSE, localvals = TRUE)
 
@@
 
-defplot(rho, C(rho, 1, set = list( v = 0.5)), ylab = "mortality")
-lines(rho,C(rho, 0.75, set = list( v = 0.5)))
-lines(rho,C(rho, 0.5, set = list( v = 0.5)))
-lines(rho,C(rho, 0.25, set = list( v = 0.5)))
-lines(rho,C(rho, 0, set = list( v = 0.5)))
-
-lines(rho, G(rho, set = list( b = 0.6)), col = "green3")
+parms <- defparms
+parms$v = 0.9
+attractor(parms, pairapprox = FALSE, meanfield = FALSE, localvals = TRUE)
 
@@
 
-defplot(rho, C(rho, 1, set= list(p = 0.8, L = 10)), ylab = "mortality")
-lines(rho,C(rho, 0.75, set= list(p = 0.8, L = 10)))
-lines(rho,C(rho, 0.5, set= list(p = 0.8, L = 10)))
-lines(rho,C(rho, 0.25, set= list(p = 0.8, L = 10)))
-lines(rho,C(rho, 0, set= list(p = 0.8, L = 10)))
-
-lines(rho, G(rho, set= list(b = 0.6)), col = "green3")
+parms <- defparms
+parms$p = 0.8
+attractor(parms, pairapprox = FALSE, meanfield = FALSE, localvals = TRUE)
 
@@
 
-defplot(rho, C(rho, 1, set = list( p = 0.8, v = 0.6, L = 10, q = .5)), ylab = "mortality")
-lines(rho,C(rho, 0.75, set = list( p = 0.8, v = 0.6, L = 10, q = .5)))
-lines(rho,C(rho, 0.5, set = list( p = 0.8,v = 0.6,  L = 10, q = .5)))
-lines(rho,C(rho, 0.25, set = list( p = 0.8,v = 0.6,  L = 10, q = .5)))
-lines(rho,C(rho, 0, set = list( p = 0.8,v = 0.6,  L = 10, q = .5)))
-
-lines(rho, G(rho, 1, set = list(c = 0.3, f = 0.9, b = 0.3, alpha = 0.5)), col = "green3")
-lines(rho,G(rho, 0.75, set = list( c = 0.3,f = 0.9,  b = 0.3, alpha = 0.5)), col = "green3")
-lines(rho,G(rho, 0.5, set = list(c = 0.3,f = 0.9,  b = 0.3, alpha = 0.5)), col = "green3")
-lines(rho,G(rho, 0.25, set = list( c = 0.3, f = 0.9, b = 0.3, alpha = 0.5)), col = "green3")
-lines(rho,G(rho, 0, set = list(c = 0.3,f = 0.9,  b = 0.3, alpha = 0.5)), col = "green3")
+parms <- defparms
+parms$p = 0.8
+parms$v = 0.6
+parms$q = 0.5
+parms$L = 6
+parms$c = 0.3
+parms$f = 0.9
+parms$b = 0.2
+attractor(parms, pairapprox = FALSE, meanfield = FALSE, localvals = TRUE)
 
 ```
+
+## models
+
+These mechanisms can be modelled in different ways: The simplest is to approximate them by a mean field model, assuming that spatial structure is negligible.
+Second, we can use a pair-approximation model that includes the cover of pairs of cells, thereby keeping track of the local vegetation cover.
+Finally, numerical simulations provide the highest realism, where $G$ and $C$ can be implemented as the transition rates in a two-state spatial model depending on global and local vegetation cover.
+
+We are developing all three approaches and compare their predictive possibilities.
+
+### mean field model
+
+As a simple approximation of the model serves a mean field model, assuming that the conditional local vegetation cover of empty cells $q_{0|1}$ and of vegetated cells $q_{1|1}$ both are well represented by the global vegetation cover $\rho_q$.
+Of course $\rho_0 = 1-\rho_1$.
+The change in cover derives from the growth and mortality terms
+
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = \rho_{0}w_{0,1} - \rho_{1}w_{1,0}   $$
+
+Thus, we can substitute $C(\rho) = w_{1,0}$ and $G(\rho) = w_{0,1}$ and obtain
+
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b + (1 - b)  f \rho_1)) (1-\frac{\rho_1}{K(1 - c \rho_1}) - \rho_{1}m - \frac{(a  + v \rho_1)L (1 - p \rho_1)\rho_1^{1+q}}{1+(a  + v \rho_1)h\rho_1^{1+q}}  $$
+
+Plotting the meanfield growth and mortality over vegetation cover is straigthforward now.
+
+$$ G = r \rho_1^{1+\alpha} (b + (1 - b)  f \rho_1)) (1-\frac{\rho_1}{K(1 - c \rho_1}) $$
+
+$$ C = \rho_{1}m + \frac{(a  + v \rho_1)L (1 - p \rho_1)\rho_1^{1+q}}{1+(a  + v \rho_1)h\rho_1^{1+q}}$$
+
+The alternative steady states and unstable equilibria can be simulated by solving the ODE.
+
+```{r, echo=FALSE}
+
+parms <- defparms
+parms$f = 0.9
+parms$b = 0.4
+parms$L = 5
+parms$p = 0.8
+parms$m = 0.01
+attractor(parms, pairapprox = FALSE, localvals = TRUE)
+
+```
+
+But the steady state also can be solved analytically.
+
+Therefore, we solve for the equilibrium case
+
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = 0 $$
+
+$$  0 = r \rho_1^{1+\alpha} (b + (1 - b)  f \rho_1)) (1-\frac{\rho_1}{K(1 - c \rho_1}) - \rho_{1}m - \frac{(a  + v \rho_1)L (1 - p \rho_1)\rho_1^{1+q}}{1+(a  + v \rho_1)h\rho_1^{1+q}}  $$
 
@@
 plot(spexmodel$rho_one ~ spexmodel$time, 
-     ylim = c(0,1), ylab = expression( rho["+"]), 
+     ylim = c(0,1), ylab = expression( rho["1"]), 
      xlab = "time [yr]", type = "l")
@@
 ```
-
-### mean field model
-
-As a simple approximation of the model serves a mean field model, assuming that the conditional local vegetation cover of empty cells $q_{0|1}$ and of vegetated cells $q_{1|1}$ both are well represented by the global vegetation cover $\rho_q$.
-Of course $\rho_0 = 1-\rho_1$.
-The change in cover derives from the growth and mortality terms
-
-$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = \rho_{0}w_{0,1} - \rho_{1}w_{1,0}   $$
-
-Thus, we can substitute $w_{1,0}$ and $w_{0,1}$ and obtain
-
-$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b + (1 - b)  f \rho_1)) (1-\frac{\rho_1}{K(1 - c \rho_1}) - \rho_{1}m - \frac{(a  + v \rho_1)L (1 - p \rho_1)\rho_1^{1+q}}{1+(a  + v \rho_1)h\rho_1^{1+q}}  $$
-
-Plotting the meanfield growth and mortality over vegetation cover is straigthforward now.
-
-$$ G = r \rho_1^{1+\alpha} (b + (1 - b)  f \rho_1)) (1-\frac{\rho_1}{K(1 - c \rho_1}) $$
-
-$$ C = \rho_{1}m + \frac{(a  + v \rho_1)L (1 - p \rho_1)\rho_1^{1+q}}{1+(a  + v \rho_1)h\rho_1^{1+q}}$$
-
-The steady states can be simulated by solving the ODE.
-
-```{r, echo=FALSE}
-
-parms <- defparms
-parms$f = 0.9
-parms$b = 0.4
-parms$L = 5
-
-attractor(parms, pairapprox = FALSE, localvals = TRUE)
-
-```
-
-But the steady state also can be solved analytically.
-
-Therefore, we solve for the equilibrium case
-
-$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = 0 $$
-
-$$  0 = r \rho_1^{1+\alpha} (b + (1 - b)  f \rho_1)) (1-\frac{\rho_1}{K(1 - c \rho_1}) - \rho_{1}m - \frac{(a  + v \rho_1)L (1 - p \rho_1)\rho_1^{1+q}}{1+(a  + v \rho_1)h\rho_1^{1+q}}  $$
 
@@
 
-Visualising the attractor of the system is not as straightforward as for the models that are determined by global cover only, where the intersections of growth and mortality defined the equilibria.
+Visualising the attractor of the system is not as straightforward as for the mean-field models that are determined by global cover only, where the intersections of growth and mortality defined the equilibria.
 In the spatially-explict pair-approximation model the attractor can be visualised if we assume that clustering in the initial state is non-random, and we assign starting values for $\rho_1$ and global mortality of the population $m_1$.
@@
 ```{r, echo=FALSE}
-
-# provides parallel backend
-library(foreach)
-library(doSNOW)
-
-workstation <-  list(host = "162.38.184.118", user = "schneider",
-                     rscript = "/usr/lib/R/bin/Rscript",
-                     snowlib = "/usr/lib/R/library")
-
-workerlist <- c(rep("localhost", times = 10))
-
-cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')
-
-registerDoSNOW(cl)
 
@@
 
-stopCluster(cl)
-
 ```
@@
 iterations$f[iterations$f < iterations$b] <- iterations$b[iterations$f < iterations$b]
-
-# provides parallel backend
-library(foreach)
-library(doSNOW)
-
-workstation <-  list(host = "162.38.184.118", user = "schneider",
-                     rscript = "/usr/lib/R/bin/Rscript",
-                     snowlib = "/usr/lib/R/library")
-
-workerlist <- c(rep("localhost", times = 10))
-
-cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')
-
-registerDoSNOW(cl)
 

### 2015-03-04 documentation/interactive_spexmodel.Rmd
update interactive timeseries (Shiny)

@@
 ---
-title: "global-model (ODE)"
+title: "spatially-explicit model"
 author: "Florian Schneider"
@@
 
-## developping the null model
+The model includes global accelerating and decelerating feedbacks on mortality and grazing.
 
-Non-spatial global model solved as ordinary differential equations. It includes global accelerating and decelerating feedbacks on mortality and grazing.
+This panel runs all three models
+
+- cellular automata (grey line)
+- pair-approximation (black line)
+- mean-field model (red line)
+
+The pair-approximation and mean field are  solved as ordinary differential equations.
 
@@
 
-  ini_rho <- function(rho_1) {
-    c(
-    rho_1 = rho_1,
-    rho_11 = rho_1*rho_1,
-    rho_10 = rho_1*(1-rho_1)*2,
-    rho_00 = (1-rho_1)*(1-rho_1),
-    rho_0 = 1-rho_1
+defparms <- list(
+    m = 0.05,
+    r = 1,
+    b = 0.4, 
+    K = 0.9, 
+    a = 0.3,
+    h = 50,
+    L = 5, 
+    alpha = 0,
+    q = 0,
+    c = 0,
+    f = 0,
+    v = 0,
+    p = 0
   )
-
-  }
-
-d_rho_1 <- function(rho, parms) { 
-  with(parms,
-       r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) - m * rho[1] - ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q))
-       )  
-  }
-
-d_rho_11 <- function(rho,  parms) { 
-  with(parms,
-       2* (rho[1] - rho[2]) * r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) / (1-rho[1]) - 2 * rho[2] * m  - 2 * rho[2] * ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q))
-       )  
-  }
-
-odesys_spex <- function(t, rho, parms = model_parms) {
-  rho_1 <- d_rho_1(rho, parms)
-  rho_11 <- d_rho_11(rho, parms)
-  list(c(rho_1, rho_11, rho_1-rho_11, 1-rho_1-(rho_1-rho_11),1-rho_1  
-               ) )
-}
 
@@
 As an accelerating feedback we add water runoff, which reduces growth at low cover. 
+As local feedbacks, we assume local competition, reducing the carrying capacity at high local cover, and local facilitation, compensating for poor environmental quality (a value of one restituates the maximal growth rate)
 
@@
   sliderInput("r", label = "max. growth rate:",
-              min = 0.0, max = 3, value = 1, step = 0.01),
+              min = 0.0, max = 3, value = defparms$r, step = 0.01),
   sliderInput("b", label = "environmental quality:",
-              min = 0.0, max = 1, value = 0.1, step = 0.01),
+              min = 0.0, max = 1, value = defparms$b, step = 0.0001),
   sliderInput("K", label = "carrying capacity:",
-              min = 0.0, max = 1, value = 0.9, step = 0.01),
+              min = 0.0, max = 1, value = defparms$K, step = 0.01),
   sliderInput("alpha", label = "water runoff:",
-              min = 0.0, max = 1, value = 0, step = 0.01),
+              min = 0.0, max = 1, value = defparms$alpha, step = 0.01),
   sliderInput("f", label = "local facilitation:",
-              min = 0.0, max = 1, value = 0.9, step = 0.01),
+              min = 0.0, max = 1, value = defparms$f, step = 0.01),
   sliderInput("c", label = "local competition:",
-              min = 0.0, max = 1, value = 0.1, step = 0.01)
+              min = 0.0, max = 1, value = defparms$c, step = 0.01)
 )
@@
 
+Local feedbacks are local attractivity of vegetated patches, increasing the search efficiency, and local protection (associational resistance) which reduces the maximum feeding rate.
+
 ```{r, echo=FALSE}
@@
   sliderInput("m", label = "intrinsic mortality:",
-              min = 0.0, max = 1, value = 0.05, step = 0.01),
+              min = 0.0, max = 1, value = defparms$m, step = 0.01),
   sliderInput("a", label = "search efficiency:",
-              min = 0.0, max = 1, value = 0.6, step = 0.01), 
+              min = 0.0, max = 1, value = defparms$a, step = 0.01), 
   sliderInput("h", label = "handling time:",
-              min = 0.0, max = 200, value = 50, step = 1), 
+              min = 0.0, max = 200, value = defparms$h, step = 1), 
   sliderInput("L", label = "livestock units:",
-              min = 0.0, max = 100, value = 6, step = 1),
+              min = 0.0, max = 100, value = defparms$L, step = 1),
   sliderInput("q", label = "hill coefficient:",
-              min = 0.0, max = 1, value = 0, step = 0.01),
+              min = 0.0, max = 1, value = defparms$q, step = 0.01),
   sliderInput("v", label = "local attractivity:",
-              min = 0.0, max = 1, value = 0.0, step = 0.01),
+              min = 0.0, max = 1, value = defparms$v, step = 0.01),
   sliderInput("p", label = "local protection:",
-              min = 0.0, max = 1, value = 0.8, step = 0.01)
+              min = 0.0, max = 1, value = defparms$p, step = 0.01)
 )
@@
 renderPlot({
+
+  source("../code/simfunctions.r")
 
@@
 # running the ode-solver
-ODE_spex <- ode(y = ini_rho(input$rho_1), func = odesys_spex, times = 1.1^(1:50), parms = parms)
-
-# transfer into ouput and calculate missing rho values
-ODE_spex <- as.data.frame(ODE_spex )
+ODE_spex <-  as.data.frame(ode(y = ini_rho(input$rho_1), func = odesys_spex, times = exp(seq(0,4,length = 100))-1, parms = parms))
 
 par(mar = c(4,4,1,3)+0.1)
-plot(rho_1 ~ time , data = ODE_spex , type = "l", ylim = c(0,1))
+
+plot(rho_1 ~ time , data = ODE_spex , type = "l", ylim = c(0,1), xlim = c(0,50))
 axis(4, at = round(tail(ODE_spex,1)$rho_1,2),cex.axis = 0.7, las = 1 )
+
+spexmodel <- runCA(input$rho_1, parms, delta = 0.2, t_max = 60, t_min = 50, t_eval = 5, isstable = 0.001, saveeach = 1)
+
+lines(spexmodel$rho_one ~ spexmodel$time, col = "grey50")
+
+meanfield <-as.data.frame( ode(ini_rho(input$rho_1), func = odesys_mean, times = exp(seq(0,4,length = 100))-1, parms = parms))
+lines(rho_1 ~ time , data = meanfield, col = "red3")
 

### 2015-03-04 code/simfunctions.r
fix growth and mortality criteria

@@ -257,6 +257,7 @@ runCA <- function(init, parms, width = 100, height = 100, delta = 0.1, t_max = 1
   
   # initialize landscape: 
   parms_temp <- parms   # set temporary parameter object
+  
   parms_temp$rho_one <- init   # draw initial vegetation cover
   
   init_plant <- as.integer(width*height*parms_temp$rho_one) # how many initial plants
@@ -377,7 +378,7 @@ odesys <- function (t, rho, parms = model_parms) {
 
 d_rho_mean <- function(rho, parms) { 
   growth <- with(parms, r *  rho^( 1 + alpha) * (b + (1-b) * f * rho)  * (1 - rho/(K * (1-c*rho) ) ))
-  if(growth <= 0) {growth <- 0}
+  if(growth <= 0| is.na(growth)) {growth <- 0}
   
   mortality <- with(parms,  m * rho + ((a+v*rho) * rho^( 1 + q) * L * (1-p*rho))/(1 + (a+v*rho) * h * rho^( 1 + q)) )  
   if(mortality <= 0 | is.na(mortality)) {mortality <- 0}
@@ -399,7 +400,7 @@ d_rho_1 <- function(rho, parms) {
   
   growth <- with(parms, 
                  r * (b + (1 - b) * f * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) )) 
-  if(growth <= 0) {growth <- 0}
+  if(growth <= 0 | is.na(growth)) {growth <- 0}
 
   mortality <- with(parms, m * rho[1] + ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L * (1 - p * rho[2]/rho[1]))/(1 +(a + v*rho[2]/rho[1]) * h  * rho[1]^( 1 + q)) )
   if(mortality <= 0 | is.na(mortality)) {mortality <- 0}
@@ -412,7 +413,7 @@ d_rho_11 <- function(rho,  parms) {
   
   growth <- with(parms,
[…]

### 2015-03-04 code/simfunctions.r
odesys_spex: set to zero below threshold

@@ -422,11 +422,22 @@ d_rho_11 <- function(rho,  parms) {
 }
 
 
-odesys_spex <- function(t, rho, parms = model_parms) {
-  rho_1 <- d_rho_1(rho, parms)
-  rho_11 <- d_rho_11(rho, parms)
-  list(c(rho_1, rho_11, rho_1-rho_11, 1-rho_1-(rho_1-rho_11),1-rho_1  
-  ) )
+odesys_spex <- function(t, rho, parms = model_parms) { 
+  if(rho[1] < 1e-7 ) {
+    rho <- ini_rho(0,0)
+  } 
+  
+    rho_1 <- d_rho_1(rho, parms)
+    rho_11 <- d_rho_11(rho, parms)
+ 
+  list(c(
+    rho_1 = rho_1,
+    rho_11 = rho_11,
+    #changes of other pairs and singletons are not calculated
+    rho_10 = 0, 
+    rho_00 = 0,
+    rho_0 = 0
+  )  )
 }
 
 # function to get initial rho values for pair-approximation

### 2015-03-04 code/simfunctions.r
move to runODE_spex function

@@ -530,6 +530,19 @@ defplot <- function(...) plot(...,
 
 
 
+runODE_spex <- function(starting, model_parms, times = c(0,1000))  {
+  
+    out <- as.data.frame(ode(starting, func = odesys_spex, times = times, parms = model_parms))
+    
+    names(out) <- c("time", "rho_1", "rho_11", "rho_10", "rho_00", "rho_0")
+    
+    out$rho_10 <- out$rho_1- out$rho_11
+    out$rho_00 <- 1-2*out$rho_1+out$rho_11
+    out$rho_0  <- 1- out$rho_1
+    
+  return(round(out,6))
+}
+
 # graphical visualisation of attractor
 # simulates trajectories for many differently clustered starting conditions. 
 # running pairapprox = TRUE requires parallel backend
@@ -606,15 +619,10 @@ attractor <- function(model_parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini
       rho_starting <- ini[iteration, 2:6]
       
       # running the ode-solver
-      runmodel <- ode(y = as.numeric(rho_starting), func = odesys_spex, times = seq(1,2), parms = model_parms)
+      runmodel <- runODE_spex(as.numeric(rho_starting),  model_parms, times = seq(1,2))
       
-      out <-as.data.frame(runmodel)[dim(runmodel)[1],]
-      
-      # transfer into ouput and calculate missing rho values
-      return(out)
+      return(tail(runmodel,1))
     } -> output 
-    
-    names(output) <- c("time", "rho_1", "rho_11", "rho_10", "rho_00", "rho_0")
     
     steady <- ini$rho_1 == output$rho_1
     ini <- ini[!steady,]
@@ -674,10 +682,9 @@ bifurcation <- function(parms, over, xrange, res = 201, times = c(1,1000), ini =
     model_parms <- as.list(iterations[iteration,])
     
     # running the ode-solver
-    runmodel <- ode(y = ini_rho(model_parms$rho_ini), func = odesys_spex, times = times, parms = model_parms)
-    
-    # transfer into ouput and calculate missing rho values
-    return(as.data.frame(runmodel)[dim(runmodel)[1],])
+    runmodel <- runODE_spex(ini_rho(model_parms$rho_ini), model_parms, times = times) 
+      
+    return(tail(runmodel,1))
   } -> output
   
[…]

### 2015-03-04 code/simfunctions.r
minor edits

@@ -572,7 +572,7 @@ attractor <- function(model_parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini
     for(i in 1:10) {
       
       mid <- (lo+hi)/2 
-      runmodel_mid <- as.data.frame(ode(mid, func = odesys_mean, times = c(1,1.01), parms = model_parms))
+      runmodel_mid <- as.data.frame(ode(mid, func = odesys_mean, times = c(1,1.2), parms = model_parms))
       
       if(runmodel_mid[2,2] > mid) { hi <- mid } else { lo <- mid}
       
@@ -609,7 +609,7 @@ attractor <- function(model_parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini
     ini$m_ini <- C(ini$rho_1, ini$rho_11/ini$rho_1, model_parms)
     ini$g_ini <- G(ini$rho_1, (ini$rho_1-ini$rho_11)/(1-ini$rho_1), model_parms)
     
-    ini <- subset(ini, !is.na(m_ini))
+    ini <- subset(ini, !is.na(m_ini) & !is.na(ini$g_ini) &  ini$g_ini >= 0)
     ini <- cbind(ID = 1:nrow(ini),ini)
     
     
@@ -665,7 +665,7 @@ attractor <- function(model_parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini
 }
 
 
-bifurcation <- function(parms, over, xrange, res = 201, times = c(1,1000), ini = c(0.99, 0.0001) , meanfield = TRUE, pairapprox = FALSE, numerical = FALSE ) {
+bifurcation <- function(parms, over, xrange, res = 201, times = c(0,1000), ini = c(0.9, 0.0001) , meanfield = TRUE, pairapprox = FALSE, numerical = FALSE ) {
   
   require(foreach)
   

### 2015-03-04 documentation/interactive_spexmodel_attractor.Rmd
update interactive models (Shiny)

--- 2015-02-02 documentation/interactive_spexmodel_bifurcations.Rmd
@@
 
-  ini_rho <- function(rho_1) {
-    c(
-    rho_1 = rho_1,
-    rho_11 = rho_1*rho_1,
-    rho_10 = rho_1*(1-rho_1)*2,
-    rho_00 = (1-rho_1)*(1-rho_1),
-    rho_0 = 1-rho_1
+defaultparms <- list(
+    m = 0.05,  # intrinsic mortality of plants 
+    r = 1,   # growth rate
+    f = 0.8, #  local facilitation
+    b = 0.25,  # environmental quality inverse of aridity
+    K = 0.9,  # carrying capacity
+    alpha = 0.1, # water runoff
+    c = 0.2, # local competition
+    a = 0.6, # search efficiency
+    v = 0.1, # attraction effect
+    h = 50, # handling time (one individual on landscape unit)
+    p = 0.8, # protection effect
+    L = 6, # livestock units per landscape unit 
+    q = 1 # search ineffiency at low cover
   )
-
-  }
-
-d_rho_1 <- function(rho, parms) { 
-  with(parms,
-       r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) - m * rho[1] - ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q))
-       )  
-  }
-
-d_rho_11 <- function(rho,  parms) { 
-  with(parms,
-       2* (rho[1] - rho[2]) * r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) / (1-rho[1]) - 2 * rho[2] * m  - 2 * rho[2] * ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q))
-       )  
-  }
-
-odesys_spex <- function(t, rho, parms = model_parms) {
-  rho_1 <- d_rho_1(rho, parms)
-  rho_11 <- d_rho_11(rho, parms)
-  list(c(rho_1, rho_11, rho_1-rho_11, 1-rho_1-(rho_1-rho_11),1-rho_1  
-               ) )
-}
 
@@
 ### initial conditions
-```{r, echo=FALSE}
+```{r, echo=FALSE, fig.height=7.5}
 
 renderPlot({
+
+  source("../code/simfunctions.r")
 
@@
 
-parms$b <- seq(0,1,.005) #seq(0,1,.002)+0.001
-parms$rho_ini <- c(0.9, 0.01)
+# if(input$gradient == "b") parms$b <- seq(0,1,.005) 
+# if(input$gradient == "L") parms$L <- seq(0,20,.1)
 
-iterations <- expand.grid(parms)
-iterations <- cbind(ID = 1:dim(iterations)[1],iterations)
+#iterations$f[iterations$f < iterations$b] <- iterations$b[iterations$f < iterations$b]
 
-iterations$f[iterations$f < iterations$b] <- iterations$b[iterations$f < iterations$b]
+attractor(parms)
 
-foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
-  source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")
-
-  model_parms <- as.list(iterations[iteration,])
-
-  # running the ode-solver
-  runmodel <- ode(y = ini_rho(model_parms$rho_ini), func = odesys_spex, times = c(1,1000), parms = model_parms)
-
-  # transfer into ouput and calculate missing rho values
-  return(as.data.frame(runmodel)[dim(runmodel)[1],])
-  } -> output
-
-output <- cbind(output, iterations)
-
-plot(rho_1 ~ b, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
+# if(input$gradient == "b") plot(rho_1 ~ b, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
+# if(input$gradient == "L") plot(rho_1 ~ L, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
 

### 2015-03-04 documentation/interactive_spexmodel_bifurcations.Rmd
update interactive models (Shiny)

--- 2015-02-02 documentation/interactive_spexmodel_bifurcations.Rmd
@@
 registerDoSNOW(cl)
-
-library("deSolve")
-
-  ini_rho <- function(rho_1) {
-    c(
-    rho_1 = rho_1,
-    rho_11 = rho_1*rho_1,
-    rho_10 = rho_1*(1-rho_1)*2,
-    rho_00 = (1-rho_1)*(1-rho_1),
-    rho_0 = 1-rho_1
-    )
-
-  }
-
-d_rho_1 <- function(rho, parms) { 
-  with(parms,
-       r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) - m * rho[1] - ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q))
-       )  
-  }
-
-d_rho_11 <- function(rho,  parms) { 
-  with(parms,
-       2* (rho[1] - rho[2]) * r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) / (1-rho[1]) - 2 * rho[2] * m  - 2 * rho[2] * ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q))
-       )  
-  }
-
-odesys_spex <- function(t, rho, parms = model_parms) {
-  rho_1 <- d_rho_1(rho, parms)
-  rho_11 <- d_rho_11(rho, parms)
-  list(c(rho_1, rho_11, rho_1-rho_11, 1-rho_1-(rho_1-rho_11),1-rho_1  
-               ) )
-}
 
@@
 
+inputPanel(
+  selectInput("gradient", label = "select gradient:",             choices = c("b", "L"), selected = "b")
+  )
+
 renderPlot({
+
+  source("../code/simfunctions.r")
 
@@
 
-parms$b <- seq(0,1,.005) #seq(0,1,.002)+0.001
-parms$rho_ini <- c(0.9, 0.01)
-
-iterations <- expand.grid(parms)
-iterations <- cbind(ID = 1:dim(iterations)[1],iterations)
-
-iterations$f[iterations$f < iterations$b] <- iterations$b[iterations$f < iterations$b]
-
-foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
-  source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")
-
-  model_parms <- as.list(iterations[iteration,])
-
-  # running the ode-solver
-  runmodel <- ode(y = ini_rho(model_parms$rho_ini), func = odesys_spex, times = c(1,1000), parms = model_parms)
-
-  # transfer into ouput and calculate missing rho values
-  return(as.data.frame(runmodel)[dim(runmodel)[1],])
-  } -> output
-
-output <- cbind(output, iterations)
-
-plot(rho_1 ~ b, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
+ if(input$gradient == "b") bifurcation(parms,"b", c(0,1))
+ if(input$gradient == "L") bifurcation(parms, "L", c(0,12))
 

### 2015-03-04 documentation/model_formulation.Rmd
model_formulation.Rmd: apply bifurcation() and attractor()

@@
 ```{r, echo=FALSE}
-defplot(rho,  C(rho, m = parms$m, L = parms$L, h = parms$h, a = parms$a ) )
+defplot(rho,  C(rho, m = parms$m, L = parms$L, h = parms$h, a = parms$a ), ylab = "growth/mortality" )
 
@@
 
-defplot(rho,  C(rho, m = parms$m, L = parms$L, h = parms$h, a = parms$a ) )
+defplot(rho,  C(rho, m = parms$m, L = parms$L, h = parms$h, a = parms$a), ylab = "growth/mortality" )
 
@@
     L = 6, 
-    alpha = 0,
-    q = 0,
+    alpha = 0.3,
+    q = 0.5,
     c = 0,
@@
 
-rho <- c(rho_1 = 0.9, rho_11 = 0.8, rho_10 = 0.1, rho_00 = 0, rho_0 = 0.1)
-
-d_rho_1 <- function(rho, parms) { 
-  with(parms,
-       r * (b + (1 - b) * f * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) - m * rho[1] - ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L * (1 - p * rho[2]/rho[1]))/(1 +(a + v*rho[2]/rho[1]) * h  * rho[1]^( 1 + q))
-       )  
-  }
-
-d_rho_11 <- function(rho,  parms) { 
-  with(parms,
-       2* (rho[1] - rho[2]) * r * ( b + (1 - b) * f * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) / (1-rho[1]) - 2 * rho[2] * m  - 2 * rho[2] * ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L * (1 - p * rho[2]/rho[1]))/(1 +(a + v*rho[2]/rho[1]) * h  * rho[1]^( 1 + q))
-       )  
-  }
-
-odesys_spex <- function(t, rho, parms) {
-  rho_1 <- d_rho_1(rho, parms)
-  rho_11 <- d_rho_11(rho, parms)
-  list(c(rho_1, rho_11, rho_1-rho_11, 1-rho_1-(rho_1-rho_11),1-rho_1  
-               ) )
-}
-
 # running the ode-solver
-ODE_spex <- ode(y = rho, func = odesys_spex, times = 1.1^(1:50), parms = parms)
+ODE_spex <- ode(ini_rho(0.9,0.8), func = odesys_spex, times = 1.1^(1:50), parms = parms)
 
@@
 lines(spexmodel$rho_one ~ spexmodel$time, col = "grey50")
+
+meanfield <- as.data.frame( ode(ini_rho(0.9,0.8)[1], func = odesys_mean, times = 1.1^(1:50), parms = parms))
+lines(rho_1 ~ time , data = meanfield, col = "red3")
 
@@
 ```{r, echo=FALSE}
-
 attractor(parms)
-
 ```
@@
     f = 0.8, #  local facilitation
-    b = 0.25,  # environmental quality inverse of aridity
+    b = 0.3,  # environmental quality inverse of aridity
     K = 0.9,  # carrying capacity
-    alpha = 0.1, # water runoff
+    alpha = 0, # water runoff
     c = 0.2, # local competition
@@
 ```{r, echo=FALSE, fig.height = 3}
+par(mfrow = c(1,2))
+bifurcation(parms, "L", c(0,12))
 
-parms$b <- seq(0,1,.005) #seq(0,1,.002)+0.001
-parms$rho_ini <- c(0.9, 0.01)
-
-iterations <- expand.grid(parms)
-iterations <- cbind(ID = 1:dim(iterations)[1],iterations)
-
-iterations$f[iterations$f < iterations$b] <- iterations$b[iterations$f < iterations$b]
-
-par(mfrow= c(1,2), bty = "n", mar = c(4,4,1,0))
-
-foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
-  source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")
-
-  model_parms <- as.list(iterations[iteration,])
-
-  # running the ode-solver
-  runmodel <- ode(y = ini_rho(model_parms$rho_ini), func = odesys_spex, times = c(1,1000), parms = model_parms)
-
-  # transfer into ouput and calculate missing rho values
-  return(as.data.frame(runmodel)[dim(runmodel)[1],])
-  } -> output
-
-output <- cbind(output, iterations)
-
-plot(rho_1 ~ b, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1), cex = 0.5)
-
-parms$L <- seq(0,20,0.2) #seq(0,1,.002)+0.001
-parms$b <- 0.4
-parms$rho_ini <- c(0.9, 0.01)
-
-iterations <- expand.grid(parms)
-iterations <- cbind(ID = 1:dim(iterations)[1],iterations)
-
-iterations$f[iterations$f < iterations$b] <- iterations$b[iterations$f < iterations$b]
-
-foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
-  source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")
-
-  model_parms <- as.list(iterations[iteration,])
-
-  # running the ode-solver
-  runmodel <- ode(y = ini_rho(model_parms$rho_ini), func = odesys_spex, times = c(1,1000), parms = model_parms)
-
-  # transfer into ouput and calculate missing rho values
-  return(as.data.frame(runmodel)[dim(runmodel)[1],])
-  } -> output
-
-output <- cbind(output, iterations)
-
-plot(rho_1 ~ L, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1), cex = 0.5)
+bifurcation(parms,"b", c(0,1))
 

### 2015-03-06 documentation/model_exploration.Rmd
model exploration edits

@@
 This means, they can shift their attractor between a vegetated state and a degraded state, if they experience a sufficient perturbation.
-The question is how resilience thinking in managed rangelands can be improved [@folke06; @walker04; @walker12; @folke10].
+The question is how resilience thinking in managed rangelands in arid and semi-arid ecosystems can be improved [@folke06; @walker04; @walker12; @folke10].
 How can the ecosystem be put in a state that is far from the tipping point or the critical thresholds [@may75; @westoby89; @walker81]?
 
-It aims at an integration of spatial grazing mortality into a graphical population-dynamics framework as developed by @noy-meir75 and extended by @vandekoppel97 and @rietkerk97.
-The model is of conceptual character and identifies how certain management practices or characteristics of vegetation affect the thresholds of catastrophic shifts.
+It aims at an integration of local feedback mechanisms in drylands into a graphical population-dynamics framework as developed by @noy-meir75 and extended by @vandekoppel97 and @rietkerk97.
+The model identifies how certain mechanisms of vegetation affect the thresholds of catastrophic shifts.
 This kind of model can be used to inform state-and-transition models that are of high relevance in decisionmaking in the management of rangelands [@westoby89; @briske05; @bestelmeyer03; @briske08].
@@
 The model is neutral in terms of species and assumes only one relevant type of vegetation and one population of grazer.
-Noy-meir notes that the model can apply also if there is one plant species that dominates the ecological dynamics of the entire system.
-[@noy-meir75]
+Noy-meir notes that the model can apply also if there is one plant species that dominates the ecological dynamics of the entire system [@noy-meir75].
 We stick to this approach of neglecting species-specific differences.

### 2015-03-12 documentation/model_exploration.Rmd
update model exploration

@@
 par(mfrow = c(2,2), bty ="n", las = 1 )
+
 defplot(rho, G(rho), ylab = "plant mortality/growth", col = colpal$grow[1])
@@
 parms <- defparms
-parms$f = 0.9
-parms$b = 0.4
+parms$f = 0.8
+parms$b = 0.3
 
-attractor(parms)
+attractor(parms, pairapprox = FALSE, localvals = TRUE)
 
-bifurcation(parms, "b", c(0,1))
+bifurcation(parms, pairapprox = TRUE, "b", c(0,1))
 
@@
 
-attractor(parms)
+attractor(parms,  pairapprox = FALSE, localvals = TRUE)
 
-bifurcation(parms, "b", c(0,1))
+bifurcation(parms, pairapprox = TRUE, "b", c(0,1))
 
@@
 parms$L = 8
-parms$p = 0.8
+parms$p = 0.7
 
-attractor(parms)
+attractor(parms, pairapprox = FALSE, localvals = TRUE)
 
-bifurcation(parms, "L", c(0,12))
+bifurcation(parms, pairapprox = TRUE, "L", c(0,12))
 
@@
 parms$L = 6
-parms$q = 0
+parms$q = 0.5
 parms$v = 0.9
 
-attractor(parms)
+attractor(parms, pairapprox = FALSE, localvals = TRUE)
 
-bifurcation(parms, "L", c(0,12))
+bifurcation(parms, pairapprox = FALSE, "L", c(0,12))
 

### 2015-03-12 documentation/model_exploration.Rmd
add description to model exploration

--- 2015-03-06 documentation/model_exploration.Rmd
@@
 This means, they can shift their attractor between a vegetated state and a degraded state, if they experience a sufficient perturbation.
-The question is how resilience thinking in managed rangelands in arid and semi-arid ecosystems can be improved [@folke06; @walker04; @walker12; @folke10].
+The question is how resilience thinking [@folke06; @walker04; @walker12; @folke10]  in managed rangelands in arid and semi-arid ecosystems can be improved.
 How can the ecosystem be put in a state that is far from the tipping point or the critical thresholds [@may75; @westoby89; @walker81]?
 
-It aims at an integration of local feedback mechanisms in drylands into a graphical population-dynamics framework as developed by @noy-meir75 and extended by @vandekoppel97 and @rietkerk97.
+Therefore, this study ims at an integration of local feedback mechanisms in drylands into a graphical population-dynamics framework as developed by @noy-meir75 and extended by @vandekoppel97 and @rietkerk97.
 The model identifies how certain mechanisms of vegetation affect the thresholds of catastrophic shifts.
@@
 
-## spatially-explicit mechanisms
+## local mechanisms
 
-The spatial extend at which ecological feedback mechanisms act often differ between small local scales and the entire landscape.
-Previous models have shown that feedbacks acting simultaneously but on different spatial scales emerge and alter the stability properties of the ecosystem [@kefi07a].
-Therefore, we are going to implement different kinds of local feedback, complementary to the global feedbacks described above, into the model framework.
+The graphical approach pursued by @noy-meir75 and @vandekoppel97 is limited to systems that are spatially homogeneous.
+The functions of vegetation growth and mortality are well defined by only one parameter, vegetation cover.
+However, in arid and semi-arid ecosystems vegetation cover turns out to be insufficient to describe the state of the ecosystem.
+Moreover, in a landscape where vegetation grows in aggregated clusters, spatial vegetation structure is important in determining the amount of growth or mortality.
+This is due to the dependence of the underlying feedback mechanisms on the locally present vegetation, rather than on total cover on the landscape scale.
+We subsequently describe positive feedback mechanisms that are known to play a role in the determination of growth and mortality in arid and semi-arid rangelands.
 
-The graphical approach pursued by @noy-meir75 and @vandekoppel97 is limited by it's dimensionality.
-The functions of vegetation growth and mortality are well defined by only one parameter, vegetation cover.
-Especially in arid and semi-arid ecosystems vegetation cover turned out to be insufficient to describe the state of the ecosystem.
+The simplest approach to explore these models is by mean-field approximation, which assumes that spatial structure of vegetation is random and that local cover equals total vegetation cover.
+This allows the estimation of the steady states of the ecosystem under a given set of parameters, or along a parameter gradient.
+
+<!-- The spatial extend at which ecological feedback mechanisms act often differ between small local scales and the entire landscape.
+Previous models have shown that feedbacks acting simultaneously but on different spatial scales emerge and alter the stability properties and spatial structure of the ecosystem [@kefi07a].
+-->
 
@@
 
-##
+We conclude that the different kinds of local feedbacks can fundamentally alter the critical thresholds of arid rangelands.
+Knowing about their relative importance is vital for the management of arid ecosystems under grazing pressure.
+
+But how can this knowledge be put to practice?
+The acknowledgement of alternative stable states in rangelands led to the development of state-and-transition models, i.e.
+descriptions of the ecosystems states and the probabilities and conditions for transitions between them.
+
+## trajectories
+
+The mean-field approximation allows us to estimate the alternative states of an ecosystem under a given set of circumstances, it is not very informative about which starting conditions would lead to the one or the other alternative steady state.
+
+Here, pair-approximation models, that keep track not only of the total vegetation cover but also of the local cover, provide more information and can help to estimate the boundary between the alternative steady state attractors.
+
+We apply this approach to different scenarios, to discuss under which circumstances an ecosystem might respond positively to a given management condition (e.g. the increase or decrease of livestock)
+
+## spatial structure as state indicator
+
+## Discussion
 

### 2015-03-12 code/simfunctions.r
update associational resistance in simfunctions.r

@@ -313,7 +313,7 @@ runCA <- function(init, parms, width = 100, height = 100, delta = 0.1, t_max = 1
     
     growth[growth < 0] <- 0
     
-    death <- with(parms_temp, (m + ( (a + v*q_one_one) * L * (1-p*q_one_one) * rho_one^(q) )/( 1 + (a + v*q_one_one) * h * rho_one^(1+q) )) *delta)   # set probability of death for each cell
+    death <- with(parms_temp, (m + ( (a * (1-p*q_one_one)) * L * rho_one^(q) )/( 1 + ((a * (1-p*q_one_one))  + v*q_one_one) * h * rho_one^(1+q) )) *delta)   # set probability of death for each cell
     
     death[death < 0] <- 0
     
@@ -380,7 +380,7 @@ d_rho_mean <- function(rho, parms) {
   growth <- with(parms, r *  rho^( 1 + alpha) * (b + (1-b) * f * rho)  * (1 - rho/(K * (1-c*rho) ) ))
   if(growth <= 0| is.na(growth)) {growth <- 0}
   
-  mortality <- with(parms,  m * rho + ((a+v*rho) * rho^( 1 + q) * L * (1-p*rho))/(1 + (a+v*rho) * h * rho^( 1 + q)) )  
+  mortality <- with(parms,  m * rho + ((a* (1-p*rho)+v*rho) * rho^( 1 + q) * L )/(1 + (a* (1-p*rho)+v*rho) * h * rho^( 1 + q)) )  
   if(mortality <= 0 | is.na(mortality)) {mortality <- 0}
   
   return(growth - mortality)
@@ -402,7 +402,7 @@ d_rho_1 <- function(rho, parms) {
                  r * (b + (1 - b) * f * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) )) 
   if(growth <= 0 | is.na(growth)) {growth <- 0}
 
-  mortality <- with(parms, m * rho[1] + ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L * (1 - p * rho[2]/rho[1]))/(1 +(a + v*rho[2]/rho[1]) * h  * rho[1]^( 1 + q)) )
[…]

### 2015-03-12 documentation/model_formulation.Rmd
update associational resistance in model_formulation.Rmd

@@
 
+**Update 12.03.2015:** This is not a straightforward logic.
+If plants are attractive to grazers, they increase local presence of grazers.
+Thus the model should ensure that attractivity redistributes grazing to plants with high local cover.
+Thus a value of $v = 0$ means that grazing distributes homogeneously, whereas $v = 1$ means that grazing goes with maximal intensity on plants with many neighbors.
+But it would need to make sure that grazing is not exploding at high cover.
+It is difficult to define this properly and in a simple way, though.
+
 ```{r, echo=FALSE}
@@
 $$ L = L^* (1 - p  q_{1|1}) $$
+
+**Update 12.03.2015:** since the effect is of less relevance at high cover, since the development of thorns is only induced if local pressure is already high,  it makes more sense to implement associational protection as an effect on search efficiency.
+
+$$ a = a^* (1- p q_{1|1}) $$
+
+This causes a to fall to its minimal value ($1-p$) at full local cover.
 

### 2015-03-12 documentation/model_exploration.Rmd
update model exploration

@@
     fig_width: 7
+    fig_caption: yes
+    fig_crop: no
+    includes:
+      in_header: include.tex
 fontsize: 12pt
@@
 
-# alternative stable states due to feedbacks
+## alternative stable states due to global feedbacks
 
@@
 
-## local mechanisms
+## local feedback mechanisms
 
@@
 
-### facilitation
+#### facilitation
 
@@
 Such mechanisms can be species-pair specific but they also play a general abiotic role as the cause of local amelioration of the environmental conditions of the neighborhood of a 'nurse' plant.
+The mechanisms commonly subsumised under local facilitation are shading effects, soil water retention, aggregation of litter and fixation of organic matter in top soils [@bautista].
 
-```{r, echo = FALSE}
+We implement local facilitation as a mechanism that increases growth rates on empty ground locally with increasing local plant cover, $q_{0|1}$.
+
+```{r, echo = FALSE, fig.cap="a) mean field approximation of density dependent mortality and growth of vegetation with local facilitation.
+As plant cover increases, the effect of facilitation is stronger.
+solid points show steady states, open circle shows unstable equilibrium.
+b) bifurcation diagram of vegetation cover over a gradient of environmental quality (black: steady states, grey: unstable equilibria).
+High facilitation has strong stabilising effects on the vegetated state.
+"}
 
@@
 
-bifurcation(parms, pairapprox = TRUE, "b", c(0,1))
+parms$f = 0.8
+bifurcation(parms, pairapprox = FALSE, "b", c(0,1))
+
+parms$f = 0.4
+bifurcation(parms, pairapprox = FALSE, "b", c(0,1), add =TRUE)
+
+parms$f = 0
+bifurcation(parms, pairapprox = FALSE, "b", c(0,1), add =TRUE)
+
+#parms$b = 0.5
+#parms$f = 0.8
+#bifurcation(parms, pairapprox = FALSE, "L", c(0,12))
+#
+#parms$f = 0.4
+#bifurcation(parms, pairapprox = FALSE, "L", c(0,12), add = TRUE)
+#
+#parms$f = 0
+#bifurcation(parms, pairapprox = FALSE, "L", c(0,12), add = TRUE)
 
@@
 
-### local competition
+Facilitation has a strong stabilizing effect on the vegetated state of the ecosystem and extends the range where vegetation can persist far towards very harsh environmental conditions.
+
+#### local competition
+
+Local competition is a mechanism counteracting local facilitation.
+If plants grow close together they exploit the locally available resources, like water or nutrients.
+Thus, the perceived carrying capacity of an empty space is reduced as the neighboring plant cover $q_{1|1}$ increases.
 
@@
 
-parms$c = 0.3
+parms$c = 0.6
 
@@
 
-bifurcation(parms, pairapprox = TRUE, "b", c(0,1))
+parms$c = 0
+bifurcation(parms, pairapprox = FALSE, "b", c(0,1))
+
+parms$c = 0.3
+bifurcation(parms, pairapprox = FALSE, "b", c(0,1), add = TRUE)
+
+parms$c = 0.6
+bifurcation(parms, pairapprox = FALSE, "b", c(0,1), add = TRUE)
 
@@
 
-### associational resistance
+Where local competition is important, the total vegetation cover is reduced, since competition affects growth primarily at high cover.
+The critical thresholds of sustaineable environmental quality are hardly affected.
+
+#### associational resistance
+
+In arid rangelands many plants develop protective physical structures, like thorns or cushion growth, or secondary plant metabolites to reduce the impact of grazing.
+This investment usually not only benefits the plant itself but also other plants in the direct neighborhood, since it reduces the accessibility of space or acts as a repellent for herbivores.
+This associational resistance is providing a local decrease of search efficiency (or an increase of handling effort) for grazers, which is reducing total consumption, and thus acts as a positive local feedback on vegetation cover.
 
@@
 
-parms$L = 8
-parms$p = 0.7
+parms$L = 6
+parms$p = 0.6
 
@@
 
-bifurcation(parms, pairapprox = TRUE, "L", c(0,12))
+parms$p = 0.0
+bifurcation(parms, pairapprox = FALSE, "L", c(0,12))
+text(3,0.7, "0.0")
+
+parms$p = 0.6
+bifurcation(parms, pairapprox = FALSE,add = TRUE, "L", c(0,12))
+text(7,0.65, "0.6")
+
+parms$p = 1
+bifurcation(parms, pairapprox = FALSE,add = TRUE, "L", c(0,12))
+text(8,0.81, "1.0")
+
+#parms$p = 0.0
+#bifurcation(parms, pairapprox = FALSE, "b", c(0,1))
+#text(3,0.7, "0.0")
+
+#parms$p = 0.6
+#bifurcation(parms, pairapprox = FALSE,add = TRUE, "b", c(0,1))
+#text(7,0.65, "0.6")
+
+#parms$p = 1
+#bifurcation(parms, pairapprox = FALSE,add = TRUE, "b", c(0,1))
+#text(8,0.81, "1.0")
 
@@
 
-### attractant-decoy
+Associational resistance is a stabilizing feature of vegetation cover.
+It greatly extends the range of persistent vegetation cover under high levels of livestock grazing, since it effectively reduces grazing impact at high vegetation cover.
+
+#### attractant-decoy
+
+As an opposing mechanism to associational resistance, palatable plants attract grazers and
 
@@
 
+parms$v = 0.9
+
 bifurcation(parms, pairapprox = FALSE, "L", c(0,12))
+
+text(8,0.75, "0.9")
+
+parms$v = 0.5
+
+bifurcation(parms, pairapprox = FALSE, "L", c(0,12), add = TRUE)
+
+text(8,0.75, "0.5")
+
+parms$v = 0
+
+bifurcation(parms, pairapprox = FALSE, "L", c(0,12), add = TRUE)
+
+text(8,0.75, "0")
 
@@
 
-We conclude that the different kinds of local feedbacks can fundamentally alter the critical thresholds of arid rangelands.
+Note that competition too could cause increased mortality due to high local cover.
+For reasons of simplicity we do not mix those effects here, but consider it of similar effect as an inverse of the associational resistance.
+
+### conclusion
+
+We conclude that the different kinds of local feedbacks can fundamentally alter the stability properties of arid rangelands.
 Knowing about their relative importance is vital for the management of arid ecosystems under grazing pressure.

### 2015-04-14 documentation/model_formulation.Rmd
update attractant-decoy and associational resistance

@@
 
+defparms <- list(  
+    m = 0.05,
+    r = 1,
+    b = 0.9, 
+    K = 0.9, 
+    a = 0.6,
+    h = 100,
+    L = 12, 
+    alpha = 0,
+    q = 0,
+    c = 0,
+    f = 0,
+    v = 0,
+    p = 0
+
+  )
 parms <- defparms
@@
 parms <- defparms
-parms$v = 0.9
+parms$v = 1
+parms$q = 1
 attractor(parms, pairapprox = FALSE, meanfield = FALSE, localvals = TRUE)
@@
 
-**Update 12.03.2015:** since the effect is of less relevance at high cover, since the development of thorns is only induced if local pressure is already high,  it makes more sense to implement associational protection as an effect on search efficiency.
+**Update 12.03.2015:** the effect is of less relevance at high cover, since the development of thorns is only induced if local pressure is already high.
+At high cover it is the handling time limitation that acts as an individual protection.
+Thus, it makes more sense to implement associational protection as an effect on search efficiency.
 
@@
 
-This causes a to fall to its minimal value ($1-p$) at full local cover.
+This causes $a$ to fall to its minimal value ($1-p$) at full local cover.
+If $p=1$ the feeding is completely neutralized under high local cover, which also causes a drop in grazing mortality if cover is really high.
+This might be unrealistic, for the biological reasons named above.
+So, values of $p \leq 0.9$ seem to make sense.
+
+**Update 26.03.2015:** Also, the formulations of local effects on mortality are unclear about their compatibility, since the attractant effect it is additive while the protective effect is multiplied.
+Thus, I have to specify
+
+$$ a =  (a^*  + v q_{1|1} ) (1- p q_{1|1}) $$
 
@@
 parms <- defparms
-parms$p = 0.8
+parms$p = 0.99
 attractor(parms, pairapprox = FALSE, meanfield = FALSE, localvals = TRUE)
@@
 
-$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b + (1 - b)  f \rho_1)) (1-\frac{\rho_1}{K(1 - c \rho_1}) - \rho_{1}m - \frac{(a  + v \rho_1)L (1 - p \rho_1)\rho_1^{1+q}}{1+(a  + v \rho_1)h\rho_1^{1+q}}  $$
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b + (1 - b)  f \rho_1)) (1-\frac{\rho_1}{K(1 - c \rho_1}) - \rho_{1}m - \frac{(a  + v \rho_1)(1 - p \rho_1)L \rho_1^{1+q}}{1+(a  + v \rho_1)h\rho_1^{1+q}}  $$
 
@@
 
-$$ C = \rho_{1}m + \frac{(a  + v \rho_1)L (1 - p \rho_1)\rho_1^{1+q}}{1+(a  + v \rho_1)h\rho_1^{1+q}}$$
+$$ C = \rho_{1}m + \frac{(a  + v \rho_1)(1 - p \rho_1)L \rho_1^{1+q}}{1+(a  + v \rho_1)h\rho_1^{1+q}}$$
 

### 2015-04-14 code/simfunctions.r

@@ -313,7 +313,7 @@ runCA <- function(init, parms, width = 100, height = 100, delta = 0.1, t_max = 1
     
     growth[growth < 0] <- 0
     
-    death <- with(parms_temp, (m + ( (a * (1-p*q_one_one)) * L * rho_one^(q) )/( 1 + ((a * (1-p*q_one_one))  + v*q_one_one) * h * rho_one^(1+q) )) *delta)   # set probability of death for each cell
+    death <- with(parms_temp, (m + ( (a+ v*q_one_one) * (1-p*q_one_one) * L * rho_one^(q) )/( 1 + (a+ v*q_one_one) * (1-p*q_one_one) * h * rho_one^(1+q) )) *delta)   # set probability of death for each cell
     
     death[death < 0] <- 0
     
@@ -380,7 +380,7 @@ d_rho_mean <- function(rho, parms) {
   growth <- with(parms, r *  rho^( 1 + alpha) * (b + (1-b) * f * rho)  * (1 - rho/(K * (1-c*rho) ) ))
   if(growth <= 0| is.na(growth)) {growth <- 0}
   
-  mortality <- with(parms,  m * rho + ((a* (1-p*rho)+v*rho) * rho^( 1 + q) * L )/(1 + (a* (1-p*rho)+v*rho) * h * rho^( 1 + q)) )  
+  mortality <- with(parms,  m * rho + ( (a+v*rho) * (1-p*rho) * rho^( 1 + q) * L )/(1 + (a+v*rho) * (1-p*rho) * h * rho^( 1 + q)) )  
   if(mortality <= 0 | is.na(mortality)) {mortality <- 0}
   
   return(growth - mortality)
@@ -402,7 +402,7 @@ d_rho_1 <- function(rho, parms) {
                  r * (b + (1 - b) * f * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) )) 
   if(growth <= 0 | is.na(growth)) {growth <- 0}
 
[…]

### 2015-04-14 code/simfunctions.r
fix stability issues, change to runge kutta solver

@@ -546,10 +546,11 @@ runODE_spex <- function(starting, model_parms, times = c(0,1000))  {
 
 runODE_meanfield <- function(starting, model_parms, times = c(0,1000))  {
   
-  out <- as.data.frame(ode(starting, func = odesys_mean, times = times, parms = model_parms))
+  out <- as.data.frame(ode(starting, func = odesys_mean, times = times, parms = model_parms, method = "ode45"))
   
   names(out) <- c("time", "rho_1")
   
+  out[out < 1e-6] <- 0
   return(round(out,6))
 }
 
@@ -568,11 +569,11 @@ attractor <- function(model_parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini
     lines(rho,G(rho, rho, model_parms), col = colpal$grow[1], lwd = 2)
     
     
-    runmodel_high <- as.data.frame(ode(y = 0.99, func = odesys_mean, times = c(1,1000), parms = model_parms))
+    runmodel_high <- as.data.frame(ode(y = 0.99, func = odesys_mean, times = c(0,1000), parms = model_parms, method = "ode45"))
     
     points(runmodel_high[2,2],G(runmodel_high[2,2], runmodel_high[2,2], model_parms), xpd = TRUE, pch = 20, cex = 2)
     
-    runmodel_low <- as.data.frame(ode(y = 0.0001, func = odesys_mean, times = c(1,1000), parms = model_parms))
+    runmodel_low <- as.data.frame(ode(y = 0.0001, func = odesys_mean, times = c(0,1000), parms = model_parms, method = "ode45"))
     
     points(runmodel_low[2,2],G(runmodel_low[2,2],runmodel_low[2,2],model_parms), xpd = TRUE, pch = 20, cex = 2)
     
@@ -582,7 +583,7 @@ attractor <- function(model_parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini
     for(i in 1:10) {
       
       mid <- (lo+hi)/2 
-      runmodel_mid <- as.data.frame(ode(mid, func = odesys_mean, times = c(1,1.2), parms = model_parms))
+      runmodel_mid <- as.data.frame(ode(mid, func = odesys_mean, times = c(0,0.2), parms = model_parms, method = "ode45"))
       
       if(runmodel_mid[2,2] > mid) { hi <- mid } else { lo <- mid}
       
[…]

### 2015-04-14 code/simfunctions.r
update bifurcation plot: return values

@@ -680,13 +680,15 @@ bifurcation <- function(parms, over, xrange, res = 201, times = c(0,1000), ini =
   
   require(foreach)
   
+  out <- list() # initialise output object
+  
   parms[[over]] <- seq(xrange[1],xrange[2],length = res)
   
   parms$rho_ini <- ini
   
   iterations <- expand.grid(parms)
   iterations <- cbind(ID = 1:dim(iterations)[1],iterations)
-      
+  iterations$b <- as.numeric(as.character(iterations$b))
   
   if(pairapprox) {
     
@@ -703,6 +705,9 @@ bifurcation <- function(parms, over, xrange, res = 201, times = c(0,1000), ini =
     
     output <- cbind(iterations,output)
     
+    out$pairapprox <- output
+    out$pairapprox$q_11 <- out$pairapprox$rho_11/out$pairapprox$rho_1
+    
     upper <- output[output$rho_ini == ini[1],][which(round(output[output$rho_ini == ini[1],]$rho_1,4) != round(output[output$rho_ini == ini[2],]$rho_1,4)),]
     lower <- output[output$rho_ini == ini[2],][which(round(output[output$rho_ini == ini[2],]$rho_1,4) != round(output[output$rho_ini == ini[1],]$rho_1,4)),]
     
@@ -711,7 +716,7 @@ bifurcation <- function(parms, over, xrange, res = 201, times = c(0,1000), ini =
         
         source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")
         
-        model_parms <- upper[upper[, over] == i,]
+        model_parms <- upper[ upper[, over] == i,]
         
         hi_1 <- upper[upper[, over] == i,]$rho_1
         lo_1 <- lower[lower[, over] == i,]$rho_1
@@ -742,17 +747,22 @@ bifurcation <- function(parms, over, xrange, res = 201, times = c(0,1000), ini =
       
       output_unstable <- cbind(upper[,1:16],output_unstable)
       
+      out$pairapprox_unstable <- output_unstable
     }
     
     if(!add) {
-      plot(output$rho_1 ~ output[,over], xlab = over, ylab = "vegetation cover", type = "p", pch  = 20, ylim = c(0,1), cex = 0.5, yaxp = c(0,1,2))
-    } else {
-      points(output[,over],output$rho_1, pch  = 20,  cex = 0.5)
-    }
+      plot(NA,NA, 
[…]

### 2015-04-14 code/simfunctions.r
add bifurcation 3D

@@ -856,4 +856,95 @@ bifurcation <- function(parms, over, xrange, res = 201, times = c(0,1000), ini =
  return(out)
 }
   
+
+
+bifurcation3d <- function(parms, over, xrange, res = 201, times = c(0,1000), ini = c(0.9, 0.0001) , add = FALSE ) {
   
+  require(foreach)
+  
+  parms[[over]] <- seq(xrange[1],xrange[2],length = res)
+  
+  parms$rho_ini <- ini
+  
+  iterations <- expand.grid(parms)
+  iterations <- cbind(ID = 1:dim(iterations)[1],iterations)
+  iterations$b <- as.numeric(as.character(iterations$b))
+  
+    foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
+      source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")
+      
+      model_parms <- as.list(iterations[iteration,])
+      
+      # running the ode-solver
+      runmodel <- runODE_spex(ini_rho(model_parms$rho_ini), model_parms, times = times) 
+      
+      return(tail(runmodel,1))
+    } -> output
+    
+    output <- cbind(iterations,output)
+    
+    upper <- output[output$rho_ini == ini[1],][which(round(output[output$rho_ini == ini[1],]$rho_1,4) != round(output[output$rho_ini == ini[2],]$rho_1,4)),]
+    lower <- output[output$rho_ini == ini[2],][which(round(output[output$rho_ini == ini[2],]$rho_1,4) != round(output[output$rho_ini == ini[1],]$rho_1,4)),]
+    
+    if(nrow(upper)>0) {
+      foreach(i = upper[,over], .combine = rbind, .packages = c("deSolve") ) %dopar% {
+        
+        source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")
+        
+        model_parms <- upper[ upper[, over] == i,]
+        
+        hi_1 <- upper[upper[, over] == i,]$rho_1
+        lo_1 <- lower[lower[, over] == i,]$rho_1
+        hi_11 <- upper[upper[, over] == i,]$rho_11
+        lo_11 <- lower[lower[, over] == i,]$rho_11
+        
+        for(j in 1:10) {
+          
[…]

### 2015-04-14 code/simfunctions.r
update attractor: smooth lines and arrows, fix plotting mismatch

@@ -560,7 +560,8 @@ runODE_meanfield <- function(starting, model_parms, times = c(0,1000))  {
 
 attractor <- function(model_parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini = seq(0,1, length = 11), meanfield = TRUE, pairapprox = TRUE, localvals = FALSE) {
   
-  defplot(NA,NA, ylab = "plant mortality/growth")
+  plot(NA,NA, ylab = "plant mortality/growth", xlim = c(0,1), ylim= c(0,.25), bty = "n", xaxs = "i" , yaxs = "i")
+  
   rho <- seq(0,1,length = 100)
   
   if(meanfield) {
@@ -597,51 +598,65 @@ attractor <- function(model_parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini
   
   if(pairapprox) {
     
-    ini <- list(
-      rho_1 = rho_1_ini,
-      rho_11 = rho_11_ini,
-      rho_10 = NA,
-      rho_00 = NA,
-      rho_0 = NA
-    )
+ ini <- list(
+   rho_1 = rho_1_ini,
+   rho_11 = c(0, NA),
+   rho_10 = NA,
+   rho_00 = NA,
+   rho_0 = NA
+ )
+ 
+ ini <- expand.grid(ini)
+ 
+ ini$rho_11[is.na(ini$rho_11)] <- ini$rho_1[is.na(ini$rho_11)]
+ 
+ 
+ for(x in 1:nrow(ini)) {
+   temp <- ini_rho(ini[x,]$rho_1, ini[x,]$rho_11)
+   
+   #if(ini[x,]$rho_11 == 0) {
+      while(any(is.na(temp)) & ini[x,]$rho_11 <=ini[x,]$rho_1) {
+        ini[x,]$rho_11 <-  ini[x,]$rho_11+0.005
+        temp <- ini_rho(ini[x,]$rho_1, ini[x,]$rho_11)
+      }
+  
+   ini[x,]$rho_1 <- temp[1]
+   ini[x,]$rho_11 <- temp[2]
+   ini[x,]$rho_10 <- temp[3]
+   ini[x,]$rho_00 <- temp[4]
+   ini[x,]$rho_0 <- temp[5]
+ } 
+ 
+ 
+ ini$m_ini <- C(ini$rho_1, ini$rho_11/ini$rho_1, model_parms)
+ ini$g_ini <- G(ini$rho_1, ini$rho_10/1-ini$rho_0, model_parms)
+ 
[…]

### 2015-04-14 documentation/model_formulation.Rmd
fix initial offset for ODE (starting at 0 instead 1)

@@
   with(parms, r* b* rho^( 1 + alpha) * (1 - rho/K) - m * rho - (a * rho^( 1 + q) * L)/(1 + a * h * rho^( 1 + q)) )  
-  }
+  } # instead use function from simfunctions.r?
 
 # running the ode-solver
-ODEglobal_high <- ode(y = 0.9, func = odesys, times = 1.1^(1:50), parms = parms)
+ODEglobal_high <- ode(y = 0.9, func = odesys, times = 1.1^(0:50)-1, parms = parms)
 
@@
 # running the ode-solver
-ODEglobal_low <- ode(y = 0.1, func = odesys, times = 1.1^(1:50), parms = parms)
+ODEglobal_low <- ode(y = 0.1, func = odesys, times = 1.1^(0:50)-1, parms = parms)
 

### 2015-04-14 documentation/model_formulation.Rmd
add lattice

@@
 
+```{r, eval = FALSE, dev='png', echo = FALSE, fig.show='animate', fig.height = 4, fig.width = 4, interval = 0.01}
+if(Sys.info()[['sysname']] == "Linux") X11.options(antialias = "none") #for Linux Systems to enable pixel-wise plotting in (animated) gif-files. 
+  if(Sys.info()[['sysname']] == "Windows") windows.options(antialias = "none") #for Windows Systems to enable pixel-wise plotting in (animated) gif-files. 
+  width = nullmodel$timeseries[[1]]$dim[1]
+  height = nullmodel$timeseries[[1]]$dim[2]
+
+    for(i in 1:length(nullmodel$timeseries) ) {
+      par(mar = c(0,0,0,0))
+      plot(nullmodel$timeseries[[i]], grid = FALSE, ani = TRUE) 
+    }
+```
+
 ```{r, eval = FALSE, echo = FALSE, warning = FALSE}

### 2015-04-14 documentation/model_formulation.Rmd
update model formulation: bifurcation section

@@
     a = 0.6,
-    h = 50,
-    L = 6, 
-    alpha = 0.3,
-    q = 0.5,
+    h = 100,
+    L = 16, 
+    alpha = 0.1,
+    q = 1,
     c = 0,
@@
 
-If those effects coincide, potentially even with the global feedbacks, the intersections are difficult to estimate.
+If those effects coincide, potentially even with the global feedbacks, the intersections are difficult to estimate visually.
 
@@
     K = 0.9,  # carrying capacity
-    alpha = 0.1, # water runoff
+    alpha = 0, # water runoff
     c = 0.2, # local competition
-    a = 0.6, # search efficiency
-    v = 0.1, # attraction effect
+    a = 0.2, # search efficiency
+    v = 0.8, # attraction effect
     h = 50, # handling time (one individual on landscape unit)
-    p = 0.8, # protection effect
-    L = 6, # livestock units per landscape unit 
-    q = 1 # search ineffiency at low cover
+    p = 0.99, # protection effect
+    L = 5, # livestock units per landscape unit 
+    q = 0 # search ineffiency at low cover
   )
@@
 
-```{r, echo=FALSE}
+### Bifurcation diagrams
 
-parms <- list(
-    m = 0.05,  # intrinsic mortality of plants 
-    r = 1,   # growth rate
-    f = 0.8, #  local facilitation
-    b = 0.3,  # environmental quality inverse of aridity
-    K = 0.9,  # carrying capacity
-    alpha = 0, # water runoff
-    c = 0.2, # local competition
-    a = 0.6, # search efficiency
-    v = 0.1, # attraction effect
-    h = 50, # handling time (one individual on landscape unit)
-    p = 0.8, # protection effect
-    L = 6, # livestock units per landscape unit 
-    q = 1 # search ineffiency at low cover
-  )
+The following shows bifurcation diagrams of the mean field approximation over an environmental gradient, $b$, as well as over a gradient of livestock density, $L$.
 
-attractor(parms)
+```{r, echo=FALSE, fig.height = 3 }
+par(mfrow = c(1,2))
+parms$q = 0.4
+bifurcation(parms, "L", c(0,12)) -> over_L
+
+bifurcation(parms,"b", c(0,1)) -> over_b
 
@@
 
-### Bifurcation diagrams
+A dynamic Shiny implementation of that can be found under 'documentation/interactive_spexmodel_bifurcations.Rmd' ( run in RStudio, specify parallel backend.).
 
-The following shows bifurcation diagrams over an environmental gradient, $b$, as well as over a gradient of livestock density, $L$.
+The same graph using the pair-approximation.
 
@@
 par(mfrow = c(1,2))
-bifurcation(parms, "L", c(0,12))
+bifurcation(parms, "L", pairapprox = TRUE, meanfield = FALSE, c(0,12)) -> over_L
 
-bifurcation(parms,"b", c(0,1))
+bifurcation(parms,"b", pairapprox = TRUE, meanfield = FALSE, c(0,1)) -> over_b
+
+```
+
+The bifurcation diagram visualises the attractor only along the total cover which is only an incomplete visualisation, since the steady state is described by both, global vegetation cover $\rho_1$ and average local cover $q_{1|1}$ (or alternatively the cover of vegetated pairs $\rho_{11}$ but this is rather unintuitive).
+To visualise this, a three dimensional graph is necessary.
+
+```{r, echo=FALSE, fig.height = 5}
+
+persp(0:1,0:1, matrix(c(0,0,0,0), ncol = 2), zlim = c(0,1), ticktype = "detailed", zlab = "vegetation cover", ylab = "local cover" , xlab = "b", theta = -65, phi = 19, r = sqrt(3), d = 1) -> p
+
+  points(trans3d(over_b$pairapprox$b, over_b$pairapprox$rho_1, rep(0, length(over_b$pairapprox$rho_1)), pmat = p), col = "#D0D0D0", pch = 20,  cex = 0.5)
+  points(trans3d(over_b$pairapprox$b,rep(1, length(over_b$pairapprox$rho_1)), over_b$pairapprox$rho_11/over_b$pairapprox$rho_1,  pmat = p), col = "#D0D0D0", pch = 20,  cex = 0.5)  
+  points(trans3d(rep(1, length(over_b$pairapprox$rho_1)),over_b$pairapprox$rho_1, over_b$pairapprox$rho_11/over_b$pairapprox$rho_1,  pmat = p), col = "#D0D0D0", pch = 20,  cex = 0.5)
+
+  points(trans3d(over_b$pairapprox$b, over_b$pairapprox$rho_1, over_b$pairapprox$rho_11/over_b$pairapprox$rho_1, pmat = p), pch = 20,  cex = 0.5)
+
+```
+
+Decomposing it into its dimensions is easier to print.
+
+```{r, echo = FALSE}
+over_b$pairapprox$q_11 <- over_b$pairapprox$rho_11/over_b$pairapprox$rho_1
+over_b$pairapprox$q_11[is.na(over_b$q_11)] <- 0
+
+par(mfrow =c(1,2))
+  plot(over_b$pairapprox$b, over_b$pairapprox$rho_1, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
+  plot(over_b$pairapprox$b, over_b$pairapprox$q_11, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
+```
+
+The direct comparison of global cover vs. local cover indicates the amount of clustering along the stress gradient and at the critical thresholds. 
+We are going to compare these plots for a range of models to see how the spatial configuration changes with increasing stress under different mechanism scenarios.
+
+```{r, echo=FALSE, fig.width = 3}
+par(mfrow = c(1,1))
+  plot(over_b$pairapprox$rho_11, over_b$pairapprox$q_11, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
+abline(a = 0, b = 1, lty = 3, col = "#D0D0D0")
+points(0,0, pch=20, col = "white", cex = 2)
+ points(0,0, pch=20, cex = 0.5)
+
+```
+
+```{r}
 
@@
 
-A dynamic Shiny implementation of that can be found under 'documentation/interactive_spexmodel_bifurcations.Rmd' ( run in RStudio, specify parallel backend.).
+## analytical solutions
 
 The simulation backend is now all set up. As a final step, we want to check out if the model can be solved analytically.
-
-## analytical solutions
 

### 2015-04-14 documentation/model_exploration.Rmd
update model exploration: add attractor and bifurcation plots

@@
     K = 0.9, 
-    a = 0.6,
+    a = 0.2,
     h = 50,
@@
 The model is neutral in terms of species and assumes only one relevant type of vegetation and one population of grazer.
-Noy-meir notes that the model can apply also if there is one plant species that dominates the ecological dynamics of the entire system [@noy-meir75].
+Noy-meir emphasizes that the model can apply also if there is one plant species that dominates the ecological dynamics of the entire system [@noy-meir75].
 We stick to this approach of neglecting species-specific differences.
@@
 Such mechanisms can be species-pair specific but they also play a general abiotic role as the cause of local amelioration of the environmental conditions of the neighborhood of a 'nurse' plant.
-The mechanisms commonly subsumised under local facilitation are shading effects, soil water retention, aggregation of litter and fixation of organic matter in top soils [@bautista].
+The mechanisms commonly subsumised under local facilitation are shading effects, soil water retention, aggregation of litter and fixation of organic matter in top soils [@bautista04, @milchunas02].
 
-We implement local facilitation as a mechanism that increases growth rates on empty ground locally with increasing local plant cover, $q_{0|1}$.
+We implement local facilitation as a mechanism that increases local growth rates on empty ground with increasing local plant cover, $q_{1|0}$.
 
@@
 
-Facilitation has a strong stabilizing effect on the vegetated state of the ecosystem and extends the range where vegetation can persist far towards very harsh environmental conditions.
+The mean-field model shows that facilitation has a strong stabilizing effect on the vegetated state of the ecosystem and extends the range where vegetation can persist far towards very harsh environmental conditions [@kefi07a].
 
@@
 If plants grow close together they exploit the locally available resources, like water or nutrients.
-Thus, the perceived carrying capacity of an empty space is reduced as the neighboring plant cover $q_{1|1}$ increases.
+Thus, the perceived carrying capacity of an empty space is reduced as the neighboring plant cover $q_{1|0}$ increases.
 
@@
 Where local competition is important, the total vegetation cover is reduced, since competition affects growth primarily at high cover.
-The critical thresholds of sustaineable environmental quality are hardly affected.
+The critical thresholds of sustainable environmental quality are only marginally affected.
+Primarily, local competition affects the potential vegetation cover.
 
@@
 This investment usually not only benefits the plant itself but also other plants in the direct neighborhood, since it reduces the accessibility of space or acts as a repellent for herbivores.
-This associational resistance is providing a local decrease of search efficiency (or an increase of handling effort) for grazers, which is reducing total consumption, and thus acts as a positive local feedback on vegetation cover.
+This associational resistance [@milchunas02] is providing a local decrease of search efficiency (or an increase of handling effort) for grazers, which is reducing total consumption, and thus acts as a positive local feedback on vegetation cover.
 
@@
 parms$L = 6
-parms$p = 0.6
+parms$p = 1
 
@@
 Associational resistance is a stabilizing feature of vegetation cover.
-It greatly extends the range of persistent vegetation cover under high levels of livestock grazing, since it effectively reduces grazing impact at high vegetation cover.
+It greatly extends the range of persistent vegetation cover, especially under high levels of livestock grazing, since it effectively reduces grazing impact at high vegetation cover.
+At low cover, grazing effects are relatively strong because most plants are not associated in patches.
+The establishment of patches is facilitated, but is regularly broken by the plants intrinsic mortality.
 
@@
 
-As an opposing mechanism to associational resistance, palatable plants attract grazers and
+As an opposing mechanism to associational resistance, volatiles and visual impression of palatable plants attract grazers to spaces with many plants, which renders the grazers search more efficient.
+Thus, search efficiency of grazers is locally increased if a plant has many neighbors.
 
@@
 parms$L = 6
-parms$q = 0.5
-parms$v = 0.9
+parms$q = 0
+parms$a = 0.1
+parms$v = 1
 
@@
 
-parms$v = 0.9
+parms$v = 1
 
@@
 Knowing about their relative importance is vital for the management of arid ecosystems under grazing pressure.
+Four types of positive feedbacks were discussed here.
+They might be accelerating mortality at low local cover (attractant-decoy), decelerating mortality at high local cover (associational protection), accelerating growth at high cover (facilitation), or decelerating growth at high local cover (competition).
 
 But how can this knowledge be put to practice?
-The acknowledgement of alternative stable states in rangelands led to the development of state-and-transition models, i.e.
-descriptions of the ecosystems states and the probabilities and conditions for transitions between them.
+The acknowledgement of alternative stable states in rangelands led to the descriptions of the ecosystems states and the probabilities and conditions for transitions between them (i.e., state-and-transition models).
+
+## attractor of spatial structure
+
+The pair-approximation model is much more informative about the characteristics of the vegetated and the degraded state, since it describes the attractor as a function of both global and local cover.
 
@@
 
-The mean-field approximation allows us to estimate the alternative states of an ecosystem under a given set of circumstances, it is not very informative about which starting conditions would lead to the one or the other alternative steady state.
+The mean-field approximation allows us to estimate the alternative stable states of an ecosystem under a given set of circumstances, it is not very informative about which spatial configuration would develop into the one or the other alternative steady state.
 
-Here, pair-approximation models, that keep track not only of the total vegetation cover but also of the local cover, provide more information and can help to estimate the boundary between the alternative steady state attractors.
+Here, pair-approximation models, that keep track not only of the total vegetation cover but also of the local cover, provide more information about the transient dynamics and the domain of attraction and can help to estimate the boundary between the alternative steady state attractors.
 
@@
 
-## spatial structure as state indicator
+### facilitation and protection
+
+```{r, echo = FALSE, fig.cap="", fig.height = 2}
+
+par(mfrow = c(1,3), bty ="n", las = 1 ) 
+parms <- defparms
+parms$f = 1
+parms$b = 0.3
+parms$L = 6
+parms$p = 1
+
+attractor(parms, pairapprox = TRUE, localvals = FALSE, meanfield = FALSE)
+
+parms$f = 1
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1)) -> out_1
+
+parms$f = 0.5
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1), add =TRUE) -> out_2
+
+parms$f = 0
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1), add =TRUE) -> out_3
+
+###
+
+parms$f = 1
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "L", c(0,12))-> out_4
+
+parms$f = 0.5
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "L", c(0,12), add =TRUE)-> out_5
+
+parms$f = 0
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "L", c(0,12), add =TRUE)-> out_6
+
+```
+
+### facilitation and attraction
+
+```{r, echo = FALSE, fig.cap="", fig.height = 2}
+
+par(mfrow = c(1,3), bty ="n", las = 1 ) 
+parms <- defparms
+parms$f = 1
+parms$b = 0.3
+parms$L = 6
+parms$v = 1
+
+attractor(parms, pairapprox = TRUE, localvals = FALSE, meanfield = FALSE)
+
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1))
+
+parms$f = 0.5
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1), add =TRUE)
+
+parms$f = 0
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1), add =TRUE)
+
+parms$f = 1
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "L", c(0,12))
+
+parms$f = 0.5
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "L", c(0,12), add =TRUE)
+
+parms$f = 0
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "L", c(0,12), add =TRUE)
+
+```
+
+### competition and protection
+
+```{r, echo = FALSE, fig.cap="", fig.height = 2}
+
+par(mfrow = c(1,3), bty ="n", las = 1 ) 
+parms <- defparms
+parms$c = 1
+parms$L = 6
+parms$p = 1
+
+attractor(parms, pairapprox = TRUE, localvals = FALSE, meanfield = FALSE)
+
+bifurcation(parms,  pairapprox = TRUE, meanfield = FALSE, "b", c(0,1))
+
+parms$c = 0.5
+bifurcation(parms,  pairapprox = TRUE, meanfield = FALSE, "b", c(0,1), add =TRUE)
+
+parms$c = 0
+bifurcation(parms,  pairapprox = TRUE, meanfield = FALSE, "b", c(0,1), add =TRUE)
+
+parms$c = 1
+bifurcation(parms,  pairapprox = TRUE, meanfield = FALSE, "L", c(0,12))
+
+parms$c = 0.5
+bifurcation(parms,  pairapprox = TRUE, meanfield = FALSE, "L", c(0,12), add =TRUE)
+
+parms$c = 0
+bifurcation(parms,  pairapprox = TRUE, meanfield = FALSE, "L", c(0,12), add =TRUE)
+
+```
+
+### competition and attraction
+
+```{r, echo = FALSE, fig.cap="", fig.height = 2}
+
+par(mfrow = c(1,3), bty ="n", las = 1 ) 
+parms <- defparms
+parms$c = 1
+parms$L = 6
+parms$v = 1
+
+attractor(parms, pairapprox = TRUE, localvals = FALSE)
+
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1))
+
+parms$c = 0.5
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1), add =TRUE)
+
+parms$c = 0
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1), add =TRUE)
+
+parms$c = 1
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "L", c(0,12))
+
+parms$c = 0.5
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "L", c(0,12), add =TRUE)
+
+parms$c = 0
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "L", c(0,12), add =TRUE)
+
+```
 

### 2015-04-22 documentation/interactive_spexmodel_attractor.Rmd
interactive spexmodel attractor; use default data, plot pairapproximation

@@
 
-## developping the null model
+## developing the null model
 
@@
     r = 1,   # growth rate
-    f = 0.8, #  local facilitation
+    f = 1, #  local facilitation
     b = 0.25,  # environmental quality inverse of aridity
@@
     alpha = 0.1, # water runoff
-    c = 0.2, # local competition
-    a = 0.6, # search efficiency
+    c = 0, # local competition
+    a = 0.3, # search efficiency
     v = 0.1, # attraction effect
@@
     p = 0.8, # protection effect
-    L = 6, # livestock units per landscape unit 
-    q = 1 # search ineffiency at low cover
+    L = 5, # livestock units per landscape unit 
+    q = 0 # search ineffiency at low cover
   )
@@
   sliderInput("r", label = "max. growth rate:",
-              min = 0.0, max = 3, value = 1, step = 0.01),
+              min = 0.0, max = 3, value = defaultparms$r, step = 0.01),
   sliderInput("b", label = "environmental quality:",
-              min = 0.0, max = 1, value = 0.1, step = 0.01),
+              min = 0.0, max = 1, value = defaultparms$b, step = 0.01),
   sliderInput("K", label = "carrying capacity:",
-              min = 0.0, max = 1, value = 0.9, step = 0.01),
+              min = 0.0, max = 1, value = defaultparms$K, step = 0.01),
   sliderInput("alpha", label = "water runoff:",
-              min = 0.0, max = 1, value = 0, step = 0.01),
+              min = 0.0, max = 1, value = defaultparms$alpha, step = 0.01),
   sliderInput("f", label = "local facilitation:",
-              min = 0.0, max = 1, value = 0.9, step = 0.01),
+              min = 0.0, max = 1, value = defaultparms$f, step = 0.01),
   sliderInput("c", label = "local competition:",
-              min = 0.0, max = 1, value = 0.1, step = 0.01)
+              min = 0.0, max = 1, value = defaultparms$c, step = 0.01)
 )
@@
   sliderInput("m", label = "intrinsic mortality:",
-              min = 0.0, max = 1, value = 0.05, step = 0.01),
+              min = 0.0, max = 1, value = defaultparms$m, step = 0.01),
   sliderInput("a", label = "search efficiency:",
-              min = 0.0, max = 1, value = 0.6, step = 0.01), 
+              min = 0.0, max = 1, value = defaultparms$a, step = 0.01), 
   sliderInput("h", label = "handling time:",
-              min = 0.0, max = 200, value = 50, step = 1), 
+              min = 0.0, max = 200, value = defaultparms$h, step = 1), 
   sliderInput("L", label = "livestock units:",
-              min = 0.0, max = 100, value = 6, step = 1),
+              min = 0.0, max = 100, value = defaultparms$L, step = 1),
   sliderInput("q", label = "hill coefficient:",
-              min = 0.0, max = 1, value = 0, step = 0.01),
+              min = 0.0, max = 1, value = defaultparms$q, step = 0.01),
   sliderInput("v", label = "local attractivity:",
-              min = 0.0, max = 1, value = 0.0, step = 0.01),
+              min = 0.0, max = 1, value = defaultparms$v, step = 0.01),
   sliderInput("p", label = "local protection:",
-              min = 0.0, max = 1, value = 0.8, step = 0.01)
+              min = 0.0, max = 1, value = defaultparms$p, step = 0.01)
 )
@@
 
-attractor(parms)
+attractor(parms, localvals = FALSE, pairapprox = TRUE, meanfield = FALSE, rho_1_ini = seq(0,1, length = 21) )
 

### 2015-04-22 documentation/model_exploration.Rmd
model_exploration.Rmd; return simulation data and plot attractor

@@
 
-```{r, echo = FALSE, fig.cap="", fig.height = 2}
+```{r, echo = FALSE, fig.cap="", fig.height = 2, warning=FALSE}
 
@@
 parms$f = 1
-bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1)) -> out_1
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1), returnoutput = TRUE) -> out_1
 
 parms$f = 0.5
-bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1), add =TRUE) -> out_2
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1), add =TRUE, returnoutput = TRUE) -> out_2
 
 parms$f = 0
-bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1), add =TRUE) -> out_3
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1), add =TRUE, returnoutput = TRUE) -> out_3
 
@@
 parms$f = 1
-bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "L", c(0,12))-> out_4
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "L", c(0,12), returnoutput = TRUE)-> out_4
 
 parms$f = 0.5
-bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "L", c(0,12), add =TRUE)-> out_5
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "L", c(0,12), add =TRUE, returnoutput = TRUE)-> out_5
 
 parms$f = 0
-bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "L", c(0,12), add =TRUE)-> out_6
+bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "L", c(0,12), add =TRUE, returnoutput = TRUE)-> out_6
+
+  plot(out_1$pairapprox$rho_11, out_1$pairapprox$q_11, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
+ points(out_2$pairapprox$rho_11, out_2$pairapprox$q_11, pch = 20,  cex = 0.5)
+ points(out_3$pairapprox$rho_11, out_3$pairapprox$q_11, pch = 20,  cex = 0.5)
+ points(out_4$pairapprox$rho_11, out_4$pairapprox$q_11, pch = 20,  cex = 0.5)
+ points(out_5$pairapprox$rho_11, out_5$pairapprox$q_11, pch = 20,  cex = 0.5)
+ points(out_6$pairapprox$rho_11, out_6$pairapprox$q_11, pch = 20,  cex = 0.5)
+abline(a = 0, b = 1, lty = 3, col = "#D0D0D0")
+points(0,0, pch=20, col = "white", cex = 2)
+ points(0,0, pch=20, cex = 0.5)
 

### 2015-07-01 documentation/model_formulation.Rmd
update model formulation

@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 
@@
 
-workstation <-  list(host = "162.38.184.118", user = "schneider",
+workstation <-  list(host = "162.38.184.118", user = "florian",
                      rscript = "/usr/lib/R/bin/Rscript",
@@
 
-workerlist <- c(rep("localhost", times = 10))
+workerlist <- c(rep("localhost", times = 7))
 
-cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')
+cl <- makeSOCKcluster(workerlist,  outfile='out_messages.txt')
 
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 C <- function(rho, L = 5, g = 0.01) (g*rho*L)
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 C <- function(rho, L = 5, g = 0.01) g * L + rho*0
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 C <- function(rho, L = 5, a = 0.6, h = 100) (a*rho*L)/(1+a*h*rho)
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 C <- function(rho, L = 5, a = 0.6, h = 100, q = 1) (a*rho^(1+q)*L)/(1+a*h*(rho^(1+q)) )
@@
 Given a parameter set  
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 defplot(rho,  C(rho, m = parms$m, L = parms$L, h = parms$h, a = parms$a ), ylab = "growth/mortality" )
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 d_rho <- function(rho, parms, z = 4) { 
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 par(mar = c(5,4,2,3)+0.1)
@@
 
-```{r, echo = FALSE, warning = FALSE}
+```{r, echo = FALSE, warning = FALSE, warning=FALSE}
 
@@
 
-```{r}
+```{r, warning=FALSE}
 
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 plot(I(rho_11/rho_1^2) ~ time , data = ODE_spex , type = "l", ylim = c(0,2))
@@
 In the spatially-explict pair-approximation model the attractor can be visualised if we assume that clustering in the initial state is non-random, and we assign starting values for $\rho_1$ and global mortality of the population $m_1$.
-To do this, we need to solve the mortality function towards $\rho{1\,,1}$ to obtain initial covers of all the different pairs and singletons.
 
-Thus,
+I assign inital values of mortality equally along the x-axis for the assumption of highest and lowest clustering.
+The development of global and local cover yields timeseries of average growth and mortality for all those virtualised cases which can be plotted as arrowstowards the attractors.
 
-$$ m_1 = \rho_1 m_0 + \frac{(a+v\frac{\rho_{11}}{\rho_1})L(1-p\frac{\rho_{11}}{\rho_1}\rho_1^{1+q})}{1+(a+v\frac{\rho_{11}}{\rho_1})h\rho_1^{1+q}} $$
-
-rearranging yields:
-
-$$ 0 = m_1 - m_0\rho_1+m_qah\rho_1^{1+q}-aL\rho_1^{1+q}-m_0ah\rho_1^{1+q} + (m_1v\rho_1^{q}+aLp\rho_1^{q}-vL\rho_1^{q}-m_0v\rho_1^{1+q})\rho_{11} + vLp\rho_1^{q-1}\rho_{11}^2$$
-Polynomials of second order of the form $0 =  c+ bx +ax^2$ can be solved by the quadratic formula
-
-$$x = \frac{-b\pm \sqrt{b^2-4ac} }{2a}$$
-
-**apparently this solution is wrong!! Instead, I draw  initial $rho_11$ values systematically and elimitate those that produce invalid $m_ini$**
-
-This works quite well, but produces unequal distributions of arrows along the y-axis.
-
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 attractor(parms)
 ```
+
+This nicely illustrates that the unstable equilibrium now turns from a simple threshold point to a line or surface.
 
@@
 
-```{r, echo=FALSE, fig.height = 3 }
+```{r, echo=FALSE, fig.height = 3, warning=FALSE }
 par(mfrow = c(1,2))
 parms$q = 0.4
-bifurcation(parms, "L", c(0,12)) -> over_L
+bifurcation(parms, "L", c(0,12), returnoutput = TRUE) -> over_L
 
-bifurcation(parms,"b", c(0,1)) -> over_b
+bifurcation(parms,"b", c(0,1), returnoutput = TRUE) -> over_b
 
@@
 
-```{r, echo=FALSE, fig.height = 3}
+```{r, echo=FALSE, fig.height = 3, warning=FALSE}
 par(mfrow = c(1,2))
-bifurcation(parms, "L", pairapprox = TRUE, meanfield = FALSE, c(0,12)) -> over_L
+bifurcation(parms, "L", pairapprox = TRUE, meanfield = FALSE, c(0,12), returnoutput = TRUE) -> over_L
 
-bifurcation(parms,"b", pairapprox = TRUE, meanfield = FALSE, c(0,1)) -> over_b
+bifurcation(parms,"b", pairapprox = TRUE, meanfield = FALSE, c(0,1), returnoutput = TRUE) -> over_b
 
@@
 
-```{r, echo=FALSE, fig.height = 5}
+```{r, echo=FALSE, fig.height = 5, warning=FALSE}
 
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, warning=FALSE}
 over_b$pairapprox$q_11 <- over_b$pairapprox$rho_11/over_b$pairapprox$rho_1
@@
 par(mfrow =c(1,2))
-  plot(over_b$pairapprox$b, over_b$pairapprox$rho_1, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
-  plot(over_b$pairapprox$b, over_b$pairapprox$q_11, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
+  plot(over_b$pairapprox$b, over_b$pairapprox$rho_1, 
+       pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1),
+       xlab = "environmental quality", ylab = "global cover"
+       )
+  plot(over_b$pairapprox$b, over_b$pairapprox$q_11, 
+       pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1),
+       xlab = "environmental quality", ylab = "local cover")
 ```
@@
 
-```{r, echo=FALSE, fig.width = 3}
-par(mfrow = c(1,1))
-  plot(over_b$pairapprox$rho_11, over_b$pairapprox$q_11, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
+```{r, echo=FALSE, fig.width = 3, fig.height = 3, warning=FALSE}
+par(mfrow = c(1,1), mar = c(4,4,1,1))
+  plot(over_b$pairapprox$rho_1, over_b$pairapprox$q_11, 
+       pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1),
+       xlab = "global cover", ylab = "local cover")
 abline(a = 0, b = 1, lty = 3, col = "#D0D0D0")
@@
 
-```{r}
+```{r, echo=FALSE, warning=FALSE}
 

### 2015-07-01 code/simfunctions.r
update refs in simfunctions.r

@@ -639,7 +639,7 @@ attractor <- function(model_parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini
  
     
     foreach(iteration = ini$ID, .packages = c("deSolve")) %dopar% { 
-      source("C:/Users/florian.schneider/Documents/projects/CAS02_livestock/code/simfunctions.r")
+      source("/media/flo/Windows/Users/florian.schneider/Documents/projects/CAS02_livestock/code/simfunctions.r")
       
       rho_starting <- ini[iteration, 2:6]
       
@@ -711,7 +711,7 @@ bifurcation <- function(parms, over, xrange, res = 201, times = c(0,1000), ini =
   if(pairapprox) {
     
     foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
-      source("C:/Users/florian.schneider/Documents/projects/CAS02_livestock/code/simfunctions.r")
+      source("/media/flo/Windows/Users/florian.schneider/Documents/projects/CAS02_livestock/code/simfunctions.r")
       
       model_parms <- as.list(iterations[iteration,])
       
@@ -732,7 +732,7 @@ bifurcation <- function(parms, over, xrange, res = 201, times = c(0,1000), ini =
     if(nrow(upper)>0) {
       foreach(i = upper[,over], .combine = rbind, .packages = c("deSolve") ) %dopar% {
         
-        source("C:/Users/florian.schneider/Documents/projects/CAS02_livestock/code/simfunctions.r")
+        source("/media/flo/Windows/Users/florian.schneider/Documents/projects/CAS02_livestock/code/simfunctions.r")
         
         model_parms <- upper[ upper[, over] == i,]
         
@@ -788,7 +788,7 @@ bifurcation <- function(parms, over, xrange, res = 201, times = c(0,1000), ini =
     
     
     foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
-      source("C:/Users/florian.schneider/Documents/projects/CAS02_livestock/code/simfunctions.r")
+      source("/media/flo/Windows/Users/florian.schneider/Documents/projects/CAS02_livestock/code/simfunctions.r")
       
       model_parms <- as.list(iterations[iteration,])
       
@@ -808,7 +808,7 @@ bifurcation <- function(parms, over, xrange, res = 201, times = c(0,1000), ini =
     if(nrow(upper)>0) {
[…]

### 2015-07-01 documentation/model_exploration.Rmd
update model exploration

@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 # ODE solving
@@
 
-workerlist <- c(rep("localhost", times = 10))
+workerlist <- c(rep("localhost", times = 7))
 
@@
 
-```{r, echo=FALSE, fig.height = 6}
+```{r, echo=FALSE, fig.height = 6, warning=FALSE}
 colpal <- list(mort = c("#000000","#00000040","#00000030","#00000020"), grow = c("#009933","#00993340","#00993330","#00993320") )
@@
 High facilitation has strong stabilising effects on the vegetated state.
-"}
+", warning=FALSE}
 
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, warning=FALSE}
 

### 2015-07-22 documentation/socialecological.Rmd
notes on social ecological models

---
title: "Notes on socio-economic modelling of rangelands"
author: Florian D. Schneider
date: "July 22nd, 2015"
output: 
  rmarkdown::tufte_handout:
    fig_width: 4
    fig_height: 4
---

\marginnote{Institut des Sciences de l'Evolution, Université Montpellier 2, CNRS, IRD, CC065, Place Eugène Bataillon, 34095 Montpellier Cedex 05, France}

\vspace{3cm}

\newthought{Last Thursday}, Diana an I had a skype conversation about how to integrate the biophysical/ecological models with a socio-economic perspective. One model that we developped in WP6 seems particularly well suited to discuss such aspects: we refer to it as the 'rangeland resilience model'.

# The rangeland resilience model

```{r, echo=FALSE, warning=FALSE, message = FALSE}
# ODE solving
library(deSolve)

source("../code/simfunctions.r")

parallel = TRUE
if(parallel) {
# provides parallel backend
library(foreach)
library(doSNOW)

workstation <-  list(host = "162.38.184.118", user = "florian",
                     rscript = "/usr/lib/R/bin/Rscript",
                     snowlib = "/usr/lib/R/library")

workerlist <- c(rep("localhost", times = 7))

cl <- makeSOCKcluster(workerlist, outfile='out_messages.txt')

registerDoSNOW(cl)

}

# default parameters, global model without spatially-explicit effects

defparms <- list(
    m = 0.01,
    r = 1,
    b = 0.7, 
    K = 0.9, 
    a = 0.5,
    h = 50,
    L = 6, 
    alpha = 0,
    q = 0,
    c = 0,
    f = 0,
    v = 0,
    p = 0
  )

```

I will quickly sum up the assumptions and definitions of the model that can be described as both a simplification/approximation of the spatially explicit cellular automata model that we also work with^[Kéfi, S., M. Rietkerk, M. van Baalen & M. Loreau. 2007. Local facilitation, bistability and transitions in arid ecosystems. *Theoretical Population Biology*. 71(3): 267-400.] ^[Kéfi, S., M. Rietkerk, C. L. Alados, Y. Pueyo, A. ElAich, V. Papanastasis & P. C. de Ruiter. 2007. Spatial vegetation patterns and imminent desertification in Mediterranean arid ecosystems. *Nature*. 449(7159):213-217], and an extension of a simple consumer-resource model of bistability in rangelands by Noy-Meir^[Noy-Meir, I. 1975. Stability of grazing systems: an application of predator-prey graphs. *Journal of Ecology* 63, 459-481.].

The model by Noy-Meir is a graphical comparison of plant mortality and growth, which both are definite functions of current vegetation cover. Growth, i.e. the gain of vegetation per time, is logistic, which means it is increasing rapidly as vegetation cover increases, until a half saturation value is reached. From that point on, growth declines until carrying capacity of the system is reached where growth falls to 0 and vegetation cover reaches it's definite upper limit.
Mortality is defined as a functional response of grazers, i.e. it increases linearly as cover increases, but soon reaches saturation, where grazing is limited by the time that animals need to handle and digest their resource. 
The steady states of the system are defined as the intersections of both functions, i.e. the state where vegetation declines as fast as it grows. A range of parameters allow for two **alternative steady states**, a desert without vegetation and a vegetated landscape (Fig. 1, black points). Depending on the original vegetation cover of the landscape, it will develop into the one or the other. The threshold value is a well defined *unstable equilibrium* (Fig. 1, open points).

```{r, echo = FALSE, fig.margin = TRUE, fig.cap = "Graphical model for bistability of rangelands by Noy-Meir 1975. Growth (green) and Mortality (black) are both functions of vegetation cover. Many parameter sets produce two alternative stable states. "}

parms <- defparms

attractor(parms, pairapprox = FALSE, localvals = FALSE, las = 1, xlab = "vegetation cover")

```

Our model, being a pair-approximation of a two-state cellulara automata model, assumes that growth is depending not only on global cover, but also on the amount of vegetation in the local neighborhood, by the mechanism of local facilitation and competition. Similarly, grazing mortality is altered if plants have many neighbors due to attraction and diversion and the plants' shared investment in protective structures (i.e. associational resistance).

Those additions to the model extend the graphical analysis of Noy-Meir by the third dimension of local cover, which defines the steady states  of the system. Even if global cover is the same, the system might grow into the vegetated state or decline in the desert depending on the  average *local* configuration of the landscape, i.e. either a clustered or a homogeneous vegetation.

```{r, echo = FALSE, fig.margin = TRUE, fig.cap = "Extension of graphical model includes the local cover, i.e. high or low clustering of vegetation. Growth (green) and Mortality (black) are both functions of global and local vegetation cover. The latter can be initialised with high clustering (upper starting points of arrows) and low clustering (lower starting points) Many parameter sets produce two alternative stable states. ", warning = FALSE, message = FALSE}

parms <- defparms
parms$L = 4
parms$f = 0.8
parms$b = 0.2
parms$p = 0.9
attractor(parms, pairapprox = TRUE, meanfield = FALSE, localvals = FALSE, las = 1, xlab = "vegetation cover")

```

A possible visualisation adds two alternative starting conditions, with highly clustered and homogeneous vegetation, as trajectory arrows to the original growth *vs.* mortality plot (Fig 2). It becomes apparent that while the stable equilibiria, the attractors, are well defined points, specified by one global cover and one local cover, the unstable equilibrium becomes more fuzzy: it now is a plane in variable space.
While this is of minor importance for the investigation of the potantial alternative steady states of the system, which we traditionally investigate from the ecological perspective, this has some implications for the modelling of social-ecological dynamics as I will discuss below.

First, as we are used to do it from the ecological perspective, we would investigate the effect of a change along a parameter gradient on the existence of alternative stable states. 
The Noy-Meir model already visualizes the emergence of bistability along gradients of environmental quality or grazing pressure (Fig. 3).

```{r, echo = FALSE, fig.margin = TRUE, fig.cap = "Bifurcation diagrams along gradients of a) environmental quality $b$ and b) livestock density $L$. The stable equilibria (black) form alternative steady states only in a limited range along the gradient. The unstable equilibria (grey) cannot be visualised unambiguously, but are shown as average expectancy values.", warning = FALSE, message = FALSE, fig.height = 5.5}

parms <- defparms
parms$f = 0.8
parms$b = 0.2
parms$p = 0.9

par(las = 1, bty = "n", mfrow = c(2,1), mar = c(4,4,0,0))
bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "L", c(0,12), res = 101)
text(0.05,1, "a)")
bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1), res =101)
text(0.05,1, "b)")

```

This directly relates to the application of management methods, which can be translated as an additional term added or substracted to the parameters soil/environmental quality ($b$) or livestock density ($L$), as I discuss in the next section.

# Management methods in the model context

We were talking about management methods that affect the major parameters of the model, soil and livestock. In the first place, those methods are changes along the x-axis of the bifurcation diagram, without fundamental effects on the *shape* of the lines.

## Soil management

Tilling, mulching or watering are methods that directly would improve the variable $b$ of our model, *i.e.* the global environmental quality which incorporates seed production and establishment, thus reproduction of plants.

We can imagine the management method as an additive term to $b$, which would shift the entire system's bistability properties to the left. Or, it can be visualized as an arrow pointing from the current state of the system to a point to the right-hand side of it (Fig. 4).  
As a consequence, the attractors of the system may change. A qualitative change may be observed if the improved $b$ value now allows or disallows for alternative stable states. This can mean that a previously degraded system now potentially can be restored, or that it now is forced into a new attractor leading to its imminent recovery, or that a vegetated system now is not longer at risk of experiencing a catastrophic shift because it is not longer in the bistability domain.

```{r, echo = FALSE, fig.margin = TRUE, fig.cap = "Effect of management methods on the attractor of the landscape. A management method can move the system a) away from the threshold, reducing risk of degradation, b) out of the bistability domain, warranting recovery towards the vegetated attractor, c) across the threshold of the bistability domain making recovery possible in principle.", warning = FALSE, message = FALSE}

parms <- defparms
parms$f = 0.8
parms$b = 0.2
parms$p = 0.9
par(las = 1, bty = "n")
bifurcation(parms, pairapprox = TRUE, meanfield = FALSE, "b", c(0,1), res =101)
arrows(c(0.22,0.35,0.7),c(0.2,0.4,0.3),c(0.42,0.55,0.9),c(0.2,0.4,0.3))

text(c(0.22,0.35,0.7)-0.06,c(0.2,0.4,0.3), c("c)", "a)", "b)") )
```

In a mean-field model, those methods do not alter the system permanently, *i.e.* once the management method is interrupted, the previous attractor is restored. However, the pair-approximation approach allows for a inert alteration of the unstable equilibrium, since the short term application of a management measure would alter not only the vegetation cover but also the spatial structure, thus changing the threshold values of a catastrophic transition for the time being.
If this change of spatial structure is positive, the effect of the management method would be a sustainable,  long-term risk reduction.

Diana asks which changes would cause fundamental changes in the *shape* of the bifurcation graph. This now would require to include feedbacks of the landuse practice with some other parameter of the model and thus brings us to some multi-dimensional thinking in parameter space, for instance if livestock is responding to vegetation cover dynamically, but we exclude this from the model exploration for reasons of simplicity. Noy-Meir discusses this case in his 1975 paper.

## Livestock management

Still livestock is the second major land use management practice that we can investigate within CASCADE. A remaining task will be to parameterize the livestock model for realistic animal densities and to translate per capita feeding rates to a *per hectare* dimension. To achieve that the handling time and attack rate terms must be chosen to generate realistic mortalities per livestock individual per hectare. Our collaborator Andres Baeza has developed a model for livestock rotation that provides a potential starting point here.

Management practices that affect livestock densities or impact range from fencing to protection or supplementary fodder. We will need to define some use cases that reflect management practices that are well represented by the models' parameters. I imagine that most methods boil down to reducing or increasing the livestock density in a given area. Thus, similar to the soil management above, livestock management methods are designed to shift the systems state towards lower livestock rates, thus shifting to an attractor with an increased steady state vegetation cover.

Some mechanisms provide interesting feedbacks with the local cover, *i.e.* patchiness of the system, particularly associational resistance.
This mechanism is activated under high grazing, and negligible if individual grazing pressure is low. That is, the investment in protective structures is dynamically adjusted to the need for being protected. The implementation of the mechanism as being only of effect at low cover while the beneficial effect is zero if no grazing occurs.

Thus, the short-time application of grazing pressure on a landscape alters the pattern formation processes. If grazing is reduced, the vegetation pattern  remains as it is and might cause the systems attractor to be fundamentally altered.

# Obtaining predictions from models

We discussed how the rather fundamental and qualitative model can be put to use for a socio-economic model that needs to have some quantitative dimensions as well. The model in the first place provides conclusions of the qualitative kind: "if that mechanism is strong, then the bistability domain will be smaller/larger".

As discussed on the plenary meeting on Crete, the models are ill suited to undergo a precise parameterization for a particular field site. They also will not provide somewhat precise quantitative outcomes of the kind "if environmental quality is improved by value x, then the vegetation cover increases by value y". While Mara was very cautious that such claims should be made at all based on her rather fundamental model, I would say that we can achieve approximate quantitative estimates by incorporating parameter uncertainty. E.g. if we assume an environment has a variability in environmental quality then we can estimate the variability of the prediction value as well by simulating systems for each parameter. This allows to estimate maximum likelihood predictions for a given parameter array. In other words, we would make prior assumptions about a particular parameter to be around value x plus minus a standard deviation. From this distribution function, we can simulate a likely outcome of the vegetation cover y. The model might be invariant to some parameters' uncertainties or might respond strongly to others.

Following the discussion we had during the CASCADE plenary on Crete, we talked about the importance of communicating uncertainty of models if models are used to derive conclusions within the socio-economic workpackage. Awareness must be raised that the models are incomplete and rather simple approximations of nature, as opposed to high-detail reductionist models that include as much realism as possible.

Still, it would be interesting to relate the simple livestock model to the field sites of CASCADE, at least to those that are strongly influenced by grazing. I imagine we could turn parameters into two or three-level variables to classify at least the Cyprus, Crete and Ariola sites. They differ at least in aridity, grazing impact and even in the importance of some mechanisms as associational resistance or attractant decoy.

# Terminology: Probability in the pair-approximation model

The term of probability is confusing or does not apply at all in the context of the mean-field or pair-approximation model. The model is fully deterministic and thus, starting from one initial condition and one parameter set, the outcome is no matter of probability.
I apply the term nonetheless when speaking about how the change in a parameter alters the "probability of the system to collapse". Strictly speaking, in ordinary differential equation models, no change is going to occur for a given starting condition and parameter set. But I integrate the concept of probability in at least four senses:

- as an integration of **random starting conditions**: If the model is run repeatedly with *different starting conditions*, i.e. with a gradient of initial vegetation cover, we can speak of a probability that a random landscape will end up in one or another steady state/attractor.  
- **uncertainty in parameters**: If the model is run repeatedly with variation in one or several parameters, the probability to end up in one or the other attractor can be determined in relation to the prior range of parameters.  
- as an **interpolation of the stochastic model**: The cellular automata implementation or natural data never reach steady state equilibria, rather oscillate around the attractors, and thus will deviate from those in stochastic events, e.g. by random gain or loss of vegetation. These random events will trigger the crossing of a threshold, i.e. the sudden shift with a certain probability.
- as an **interpolation of non-formulated stochastic events**: Stochastic events like fire or weather extremes are not part of the model. Still, assuming the system is at steady state, those events would be occurring stochastically over time and/or temporarily disturb the system with a random intensity. This translates into a risk of the system to fall to such events and experience sudden transition.

Thus, even in the context of a deterministic model approach, one can translate into terms of 'probability'. Even a mathematical framework could be developed to assess those probabilities quantitatively.

# Timescales

Our models tend to assume infinite timescales and we are looking at systems at equilibrium. This means we are looking at the potential *attractors* of the system, rather than transient dynamics towards a particular state.

This is important when talking about predictions. The model are employed to  estimate potential steady states for a given set of parameters, and the threshold values leading to one or the other steady state. We compare the overlap of vegetated and degraded state along parameter gradients, but do not investigate how events leading to the one or the other differ. From this background, the history of a landscape is almost negligible. We only differentiate two starting conditions, extremely degraded and well vegetated, because over infinite time they eventually will end up in the one or the other state.

If thinking about economics, like yield or investment costs, the timescale usually matters, e.g. 'how long must a management practice be applied?', 'how long does it take until we restored resilience of the system?'.

The models could be used to estimate relative times or velocities of developing towards a certain attractor. This can be used to describe transient dynamics of systems even at intermediate states, i.e. between extremely degraded and well vegetated. Also the history of a landscape along a schedule of management practices can be recorded, e.g. application of a management practice with a certain frequency or over a certain period. Given stochastic events of disturbance, the questions then can be : 'What is the risk of a landscape with a given spatial pattern to degrade in the next two/ten/hundred years?', 'How long needs a management method be applied to lift the landscapes spatial pattern out of a high risk state?', 'What is the chance of recovery within the next 10/100 years if this method is applied compared to that method?'.

# Closing thoughts

The livestock model that we formulated so far has many links that would allow it's use within an social-ecological model. I propose that we

1) define "risk" of degradation and "chance" of regeneration as target measures of an economic model  
2) translate economic "yield" and "investment" into the framework of our ecological model  
3) think about a framework for describing and measuring the quality and velocity of transient dynamics.

This can translate the model into a social-ecological framework. Models on land-user decisionmaking or scenarios for management practice can be added as further layers of model complexity that define how $L$ or $b$ behave as a response to spatial structure or global cover.

```{r, echo=FALSE}
if(parallel) {
stopCluster(cl)
}
```

### 2015-07-22 code/simfunctions.r

@@ -561,9 +561,9 @@ home_simfunction <- paste0(getwd(),"/code/simfunctions.r")
 # simulates trajectories for many differently clustered starting conditions. 
 # running pairapprox = TRUE requires parallel backend
 
-attractor <- function(model_parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini = seq(0,1, length = 11), meanfield = TRUE, pairapprox = TRUE, localvals = FALSE) {
+attractor <- function(model_parms, rho_1_ini = seq(0,1, length = 21), rho_11_ini = seq(0,1, length = 11), meanfield = TRUE, pairapprox = TRUE, localvals = FALSE, ...) {
   
-  plot(NA,NA, ylab = "plant mortality/growth", xlim = c(0,1), ylim= c(0,.25), bty = "n", xaxs = "i" , yaxs = "i")
+  plot(NA,NA, ylab = "plant mortality/growth", xlim = c(0,1), ylim= c(0,.25), bty = "n", xaxs = "i" , yaxs = "i", ...)
   
   rho <- seq(0,1,length = 100)
   
@@ -639,8 +639,8 @@ attractor <- function(model_parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini
  
     
     foreach(iteration = ini$ID, .packages = c("deSolve")) %dopar% { 
-      source("/media/flo/Windows/Users/florian.schneider/Documents/projects/CAS02_livestock/code/simfunctions.r")
-      
+      #source("/media/flo/Windows/Users/florian.schneider/Documents/projects/CAS02_livestock/code/simfunctions.r")
+      source("E:/Eigene Dokumente/Uni/projects/CAS02_livestock/code/simfunctions.r")
       rho_starting <- ini[iteration, 2:6]
       
       # running the ode-solver
@@ -711,8 +711,8 @@ bifurcation <- function(parms, over, xrange, res = 201, times = c(0,1000), ini =
   if(pairapprox) {
     
     foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
-      source("/media/flo/Windows/Users/florian.schneider/Documents/projects/CAS02_livestock/code/simfunctions.r")
-      
+      #source("/media/flo/Windows/Users/florian.schneider/Documents/projects/CAS02_livestock/code/simfunctions.r")
+      source("E:/Eigene Dokumente/Uni/projects/CAS02_livestock/code/simfunctions.r")
[…]

### 2015-09-16 documentation/interactive_spexmodel.Rmd
update shiny docs

@@
 inputPanel(
+
+  checkboxInput("meanfield", "Meanfield model", FALSE),
+  checkboxInput("pair", "Pair-approximation model", TRUE),
+  checkboxInput("ca", "Cellular automata", TRUE),
   sliderInput("rho_1", label = "initial plant cover:",
@@
 
+if(input$ca) {
   spexmodel <- runCA(input$rho_1, parms, delta = 0.2, t_max = 60, t_min = 50, t_eval = 5, isstable = 0.001, saveeach = 1)
+  lines(spexmodel$rho_one ~ spexmodel$time, col = "grey50")
+}
 
-lines(spexmodel$rho_one ~ spexmodel$time, col = "grey50")
-
+if(input$meanfield) {
   meanfield <-as.data.frame( ode(ini_rho(input$rho_1), func = odesys_mean, times = exp(seq(0,4,length = 100))-1, parms = parms))
   lines(rho_1 ~ time , data = meanfield, col = "red3")
+}
 

### 2015-09-16 documentation/interactive_spexmodel_attractor.Rmd
update shiny docs

@@
 
-workerlist <- c(rep("localhost", times = 10))
+workerlist <- c(rep(workstation, times = 7))
 
-cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')
+cl <- makeSOCKcluster(workerlist, outfile='out_messages.txt')
 
@@
 
+inputPanel(
+  selectInput("model",
+      label = "model type:",
+      choices = c("meanfield", "pairapprox"),
+      selected = 1
+  )
+)
+
 ```
@@
 
-  source("../code/simfunctions.r")
+  source("C:/Users/florian.schneider/Documents/projects/CAS02_livestock/code/simfunctions.r")
 
@@
 #iterations$f[iterations$f < iterations$b] <- iterations$b[iterations$f < iterations$b]
+if(input$model == "meanfield") attractor(parms, localvals = TRUE, pairapprox = FALSE, meanfield = TRUE, rho_1_ini = seq(0,1, length = 11) )
 
-attractor(parms, localvals = FALSE, pairapprox = TRUE, meanfield = FALSE, rho_1_ini = seq(0,1, length = 21) )
-
+if(input$model == "pairapprox") attractor(parms, localvals = FALSE, pairapprox = TRUE, meanfield = FALSE, rho_1_ini = seq(0,1, length = 11) )
 # if(input$gradient == "b") plot(rho_1 ~ b, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))

### 2015-09-16 documentation/interactive_spexmodel_bifurcations.Rmd
update shiny docs

@@
 
+source("../code/simfunctions.r")
+
 # provides parallel backend
@@
 
- if(input$gradient == "b") bifurcation(parms,"b", c(0,1))
- if(input$gradient == "L") bifurcation(parms, "L", c(0,12))
+ if(input$gradient == "b") { 
+#par(mfrow =c(2,2))
+
+   bifurcation(parms,"b", c(0,1)) -> over_b
+
+#over_b$pairapprox$q_11 <- over_b$pairapprox$rho_11/over_b$pairapprox$rho_1
+#over_b$pairapprox$q_11[is.na(over_b$q_11)] <- 0
+
+#  plot(over_b$pairapprox$b, over_b$pairapprox$q_11, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
+
+ # plot(over_b$pairapprox$rho_11, over_b$pairapprox$q_11, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
+#abline(a = 0, b = 1, lty = 3, col = "#D0D0D0")
+#points(0,0, pch=20, col = "white", cex = 2)
+# points(0,0, pch=20, cex = 0.5)
+
+ } 
+ if(input$gradient == "L") {
+
+#par(mfrow =c(2,2))
+
+   bifurcation(parms, "L", c(0,12)) -> over_b
+
+#over_b$pairapprox$q_11 <- over_b$pairapprox$rho_11/over_b$pairapprox$rho_1
+#over_b$pairapprox$q_11[is.na(over_b$q_11)] <- 0
+
+#  plot(over_b$pairapprox$b, over_b$pairapprox$q_11, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
+
+#  plot(over_b$pairapprox$rho_11, over_b$pairapprox$q_11, pch = 20,  cex = 0.5, ylim = c(0,1), xlim = c(0,1))
+#abline(a = 0, b = 1, lty = 3, col = "#D0D0D0")
+#points(0,0, pch=20, col = "white", cex = 2)
+# points(0,0, pch=20, cex = 0.5)
+
+   }
 
 } )
+
 ```

### 2016-06-23 manuscript/manuscript.Rmd
add new manuscript draft (Notes version)

---
title: "the implications of local interactions on the resilience of arid rangelands"
author: "Florian D. Schneider"
date: "20 Juni 2016"
output: pdf_document
---

```{r, echo = FALSE}

library(livestock)

```

# Introduction

- The conceptualisation of rangelands as systems that can experience abrupt shifts between alternative stable states has been happening since the 70s (Noy Meir, Holling, May). 
- The responsible mechanisms are feedbacks on the landscape scale, i.e. dependent on the 'global' density of vegetation and livestock, as well as global environmental parameters, such as aridity.
- these studies had hughe implications for rangeland management and were well received in applied, fundamental and theoretical research. 
   - applied concepts such as state and transition models, i.e. descriptions of landscapes as systems with multiple potential states and probabilities and conditions for the transition between those. 
   - fundamental research addressed the existence of alternative stable states in observational and experimental studies (examples in rangelands?? arid shrublands? grasslands?). It relates to the study of successional states. 
   - theoretical studies were addressing mechanisms and behaviour of alternative stable states along gradients of pressure, where the rangeland system provided a good example. (Rietkerk, van de Koppel, Scheffer, ...) Most recently, the study of social-ecological systems addressed the catastrophic behaviour of ecosystems in the context of anthropogenic activities. (Carpenter, Schlüter)

Figure 1 | Citations of Noy-Meir 1975 in publications of theoretical, applied and fundamental ecology journals.

- While these studies investigate feedbacks of global vegetation cover only, applying mean field approximations of vegetation, recent studies discuss the importance of spatial structure in forming stable vegetation even under high levels of pressure. (von Hardenberg, Rietkerk, Kefi)
- the local interactions of plants are providing feedbacks on plant growth and mortality in the close surroundings of plants, for instance via providing shade and protection for neighboring plants. Growth and mortality as functions of global environment, but if looking at the local scale environmental quality might be substantially altered by the conditions in the plants direct vicinity. 
- This leads to the formation of  the pattern of vegetation along gradients of stress. Examples in shrublands, etc. (Rietkerk) 
- Spatial structure is also suggested as an early indicator of shifts (Kefi, Dakos, Scheffer)

- In this paper, we will discuss the local interactions that lead to feedbacks on the formation of vegetation and alter the steaady state / attractors. 
- Knowledge of local interactions is mandated if one wants to understand thresholds and conditions for critical transitions, for instance for the description of state and transition models.

# describing the state of a rangeland system

- the framework proposed by Noy Meir 1975 looks at simple graphical analysis of the growth vs mortality of vegetation. This simple population dynamic approach identifies the stable and unstable equilibria of an ecosystem/population at the intersections of the growth and mortality functions over vegetation cover.
- We observe net increase in vegetation cover, where the growth rate is higher than mortality and net decline in vegetation cover in the opposite case. The alternative steady states arise in cases where the two functions intersect more than once. (equations of growth and mortality as functions of vegetation cover $$ $$)
-  growth function depends on environmental quality and the mortality function depends on livestock densities. If observing the alternative steady states /attractors along a gradient of pressure, this can be plotted as a bifurcation graph of alternative steady states. Illustrating the thresholds of pressure and vegetation cover that will lead to an ecosystem shift.

```{r, echo = FALSE, fig.cap= "Noy Meir graphs a) M vs G function, b) bifurcation along livestock density"}
plot_meanfield(livestock)

```

Figure 2 | Noy Meir graphs a) M vs G function, b) bifurcation along livestock density

- This description assumes that vegetation cover is a sufficient descriptor for the feedbacks occuring on mortality and growth of vegetation, resulting in non-linear functions of vegetation cover. Other models departed from this basic model and integrated more elaborate mechanisms that took other constants into account such as water runoff etc. (Rietkerk, van de Koppel?)
- The recent studies have shown that vegetation structure is a major factor in understanding the stability of arid and semi-arid vegetation.
- To integrate spatial structure into the framework of Noy Meir, we define the both growth and mortality as functions of vegetation cover AND local vegetation density / clustering coefficient. This provides information about the strength of local interactions and alters growth and mortality. 
- (equations for growth and mortality as  functions of global and local cover)
- The visualisation of this  case is difficult because it adds a third dimension to the graph of Noy-Meir: in the z-axis we visualise the local cover of vegetation, i.e. the average cover in the direct neighborhood of a plant $\rho_{1|1}$. However, the attractors reduce to a single point.

Figure 3 | 3D graph of growth vs. mortality

- If illustrated along the gradient of pressure, a bifurcation graph visualises the location of the attractor at any given parameter combination. In addition to the vegetation cover, which is visualised on the x axis, we require the local vegetation cover on the z axis to provide a full description of the steady state ecosystem.

Figure 4 | bifurcation along a) livestock density and b) local cover

# the role of local interactions for transitions

- To investigate the role of local interactions between plants, we implement these mechanisms into the model. This is complemented with figures.

### 2016-06-24 manuscript/manuscript.Rmd
update manusript draft: add code for livestock package; add structure for analysis and discussion

@@
 library(livestock)
+library(knitr)
+library(rgl)
+library(foreach)
+library(doSNOW)
+
+knit_hooks$set(rgl = hook_rgl)
+
+workerlist <- c(rep("localhost", times = 3))
+cl <- makeSOCKcluster(workerlist)
+ registerDoSNOW(cl)
 
@@
 
-- The conceptualisation of rangelands as systems that can experience abrupt shifts between alternative stable states has been happening since the 70s (Noy Meir, Holling, May). 
+- The conceptualisation of rangelands as systems that can experience abrupt shifts between alternative stable states has been develloped since the 70s (Noy Meir, Holling, May). 
 - The responsible mechanisms are feedbacks on the landscape scale, i.e. dependent on the 'global' density of vegetation and livestock, as well as global environmental parameters, such as aridity.
@@
 
-```{r, echo = FALSE, fig.cap= "Noy Meir graphs a) M vs G function, b) bifurcation along livestock density"}
-plot_meanfield(livestock)
-
+```{r, echo = FALSE, fig.cap="Noy Meir graphs a) M vs G function, b) bifurcation along livestock density"}
+par(mfrow = c(1,2))
+p <- set_parms(livestock$defparms, set = list(b = 0.5))
+plot_meanfield(livestock, new = TRUE, parms = p)
+plot_bifurcation(livestock, over = "L", xrange = c(0,15), res = 51, parms = p)
 ```
@@
 
-Figure 3 | 3D graph of growth vs. mortality
+```{r, echo = FALSE, fig.cap = "3D graph of growth vs. mortality" , rgl=TRUE}
+p <- set_parms(livestock$defparms, set = list(b = 0.3, c = 0.2, f = 0.9, p = 0.99, alpha = 0.2))
+plot_pairapproximation3D(livestock, parms = p)
+```
+
+Figure 3 |
 
@@
 
-- To investigate the role of local interactions between plants, we implement these mechanisms into the model. This is complemented with figures.
+- To investigate the role of local interactions between plants, we subsequently implement these mechanisms into the model.
+We investigate their impact on the ecosystem state along gradients of livestock pressure (i.e.
+bifurcation diagrams) under a given environmental quality, that leads to the expression of catastrophic shifts.
+We discuss positive and negative effects on plant growth and mortality.
+
+## local facilitation
+
+- local facilitation is the positive local effect of plants on the growth of neighboring plants.
+Plants in arid environments provide shade and increase water retention, as well as improving soil organic matter by adding and capturing litter.
+This leads to an improvement of the local environmental quality, which is beneficial for seed germination and early survival.
+It therefore provides habitat for certain species that could not survive under the given conditions of the surroundings.
+- we implement local facilitation as a positive effect on cell colonisation. 
+- [equation] 
+- Thus, the likelihood of 
+- Thus, the positive effect is of great importance in arid environments.
+
+Figure |
+
+## local competition
+
+- plants compete for resources like nutrients and water with other plants in their close vicinity.
+This effect leads to decellerated growth and eventually outcompetition of the weaker competitor on the local patch scale.
+Plants will benefit from competitor free space around their location, if the environment allows to grow there.
+- We implement local competition as a negative effect on growth. 
+- [equation]
+- This affects the local carrying capacity and therefore the likelihood of establishment of the system linearly with increasing density of plants in the surroundings.
+
+Figure |
+
+## associational resistance
+
+- In rangelands, plants are subject to feeding pressure by livestock.
+In drylands, many plants have evolved defence mechanisms that protect them from being eaten, such as thorns or indigestible plant metabolites.
+Plants that grow next to each other share this investment and benefit from the structures provided by their neighbors.
+- We implement this as a reduction of plant mortality due to livestock feeding with increasing plant cover in the direct neighborhood. 
+- [equation]
+-
+
+Figure |
+
+## attractant decoy
+
+- In contrast to the protective effects of plants onto each other,  accumulations of plants will also attract livestock and increase feeding pressure locally. 
+- We implement this mechanism as a positive effect of local plant cover on the feeding rate of grazers, rendering more efficient. 
+- [equation]
+
+# Discussion
+
+The different mechanisms alter the possible ecosystem states observed along a gradient of pressure, but they also alter the trajectories towards those attractors departing from weakly or highly clustered landscapes.
+
+# Methods
+
+We implement the equations for mortality and growth outlined above in a pair-approximation model of a spatially-explicit cellular automata framework.
+We contrast this to a model that does not account for spatial structure and assumes that the local neighborhood cover of plants is equal to the global vegetation cover, i.e.
+a mean field model.
+
+## cellular automata model
+
+## pair-approximation model
+
+## mean-field model
+
+```{r}
+stopCluster(cl)
+```
+
+# References

### 2016-07-05 manuscript/manuscript.Rmd
update manuscript.Rmd: add methods description and r chunks for figures (no nice display yet)

@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
 
@@
 
+r3dDefaults$userMatrix <- matrix(c(0.7849265,0.1177803,-0.6082912,0,0.008146744,0.9797189,0.2002101,0,0.6195351,-0.1621057,0.7680479,0,0,0,0,1), nrow = 4) 
 knit_hooks$set(rgl = hook_rgl)
+head(hook_rgl)
 
@@
 ```
+
+# Abstract
+
+The conceptualisation of rangelands as systems that can experience abrupt shifts between alternative stable states due to positive feedback mechanisms has been developed since the 70s.
+While these studies investigate feedbacks of a homogeneous, global vegetation cover, recent studies discuss the importance of heterogeneous spatial structure in forming stable vegetation even under high levels of pressure.
+This study proposes an approximation to a spatially-explicit consumer-resource model as a framework to investigate the stability and resilience of arid rangelands under different management scenarios.
+Using this model, we explore how plant-plant interactions on the local scale, in combination with global management practices and environmental pressure, shape the thresholds of catastrophic shifts.
+Knowledge of local interactions is mandated if one wants to understand thresholds and conditions for critical transitions, for instance for the description of state and transition models.
 
@@
 Figure 1 | Citations of Noy-Meir 1975 in publications of theoretical, applied and fundamental ecology journals.
+
+<!-- Difficult to get these data.
+The paper is not on ISI or Pubmed, Google citations cannot be extracted except by hand.
+I will continue to try, but since this is not essential for the paper I would skip it if it is to much work -->
 
@@
 
-```{r, echo = FALSE, fig.cap="Noy Meir graphs a) M vs G function, b) bifurcation along livestock density"}
+```{r, echo = FALSE, fig.cap="Noy Meir graphs a) the intersections of the growth and mortality functions define the equilibrium points (a.k.a.
+attractors) of the system; if the functions intersect this leads to two stable equilibria (black), and one unstable equilibrium (white).
+b) The bifurcation graph along livestock density illustrates the potential states of the ecosystem (i.e.
+it's cover) under a given parameter.", cache = TRUE, warning = FALSE, message = FALSE, error = FALSE}
 par(mfrow = c(1,2))
-p <- set_parms(livestock$defparms, set = list(b = 0.5))
+p <- set_parms(livestock$parms, set = list(b = 0.5,  L = 0.6))
 plot_meanfield(livestock, new = TRUE, parms = p)
-plot_bifurcation(livestock, over = "L", xrange = c(0,15), res = 51, parms = p)
+plot_bifurcation(livestock, over = "L", xrange = c(0,2), res = 201, parms = p) -> out
 ```
-
-Figure 2 | Noy Meir graphs a) M vs G function, b) bifurcation along livestock density
 
@@
 Other models departed from this basic model and integrated more elaborate mechanisms that took other constants into account such as water runoff etc.
-(Rietkerk, van de Koppel?)
+(Rietkerk, van de Koppel?), however they still used mean-field approximations i.e.
+the assumption that the local cover is at all times and in all locations equal to the global cover.
 - The recent studies have shown that vegetation structure is a major factor in understanding the stability of arid and semi-arid vegetation.
@@
 
-```{r, echo = FALSE, fig.cap = "3D graph of growth vs. mortality" , rgl=TRUE}
-p <- set_parms(livestock$defparms, set = list(b = 0.3, c = 0.2, f = 0.9, p = 0.99, alpha = 0.2))
-plot_pairapproximation3D(livestock, parms = p)
+```{r,echo = FALSE, fig.cap = "3D graph of growth vs. mortality", rgl=TRUE, results = FALSE, warning = FALSE, message = FALSE, error = FALSE}
+p <- set_parms(livestock$parms, set = list(b = 0.3, c = 0.4, f = 0.9, p = 0.99, v = 0.99))
+plot_pairapproximation3D(livestock, parms = p) -> out
 ```
@@
 Figure 4 | bifurcation along a) livestock density and b) local cover
+
+```{r, echo = FALSE, fig.cap = "3D graph of growth vs. mortality", rgl=TRUE, results = FALSE, warning = FALSE, message = FALSE, error = FALSE}
+p <- set_parms(livestock$parms, set = list(b = 0.3, c = 0.2, f = 0.9, p = 0.99))
+plot_pairapproximation3D(livestock, parms = p) -> out
+```
 
@@
 
-Figure |
+```{r,echo = FALSE, fig.cap = "3D graph of growth vs. mortality", rgl=TRUE, results = FALSE, warning = FALSE, message = FALSE, error = FALSE}
+p <- set_parms(livestock$parms, set = list(b = 0.3, f = 0.9))
+plot_pairapproximation3D(livestock, parms = p, meanfield = TRUE) -> out
+
+```
 
@@
 
-Figure |
+```{r,echo = FALSE, fig.cap = "3D graph of growth vs. mortality", rgl=TRUE, results = FALSE, warning = FALSE, message = FALSE, error = FALSE}
+p <- set_parms(livestock$parms, set = list(b = 0.7, c = 0.7))
+plot_pairapproximation3D(livestock, parms = p, meanfield = TRUE) -> out
+
+```
 
@@
 
-Figure |
+```{r,echo = FALSE, fig.cap = "3D graph of growth vs. mortality", rgl=TRUE, results = FALSE, warning = FALSE, message = FALSE, error = FALSE}
+p <- set_parms(livestock$parms, set = list(b = 0.6, p = 0.99, L = 1.3))
+plot_pairapproximation3D(livestock, parms = p, meanfield = TRUE) -> out
+
+```
 
@@
 - We implement this mechanism as a positive effect of local plant cover on the feeding rate of grazers, rendering more efficient.
-- [equation]
+
+```{r,echo = FALSE, fig.cap = "3D graph of growth vs. mortality", rgl=TRUE, results = FALSE, warning = FALSE, message = FALSE, error = FALSE}
+p <- set_parms(livestock$parms, set = list(b = 0.7, v = 0.9))
+plot_pairapproximation3D(livestock, parms = p, meanfield = TRUE) -> out
+
+```
 
 # Discussion
+
+<!-- description of attractors and trajectories -->
 
@@
 
-We implement the equations for mortality and growth outlined above in a pair-approximation model of a spatially-explicit cellular automata framework.
+\small
+
+We implement the equations for plant mortality and growth outlined above in a pair-approximation model of a spatially-explicit cellular automata framework.
 We contrast this to a model that does not account for spatial structure and assumes that the local neighborhood cover of plants is equal to the global vegetation cover, i.e.
@@
 
-## cellular automata model
+Subsequently, we describe a basic form of the model that has been used by Noy-Meir and substitute it's elements with specific models of local plant-plant interactions.
+A system of ordinary differential equations is implemented in the framework of a mean-field approximation as well as a pair-approximation of a stochastic cellular automata model.
+We provide code for the mean-field approximation, the pair-approximation and the cellular automata online at https://cascade-wp6.github.com/livestock.
+
+## basic model
+
+The framework provided by Noy-Meir is a graphical assessment of the attractors of the population dynamics in a simple predator-prey system.
+It describes mortality and growth as functions of global vegetation cover, $\rho_1$.
+
+For vegetation mortality, $M$, we assume intrinsic plant mortality, i.e. the inverse of average lifespan, and add the mortality due to grazing risk as a classic type II functional response.
+
+$$ M(\rho_1) = m \rho_1 + \frac{a \rho_1 L}{1+a h \rho_1} \,,$$
+
+where $m$ denotes the intrinsic mortality rate, $a$ and $h$ are the search efficiency and handling time of the nonlinear functional response (saturating at $1/h$) which is multiplied with the livestock density, $L$.
+Note that livestock density is defined in the unit number of grazers per area $[\frac{ind.}{ha}]$.
+Search efficiency $a$ is given in area browsed in per time and grazer individual $[\frac{ha}{y \times ind.}]$, and handling time $h$ is given in time per area consumed $[\frac{y}{ha}]$, which alltogether defines feeding rate as an area consumed per year by a single individual $[\frac{ha}{y \times ind.}]$.
+The overall change of cover, $M(\rho)$, is thus independent from the size and unit of the observed area $[\frac{1}{y}]$.
+
+The vegetation growth, $G$, is defined as a logistic growth function.
+
+$$ G(\rho_1) = r\rho_1(1-\frac{\rho_1}{K}) \,,$$
+
+where the intrinsic growth rate, $r$, is neutralized when cover approaches the global carrying capacity, $K$.
+This reflects the global competition of plants fo  resources of homoegeneous distribution, such as space: when all habitable space is occupied, plant growth falls to zero.
+
+<!-- global feedback mechanisms such as water runoff (negative effect on growth at low cover) or plant refuges against grazing for instance due to fencing (type III funuctional response) might be added here.
+-->
+
+## local feedbacks
+
+The basic model assumes mortality and growth to be of homogeneous distribution, and neglects spatially-explicit effects.
+Here, we add plant-plant interactions at the local scale by taking the cover, $q$, in the direct vicinity of plants into account.
+By adding such interactions at the local level into the functions of growth and mortality, patterns emerge that act as positive or negative feedbacks on vegetation cover.
+
+For vegetation growth we assume that the local environmental suitability is enhanced by local facilitation (Milchunas et al, Kéfi et al 2008).
+The reduction in growth rate, $r$, by aridity, i.e.
+the inverse of environmental quality $b$, is compensated by a function of $q_{0|1}$, the local density of cells in state 1 (i.e.
+vegetation) given that the focal cell is in state 0 (i.e.
+empty), and maximizes to one if the cell has four neighbors.
+
+$$ b = b^* + (1 - b^*)  f q_{0|1} \,,$$
+
+with the effect of aridity, $a_0$ in absence of local vegetation being increased to the value of $f$ if additive facilitation due to the presence of neighboring plants occurs.
+This term gradually determines the enhancement of the growth with an increasing local vegetation cover (maximizing at $q_{0|1} = 1$ if the cell has 4 neighbors).
+
+As an opposing effect, we assume that local competition of plants inhibits colonisation locally by depleting nutrients or light.
+As cover increases and the interspace areas are closing in, the space available for rejuvenation approaches zero.
+This is assuming that growth is diminished locally by competition, $c$. This affects carrying capacity rather than growth rate itself.
+
+$$ K = K^* (1 - c  q_{0|1})$$
+
+That is, if a cell has a fully vegetated local neighborhood, it's carrying capacity will be reduced by value $c$ ($c \geq K$).
+
+Regarding plant mortality, we implement two interactions into the model.
+In sparse environments, high local cover will be attracting grazers and concentrate plant mortality at locations with high local density.
+We assume that low plant cover reduces search efficiency, $a$, locally by the value $v$
+
+$$ a = a^* + (v \frac{q_{1|1}}{\rho_1}) $$
+
+Thus, at low local cover feeding will be high on plants with neighbors, whereas at high cover the term has no effect since handling time is limiting consumption.
+An enhanced search efficiency will raise the critical thresholds for a collapse.
+
+Opposing to that mechanism, plants in grazed habitats develop protective traits, such as thorns or cushion growth.
+They thereby provide protected habitat to their direct neighborhood, or share the investments in those traits with their neighbors.
+Overall grazing mortality thus is reduced through an increase in handling time by associational resistance, $p$.
+
+$$ a = a^{**} (1- p q_{1|1}) $$
+
+This causes $a$ to fall to its minimal value ($1-p$) at full local cover.
+If $p=1$ the feeding is completely neutralized under high local cover, which also causes a drop in grazing mortality if cover is really high.
+
+Combining the attractant decoy and the associational resistance effect yields
+<!-- $$ h = h^*  / (1 - p  q_{1|1}) $$
+
+An alternative formulation should be found. The term is highly non-linear and provokes relevant changes only at values above 0.9. It is defined within $(0,1($ 
+Instead, we could define associational protection as a reduced impact due to a reduced local density of grazers.
+
+$$ L = L^* (1 - p  q_{1|1}) $$
+
+**Update 12.03.2015:** the effect is of less relevance at high cover, since the development of thorns is only induced if local pressure is already high.
+At high cover it is the handling time limitation that acts as an individual protection.
+Thus, it makes more sense to implement associational protection as an effect on search efficiency.
+
+$$ a = a^* (1- p q_{1|1}) $$
+
+This causes $a$ to fall to its minimal value ($1-p$) at full local cover.
+If $p=1$ the feeding is completely neutralized under high local cover, which also causes a drop in grazing mortality if cover is really high.
+This might be unrealistic, for the biological reasons named above.
+So, values of $p \leq 0.9$ seem to make sense.
+
+**Update 26.03.2015:** Also, the formulations of local effects on mortality are unclear about their compatibility, since the attractant effect it is additive while the protective effect is multiplied.
+Thus, I have to specify-->
+
+$$ a = (a^*  + v q_{1|1} ) (1- p q_{1|1}) $$
+
+Since $q_{1|1}$ and $q_{1|0}$ are spatially explicit expressions, these equations are referring to the situation at a particular location in space which is currently in one state or the other.
+The substitution into $M(\rho)$ and $G(\rho)$ yields transition probabilities, i.e.
+the probability for death and colonization for the given location, based on the global vegetation cover $\rho_1$ and the local vegetation cover $q_{1|1}$ or $q_{1|0}$ for vegetated or empty locations, respectively:
+
+$$ M(\rho_1,q_{1|1}) = w_{1,0} = m + \frac{(a^*  + v q_{1|1} ) (1- p q_{1|1}) L \rho_1}{1+ (a^*  + v q_{1|1} ) (1- p q_{1|1}) h \rho_1} $$
+
+$$ G(\rho_1,q_{1|0}) = w_{0,1} = \frac{r \rho_1 (b^* + (1 - b^*) f q_{0|1}) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})})}{1-\rho_1} $$
+
+Default parameters for the model are provided in table 1.
+
+parameter | default value | unit   | definition
+-----| -----| -------- | ----------------------------------------------
+  r  |  1.0 | $[\frac{1}{y}]$  | max. reproduction rate of vegetation per year
+  b  |  0.5 | unitless    |environmental quality, a factor that indicates  \newline quality of environment as compared to the best case scenario
+  K  |  0.9 | unitless   | carrying capacity of the system, a landscape  \newline specific value that defines the max. potential cover
+  f  |  0 | unitless     |  local facilitation, positive effect of plants  \newline  on the colonization probability in their direct neighborhood
+  c  |  0 | unitless   | local competition, negative effect of plants on \newline   colonization probability in their direct neighborhood
+  m  |  0.05| $[\frac{1}{y}]$   | intrinsic mortality of plants (inverse of av. lifespan)
+  v  |  0 | unitless   | attractant-decoy effect, negative effect of plants  \newline  on mortality in their direct neighborhood
+  p  |  0 | unitless   | associational resistance, positive protection effects  \newline  of plants on each other against mortality due to grazing
+  L  |  2   | $[\frac{ind.}{ha}]$   | Livestock density
+  h  |  10  | $[\frac{y}{ha}]$   | handling time, time required to consume one hectar  \newline  of vegetation 
+  a  |  5 | $[\frac{ha}{y \times ind.}]$   | search efficiency of livestock,
 
@@
 
+Subsequently, we implement these functions into a pair-approximation of a spatially-explicit model. 
+The pair-approximation framework [@matsuda92] is defined as a system of ordinary differential equations, describing the populations of *pairs* of neighboring locations.
+In a two state system, three pairs can be defined with population densities $\rho_{1,1}$ , $\rho_{0,0}$, and $\rho_{0,1}$.
+Besides, there are the population densities of the singletons, $\rho_0$ and $\rho_1$.
+
+The following conservation equations allow to calculate population densities of one pair based on the other pairs and singletons:
+
+$$ \rho_1 + \rho_0 = 1$$
+
+$$ \rho_{1,1} + \rho_{0,1} = \rho_1$$
+
+$$ \rho_{0,1} + \rho_{0,0} = \rho_0$$
+
+We now can use the transition probabilities of the individual locations to describe the full system of differential equations.
+
+$$  \frac{\mathrm{d}\rho_{1,1}}{\mathrm{d}t} = 2 \rho_{1,0}w_{0,1} - 2 \rho_{1,1} w_{1,0}  $$
+
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = \rho_{0}w_{0,1} - \rho_{1}w_{1,0}   $$
+
+$$  \frac{\mathrm{d}\rho_{1,0}}{\mathrm{d}t} = 2 \rho_{0,0}w_{0,1} + 2 \rho_{1,1}w_{1,0} - \rho_{1,0}w_{0,1}    $$
+
+$$  \frac{\mathrm{d}\rho_{0}}{\mathrm{d}t} = \rho_{1}w_{1,0} - \rho_{0}w_{0,1} $$
+
+$$  \frac{\mathrm{d}\rho_{0,0}}{\mathrm{d}t} = 2 \rho_{1,0}w_{1,0} - 2 \rho_{0,0}w_{0,1}  $$
+
+Due to the conservation equations, two of these are sufficient to fully describe the dynamics of the system.
+
+We chose to keep the first two, substitute and simplify
+<!--
+$$  \frac{\mathrm{d}\rho_{11}}{\mathrm{d}t} = 2 \rho_{10} \frac{r \rho_1^{1+\alpha} (b^* + (1 - b) f q_{0|1})) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a^*  + v q_{1|1})L^* (1 - p  q_{1|1})\rho_1^{q}}{1+(a^*  + v q_{1|1})h\rho_1^{1+q}}  $$
+
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = \rho_{0}\frac{r \rho_1^{1+\alpha} (b^* + (1 - b)  f q_{0|1})) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})})}{1-\rho_1} - \rho_{1}m - \frac{(a^*  + v q_{1|1})L^* (1 - p  q_{1|1})\rho_1^{1+q}}{1+(a^*  + v q_{1|1})h\rho_1^{1+q}}   $$
+
+they can be simplified to 
+-->
+
+$$  \frac{\mathrm{d}\rho_{11}}{\mathrm{d}t} = 2 \rho_{10} \frac{r \rho_1^{1+\alpha} (b^* + (1 - b)  f q_{0|1})) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a^*  + v q_{1|1})L^* (1 - p  q_{1|1}) \rho_1^{q}}{1+(a^*  + v q_{1|1})h)\rho_1^{1+q}}  $$
+
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b^* + (1 - b)  f q_{0|1})) (1-\frac{\rho_1}{K^* (1 - c q_{0|1})}) - \rho_{1}m - \frac{(a^*  + v q_{1|1})L^* (1 - p  q_{1|1})\rho_1^{1+q}}{1+(a^*  + v q_{1|1})h\rho_1^{1+q}}   $$
+
+<!--
+as a matter of fact the equation for $\frac{\mathrm{d}\rho_{1}}{\mathrm{d}t}$ will re-convert into the regular population equation. 
+-->
+Since formally $q_{i|j}$ is a conditional probability (given that one neighbor is in state $i$, what is the probability to find the other in state $j$) we can express it in terms of densities of pairs and ‘singletons,’ as
+
+$$ q_{i|j} = \frac{\rho_{ij}}{\rho_i} $$
+
+thus,
+
+$$ q_{0|1} = \frac{\rho_{10}}{\rho_0} =  \frac{\rho_{1} - \rho_{11}}{1 - \rho_1} $$
+
+$$ q_{1|1} = \frac{\rho_{11}}{\rho_1} $$
+<!--
+and substituting yields
+
+$$  \frac{\mathrm{d}\rho_{11} }{\mathrm{d}t} = 2 (\rho_1 - \rho_{11}) \frac{r \rho_1^{1+\alpha} (b^* + (1 - b)  f \frac{\rho_{1} - \rho_{11}}{1 - \rho_1} )) (1-\frac{\rho_1}{K^* (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a^*  + v \frac{\rho_{11}}{\rho_1})L^* (1 - p \frac{\rho_{11}}{\rho_1})\rho_1^{q}}{1+(a^*  + v \frac{\rho_{11}}{\rho_1})h\rho_1^{1+q}}  $$
+
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b^* + (1 - b) f  \frac{\rho_{1} - \rho_{11}}{1 - \rho_1}) (1-\frac{\rho_1}{K^* (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})}) - \rho_{1}m -  \frac{(a^*  + v \frac{\rho_{11}}{\rho_1})L^* (1 - p \frac{\rho_{11}}{\rho_1})\rho_1^{1+q}}{1+(a^*  + v \frac{\rho_{11}}{\rho_1})h\rho_1^{1+q}}   $$
+-->
+
 ## mean-field model
+
+The change in cover derives from the growth and mortality terms
+
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = \rho_{0}w_{0,1} - \rho_{1}w_{1,0}   $$
+
+Thus, we can substitute $C(\rho) = w_{1,0}$ and $G(\rho) = w_{0,1}$ and obtain
+
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b + (1 - b)  f \rho_1)) (1-\frac{\rho_1}{K(1 - c \rho_1}) - \rho_{1}m - \frac{(a  + v \rho_1)(1 - p \rho_1)L \rho_1^{1+q}}{1+(a  + v \rho_1)h\rho_1^{1+q}}  $$
+
+## Online supplementary material
+
+### Github
+
+The ODEs are solved using R package deSolve [@soetart10] applying the algorithm Dormand-Prince 4(5) of local order 7 (calling `deSolve::ode(..., method = "ode45")`).
+The full code for calling single simulations as well as replicated simulations for bifurcation graphs or attractor plots is provided as a functional R package available at https://cascade-wp6.github.com/livestock (v1.0, doi: zenodo/xxxx).
+
+### cellular automata implementation
+
+It is straightforward to implement the mechanisms described above in the syntax of a cellular automata model, where the probabilities of transition can not only depend on the global cover $\rho_1$ but also on the local cover in the direct neighborhood of the cell $\nu_1$ (closest 4 cells, von-Neumann-neighborhood).
+Cellular automata simulations are available via the R package caspr (https://fdschneider.github.com/caspr) or as an implementation in NetLogo (https://fdschneider.github.com/livestock/docs/livestockmodel.nlogo).
 

### 2016-07-21 manuscript/manuscript.Rmd
update manuscript.Rmd: minor edits

@@
 ---
-title: "the implications of local interactions on the resilience of arid rangelands"
+title: "The implications of local interactions on the resilience of arid rangelands"
 author: "Florian D. Schneider"
@@
 plot_meanfield(livestock, new = TRUE, parms = p)
-plot_bifurcation(livestock, over = "L", xrange = c(0,2), res = 201, parms = p) -> out
+plot_bifurcation(livestock, over = "L", xrange = c(0,3), res = 201, parms = p) -> out
 ```
 
-- This description assumes that vegetation cover is a sufficient descriptor for the feedbacks occuring on mortality and growth of vegetation, resulting in non-linear functions of vegetation cover.
+- This assumes that vegetation cover is a sufficient descriptor for the feedbacks occuring on mortality and growth of vegetation, resulting in non-linear functions of vegetation cover.
 Other models departed from this basic model and integrated more elaborate mechanisms that took other constants into account such as water runoff etc.
@@
 ```{r,echo = FALSE, fig.cap = "3D graph of growth vs. mortality", rgl=TRUE, results = FALSE, warning = FALSE, message = FALSE, error = FALSE}
-p <- set_parms(livestock$parms, set = list(b = 0.3, c = 0.4, f = 0.9, p = 0.99, v = 0.99))
-plot_pairapproximation3D(livestock, parms = p) -> out
+p <- set_parms(livestock$parms, set = list(b = 0.3, f = 0.9, p = 0.9 , v = 0.6))
+run <- sim_trajectories(livestock, parms = p)
+
+plot_pairapproximation3D(livestock, parms = p)
+
+layout(matrix(c(1,2,3), ncol = 3), widths = c(2,2,1))
+plot_pairapproximation(livestock, parms = p, new = TRUE, side = "rho")
+plot_pairapproximation(livestock, parms = p, new = TRUE, side = "q")
+plot_pairapproximation(livestock, parms = p, new = TRUE, side = "plain")
 ```
@@
 p <- set_parms(livestock$parms, set = list(b = 0.3, c = 0.2, f = 0.9, p = 0.99))
-plot_pairapproximation3D(livestock, parms = p) -> out
+
+#bifurcation plot
+
 ```
@@
 They thereby provide protected habitat to their direct neighborhood, or share the investments in those traits with their neighbors.
-Overall grazing mortality thus is reduced through an increase in handling time by associational resistance, $p$.
+Overall grazing mortality thus is reduced through a decrease in local livestock densities by associational resistance, $p$.
 
+<!--
 $$ a = a^{**} (1- p q_{1|1}) $$
@@
 Instead, we could define associational protection as a reduced impact due to a reduced local density of grazers.
+-->
 
@@
 
+<!--
 **Update 12.03.2015:** the effect is of less relevance at high cover, since the development of thorns is only induced if local pressure is already high.
@@
 **Update 26.03.2015:** Also, the formulations of local effects on mortality are unclear about their compatibility, since the attractant effect it is additive while the protective effect is multiplied.
-Thus, I have to specify-->
+Thus, I have to specify
 
 $$ a = (a^*  + v q_{1|1} ) (1- p q_{1|1}) $$ 
+-->
 
@@
 
+$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = \rho_{0}w_{0,1} - \rho_{1}w_{1,0}   $$
+
 $$  \frac{\mathrm{d}\rho_{1,1}}{\mathrm{d}t} = 2 \rho_{1,0}w_{0,1} - 2 \rho_{1,1} w_{1,0}  $$
-
-$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = \rho_{0}w_{0,1} - \rho_{1}w_{1,0}   $$
 
@@
 
+$$  \frac{\mathrm{d}\rho_{0,0}}{\mathrm{d}t} = 2 \rho_{1,0}w_{1,0} - 2 \rho_{0,0}w_{0,1}  $$
+
 $$  \frac{\mathrm{d}\rho_{0}}{\mathrm{d}t} = \rho_{1}w_{1,0} - \rho_{0}w_{0,1} $$
 
-$$  \frac{\mathrm{d}\rho_{0,0}}{\mathrm{d}t} = 2 \rho_{1,0}w_{1,0} - 2 \rho_{0,0}w_{0,1}  $$
+Due to the conservation equations, two of these are sufficient to fully describe the dynamics of the system. 
+We chose to keep the first two representing the density of vegetated cells $\rho_1$ and the density of pairs of vegetated cells $\rho_{1,1}$.
 
-Due to the conservation equations, two of these are sufficient to fully describe the dynamics of the system.
-
-We chose to keep the first two, substitute and simplify
 <!--
@@
 they can be simplified to
--->
 
@@
 
-<!--
 as a matter of fact the equation for $\frac{\mathrm{d}\rho_{1}}{\mathrm{d}t}$ will re-convert into the regular population equation. 

### 2016-11-08 manuscript/manuscript.Rmd
manuscript drafting

@@
 ---
-title: "The implications of local interactions on the resilience of arid rangelands"
+title: "Combining landscape and local scale to predict ecosystem resilience"
 author: "Florian D. Schneider"
@@
 
-r3dDefaults$userMatrix <- matrix(c(0.7849265,0.1177803,-0.6082912,0,0.008146744,0.9797189,0.2002101,0,0.6195351,-0.1621057,0.7680479,0,0,0,0,1), nrow = 4) 
+r3dDefaults$userMatrix <- matrix(c(0.9321498, 0.08271866, -0.3524974, 0, -0.3600706, 0.3140252, -0.878486, 0, 0.03802587, 0.9458043, 0.3225031, 0, 0, 0, 0, 1), nrow = 4) 
 knit_hooks$set(rgl = hook_rgl)
@@
 - The responsible mechanisms are feedbacks on the landscape scale, i.e. dependent on the 'global' density of vegetation and livestock, as well as global environmental parameters, such as aridity.
-- these studies had hughe implications for rangeland management and were well received in applied, fundamental and theoretical research. 
+- these studies had huge implications for rangeland management and were well received in applied, fundamental and theoretical research. 
    - applied concepts such as state and transition models, i.e. descriptions of landscapes as systems with multiple potential states and probabilities and conditions for the transition between those. 
@@
 arid shrublands?
-grasslands?).
-It relates to the study of successional states.
+grasslands?, refs).
+It introduced discontinuity and irreversibility into the study of successional states (refs).
+   - A change in climatic pressure or anthropogenic drivers may not only alter one state variable of the system, but multiple. 
    - theoretical studies were addressing mechanisms and behaviour of alternative stable states along gradients of pressure, where the rangeland system provided a good example.
-(Rietkerk, van de Koppel, Scheffer, ...) Most recently, the study of social-ecological systems addressed the catastrophic behaviour of ecosystems in the context of anthropogenic activities.
-(Carpenter, Schlüter)
+(ref Rietkerk, van de Koppel, Scheffer, ...) Most recently, the study of social-ecological systems addressed the catastrophic behaviour of ecosystems in the context of anthropogenic activities.
+(ref Carpenter, Schlüter) Conceptualisations like this require a systemic understanding of ecosystems, incorporating mechanistics such as feedback loops and nonlinearities.
 
-Figure 1 | Citations of Noy-Meir 1975 in publications of theoretical, applied and fundamental ecology journals.
+<!--Figure 1 | Citations of Noy-Meir 1975 in publications of theoretical, applied and fundamental ecology journals.
 
-<!-- Difficult to get these data.
+ Difficult to get these data.
 The paper is not on ISI or Pubmed, Google citations cannot be extracted except by hand.
@@
 
-- the framework proposed by Noy Meir 1975 looks at simple graphical analysis of the growth vs mortality of vegetation.
-This simple population dynamic approach identifies the stable and unstable equilibria of an ecosystem/population at the intersections of the growth and mortality functions over vegetation cover.
-- We observe net increase in vegetation cover, where the growth rate is higher than mortality and net decline in vegetation cover in the opposite case.
-The alternative steady states arise in cases where the two functions intersect more than once.
-(equations of growth and mortality as functions of vegetation cover $$ $$)
--  growth function depends on environmental quality and the mortality function depends on livestock densities.
-If observing the alternative steady states /attractors along a gradient of pressure, this can be plotted as a bifurcation graph of alternative steady states.
-Illustrating the thresholds of pressure and vegetation cover that will lead to an ecosystem shift.
+- the framework proposed by Noy-Meir in 1975 looks at simple graphical analysis of the growth vs mortality of vegetation.
+This population dynamic approach identifies the stable and unstable equilibria of an ecosystem/population at the intersections of the growth and mortality functions along the state variable of vegetation cover [@noymeir75].
+- Where the growth rate is higher than mortality, we would observe net increase in vegetation cover.
+A net decline in vegetation cover is expected if the opposite is the case.
+Equilibrium is reached where growth and mortality cancel each other out.
+These equilibria are stable, if the state variable after a small disturbance would return to their original value (Fig 1a, black points).
+- Alternative stable states arise in cases where the two functions of growth and mortality intersect more than once.
+Noy-Meir showed that this is regularly the case in grazing systems, due to the intersection of a saturating feeding curve with a logistic growth rate.
+In those cases, the vegetated state can transition into a permanent and irreversible degradet state close to zero vegetation cover.
+-  In this framework, the growth function depends on environmental quality and is strictly limited by a carrying capacity (logistic growth, fig 1a, green curve).
+The mortality function is based on the assumption of saturating feeding rates with increasing resource density, so called functional response, and scales with livestock densities (Fig 1a, black curve).
+Those curves intersect three times along the state variable: First, they both take value 0 at vegetation cover 0, which marks the degraded state of the grazing system; second, they intersect in the rising branch at intermediate vegetation cover; third, they intersect at the plateau of the mortality function and the lowering part of the growth function.
+While the second equilibrium is unstable (i.e.
+a disturbance would cause the system to move away), the third intersection is the stable quilibrium of the vegetated state, a.k.a.
+its attractor.
+
+- If observing the alternative steady states /attractors along a gradient of pressure, e.g.
+Livestock density, this can be plotted as a bifurcation graph of alternative steady states.
+Illustrating the thresholds of pressure and vegetation cover that will lead to an ecosystem shift (Fig 1b).
+- This graphical visualisation allows the identification of two kinds of critical thresholds.
+First, the distance of the stable attractors to the unstable one represents the supported disturbance amplitude.
+Second, the distance of the current value of pressure, to the point where the vegetated state perseveres, signifies the degree of pressure that the system can additionally bear before collapsing.
+In practice, both quantities must be regarded, since in the face of stochastic disturbance events like droughts or erosion, they are composites of a trade-off between security and productivity of the rangeland system (ref).
 
@@
 p <- set_parms(livestock$parms, set = list(b = 0.5,  L = 0.6))
-plot_meanfield(livestock, new = TRUE, parms = p)
+
+plot_meanfield(livestock, parms = set_parms(p, set = list(L=1.9)), new = TRUE, col =  c("#888888","#009933"))
+plot_meanfield(livestock, new = FALSE, parms = p)
+#mtext(paste("L =", p$L), line = -1, at = 0.1, adj = 0) 
+text(0.5,0.03, paste("L =", p$L), adj = 0)
+arrows(0.5,mortality(ini_rho(0.5), parms = set_parms(p, set = list(L=0.6))),0.5,mortality(ini_rho(0.5), parms = set_parms(p, set = list(L=1.9))) , length = 0.05)
+text(0.55,0.15, paste("L = 1.9"), adj = 0)
+
 plot_bifurcation(livestock, over = "L", xrange = c(0,3), res = 201, parms = p) -> out
+
+arrows(0.6,-0.35,1.9,-0.35, length = 0.05, xpd = TRUE)
+
 ```
 
-- This assumes that vegetation cover is a sufficient descriptor for the feedbacks occuring on mortality and growth of vegetation, resulting in non-linear functions of vegetation cover.
-Other models departed from this basic model and integrated more elaborate mechanisms that took other constants into account such as water runoff etc.
-(Rietkerk, van de Koppel?), however they still used mean-field approximations i.e.
+- In that sense, the state variable 'vegetation cover' is a promising indicator for the risk of degradation.
+It is easy to measure over time and its temporal variation can be quantified into defined early warning signs [@dakos].
+
+## adding spatial information to models
+
+- This assumes, however, that vegetation cover is a sufficient descriptor for the feedbacks occuring on mortality and growth of vegetation, resulting in non-linear functions of vegetation cover.
+Other models departed from this basic concept and integrated more elaborate mechanisms that took further constants into account such as water runoff etc.
+(Rietkerk, van de Koppel).
+Most of these modells still used mean-field approximations, i.e.
 the assumption that the local cover is at all times and in all locations equal to the global cover.
-- The recent studies have shown that vegetation structure is a major factor in understanding the stability of arid and semi-arid vegetation.
-- To integrate spatial structure into the framework of Noy Meir, we define the both growth and mortality as functions of vegetation cover AND local vegetation density / clustering coefficient.
-This provides information about the strength of local interactions and alters growth and mortality.
-- (equations for growth and mortality as  functions of global and local cover)
-- The visualisation of this  case is difficult because it adds a third dimension to the graph of Noy-Meir: in the z-axis we visualise the local cover of vegetation, i.e.
-the average cover in the direct neighborhood of a plant $\rho_{1|1}$.
-However, the attractors reduce to a single point.
+- Indeed, ecological mechanisms behind local feedbacks on mortality and growth have been described in empirical and theoretical studies: local facilitation, and competition constitute positive and negative effects on plant establishment  and growth.
+Attractant decoy and Associational resistance are known are positive and negative effects on plant mortality.
+Thus, depending on the importance of these factors, the proximity to other plants alters growth and mortality via these ecological mechanisms.
+- Recent studies have shown that explicit vegetation structure is a major factor in understanding the stability of arid and semi-arid vegetation (Kéfi 2008; Schneider 2016, ...). 
+- To integrate spatial structure into the framework of Noy-Meir, we define both growth and mortality as functions of vegetation cover *and* local vegetation density, i.e. the clustering coefficient. 
+- This adds a layer of information about the strength of local interactions in response to local vegetation cover.
 
-```{r,echo = FALSE, fig.cap = "3D graph of growth vs. mortality", rgl=TRUE, results = FALSE, warning = FALSE, message = FALSE, error = FALSE}
+- The visualisation of this case is difficult because it adds a third dimension to the graph of Noy-Meir: in the z-axis we visualise the average local cover of vegetation in the vicinity of a plant, i.e.
+the conditional probability to find plants in the adjacent localities given that a locality is vegetated, $q_{1|1}$.
+- The functions of growth and mortality now describe planes in this three dimensional volume (Fig 3a).
+The threshold at which vegetation cover either declines or increases, i.e.
+the unstable equilibrium, now depends on global cover as well as clustering intensity of the vegetation.
+- Nevertheless, the steady state attractors, i.e. the cases where mortality and growth equal each other out, reduce to points in this three dimensional space. (needs further explanation?)
+
+```{r,echo = FALSE, fig.cap = "3D graph of growth vs. mortality. Animated version is available online at ... .", rgl=TRUE, results = FALSE, warning = FALSE, message = FALSE, error = FALSE}
 p <- set_parms(livestock$parms, set = list(b = 0.3, f = 0.9, p = 0.9 , v = 0.6))
@@
 
+mfrow3d(1, 2)
   plot_pairapproximation3D(livestock, parms = p)
 
-layout(matrix(c(1,2,3), ncol = 3), widths = c(2,2,1))
-plot_pairapproximation(livestock, parms = p, new = TRUE, side = "rho")
-plot_pairapproximation(livestock, parms = p, new = TRUE, side = "q")
-plot_pairapproximation(livestock, parms = p, new = TRUE, side = "plain")
+bgplot3d({
+  plot.new()
+  mtext("a)", adj = "0", cex = 1)
+  }
+)
+
+plot_bifurcation3D(livestock, parms = p, res = 21)
+
+bgplot3d({
+  plot.new()
+  mtext("b)", adj = "0", cex = 1)
+  }
+)
 ```
 
-Figure 3 |
-
-- If illustrated along the gradient of pressure, a bifurcation graph visualises the location of the attractor at any given parameter combination.
+- If illustrated along the gradient of pressure, a bifurcation graph visualises the location of the attractor at any given parameter combination (Fig 3b).
 In addition to the vegetation cover, which is visualised on the x axis, we require the local vegetation cover on the z axis to provide a full description of the steady state ecosystem.
 
-Figure 4 | bifurcation along a) livestock density and b) local cover
+- To simplify visualisation, we break those graphs apart into a representation that can be compared directly to the visualisation of the Noy-Meir paper (Fig. 1): 
+- A first pair of graphs illustrates the steady state attractor and unstable equilibria of the system at a given level of pressure.
+Its first panel illustrates the trajectories of growth and mortality along a gradient of vegetation cover $\rho$.
+The second panel shows the relationship between global and local vegetation cover, i.e.
+a top-view of the cube in figure 2.
+- A second pair of graphs illustrates the bifurcation of attractors along a gradient of pressure.
+In this case, we focus on a gradient of livestock density $\L$, comparable to Fig 1b.
+The first panel illustrates the steady state vegetation cover, $\rho_{1}$ of the system along the gradient of pressure.
+The second panel puts this equilibrium vegetation cover in relation to the apparent vegetation structure, i.e.
+the quantity of clustering of the vegetation in space.
+This quantity is reportet as the clustering coefficient, $c$, which is defined as the ratio between average vegetation cover found in the neighborhood of plants, $\q_{1|1}$, to the global vegetation cover $\rho_1$.
+A value of 1 would reflect random clustering, whereas values well above 1 signify a patchy vegetation landscape.
 
-```{r, echo = FALSE, fig.cap = "3D graph of growth vs. mortality", rgl=TRUE, results = FALSE, warning = FALSE, message = FALSE, error = FALSE}
-p <- set_parms(livestock$parms, set = list(b = 0.3, c = 0.2, f = 0.9, p = 0.99))
+```{r}
 
-#bifurcation plot
+layout(matrix(c(1,2,3,4), ncol = 2, byrow = FALSE), heights = c(1,1))
+par(mar = c(0,5,3,1), las = 1)
+plot_pairapproximation(livestock, parms = p, new = TRUE, side = "rho")
+#plot_pairapproximation(livestock, parms = p, new = TRUE, side = "q")
+par(mar = c(4,5,0.55,1))
+plot_pairapproximation(livestock, parms = p, new = TRUE, side = "plain")
+par(mar = c(0,5,3,1))
+plot_bifurcation(livestock, over = "L", xrange = c(0,4), parms = p)
+par(mar = c(4,5,0.55,1))
+plot_bifurcation(livestock, over = "L", type = "plain", xrange = c(0,4), parms = p)
 
@@
 
+Interpretation of these graphs is as follows:  
+- The lines in the left panels represent trajectories or growth and mortality towards the attractor under given intial conditions, that is starting from low to high global cover and minimal vs.
+maximal clustering.
+The interaction of local cover  and total vegetation cover are visualised in the lower panel.
+-
+
 # the role of local interactions for transitions
 
-- To investigate the role of local interactions between plants, we subsequently implement these mechanisms into the model.
+- To investigate the role of local interactions between plants, we subsequently implement spatially constrained mechanisms into the model.
 We investigate their impact on the ecosystem state along gradients of livestock pressure (i.e.
